---
title: "Assemble master dataframe"
author: "Jill Ashey"
date: "2025-09-09"
output: html_document
---

Assemble master csv for phys metrics (Pmax, Pgross, Ik, Rd, alpha, Fv/Fm, net calc)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
```

Read in pi curve metrics 
```{r}
pi_metrics_sept2023 <- read.csv("../output/pi_curves/Sept2023/pi_curve_metrics_Sept2023.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)
pi_metrics_jan2024 <- read.csv("../output/pi_curves/Jan2024/pi_curve_metrics_Jan2024.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)
pi_metrics_may2024 <- read.csv("../output/pi_curves/May2024/pi_curve_metrics_May2024.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)
pi_metrics_oct2024 <- read.csv("../output/pi_curves/Oct2024/pi_curve_metrics_Oct2024.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)
pi_metrics_jan2025 <- read.csv("../output/pi_curves/Jan2025/pi_curve_metrics_Jan2025.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)
pi_metrics_may2025 <- read.csv("../output/pi_curves/May2025/pi_curve_metrics_May2025.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)

pi_metrics <- rbind(pi_metrics_sept2023, pi_metrics_jan2024, pi_metrics_may2024, pi_metrics_oct2024, pi_metrics_jan2025, pi_metrics_may2025)
```

Pivot pi metric data longer 
```{r}
pi_metrics_long <- pi_metrics %>%
  pivot_longer(cols = -c(Colony_ID, Species, Date, Phenotype),
               names_to = "Parameter",
               values_to = "Value")
```


Read in Fv/Fm data 
```{r}
pam <- read_csv("../data/Master_datasheet_pam.csv")
colnames(pam)[1] <- "Colony_ID" 
```

Pivot Fv/Fm data longer and tidy
```{r}
pam_long <- pam %>%
  pivot_longer(
    cols = starts_with("Fv/Fm_"),  # all the Fv/Fm_* columns
    names_to = "Date",
    values_to = "Fv_Fm") %>%
  mutate(
    Date = sub("Fv/Fm_", "", Date))  # remove the Fv/Fm_ prefix
  
pam_long_tidy <- pam_long %>%
  pivot_longer(cols = Fv_Fm, 
               names_to = "Parameter", 
               values_to = "Value")

# Ensure spp names are correct
pam_long_tidy$Species <- ifelse(pam_long_tidy$Species == "Mcap", "Montipora capitata",
                        ifelse(pam_long_tidy$Species == "Pcomp", "Porites compressa",
                               pam_long_tidy$Species))
```

Read in calc data 
```{r}
calc <- read.csv("../output/calcification/Calcification_data_normalized.csv") %>%
  filter(Light == "Light") %>%
  select(Colony_ID, Date, Species, Phenotype, Net_Calcification_cm2)
```

Only selecting the light calc data. 

Pivot calc data longer 
```{r}
calc_long <- calc %>%
  pivot_longer(
    cols = Net_Calcification_cm2,
    names_to = "Parameter",
    values_to = "Value"
  )
```

Bind all long dfs together 
```{r}
dfs <- list(calc_long, pam_long_tidy, pi_metrics_long)
combined_df <- do.call(rbind, dfs)
```

Remove blanks
```{r}
combined_df <- combined_df %>%
  filter(!Colony_ID == "blank")
```

Save as csv 
```{r}
write.csv(combined_df, "../output/pairs_calc_pam_pi_long.csv")
```

Keep long or wide?? what is better for downstream data analysis?