---
title: "Assemble master dataframe"
author: "Jill Ashey"
date: "2025-09-09"
output: html_document
---

Assemble master csv for phys metrics (Pmax, Pgross, Ik, Rd, alpha, Fv/Fm, net calc)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(patchwork)
```

Read in pi curve metrics 
```{r}
pi_metrics_sept2023 <- read.csv("../output/pi_curves/Sept2023/pi_curve_metrics_Sept2023.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)
pi_metrics_jan2024 <- read.csv("../output/pi_curves/Jan2024/pi_curve_metrics_Jan2024.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)
pi_metrics_may2024 <- read.csv("../output/pi_curves/May2024/pi_curve_metrics_May2024.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)
pi_metrics_oct2024 <- read.csv("../output/pi_curves/Oct2024/pi_curve_metrics_Oct2024.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)
pi_metrics_jan2025 <- read.csv("../output/pi_curves/Jan2025/pi_curve_metrics_Jan2025.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)
pi_metrics_may2025 <- read.csv("../output/pi_curves/May2025/pi_curve_metrics_May2025.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)
pi_metrics_oct2025 <- read.csv("../output/pi_curves/Oct2025/pi_curve_metrics_Oct2025.csv") %>%
  select(Colony_ID, Pmax, alpha, Rd, Ik, Pgross, Date, Species, Phenotype)

pi_metrics <- rbind(pi_metrics_sept2023, pi_metrics_jan2024, pi_metrics_may2024, pi_metrics_oct2024, pi_metrics_jan2025, pi_metrics_may2025, pi_metrics_oct2025)
```

Pivot pi metric data longer 
```{r}
pi_metrics_long <- pi_metrics %>%
  pivot_longer(cols = -c(Colony_ID, Species, Date, Phenotype),
               names_to = "Parameter",
               values_to = "Value")
```

Read in PAM data 
```{r}
pam <- read_csv("../data/Master_datasheet_pam.csv")
```

Tidy pam data
```{r}
pam_long <- pam %>%
  rename(Value = Fv_Fm) %>%       # rename the PAM measurement to "Value"
  mutate(Parameter = "Fv/Fm") %>% # add a column for the parameter name
  select(Colony_ID, Date, Species, Phenotype, Parameter, Value)
```

Read in calc data 
```{r}
calc <- read.csv("../output/calcification/Calcification_data_normalized.csv") %>%
  #filter(Light == "Light") %>%
  select(Colony_ID, Date, Species, Phenotype, Net_Calcification_cm2, Light)
```

Make new col for Light/Dark calc data 
```{r}
calc_wide <- calc %>%
  mutate(New_Col = paste0("Net_Calcification_cm2_", Light)) %>%
  pivot_wider(names_from = New_Col, values_from = Net_Calcification_cm2)
```

Pivot calc data longer 
```{r}
calc_long <- calc %>%
  mutate(
    Parameter = paste0("Net_Calcification_cm2_", Light),
    Value = Net_Calcification_cm2
  ) %>%
  select(Colony_ID, Date, Species, Phenotype, Parameter, Value)
```

Bind all long dfs together 
```{r}
dfs <- list(calc_long, pam_long, pi_metrics_long)
combined_df <- do.call(rbind, dfs)
```

Remove blanks
```{r}
combined_df <- combined_df %>%
  filter(!Colony_ID == "blank")
```

Save as csv 
```{r}
write.csv(combined_df, "../output/pairs_calc_pam_pi_long.csv")
```

Keep long or wide?? what is better for downstream data analysis?

### Plots 

Add Timepoint as a column
```{r}
combined_df <- combined_df %>%
  mutate(Timepoint = case_when(
    grepl("2023-04", Date) ~ "Apr23",
    grepl("Apr23", Date) ~ "Apr23",
    grepl("2023-09", Date) ~ "Sept23",
    grepl("September_2023", Date) ~ "Sept23",
    grepl("2024-01", Date) ~ "Jan24",
    grepl("January_2024", Date) ~ "Jan24",
    grepl("2024-05", Date) ~ "May24",
    grepl("May_2024", Date) ~ "May24",
    grepl("2024-10", Date) ~ "Oct24",
    grepl("October_2024", Date) ~ "Oct24",
    grepl("2025-01", Date) ~ "Jan25",
    grepl("January_2025", Date) ~ "Jan25",
    grepl("2025-05", Date) ~ "May25",
    grepl("May_2025", Date) ~ "May25",
    grepl("2025-10", Date) ~ "Oct25",
    grepl("May_2025", Date) ~ "Oct25",
    TRUE ~ NA_character_))
```

Calculate mean, sem, and sd for light only
```{r}
combined_summary <- combined_df %>%
  group_by(Timepoint, Phenotype, Species, Parameter) %>%
  summarise(
    mean = mean(Value, na.rm = TRUE),
    sd = sd(Value, na.rm = T),
    sem = sd(Value, na.rm = T) / sqrt(length(Value)),
    .groups = "drop") %>%
  drop_na() %>%
  filter(!Timepoint == "Jan25") %>% # remove Jan25 for now 
  filter(!Timepoint == "Apr23")
```

Plot!
```{r}
# Make sure factors are set
combined_summary <- combined_summary %>%
  mutate(
    Species   = factor(Species),
    Parameter = factor(Parameter),
    Phenotype = factor(Phenotype),
    Timepoint = factor(Timepoint, levels = unique(Timepoint))
  )

# Set levels for timepoint 
combined_summary$Timepoint <- factor(combined_summary$Timepoint, levels = c(#"Apr23",
                                                                        "Sept23",
                                                                        "Jan24", 
                                                                        "May24", 
                                                                        "Oct24", 
                                                                        #"Jan25", 
                                                                        "May25",
                                                                        "Oct25"))

# Custom y-axis labels
y_labels <- c(
  "Net_Calcification_cm2_Light" = "Light net calcification (cm\u00B2 h\u207B\u00B9)",
  "Net_Calcification_cm2_Dark"  = "Dark net calcification (cm\u00B2 h\u207B\u00B9)",
  "Fv_Fm"                        = "Maximum quantum yield (Fv/Fm)",
  "Ik"                           = "Ik (\u03BCmol photons m\u207B\u00B2 s\u207B\u00B9)",
  "Pgross"                        = "Gross photosynthesis (\u03BCmol O\u2082 cm\u207B\u00B2 h\u207B\u00B9)",
  "Pmax"                         = "Maximum photosynthesis (\u03BCmol O\u2082 cm\u207B\u00B2 h\u207B\u00B9)",
  "Rd"                            = "Dark respiration (\u03BCmol O\u2082 cm\u207B\u00B2 h\u207B\u00B9)",
  "alpha"                         = "Alpha"
)

# Param levels 
param_levels <- c(
  "Net_Calcification_cm2_Light",
  "Net_Calcification_cm2_Dark",
  "Fv_Fm",
  "alpha",
  "Ik",
  "Pmax",
  "Pgross",
  "Rd")

# combined_summary <- combined_summary %>%
#   mutate(Parameter = factor(Parameter, levels = param_levels)) %>%
#   mutate(Parameter = ifelse(Parameter %in% c("Fv/Fm", "FvFm"), "Fv_Fm", Parameter)) %>%
#   filter(!is.na(Parameter)) %>%
#   mutate(Parameter = factor(Parameter, levels = param_levels))

# Shading for times of year w/ highest temps
shade_df <- data.frame(
  Timepoint = c("Sept23", "Oct24", "Oct25"),
  xmin = c(1 - 0.5, 4 - 0.5, 6 - 0.5),  # factor positions - 0.5 for left edge
  xmax = c(1 + 0.5, 4 + 0.5, 6 + 0.5))   # factor positions + 0.5 for right edge

# Function to generate plot
plot_parameter <- function(param_name, show_legend = FALSE) {
  ggplot(filter(combined_summary, Parameter == param_name),
         aes(x = Timepoint, y = mean, color = Phenotype, group = Phenotype)) +
    geom_rect(data = shade_df,
              aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
              inherit.aes = FALSE, fill = "grey80", alpha = 0.3) +
    geom_line(linewidth = 2) +
    geom_point(size = 5) +
    geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem),
                  width = 0.5, linewidth = 1.5) +  # increase error bar thickness
    facet_wrap(~ Species, scales = "fixed") +
    scale_color_manual(values = c("Resistant" = "#4b294fff",
                                  "Susceptible" = "#f3a17dff")) +
    theme_classic() +
    labs(x = "", y = y_labels[param_name], color = "Phenotype") +
    theme(
      panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
      strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
      strip.text = element_text(face = "italic", size = 22),
      legend.position = if (show_legend) "right" else "none",  # only first plot
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 18, color = "black"),
      axis.text.y = element_text(size = 18, color = "black"),
      axis.title.y = element_text(size = 20)
    )
}

# Generate plots: show legend only in the first one
# plots <- lapply(seq_along(param_levels), function(i) {
#   plot_parameter(param_levels[i], show_legend = (i == 1))
# })
plots <- lapply(seq_along(param_levels), function(i) {
  plot_parameter(param_levels[i], show_legend = (i == 1))
})

final_plot <- wrap_plots(plots, ncol = 4, guides = "collect") +
  plot_annotation(tag_levels = "A") & 
  theme(
    plot.tag = element_text(size = 28, face = "bold"),
    legend.position = "right",
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18, face = "bold")
  )

# Save 
ggsave("../output/pairs_calc_pam_pi.pdf", final_plot, width = 30, height = 18, units = "in")
ggsave("../output/pairs_calc_pam_pi.png", final_plot, width = 30, height = 18, units = "in", dpi = 300)
```

Getting errors when trying to facet...need to redo this chunk.










```{r}
selected_params <- c(
  "Net_Calcification_cm2_Light",
  "Fv_Fm",
  "Ik",
  "Pgross",
  "Pmax",
  "alpha"
)

summary_six <- combined_summary %>%
  filter(Parameter %in% selected_params) %>%
  mutate(Parameter = factor(Parameter, levels = selected_params))

library(ggplot2)

plot_six <- ggplot(summary_six, aes(x = Timepoint, y = mean, color = Phenotype, group = Phenotype)) +
  geom_line(linewidth = 1.5) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = 0.2) +
  facet_grid(Parameter ~ Species, scales = "free_y", switch = "y") +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  labs(x = "Timepoint", y = NULL, color = "Phenotype") +
  theme_classic(base_size = 16) +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "italic", size = 16),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, color = "black"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    legend.position = "right"
  ); plot_six



```

```{r}


# Custom y-axis labels
y_labels <- c(
  "Net_Calcification_cm2_Light" = "Light net calcification (cm\u00B2 h\u207B\u00B9)",
  "Net_Calcification_cm2_Dark"  = "Dark net calcification (cm\u00B2 h\u207B\u00B9)",
  "Fv_Fm"                        = "Maximum quantum yield (Fv/Fm)",
  "Ik"                           = "Ik (\u03BCmol photons m\u207B\u00B2 s\u207B\u00B9)",
  "Pgross"                        = "Gross photosynthesis (\u03BCmol O\u2082 cm\u207B\u00B2 h\u207B\u00B9)",
  "Pmax"                         = "Maximum photosynthesis (\u03BCmol O\u2082 cm\u207B\u00B2 h\u207B\u00B9)",
  "Rd"                            = "Dark respiration (\u03BCmol O\u2082 cm\u207B\u00B2 h\u207B\u00B9)",
  "alpha"                         = "Alpha"
)

# Param levels 
selected_params <- c(
  "Net_Calcification_cm2_Light",
  "Net_Calcification_cm2_Dark",
  "Fv_Fm",
  "alpha",
  "Ik",
  "Pmax",
  "Pgross",
  "Rd")

library(ggplot2)
selected_params <- names(y_labels)
plot_all <- ggplot(
    combined_summary %>% filter(Parameter %in% selected_params),
    aes(x = Timepoint, y = mean, color = Phenotype, group = Phenotype)
  ) +
  geom_rect(
    data = shade_df,
    aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
    inherit.aes = FALSE,
    fill = "grey80", alpha = 0.3
  ) +
  geom_line(linewidth = 2) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = 0.5, linewidth = 1.5) +
  facet_grid(Parameter ~ Species, scales = "free_y",
             labeller = labeller(Parameter = y_labels)) +
  scale_color_manual(values = c("Resistant" = "#4b294fff",
                                "Susceptible" = "#f3a17dff")) +
  theme_classic() +
  labs(x = "Timepoint", y = NULL, color = "Phenotype") +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 22),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 18, color = "black"),
    axis.text.y = element_text(size = 18, color = "black")
  ); plot_all

ggsave("../output/combined_parameters_faceted.pdf", plot_all, width = 30, height = 18, units = "in")
ggsave("../output/combined_parameters_faceted.png", plot_all, width = 30, height = 18, units = "in", dpi = 300)

```

```{r}
library(dplyr)
library(ggplot2)
library(patchwork)

# Shading for times of year with highest temperatures
shade_df <- data.frame(
  Timepoint = c("Sept23", "Oct24", "Oct25"),
  xmin = c(1 - 0.5, 4 - 0.5, 6 - 0.5),  # factor positions - 0.5 for left edge
  xmax = c(1 + 0.5, 4 + 0.5, 6 + 0.5)   # factor positions + 0.5 for right edge
)

# Custom y-axis labels
y_labels <- c(
  "Net_Calcification_cm2_Light"  = "Light net calcification (cm\u00B2 h\u207B\u00B9)",
  "Net_Calcification_cm2_Dark"   = "Dark net calcification (cm\u00B2 h\u207B\u00B9)",
  "Fv_Fm"                       = "Maximum quantum yield (Fv/Fm)",
  "Ik"                          = "Ik (\u03BCmol photons m\u207B\u00B2 s\u207B\u00B9)",
  "Pgross"                      = "Gross photosynthesis (\u03BCmol O\u2082 cm\u207B\u00B2 h\u207B\u00B9)",
  "Pmax"                        = "Maximum photosynthesis (\u03BCmol O\u2082 cm\u207B\u00B2 h\u207B\u00B9)",
  "Rd"                          = "Dark respiration (\u03BCmol O\u2082 cm\u207B\u00B2 h\u207B\u00B9)",
  "alpha"                       = "Alpha"
)

# Parameter levels in order
param_levels <- c(
  "Net_Calcification_cm2_Light",
  "Net_Calcification_cm2_Dark",
  "Fv_Fm",
  "alpha",
  "Ik",
  "Pmax",
  "Pgross",
  "Rd"
)

# Plotting function for a single parameter with droplevels to fix faceting error
plot_parameter <- function(param_name, show_legend = FALSE) {
  dat <- combined_summary %>%
    filter(Parameter == param_name) %>%
    droplevels()

  ggplot(dat, aes(x = Timepoint, y = mean, color = Phenotype, group = Phenotype)) +
    geom_rect(
      data = shade_df,
      aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
      inherit.aes = FALSE,
      fill = "grey80",
      alpha = 0.3
    ) +
    geom_line(size = 1) +
    geom_point(size = 2.5) +
    geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = 0.3, size = 0.9) +
    facet_wrap(~ Species, scales = "fixed") +
    scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
    labs(
      x = "",
      y = y_labels[param_name],
      color = if (show_legend) "Phenotype" else NULL
    ) +
    theme_classic(base_size = 14) +
    theme(
      panel.border = element_rect(color = "black", fill = NA, size = 1),
      strip.background = element_rect(fill = "white", color = "black", size = 1),
      strip.text = element_text(face = "italic", size = 16),
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 12, color = "black"),
      axis.text.y = element_text(size = 12, color = "black"),
      axis.title.y = element_text(size = 14),
      legend.position = if (show_legend) "right" else "none"
    )
}

# Generate a list of plots, showing legend only on the first plot
plots <- lapply(seq_along(param_levels), function(i) {
  plot_parameter(param_levels[i], show_legend = (i == 1))
})

# Remove any NULL plots (in case of missing data)
plots <- plots[!sapply(plots, is.null)]

# Combine all plots with labels A, B, C, ... using patchwork
final_plot <- wrap_plots(plots, ncol = 4, guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(
    plot.tag = element_text(size = 18, face = "bold"),
    legend.position = "right",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14, face = "bold")
  )

# Save the combined figure
ggsave("../output/combined_parameters_labeled.pdf", final_plot, width = 30, height = 18, units = "in")
ggsave("../output/combined_parameters_labeled.png", final_plot, width = 30, height = 18, units = "in", dpi = 300)


```

this is still not working 
