---
title: "Net calcification - TCB"
author: "Jill Ashey"
date: "2025-09-09"
output: html_document
---

This script will calculate net calcification from the TCB corals (Sept 2023) and compare to the field corals from Sept 2023. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

Read in field calcification data to get information for blanks. This data did not have the raw alkalinity or blank values (Kristen might have them).
Read in data 
```{r}
calc <- read.csv("../data/Calcification_data_field.csv", header = T)
```

Add Timepoint as a column
```{r}
calc <- calc %>%
  mutate(Timepoint = case_when(
    grepl("2023-04", Date) ~ "Apr23",
    grepl("2023-09", Date) ~ "Sept23",
    grepl("2024-01", Date) ~ "Jan24",
    grepl("2024-05", Date) ~ "May24",
    grepl("2024-10", Date) ~ "Oct24",
    grepl("2025-01", Date) ~ "Jan25",
    grepl("2025-05", Date) ~ "May25",
    TRUE ~ NA_character_))
```

Calculate delta alkalinity 
```{r}
calc <- calc %>%
  mutate(Delta_TA_raw = Starting_alk_value - Alk_avg)
```

Calculate blanks 
```{r}
blank_delta_ta <- calc %>%
  filter(ColonyID == "blank" | Notes == "blank") %>%
  group_by(Timepoint) %>%
  summarise(mean_blank_delta_ta = mean(Delta_TA_raw, na.rm = TRUE))# %>%
  #pull(mean_blank_delta_ta)
print(blank_delta_ta)
```

Use the Oct24 blank as a substitute for the TCB blank. Oct24 blank is -5.3. 

Read in TCB data 
```{r}
calc_tcb <- read.csv("../data/Calcification_data_TCB.csv", header = T)
```

Calculate incubation time 
```{r}
calc_tcb <- calc_tcb %>%
  mutate(
    start = hm(Time_start),
    end   = hm(Time_end),
    elapsed_mins  = as.numeric(end - start, units = "mins"),
    elapsed_hours = elapsed_mins / 60)
```

Calculate delta alkalinity 
```{r}
calc_tcb <- calc_tcb %>%
  mutate(Delta_TA_raw = Starting_alk_value - Alk_avg)
```

Subtract blank and calculate net calcification
```{r}
calc_tcb <- calc_tcb %>%
  mutate(
    Delta_TA_corr = Delta_TA_raw - (-5.3),
    # Assuming seawater density 1.028 g/mL and volume in mL to L, and no coral weight (thus calc mg/L/hr)
    Net_Calcification = (Delta_TA_corr / 2) * (Sample_Volume_mL / 1000) * 1.028 / elapsed_hours) %>%
  filter(!is.na(Net_Calcification) & !is.nan(Net_Calcification))
```

Read in surface area information to normalize 
```{r}
sa_meta <- read_csv("../data/NSF_Career_SurfaceArea.csv") %>%
  filter(Project == "TCB")
```

Join calc and sa dfs
```{r}
# Make sure the join columns match
calc_tcb <- calc_tcb %>%
  rename(Colony_ID = ColonyID)

joined_tcb <- calc_tcb %>%
  left_join(sa_meta, by = c("Colony_ID"))
```

Normalize by SA
```{r}
joined_tcb <- joined_tcb %>%
  mutate(Net_Calcification_cm2 = Net_Calcification / Surface_Area_cm2)
```

Save as csv 
```{r}
write.csv(joined_tcb, "../output/calcification/Calcification_data_normalized_TCB.csv")
```

Read in field calc data from Sept23
```{r}
field <- read.csv("../output/calcification/Calcification_data_normalized.csv") %>%
  filter(Timepoint == "Sept23")
```

Subset specific cols 
```{r}
field_sub <- field %>%
  select(Date, Project, Colony_ID, Light, Species, Phenotype, Net_Calcification_cm2)

joined_tcb_sub <- joined_tcb %>%
  select(Date.x, Project.x, Colony_ID, Light, Species, Phenotype, Net_Calcification_cm2) %>%
  rename(Date = Date.x, Project = Project.x)
```

Bind dfs together 
```{r}
sept_calc <- rbind(field_sub, joined_tcb_sub)
```

Calculate mean, sem, and sd 
```{r}
sept_calc_summary <- sept_calc %>%
  group_by(Project, Phenotype, Species, Light) %>%
  summarise(
    mean_calc = mean(Net_Calcification_cm2, na.rm = TRUE),
    sd_calc = sd(Net_Calcification_cm2, na.rm = T),
    se_calc = sd(Net_Calcification_cm2, na.rm = T) / sqrt(length(Net_Calcification_cm2)),
    .groups = "drop") %>%
  drop_na()
```

Plot
```{r}
ggplot(sept_calc, aes(x = Project, y = Net_Calcification_cm2)) +
  geom_boxplot() +
  facet_grid(Species ~ Light) +
  theme_bw() +
  labs(x = "Project", y = "Net Calcification (cm2)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
