---
title: "Net calcification"
author: "Jill Ashey"
date: "2025-08-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

Read in data 
```{r}
calc <- read.csv("../data/Calcification_data.csv", header = T)
```

Calculate incubation time 
```{r}
calc <- calc %>%
  mutate(
    start = hm(Time_start),
    end   = hm(Time_end),
    elapsed_mins  = as.numeric(end - start, units = "mins"),
    elapsed_hours = elapsed_mins / 60)
```

Calculate delta alkalinity 
```{r}
calc <- calc %>%
  mutate(Delta_TA_raw = Starting_alk_value - Alk_avg)
```

Calculate blanks 
```{r}
blank_delta_ta <- calc %>%
  filter(ColonyID == "blank" | Notes == "blank") %>%
  group_by(Timepoint) %>%
  summarise(mean_blank_delta_ta = mean(Delta_TA_raw, na.rm = TRUE))# %>%
  #pull(mean_blank_delta_ta)

# Add manual substitutions for missing timepoints
blank_subs <- tibble(
  Timepoint = c("April_2023", "September_2023"),
  mean_blank_delta_ta = c(
    blank_delta_ta$mean_blank_delta_ta[blank_delta_ta$Timepoint == "May_2024"],
    blank_delta_ta$mean_blank_delta_ta[blank_delta_ta$Timepoint == "October_2024"]
  )
)

# Bind the substitutions into the blank dataset
blank_delta_ta <- bind_rows(blank_delta_ta, blank_subs)
```

Some of the timepoints (April and Sept 2023) do not have alk data for the blanks. Instead, use the May and Oct 2024 blanks instead. 



Subtract blank and calculate net calcification
```{r}
# Correct Delta_TA by subtracting blank
calc <- calc %>%
  left_join(blank_delta_ta, by = "Timepoint") 

calc <- calc %>%
  mutate(
    Delta_TA_corr = Delta_TA_raw - mean_blank_delta_ta,
    # Assuming seawater density 1.028 g/mL and volume in mL to L, and no coral weight (thus calc mg/L/hr)
    Net_Calcification = (Delta_TA_corr / 2) * (Sample_Volume_mL / 1000) * 1.028 / elapsed_hours) %>%
  filter(!is.na(Net_Calcification) & !is.nan(Net_Calcification))

```

Sub out names for timepoints 
```{r}
unique(calc$Timepoint)
calc$Timepoint <- gsub("April_2023", "Apr23", calc$Timepoint)
calc$Timepoint <- gsub("September_2023", "Sept23", calc$Timepoint)
calc$Timepoint <- gsub("January_2024", "Jan24", calc$Timepoint)
calc$Timepoint <- gsub("May_2024", "May24", calc$Timepoint)
calc$Timepoint <- gsub("October_2024", "Oct24", calc$Timepoint)
calc$Timepoint <- gsub("January_2025", "Jan25", calc$Timepoint)
calc$Timepoint <- gsub("May_2025", "May25", calc$Timepoint)
```

Read in sample metadata for surface area 
```{r}
sample_meta <- read_csv("../data/NSF_Career_SurfaceArea.csv")
```

Add timepoint col
```{r}
sample_meta <- sample_meta %>%
  mutate(Timepoint = case_when(
    grepl("2023-04", Date) ~ "Apr23",
    grepl("2023-09", Date) ~ "Sept23",
    grepl("2024-01", Date) ~ "Jan24",
    grepl("2024-05", Date) ~ "May24",
    grepl("2024-10", Date) ~ "Oct24",
    grepl("2025-01", Date) ~ "Jan25",
    grepl("2025-05", Date) ~ "May25",
    TRUE ~ NA_character_))
```

Subset specific cols 
```{r}
sample_meta_sub <- sample_meta %>%
  select(Colony_ID, Species, Phenotype, Surface_Area_cm2, Timepoint)
```

Join dfs 
```{r}
# Make sure the join columns match
calc2 <- calc %>%
  rename(Colony_ID = ColonyID)

# Join on Colony_ID and Timepoint
joined <- calc2 %>%
  left_join(sample_meta_sub, by = c("Colony_ID", "Timepoint"))
```

Normalize by SA
```{r}
joined <- joined %>%
  mutate(Net_Calcification_cm2 = Net_Calcification / Surface_Area_cm2)
```

Save as csv 
```{r}
write.csv(joined, "../output/calcification/Calcification_data_normalized.csv")
```

Calculate mean, sem, and sd 
```{r}
joined_summary <- joined %>%
  group_by(Timepoint, Phenotype, Species, Light) %>%
  summarise(
    mean_calc = mean(Net_Calcification_cm2, na.rm = TRUE),
    sd_calc = sd(Net_Calcification_cm2, na.rm = T),
    se_calc = sd(Net_Calcification_cm2, na.rm = T) / sqrt(length(Net_Calcification_cm2)),
    .groups = "drop") %>%
  drop_na() %>%
  filter(!Timepoint == "Jan25")
```

Plot! 
```{r}
# Set levels 
joined_summary$Timepoint <- factor(joined_summary$Timepoint, levels = c("Apr23",
                                                                        "Sept23",
                                                                        "Jan24", 
                                                                        "May24", 
                                                                        "Oct24", 
                                                                        #"Jan25", 
                                                                        "May25"))

ggplot(joined_summary, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species + Light) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Net calcification +/- SEM",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```

Decide what to do about the missing SAs for April 2023 and the weird ones for Jan 2025. Possible options: 

- Average SAs across timepoint for species + phenotype 
- Sub specific SAs (ie substitute SAs from May 2024 for April 2023)
- Wax dip--this will take the longest 
