---
title: "Net calcification"
author: "Jill Ashey"
date: "2025-08-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(emmeans)
library(lme4)
library(lmerTest)
library(lsmeans)
```

Read in data 
```{r}
calc <- read.csv("../data/Calcification_data_field.csv", header = T)
```

Add Timepoint as a column
```{r}
calc <- calc %>%
  mutate(Timepoint = case_when(
    grepl("2023-04", Date) ~ "Apr23",
    grepl("2023-09", Date) ~ "Sept23",
    grepl("2024-01", Date) ~ "Jan24",
    grepl("2024-05", Date) ~ "May24",
    grepl("2024-10", Date) ~ "Oct24",
    grepl("2025-01", Date) ~ "Jan25",
    grepl("2025-05", Date) ~ "May25",
    TRUE ~ NA_character_))
```

Calculate incubation time 
```{r}
calc <- calc %>%
  mutate(
    start = hm(Time_start),
    end   = hm(Time_end),
    elapsed_mins  = as.numeric(end - start, units = "mins"),
    elapsed_hours = elapsed_mins / 60)
```

Calculate delta alkalinity 
```{r}
calc <- calc %>%
  mutate(Delta_TA_raw = Starting_alk_value - Alk_avg)
```

Calculate blanks 
```{r}
blank_delta_ta <- calc %>%
  filter(ColonyID == "blank" | Notes == "blank") %>%
  group_by(Timepoint) %>%
  summarise(mean_blank_delta_ta = mean(Delta_TA_raw, na.rm = TRUE))# %>%
  #pull(mean_blank_delta_ta)

# Add manual substitutions for missing timepoints
blank_subs <- tibble(
  Timepoint = c("Apr23", "Sept23"),
  mean_blank_delta_ta = c(
    blank_delta_ta$mean_blank_delta_ta[blank_delta_ta$Timepoint == "May24"],
    blank_delta_ta$mean_blank_delta_ta[blank_delta_ta$Timepoint == "Oct24"]
  )
)

# Bind the substitutions into the blank dataset
blank_delta_ta <- bind_rows(blank_delta_ta, blank_subs)
```

Some of the timepoints (April and Sept 2023) do not have alk data for the blanks. Instead, use the May and Oct 2024 blanks instead. 

Subtract blank and calculate net calcification
```{r}
# Correct Delta_TA by subtracting blank
calc <- calc %>%
  left_join(blank_delta_ta, by = "Timepoint") 

calc <- calc %>%
  mutate(
    Delta_TA_corr = Delta_TA_raw - mean_blank_delta_ta,
    # Assuming seawater density 1.028 g/mL and volume in mL to L, and no coral weight (thus calc mg/L/hr)
    Net_Calcification = (Delta_TA_corr / 2) * (Sample_Volume_mL / 1000) * 1.028 / elapsed_hours) %>%
  filter(!is.na(Net_Calcification) & !is.nan(Net_Calcification))
```

Read in sample metadata for surface area 
```{r}
sample_meta <- read_csv("../data/NSF_Career_SurfaceArea.csv")
```

Add timepoint col
```{r}
sample_meta <- sample_meta %>%
  mutate(Timepoint = case_when(
    grepl("2023-04", Date) ~ "Apr23",
    grepl("2023-09", Date) ~ "Sept23",
    grepl("2024-01", Date) ~ "Jan24",
    grepl("2024-05", Date) ~ "May24",
    grepl("2024-10", Date) ~ "Oct24",
    grepl("2025-01", Date) ~ "Jan25",
    grepl("2025-05", Date) ~ "May25",
    TRUE ~ NA_character_))
```

Subset specific cols 
```{r}
sample_meta_sub <- sample_meta %>%
  select(Colony_ID, Species, Phenotype, Surface_Area_cm2, Timepoint)
```

Join dfs 
```{r}
# Make sure the join columns match
calc2 <- calc %>%
  rename(Colony_ID = ColonyID)

# Join on Colony_ID and Timepoint
joined <- calc2 %>%
  left_join(sample_meta_sub, by = c("Colony_ID", "Timepoint"))
```

Normalize by SA
```{r}
joined <- joined %>%
  mutate(Net_Calcification_cm2 = Net_Calcification / Surface_Area_cm2)
```

Save as csv 
```{r}
write.csv(joined, "../output/calcification/Calcification_data_normalized.csv")
```

Calculate mean, sem, and sd 
```{r}
joined_summary <- joined %>%
  group_by(Timepoint, Phenotype, Species, Light) %>%
  summarise(
    mean_calc = mean(Net_Calcification_cm2, na.rm = TRUE),
    sd_calc = sd(Net_Calcification_cm2, na.rm = T),
    se_calc = sd(Net_Calcification_cm2, na.rm = T) / sqrt(length(Net_Calcification_cm2)),
    .groups = "drop") %>%
  drop_na() %>%
  filter(!Timepoint == "Jan25") # remove Jan25 for now 
```

Plot! 
```{r}
# Set levels 
joined_summary$Timepoint <- factor(joined_summary$Timepoint, levels = c("Apr23",
                                                                        "Sept23",
                                                                        "Jan24", 
                                                                        "May24", 
                                                                        "Oct24", 
                                                                        #"Jan25", 
                                                                        "May25"))

ggplot(joined_summary, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  #facet_wrap(~ Species + Light) +
  facet_grid(rows = vars(Species), cols = vars(Light)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Net calcification +/- SEM",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```

## Stats! 

### Light

Subset by light 
```{r}
joined_light <- joined %>%
  filter(Light == "Light") %>%
  filter(!Timepoint == "Jan25") # remove Jan25 for now 
```

Look at normality 
```{r}
hist(joined_light$Net_Calcification_cm2)

# Q-Q Plot
qqnorm(joined_light$Net_Calcification_cm2)
qqline(joined_light$Net_Calcification_cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(joined_light$Net_Calcification_cm2)
```

Data is not normal, perform data transformation
```{r}
# Log10
joined_light$calc_log <- log10(joined_light$Net_Calcification_cm2 +1) # add 1 to avoid log10(0)
hist(joined_light$calc_log)
qqnorm(joined_light$calc_log)
qqline(joined_light$calc_log, col = "red")
shapiro.test(joined_light$calc_log)

# Sqrt
joined_light$calc_sqrt <- sqrt(joined_light$Net_Calcification_cm2)
hist(joined_light$calc_sqrt)
qqnorm(joined_light$calc_sqrt)
qqline(joined_light$calc_sqrt, col = "red")
shapiro.test(joined_light$calc_sqrt)

# Inverse 
joined_light$calc_inv <- 1 / (joined_light$Net_Calcification_cm2 +1) # add 1 to avoid division by zero
hist(joined_light$calc_inv)
qqnorm(joined_light$calc_inv)
qqline(joined_light$calc_inv, col = "red")
shapiro.test(joined_light$calc_inv)
```

Sqrt and inv produced normality. Proceed with sqrt. 
```{r}
mod_calc <- lmer(calc_sqrt ~ Species * Phenotype * Timepoint + (1 | Colony_ID),
                 data = joined_light,
                 REML = FALSE)
car::Anova(mod_calc, type = 3)
```

Look at posthoc comparisons 
```{r}
em_spp <- emmeans(mod_calc, pairwise ~ Species, adjust = "tukey")
summary(em_spp)

em_time <- emmeans(mod_calc, pairwise ~ Timepoint, adjust = "tukey")
summary(em_time)
```

Calculate mean, sem, and sd for light only
```{r}
joined_light_summary <- joined_light %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_calc = mean(Net_Calcification_cm2, na.rm = TRUE),
    sd_calc = sd(Net_Calcification_cm2, na.rm = T),
    se_calc = sd(Net_Calcification_cm2, na.rm = T) / sqrt(length(Net_Calcification_cm2)),
    .groups = "drop") %>%
  drop_na() %>%
  filter(!Timepoint == "Jan25") # remove Jan25 for now 
```

Plot! 
```{r}
# Set levels 
joined_light_summary$Timepoint <- factor(joined_light_summary$Timepoint, levels = c("Apr23",
                                                                        "Sept23",
                                                                        "Jan24", 
                                                                        "May24", 
                                                                        "Oct24", 
                                                                        #"Jan25", 
                                                                        "May25"))

ggplot(joined_light_summary, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species), cols = vars(Light)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Net calcification (Light)",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    legend.position = c(0.15, 0.83),  # adjust x, y for placement
    #legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```

### Dark

Subset by Dark 
```{r}
joined_dark <- joined %>%
  filter(Light == "Dark") %>%
  filter(!Timepoint == "Jan25") # remove Jan25 for now 
```

Look at normality 
```{r}
hist(joined_dark$Net_Calcification_cm2)

# Q-Q Plot
qqnorm(joined_dark$Net_Calcification_cm2)
qqline(joined_dark$Net_Calcification_cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(joined_dark$Net_Calcification_cm2)
```

Data is not normal, perform data transformation
```{r}
# Log10
joined_dark$calc_log <- log10(joined_dark$Net_Calcification_cm2 +1) # add 1 to avoid log10(0)
hist(joined_dark$calc_log)
qqnorm(joined_dark$calc_log)
qqline(joined_dark$calc_log, col = "red")
shapiro.test(joined_dark$calc_log)

# Sqrt
joined_dark$calc_sqrt <- sqrt(joined_dark$Net_Calcification_cm2)
hist(joined_dark$calc_sqrt)
qqnorm(joined_dark$calc_sqrt)
qqline(joined_dark$calc_sqrt, col = "red")
shapiro.test(joined_dark$calc_sqrt)

# Inverse 
joined_dark$calc_inv <- 1 / (joined_dark$Net_Calcification_cm2 +1) # add 1 to avoid division by zero
hist(joined_dark$calc_inv)
qqnorm(joined_dark$calc_inv)
qqline(joined_dark$calc_inv, col = "red")
shapiro.test(joined_dark$calc_inv)
```

Sqrt and inv produced normality. Proceed with sqrt. 
```{r}
mod_calc_dark <- lmer(calc_sqrt ~ Species * Phenotype * Timepoint + (1 | Colony_ID),
                 data = joined_dark,
                 REML = FALSE)
car::Anova(mod_calc_dark, type = 3)
```

Species x Phenotype significant. 

Look at post-hoc results 
```{r}
em_sp_ph <- emmeans(mod_calc_dark, ~ Species * Phenotype | Timepoint)
contrasts_sp_ph <- contrast(em_sp_ph, method = "pairwise", adjust = "tukey")
summary(contrasts_sp_ph)
```

Calculate mean, sem, and sd for dark only
```{r}
joined_dark_summary <- joined_dark %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_calc = mean(Net_Calcification_cm2, na.rm = TRUE),
    sd_calc = sd(Net_Calcification_cm2, na.rm = T),
    se_calc = sd(Net_Calcification_cm2, na.rm = T) / sqrt(length(Net_Calcification_cm2)),
    .groups = "drop") %>%
  drop_na() %>%
  filter(!Timepoint == "Jan25") # remove Jan25 for now 
```

Plot! 
```{r}
# Set levels 
joined_dark_summary$Timepoint <- factor(joined_dark_summary$Timepoint, levels = c("Apr23",
                                                                        "Sept23",
                                                                        "Jan24", 
                                                                        "May24", 
                                                                        "Oct24", 
                                                                        #"Jan25", 
                                                                        "May25"))

ggplot(joined_dark_summary, aes(x = Timepoint, y = mean_calc, color = Phenotype, group = Phenotype)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_calc - se_calc, ymax = mean_calc + se_calc, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species), cols = vars(Light)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Net calcification (Dark)",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    legend.position = c(0.85, 0.33),  # adjust x, y for placement
    #legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```


