---
title: "Physiology comparisons for 2019, 2021, 2023, and 2025"
author: "Jill Ashey"
date: "2026-03-19"
output: html_document
---

This script will compare certain physiological variables (Pmax, Rd, AFDW and symbiont density, color score) in 2019 (heatwave), 2021 (recovery+2 yrs), 2023 (recovery+4yrs), and 2025 (recovery+6 yrs). The 2019 and 2021 data is from Brown et al. 2023. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(emmeans)
library(lme4)
library(nlme)
library(Rmisc)
library(boot)
library(nlme)
library(patchwork)
library(lmerTest)
```

Run this before running any ANOVAs
```{r}
options(contrasts = c("contr.sum", "contr.poly"))
```

## Read in 2019 and 2021 data 
```{r}
data_2019 <- read_csv("../data/Physiology/904962_v1_coral_physiology.csv")
```

Select certain cols, filter by dates, rename cols to match the ones below, and assign susceptible and resistant tags 
```{r}
data_2019 <- data_2019 %>%
  filter(Date %in% as.Date(c("2019-10-02", "2021-10-08"))) %>%
  dplyr::rename(Colony_ID = Coral_ID, Phenotype = Bleach, Rd = R) %>%
  dplyr::select(Colony_ID, Pmax, Rd, Sym_density, AFDW, Date, Species, Phenotype, Sym_density, AFDW) %>%
  mutate(Phenotype = case_match(Phenotype, 
                             "Bleach" ~ "Susceptible", 
                             "Non-bleach" ~ "Resistant",
                             .default = Phenotype))
```

I STILL NEED TO DO SURFACE AREA FOR OCT 2025 

## Pmax 

Read in Pmax data 
```{r}
pi_2023 <- read_csv("../output/pi_curves/Sept2023/pi_curve_metrics_Sept2023.csv")
pi_2025 <- read_csv("../output/pi_curves/Oct2025/pi_curve_metrics_Oct2025.csv")
```

Select certain cols 
```{r}
pi_2023 <- pi_2023 %>%
  dplyr::select(Colony_ID, Pmax, Date, Species, Phenotype)

pi_2025 <- pi_2025 %>%
  dplyr::select(Colony_ID, Pmax, Date, Species, Phenotype)
```

### Bind data w/ historical data 

Select Pmax and R from 2019 and 2021
```{r}
pi_data_2019 <- data_2019 %>%
  dplyr::select(Colony_ID, Pmax, Date, Species, Phenotype)
```

Bind with 2019 and 2021 data 
```{r}
pi_data <- bind_rows(pi_data_2019, pi_2023, pi_2025)
```

Add Year column
```{r}
pi_data <- pi_data %>%
  mutate(Year = case_match(Date,
    as.Date("2019-10-02") ~ "2019",
    as.Date("2021-10-08") ~ "2021",
    as.Date("2023-09-14") ~ "2023",
    as.Date("2023-09-15") ~ "2023",
    as.Date("2025-10-02") ~ "2025",
    .default = as.character(Date)))
```

Save as csv
```{r}
write.csv(pi_data, "../output/pi_curves/pi_curve_all_years.csv")
```

### MCAP statistics comparing Pmax over time and by phenotype 

Subset by Mcap only 
```{r}
pi_data_mcap <- pi_data %>%
  filter(Species == "Montipora capitata")
```

Check for normality
```{r}
hist(pi_data_mcap$Pmax)

# Q-Q Plot
qqnorm(pi_data_mcap$Pmax)
qqline(pi_data_mcap$Pmax, col = "red")

# Shapiro-Wilk Test
shapiro.test(pi_data_mcap$Pmax)
```

Data is normal.

Run linear mixed effects model 
```{r}
pmax.model <- lmer(Pmax ~ Year * Phenotype + (1|Colony_ID), data = pi_data_mcap)
car::Anova(pmax.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(pmax.model, which = 1) 
lattice::qqmath(pmax.model)

# Post hoc tests 
emmeans(pmax.model, pairwise ~ Year)
emmeans(pmax.model, pairwise ~ Phenotype)

# Check plot to make sure results make sense
ggplot(pi_data_mcap, aes(x = Year, y = Pmax, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year and Phenotype are significant. 

### PCOMP statistics comparing Pmax over time and by phenotype 

Subset by Pcomp only 
```{r}
pi_data_pcomp <- pi_data %>%
  filter(Species == "Porites compressa")
```

Check for normality
```{r}
hist(pi_data_pcomp$Pmax)

# Q-Q Plot
qqnorm(pi_data_pcomp$Pmax)
qqline(pi_data_pcomp$Pmax, col = "red")

# Shapiro-Wilk Test
shapiro.test(pi_data_pcomp$Pmax)
```

Data is normal.

Run linear mixed effects model 
```{r}
pmax.model <- lmer(Pmax ~ Year * Phenotype + (1|Colony_ID), data = pi_data_pcomp)
car::Anova(pmax.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(pmax.model, which = 1) 
lattice::qqmath(pmax.model)

# Post hoc tests 
emmeans(pmax.model, pairwise ~ Year)
emmeans(pmax.model, pairwise ~ Phenotype)
emmeans(pmax.model, pairwise ~ Year | Phenotype)
emmeans(pmax.model, pairwise ~ Phenotype | Year)

# Check plot to make sure results make sense
ggplot(pi_data_pcomp, aes(x = Year, y = Pmax, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year, Phenotype, and Year x Phenotype are significant. 

## Symbiont density 

STILL NEED TO DO OCT 2025 

### Process 2023 and 2025 data 

Read in metadata 
```{r}
meta_sym <- read.csv("../data/Physiology/sym_counts/sym_count_metadata.csv")

# Remove extra spaces 
meta_sym <- meta_sym %>%
  mutate(
    Run = str_replace_all(Run, "\\s+", ""),  # "Run 1" -> "Run1"
    Well = str_trim(Well),
    Date = as.character(Date))
```

Read in raw data from guava
```{r}
files <- list.files(path = "../data/Physiology/sym_counts/raw_guava", pattern = "*.csv", full.names = TRUE)

read_with_metadata <- function(f) {
  df <- read.csv(f, skip = 4, header = TRUE, stringsAsFactors = FALSE)
  
  # extract filename without path and extension
  filename <- basename(f)
  filename <- str_remove(filename, "\\.csv$")

  # split filename at underscore, e.g., "20230805_Run1" -> c("20230805", "Run1")
  parts <- str_split(filename, "_", simplify = TRUE)

  # add columns for Date and Run parsed from filename
  df %>%
    mutate(
      Date = parts[1],
      Run = parts[2],
      # ensure consistent types and clean whitespace
      Well = str_trim(Well),
      Sample.ID = as.character(Sample.ID)
    )
}

all_dfs <- lapply(files, read_with_metadata)

combined_df <- bind_rows(all_dfs)

combined_df <- combined_df %>%
  mutate(
    Run = str_replace_all(Run, "\\s+", ""),
    Well = str_trim(Well),
    Date = as.character(Date)  # ensure same type
  )
```

Merge raw data and metadata
```{r}
sym_counts <- left_join(meta_sym, combined_df, by = c("Date", "Run", "Well"))
```

Remove samples marked "Do not use" and only keep Sept 2023 and Oct 2025 time points
```{r}
sym_counts <- sym_counts %>%
  filter(is.na(Use) | Use == "") %>%
  filter(Timepoint == "September_2023")
```

Calculate mean counts per sample and number of cells per sample 
```{r}
sym_counts_avg <- sym_counts %>%
  dplyr::group_by(Date, Run, Dilution, Timepoint, Pellet_Resuspend_Vol_mL, ColonyID) %>%
  dplyr::summarise(mean_cells_mL = mean(CHLOROPHYLL.POS.Concentration.Concentration.for.CHLOROPHYLL.POS...cells.mL., na.rm = T))%>%
  mutate(cells = Dilution * Pellet_Resuspend_Vol_mL * mean_cells_mL,
         ColonyID = as.character(ColonyID)) %>%
  ungroup()
```

Read in surface area df, merge with sym counts, and calculate cells/cm2
```{r}
surface_area <- read.csv("../data/Surface_Area/Surface_Area_wax_seasonal.csv")
surface_area$ColonyID <- as.character(surface_area$ColonyID)

sym_counts_avg <- sym_counts_avg %>%
  left_join(surface_area, by = c("ColonyID", "Timepoint")) %>%
  mutate(cells_cm2 = cells/surface.area.cm2,
          cells_cm2_M = cells_cm2 / 1e6) %>%
  mutate(Year = case_match(Timepoint,
                           "September_2023" ~ "2023", .default = Timepoint)) %>%
  select(ColonyID, cells_cm2_M, Species, Phenotype, Year)
```

### Bind data w/ historical data 

Select sym densities from 2019 and 2021 and format appro
```{r}
sym_data_2019 <- data_2019 %>%
  dplyr::select(Colony_ID, Sym_density, Date, Species, Phenotype) %>%
  dplyr::rename(
    ColonyID = Colony_ID,
    cells_cm2_M = Sym_density) %>%
  mutate(ColonyID = as.character(ColonyID)) %>%
  mutate(Year = case_match(Date,
    as.Date("2019-10-02") ~ "2019",
    as.Date("2021-10-08") ~ "2021",
    .default = as.character(Date))) %>%
  select(ColonyID, cells_cm2_M, Species, Phenotype, Year)
```

Bind with 2019 and 2021 data 
```{r}
sym_data <- bind_rows(sym_data_2019, sym_counts_avg)
```

Save as csv
```{r}
write.csv(sym_data, "../output/Physiology/sym_density_all_years.csv")
```

### MCAP statistics comparing sym density over time and by phenotype 

Subset by Mcap only 
```{r}
sym_data_mcap <- sym_data %>%
  filter(Species == "Montipora capitata")
```

Check for normality
```{r}
hist(sym_data_mcap$cells_cm2_M)

# Q-Q Plot
qqnorm(sym_data_mcap$cells_cm2_M)
qqline(sym_data_mcap$cells_cm2_M, col = "red")

# Shapiro-Wilk Test
shapiro.test(sym_data_mcap$cells_cm2_M)
```

Data is not normal. Transform data to achieve normality 
```{r}
sym_data_mcap$log <- log(sym_data_mcap$cells_cm2_M)
hist(sym_data_mcap$log)
qqnorm(sym_data_mcap$log)
qqline(sym_data_mcap$log, col = "red")
shapiro.test(sym_data_mcap$log)

sym_data_mcap$sqrt <- sqrt(sym_data_mcap$cells_cm2_M)
hist(sym_data_mcap$sqrt)
qqnorm(sym_data_mcap$sqrt)
qqline(sym_data_mcap$sqrt, col = "red")
shapiro.test(sym_data_mcap$sqrt)
```

Log and sqrt produced normality. Proceed with log. 

Run linear mixed effects model 
```{r}
sym.model <- lmer(log ~ Year * Phenotype + (1|ColonyID), data = sym_data_mcap)
car::Anova(sym.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(sym.model, which = 1) 
lattice::qqmath(sym.model)

# Post hoc tests 
emmeans(sym.model, pairwise ~ Year)
emmeans(sym.model, pairwise ~ Phenotype)

# Check plot to make sure results make sense
ggplot(sym_data_mcap, aes(x = Year, y = cells_cm2_M, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year and Phenotype are significant. 

### PCOMP statistics comparing sym density over time and by phenotype 

Subset by Pcomp only 
```{r}
sym_data_pcomp <- sym_data %>%
  filter(Species == "Porites compressa")
```

Check for normality
```{r}
hist(sym_data_pcomp$cells_cm2_M)

# Q-Q Plot
qqnorm(sym_data_pcomp$cells_cm2_M)
qqline(sym_data_pcomp$cells_cm2_M, col = "red")

# Shapiro-Wilk Test
shapiro.test(sym_data_pcomp$cells_cm2_M)
```

Data is not normal. Transform data to achieve normality 
```{r}
sym_data_pcomp$log <- log(sym_data_pcomp$cells_cm2_M)
hist(sym_data_pcomp$log)
qqnorm(sym_data_pcomp$log)
qqline(sym_data_pcomp$log, col = "red")
shapiro.test(sym_data_pcomp$log)

sym_data_pcomp$sqrt <- sqrt(sym_data_pcomp$cells_cm2_M)
hist(sym_data_pcomp$sqrt)
qqnorm(sym_data_pcomp$sqrt)
qqline(sym_data_pcomp$sqrt, col = "red")
shapiro.test(sym_data_pcomp$sqrt)
```

Sqrt produced normality.

Run linear mixed effects model 
```{r}
sym.model <- lmer(sqrt ~ Year * Phenotype + (1|ColonyID), data = sym_data_pcomp)
car::Anova(sym.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(sym.model, which = 1) 
lattice::qqmath(sym.model)

# Post hoc tests 
emmeans(sym.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(sym_data_pcomp, aes(x = Year, y = cells_cm2_M, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year is significant. 

## AFDW

STILL NEED TO DO OCT 2025 

### Process 2023 and 2025 data 

Read in AFDW data
```{r}
afdw <- read.csv("../data/Physiology/AFDW.csv")
afdw$ColonyID <- as.character(afdw$ColonyID)
```

Merge with surface area data (read in above) and caculate biomass in mg/ml
```{r}
homogenate_volume <- 50

afdw <- afdw %>%
  left_join(surface_area, by = c("ColonyID", "Timepoint")) %>%
  mutate(dry_mass_g_mL = Dry_mass_g * homogenate_volume, # Standardize dry mass to homogenate volume
         burnt_mass_g_mL = Burnt_mass_g * homogenate_volume, # Standardize burnt mass to homogenate volume
          dry_biomass_g = dry_mass_g_mL - Initial_mass_g, # Calculate dry biomass
          dw_mg_cm2 = (dry_biomass_g*1000)/surface.area.cm2,
          ash_biomass_g = dry_mass_g_mL - burnt_mass_g_mL,
          afdw_mg_cm2 = (ash_biomass_g*1000)/surface.area.cm2) %>%
    mutate(Year = case_match(Timepoint,
                           "September_2023" ~ "2023", .default = Timepoint)) %>%
  select(ColonyID, afdw_mg_cm2, Species, Phenotype, Year)
```

### Bind data w/ historical data 

Select afdw from 2019 and 2021 and format 
```{r}
afdw_data_2019 <- data_2019 %>%
  dplyr::select(Colony_ID, AFDW, Date, Species, Phenotype) %>%
  dplyr::rename(
    ColonyID = Colony_ID,
    afdw_mg_cm2 = AFDW) %>%
  mutate(ColonyID = as.character(ColonyID)) %>%
  mutate(Year = case_match(Date,
    as.Date("2019-10-02") ~ "2019",
    as.Date("2021-10-08") ~ "2021",
    .default = as.character(Date))) %>%
  select(ColonyID, afdw_mg_cm2, Species, Phenotype, Year)
```

Bind with 2019 and 2021 data 
```{r}
afdw_data <- bind_rows(afdw_data_2019, afdw)
```
Save as csv
```{r}
write.csv(afdw_data, "../output/Physiology/AFDW_all_years.csv")
```

### MCAP statistics comparing afdw over time and by phenotype 

Subset by Mcap only 
```{r}
afdw_data_mcap <- afdw_data %>%
  filter(Species == "Montipora capitata")
```

Check for normality
```{r}
hist(afdw_data_mcap$afdw_mg_cm2)

# Q-Q Plot
qqnorm(afdw_data_mcap$afdw_mg_cm2)
qqline(afdw_data_mcap$afdw_mg_cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(afdw_data_mcap$afdw_mg_cm2)
```

Data is not normal. Transform data to achieve normality 
```{r}
afdw_data_mcap$log <- log(afdw_data_mcap$afdw_mg_cm2)
hist(afdw_data_mcap$log)
qqnorm(afdw_data_mcap$log)
qqline(afdw_data_mcap$log, col = "red")
shapiro.test(afdw_data_mcap$log)

afdw_data_mcap$sqrt <- sqrt(afdw_data_mcap$afdw_mg_cm2)
hist(afdw_data_mcap$sqrt)
qqnorm(afdw_data_mcap$sqrt)
qqline(afdw_data_mcap$sqrt, col = "red")
shapiro.test(afdw_data_mcap$sqrt)
```

Log and sqrt produced normality. Proceed with log. 

Run linear mixed effects model 
```{r}
afdw.model <- lmer(log ~ Year * Phenotype + (1|ColonyID), data = afdw_data_mcap)
car::Anova(afdw.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(afdw.model, which = 1) 
lattice::qqmath(afdw.model)

# Post hoc tests 
emmeans(afdw.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(afdw_data_mcap, aes(x = Year, y = afdw_mg_cm2, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year is significant.

### PCOMP statistics comparing afdw over time and by phenotype 

Subset by Pcomp only 
```{r}
afdw_data_pcomp <- afdw_data %>%
  filter(Species == "Porites compressa")
```

Check for normality
```{r}
hist(afdw_data_pcomp$afdw_mg_cm2)

# Q-Q Plot
qqnorm(afdw_data_pcomp$afdw_mg_cm2)
qqline(afdw_data_pcomp$afdw_mg_cm2, col = "red")

# Shapiro-Wilk Test
shapiro.test(afdw_data_pcomp$afdw_mg_cm2)
```

Data is not normal. Transform data to achieve normality 
```{r}
afdw_data_pcomp$log <- log(afdw_data_pcomp$afdw_mg_cm2)
hist(afdw_data_pcomp$log)
qqnorm(afdw_data_pcomp$log)
qqline(afdw_data_pcomp$log, col = "red")
shapiro.test(afdw_data_pcomp$log)

afdw_data_pcomp$sqrt <- sqrt(afdw_data_pcomp$afdw_mg_cm2)
hist(afdw_data_pcomp$sqrt)
qqnorm(afdw_data_pcomp$sqrt)
qqline(afdw_data_pcomp$sqrt, col = "red")
shapiro.test(afdw_data_pcomp$sqrt)
```

Log and sqrt produced normality. Proceed with sqrt 

Run linear mixed effects model 
```{r}
afdw.model <- lmer(sqrt ~ Year * Phenotype + (1|ColonyID), data = afdw_data_pcomp)
car::Anova(afdw.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(afdw.model, which = 1) 
lattice::qqmath(afdw.model)

# Post hoc tests 
emmeans(afdw.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(afdw_data_pcomp, aes(x = Year, y = afdw_mg_cm2, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

## Color score 

Read in data
```{r}
color <- read.csv("../data/Physiology/color_score.csv")
```

Subset by dates of interest, remove any rows that have no data, and add year column
```{r}
color_filtered <- color %>%
  filter(Date %in% c("2019-10-2", "2021-10-8", "2023-09-13", "2025-10-1"), !is.na(Mean)) %>%
  mutate(Year = case_match(Date,
      "2019-10-2"  ~ "2019",
      "2021-10-8"  ~ "2021",
      "2023-09-13" ~ "2023",
      "2025-10-1"  ~ "2025"))
```

Calculate proportion red
```{r}
color_quant <- color_filtered %>%
  group_by(Colony.ID, Year) %>%
  mutate(
    # Use mean() to ensure we get exactly ONE value per group
    Standard = mean(Mean[Replicate == 6], na.rm = TRUE),
    prop_red = Mean / Standard
  ) %>%
  filter(Replicate != 6) %>%
  ungroup() %>%
  dplyr::select(Colony.ID, prop_red, Species, Phenotype, Year)
```

Because the higher values mean the coral is bleaching more, I am going to invert the values so that higher values = more color. 
```{r}
color_quant <- color_quant %>%
  mutate(inverted_color = 1 - prop_red) %>%
  filter(inverted_color > 0)#%>%
  #dplyr::select(Colony.ID, inverted_color, Species, Phenotype, Year)
```

Save as csv
```{r}
write.csv(color_quant, "../output/Physiology/color_score_all_years.csv")
```

### MCAP statistics comparing color score over time and by phenotype 

Subset by Mcap only 
```{r}
color_data_mcap <- color_quant %>%
  filter(Species == "Montipora capitata")
```

Check for normality
```{r}
hist(color_data_mcap$inverted_color)

# Q-Q Plot
qqnorm(color_data_mcap$inverted_color)
qqline(color_data_mcap$inverted_color, col = "red")

# Shapiro-Wilk Test
shapiro.test(color_data_mcap$inverted_color)
```

Data is normal. 

Run linear mixed effects model 
```{r}
color.model <- lmer(inverted_color ~ Year * Phenotype + (1|Colony.ID), data = color_data_mcap)
car::Anova(color.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(color.model, which = 1) 
lattice::qqmath(color.model)

# Post hoc tests 
emmeans(color.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(color_data_mcap, aes(x = Year, y = inverted_color, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year is significant. 

### PCOMP statistics comparing color score over time and by phenotype 

Subset by Pcomp only 
```{r}
color_data_pcomp <- color_quant %>%
  filter(Species == "Porites compressa")
```

Check for normality
```{r}
hist(color_data_pcomp$inverted_color)

# Q-Q Plot
qqnorm(color_data_pcomp$inverted_color)
qqline(color_data_pcomp$inverted_color, col = "red")

# Shapiro-Wilk Test
shapiro.test(color_data_pcomp$inverted_color)
```

Data is normal. 

Run linear mixed effects model 
```{r}
color.model <- lmer(inverted_color ~ Year * Phenotype + (1|Colony.ID), data = color_data_pcomp)
car::Anova(color.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(color.model, which = 1) 
lattice::qqmath(color.model)

# Post hoc tests 
emmeans(color.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(color_data_pcomp, aes(x = Year, y = inverted_color, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

## Visualization 

```{r}
# 1. Update the prep function to use "trait" instead of "Metric"
prep_df <- function(df, trait_name, value_col) {
  df %>%
    dplyr::rename(Value = !!sym(value_col)) %>%
    dplyr::mutate(trait = trait_name) %>%
    dplyr::select(Species, Phenotype, Year, Value, trait) %>%
    dplyr::mutate(Year = as.factor(Year))
}

# 2. Combine your 6 dataframes
traits_long_all <- bind_rows(
  prep_df(pi_data_mcap, "Pmax", "Pmax"),
  prep_df(sym_data_mcap, "Sym density", "cells_cm2_M"),
  prep_df(afdw_data_mcap, "AFDW", "afdw_mg_cm2"),
  prep_df(color_data_mcap, "Color Score", "inverted_color"), # Added
  prep_df(pi_data_pcomp, "Pmax", "Pmax"),
  prep_df(sym_data_pcomp, "Sym density", "cells_cm2_M"),
  prep_df(afdw_data_pcomp, "AFDW", "afdw_mg_cm2"),
  prep_df(color_data_pcomp, "Color Score", "inverted_color") # Added
) %>% dplyr::rename(estimate = Value)

# 3. Define Labels and Significance (Use exact same "trait" names as above)
# Update significance data for 4 traits per species
sig_data <- data.frame(
  trait = rep(c("Pmax", "Sym density", "AFDW", "Color Score"), each = 2),
  Species = rep(c("Montipora capitata", "Porites compressa"), 4),
  sig_label = c(
    "n.s.", "n.s.", # Pmax
    "n.s.", "n.s.", # Sym density
    "n.s.", "n.s.", # AFDW
    "Y***", "Y***"  # Color Score
  )
)

# Add the Color Score label (using bold formatting to match the others)
trait_labels <- c(
  "Pmax" = "atop(bold(Maximum~photosynthesis), bold((mu*mol~hr^-1~cm^-2)))",
  "Sym density" = "atop(bold(Symbiont~density), bold((10^6~cells~cm^-2)))",
  "AFDW" = "atop(bold(Biomass), bold((mg~cm^-2)))",
  "Color Score" = "bold(Color~Score)"
)

# 4. Calculate Shared Y-Limits (Using dplyr:: explicitly to avoid plyr errors)
trait_limits_df <- traits_long_all %>%
  dplyr::group_by(trait, Species, Year, Phenotype) %>%
  dplyr::summarise(
    m = mean(estimate, na.rm = TRUE),
    se = sd(estimate, na.rm = TRUE) / sqrt(n()),
    .groups = 'drop'
  ) %>%
  dplyr::group_by(trait) %>%
  dplyr::summarise(
    ymin = min(m - se, na.rm = TRUE),
    ymax = max(m + se, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  dplyr::mutate(ymax = ymax + (ymax - ymin) * 0.25, 
                ymin = ymin - (ymax - ymin) * 0.1)

# 5. Panel Function
create_panel_plot <- function(data, tr, spp, ylimits, is_bottom_row = FALSE, panel_tag = NULL) {
  
  plot_data <- data %>% dplyr::filter(trait == tr, Species == spp)
  sig_text <- sig_data %>% dplyr::filter(trait == tr, Species == spp) %>% dplyr::pull(sig_label)
  y_label <- parse(text = trait_labels[[tr]])

  p <- ggplot(plot_data, aes(x = factor(Year), y = estimate, color = Phenotype, group = Phenotype)) +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.15, linewidth = 1.2) +
    stat_summary(fun = mean, geom = "line", linewidth = 1.5) +
    stat_summary(fun = mean, geom = "point", size = 5) +
    annotate("text", x = Inf, y = -Inf, label = sig_text, 
             hjust = 1.1, vjust = -1.2, size = 8, color = "black") +
    scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
    theme_classic() + 
    theme(
      axis.text.y   = element_text(size = 22, color = "black"), 
      axis.text.x   = element_text(size = 22, color = "black", face = "bold"),
      axis.title.y  = element_text(size = 24),
      plot.margin   = margin(t = 20, r = 10, b = 10, l = 10),
      plot.tag      = element_text(size = 28, face = "bold"),
      plot.tag.position = c(0.05, 0.95)
    ) +
    labs(x = NULL, y = y_label, tag = panel_tag) +
    coord_cartesian(ylim = ylimits)

  if (!is_bottom_row) {
    p <- p + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  }
  
  return(p)
}

# 6. Generate and Assemble
traits <- c("Pmax", "Sym density", "AFDW", "Color Score")
tag_list <- c('A','B','C','D','E','F','G','H')
all_plots <- list()
counter <- 1

for (j in 1:length(species_list)) { 
  for (i in 1:length(traits)) {    
    current_limits <- trait_limits_df %>% 
      dplyr::filter(trait == traits[i]) %>% 
      dplyr::select(ymin, ymax) %>% 
      as.numeric()
      
    all_plots[[counter]] <- create_panel_plot(
      data = traits_long_all, 
      tr = traits[i], 
      spp = species_list[j],
      ylimits = current_limits, 
      is_bottom_row = (j == 2),
      panel_tag = tag_list[counter]
    )
    counter <- counter + 1
  }
}

mcap_header <- wrap_elements(full = grid::textGrob("Montipora capitata", 
                             x = unit(0.01, "npc"), 
                             hjust = 0,
                             gp = gpar(fontsize = 30, fontface = "bold.italic")))
                             
pcomp_header <- wrap_elements(full = grid::textGrob("Porites compressa", 
                              x = unit(0.01, "npc"), 
                              hjust = 0,
                              gp = gpar(fontsize = 30, fontface = "bold.italic")))

# 7. Final Assembly
final_phys_figure <- (
  mcap_header / 
  (all_plots[[1]] | all_plots[[2]] | all_plots[[3]] | all_plots[[4]]) /
  pcomp_header /
  (all_plots[[5]] | all_plots[[6]] | all_plots[[7]] | all_plots[[8]])
) + 
plot_layout(heights = c(0.12, 1, 0.12, 1), guides = "collect") & 
theme(
  legend.position = "right", 
  legend.title = element_text(size = 20, face = "bold"),
  legend.text = element_text(size = 18), 
  legend.key.size = unit(1.5, "cm"),
  plot.margin = margin(t = 10, r = 10, b = 10, l = 10) 
)

# 8. Save
ggsave("../output/Physiology/Physiology_panels.png", plot = final_phys_figure, 
       width = 28, height = 15, units = "in", dpi = 300)
ggsave("../output/Physiology/Physiology_panels.pdf", plot = final_phys_figure, 
       width = 28, height = 15, units = "in", dpi = 300)
```

