---
title: "Physiology"
author: "Jill Ashey"
date: "2025-11-24"
output: html_document
---

Script for all physiology (sym counts, chlorophyll, protein, MAAs, and AFDW) analyses. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(purrr)
library(readr)
library(stringr)
```

## Sym counts 

Read in metadata 
```{r}
meta_sym <- read.csv("../data/Physiology/sym_counts/sym_count_metadata.csv")

# Remove extra spaces 
meta_sym <- meta_sym %>%
  mutate(
    Run = str_replace_all(Run, "\\s+", ""),  # "Run 1" -> "Run1"
    Well = str_trim(Well),
    Date = as.character(Date))
```

Read in raw data from guava
```{r}
files <- list.files(path = "../data/Physiology/sym_counts/raw_guava", pattern = "*.csv", full.names = TRUE)

read_with_metadata <- function(f) {
  df <- read.csv(f, skip = 4, header = TRUE, stringsAsFactors = FALSE)
  
  # extract filename without path and extension
  filename <- basename(f)
  filename <- str_remove(filename, "\\.csv$")

  # split filename at underscore, e.g., "20230805_Run1" -> c("20230805", "Run1")
  parts <- str_split(filename, "_", simplify = TRUE)

  # add columns for Date and Run parsed from filename
  df %>%
    mutate(
      Date = parts[1],
      Run = parts[2],
      # ensure consistent types and clean whitespace
      Well = str_trim(Well),
      Sample.ID = as.character(Sample.ID)
    )
}

all_dfs <- lapply(files, read_with_metadata)

combined_df <- bind_rows(all_dfs)

combined_df <- combined_df %>%
  mutate(
    Run = str_replace_all(Run, "\\s+", ""),
    Well = str_trim(Well),
    Date = as.character(Date)  # ensure same type
  )
```

Merge raw data and metadata
```{r}
sym_counts <- left_join(meta_sym, combined_df, by = c("Date", "Run", "Well"))
```

Remove samples marked "Do not use"
```{r}
sym_counts <- sym_counts %>%
  filter(is.na(Use) | Use == "")
```

Calculate mean counts per sample and number of cells per sample 
```{r}
sym_counts_avg <- sym_counts %>%
  group_by(Date, Run, Dilution, Timepoint, Pellet_Resuspend_Vol_mL, ColonyID) %>%
  summarise(mean_cells_mL = mean(CHLOROPHYLL.POS.Concentration.Concentration.for.CHLOROPHYLL.POS...cells.mL., na.rm = T))%>%
  mutate(cells = Dilution * Pellet_Resuspend_Vol_mL * mean_cells_mL,
         ColonyID = as.character(ColonyID)) %>%
  ungroup()
```

Read in surface area df and add time point col 
```{r}
surface_area <- read.csv("../data/Surface_Area/fragrammetry_SurfaceArea.csv")

surface_area <- surface_area %>%
  mutate(Timepoint = case_when(
    grepl("2023-04", Date) ~ "April_2023",
    grepl("2023-09", Date) ~ "September_2023",
    grepl("2024-01", Date) ~ "January_2024",
    grepl("2024-05", Date) ~ "May_2024",
    grepl("2024-10", Date) ~ "October_2024",
    grepl("2025-01", Date) ~ "January_2025",
    grepl("2025-05", Date) ~ "May_2025",
    grepl("2025-10", Date) ~ "October_2025",
    TRUE ~ NA_character_)) %>%
  rename(ColonyID = Colony_ID) %>%
  filter(Project == "CAREER - Phys")
```

Merge with sym counts and calculate cells/cm2
```{r}
sym_counts_avg <- sym_counts_avg %>%
  left_join(surface_area, by = c("ColonyID", "Timepoint")) 

sym_counts_avg <- sym_counts_avg %>%
  mutate(cells_cm2 = cells/Surface_Area_cm2,
          cells_cm2_M = cells_cm2 / 1e6) %>%
  select(ColonyID, Timepoint, Species, Phenotype, Dilution, cells_cm2, cells_cm2_M)
```

Save as csv 
```{r}
write.csv(sym_counts_avg, "../output/Physiology/sym_density.csv")
```

Calculate mean, sem, and sd 
```{r}
sym_summary <- sym_counts_avg %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_sym = mean(cells_cm2_M, na.rm = TRUE),
    sd_sym = sd(cells_cm2_M, na.rm = T),
    se_sym = sd(cells_cm2_M, na.rm = T) / sqrt(length(cells_cm2_M)),
    .groups = "drop") %>%
  drop_na() 
```

Plot! 
```{r}
# Set levels 
sym_summary$Timepoint <- factor(sym_summary$Timepoint, levels = c("April_2023",
                                                                        "September_2023",
                                                                        "January_2024", 
                                                                        "May_2024",
                                                                        "October_2024"))

ggplot(sym_summary, aes(x = Timepoint, y = mean_sym, color = Phenotype, group = Phenotype)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_sym - se_sym, ymax = mean_sym + se_sym, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Sym density (10^6 x cells/cm2)",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```

Values look slightly lower than I would expect...maybe due to the surface area being based on the fraggameter?

## Chlorophyll 

Read in metadata and data. 
```{r}
meta_chl <- read.csv("../data/Physiology/chlorophyll/chl_metadata.csv")
data_chl <- read.csv("../data/Physiology/chlorophyll/chl_data.csv")
```

Merge raw data and metadata
```{r}
chl <- left_join(meta_chl, data_chl, by = c("Date", "Run", "Well"))
```

Calculate chlorophyll values 
```{r}
chl_values <- chl %>%
  group_by(ColonyID, Timepoint) %>%
  summarise(mean_A630 = mean(A630),
            mean_A663 = mean(A663),
            mean_A750 = mean(A750)) %>%
  mutate(chla.ug.ml = (11.43*(mean_A663-mean_A750))-(0.64*(mean_A630-mean_A750)),
         chlc.ug.ml = (24.36*(mean_A630-mean_A750))-(3.73*(mean_A663-mean_A750)))
```

Read in surface area df and add time point col 
```{r}
surface_area <- read.csv("../data/Surface_Area/fragrammetry_SurfaceArea.csv")

surface_area <- surface_area %>%
  mutate(Timepoint = case_when(
    grepl("2023-04", Date) ~ "April_2023",
    grepl("2023-09", Date) ~ "September_2023",
    grepl("2024-01", Date) ~ "January_2024",
    grepl("2024-05", Date) ~ "May_2024",
    grepl("2024-10", Date) ~ "October_2024",
    grepl("2025-01", Date) ~ "January_2025",
    grepl("2025-05", Date) ~ "May_2025",
    grepl("2025-10", Date) ~ "October_2025",
    TRUE ~ NA_character_)) %>%
  rename(ColonyID = Colony_ID) %>%
  filter(Project == "CAREER - Phys")
```

Merge with sym counts and calculate cells/cm2
```{r}
chl_homogenate_volume <- 10

chl_values <- chl_values %>%
  left_join(surface_area, by = c("ColonyID", "Timepoint")) %>%
  mutate(chla.ug.cm2 = (chla.ug.ml * chl_homogenate_volume / Surface_Area_cm2),
         chlc.ug.cm2 = (chlc.ug.ml * chl_homogenate_volume / Surface_Area_cm2)) %>%
  select(ColonyID, Timepoint, Species, Phenotype, chla.ug.cm2, chlc.ug.cm2)
```

Save as csv 
```{r}
write.csv(chl_values, "../output/Physiology/chlorophyll.csv")
```

Calculate mean, sem, and sd 
```{r}
chl_summary <- chl_values %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_chla = mean(chla.ug.cm2, na.rm = TRUE),
    sd_chla = sd(chla.ug.cm2, na.rm = T),
    se_chla = sd(chla.ug.cm2, na.rm = T) / sqrt(length(chla.ug.cm2)),
    .groups = "drop") %>%
  drop_na() 
```

Plot! 
```{r}
# Set levels 
chl_summary$Timepoint <- factor(chl_summary$Timepoint, levels = c("April_2023",
                                                                        "September_2023",
                                                                        "January_2024", 
                                                                        "May_2024"))

ggplot(chl_summary, aes(x = Timepoint, y = mean_chla, color = Phenotype, group = Phenotype)) +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_chla - se_chla, ymax = mean_chla + se_chla, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  #facet_grid(rows = vars(Species)) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Chlorophyll-a (ug/cm2)",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```



