---
title: "PI curves - trial"
author: "Jill Ashey"
date: "2025-07-07"
output: html_document
---

Trying out PI curve code 

NOTICED on Jan 2025 trip: Noticed that the Barott Lab computer time is fast by 3 min. Not sure if this affects data analysis or the files but something to note for downstream analysis 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("devtools")
#devtools::install_github("colin-olito/LoLinR")

library(tidyverse)
library(nls.multstart)
library(broom)
library(LoLinR)
library(readr)
library(lubridate)
library(fuzzyjoin)
library(stringr)
library(purrr)
library(future)
library(furrr)
library(minpack.lm)
library(lme4)
library(lmerTest)
```

```{r}
getwd()
# 1️⃣ Load the metadata ----
pi_curve_run_metadata <- read_csv("../data/pi_curves/pi_curve_metadata_run.csv")

run_meta <- read_csv("../data/pi_curves/pi_curve_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
sample_meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))


# 2️⃣ List all raw data files ----

files <- list.files(path = "../data/pi_curves/raw_data/Oct2024/", pattern = "*.csv", full.names = TRUE)


# 3️⃣ Read, process & join each file ----

all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "Ch\\d+") %>% str_remove("Ch") %>% as.numeric()

  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )

  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )

  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )

  return(raw_joined)
})


# 4️⃣ Inspect and save ----

all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

#write_csv(all_data, "all_joined_data.csv")

ggplot(all_data, aes(x = DateTime, y = Oxygen)) +
  geom_point() + 
  facet_wrap(~Colony_ID)
```

Fit local linear regressions by sample (write this in a loop ASAP)

### 205, Run 5 
Example with one sample 
```{r}
# Filter for sample AND removing Light = 0
one_sample <- all_data %>%
  filter(Colony_ID == "42", Run == 5) # , Light_Value > 0

ggplot(one_sample, aes(x = DateTime, y = Oxygen)) +
  geom_point() 
  
# Split by light interval 
intervals <- one_sample %>%
  group_by(Light_Value) %>%
  group_split()
length(intervals)  # how many intervals
map(intervals, ~ unique(.x$Light_Value))

# Define function 
fit_reg <- function(df) {
  rankLocReg(
    xall = as.numeric(df$`Delta T [min]`),
    yall = df$Oxygen,
    alpha = 0.2,
    method = "pc",
    verbose = TRUE
  )
}

# Fit regression for each interval 
fits <- map(intervals, fit_reg)
plot(fits[[1]])  # plot the first light interval
plot(fits[[6]])  
plot(fits[[4]])

# Extract slopes
micromol.L.s <- map_dbl(fits, ~ pluck(., "allRegs", "b1", 1))
light_levels <- map_dbl(intervals, ~ unique(.x$Light_Value))

one_sample_results <- tibble(
  Colony_ID = "42",
  Run = 5,
  Light_Value = light_levels,
  micromol.L.s = micromol.L.s
);one_sample_results

# Plot data by light level
ggplot(one_sample_results, aes(x = Light_Value, y = micromol.L.s)) +
  geom_point() +
  geom_line() +
  labs(
    x = "Light Level (PAR)",
    y = "Mean O₂ (μmol/L)",
    title = "PI Curve") +
  theme_minimal()
```

Hooray!! This is super computationally intense so I may have to run it as a background job on my computer OR run on the server 

Fit Platt model 
```{r}
pi_data <- one_sample_results %>%
  rename(I = Light_Value, P = micromol.L.s)

# Fit the model with Rd as an extra parameter
jp_model_Rd <- nlsLM(
  P ~ Pmax * tanh(alpha * I / Pmax) + Rd,
  data = pi_data,
  start = list(Pmax = max(pi_data$P), alpha = 0.01, Rd = -0.1)
)
summary(jp_model_Rd)

# Extract parameters
pars <- coef(jp_model_Rd)
Pmax <- pars[["Pmax"]]  # Am
alpha <- pars[["alpha"]]  # AQY
Rd <- pars[["Rd"]]  # dark respiration
Ik <- Pmax / alpha
Pgross <- Pmax + Rd
cat("Pmax (Am):", Pmax, "\n")
cat("Alpha (AQY):", alpha, "\n")
cat("Rd:", Rd, "\n")
cat("Ik:", Ik, "\n")
cat("Pgross:", Pgross, "\n")

# Plot
pi_data <- pi_data %>%
  mutate(P_fit = predict(jp_model_Rd))
ggplot(pi_data, aes(x = I, y = P)) +
  geom_point() +
  geom_line(aes(y = P_fit), color = "blue") +
  labs(x = "PAR (µmol photons m⁻² s⁻¹)",
       y = "Oxygen evolution (µmol L⁻¹ s⁻¹)") +
  annotate("text", x = max(pi_data$I)*0.7, y = max(pi_data$P)*0.7,
           label = paste0("Pmax = ", round(Pmax, 3), "\n",
                          #"Pgross = ", round(Pgross, 3), "\n",
                          "AQY = ", round(alpha, 4), "\n",
                          "Ik = ", round(Ik, 4), "\n",
                          "Rd = ", round(Rd, 4)))
```

Pgross is higher than Pmax because the Rd value is negative--how to deal with this?

```{r}
test <- read.csv("../output/all_joined_data_Oct2024.csv")
blah <- read.csv("../output/photosynthesis_fit_results_all_runs_blank_corrected_Oct2024.csv")
blah
```

Merge with sample info 
```{r}
merge <- blah %>%
  #filter(!Colony_ID == "blank") %>%
  full_join(sample_meta, by = "Colony_ID") %>%
  filter(!Colony_ID == "blank") %>%
  filter(!Colony_ID == "HOBO") 

merge_filtered <- merge %>%
  filter(!is.na(Pmax), Pmax <= 20)

summary_df <- merge_filtered %>%
  group_by(Species, Phenotype) %>%
  summarise(
    mean_Pmax = mean(Pmax, na.rm = TRUE),
    sd_Pmax = sd(Pmax, na.rm = TRUE),
    n = n(),
    se_Pmax = sd_Pmax / sqrt(n)
  )

ggplot(summary_df, aes(x = Species, y = mean_Pmax, fill = Phenotype)) +
  geom_col(position = position_dodge(width = 0.8), color = "black") +
  geom_errorbar(aes(ymin = mean_Pmax - se_Pmax, ymax = mean_Pmax + se_Pmax),
                position = position_dodge(width = 0.8), width = 0.2) +
  labs(x = "Phenotype", y = expression(P[max]~"(mean ± SE)")) +
  theme_classic() +
  scale_fill_brewer(palette = "Set1")

# Fit model: Pmax ~ Species * Phenotype + (1 | Colony_ID)
mod <- aov(Pmax ~ Species * Phenotype, data = merge_filtered)
summary(mod)

pam.lme <- lme(
  Fv_Fm ~ Species * Phenotype * Date,
  random = ~1 | Colony_ID,  # random intercept for each colony
  data = pam_long,
  na.action = na.exclude
)
summary(pam.lme)
car::Anova(pam.lme, type = 3)
```


```{r}
test <- read.csv("../output/intervals_202_run4.csv")
ggplot(test, aes(x = Light_Value, y = Slope)) +
  geom_point() +
  geom_line() +
  labs(
    x = "Light Level (PAR)",
    y = "Mean O2 (μmol/L)",
    title = "PI Curve") +
  theme_minimal()
test$Colony_ID <- as.character(test$Colony_ID)

# filter sample metadata
oct_2024 <- sample_meta %>%
  filter(grepl("^2024-10", Date))
oct_2024$Surface_Area_cm2 <- as.numeric(oct_2024$Surface_Area_cm2)

# merge with sample data info 
test_202 <- test %>%
  inner_join(oct_2024, by = "Colony_ID")

# normalize 
test_202 <- test_202 %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Slope * Chamber_Volume_L / Surface_Area_cm2) * 60
  )
ggplot(test_202, aes(x = Light_Value, y = Slope_norm)) +
  geom_point() +
  geom_line() +
  labs(
    x = "Light Level (PAR)",
    y = "μmol O₂ cm−2 h−1",
    title = "PI Curve") +
  theme_minimal()

# Fit model 
data <- test_202 %>%
  rename(I = Light_Value, P = Slope_norm)

# Fit the model with Rd as an extra parameter
jp_model_Rd <- nlsLM(
  P ~ Pmax * tanh(alpha * I / Pmax) + Rd,
  data = data,
  start = list(Pmax = max(data$P), alpha = 0.01, Rd = -0.1)
)
summary(jp_model_Rd)

# Extract parameters
pars <- coef(jp_model_Rd)
Pmax <- pars[["Pmax"]]  # Am
alpha <- pars[["alpha"]]  # AQY
Rd <- pars[["Rd"]]  # dark respiration
Ik <- Pmax / alpha
Pgross <- Pmax + Rd
cat("Pmax (Am):", Pmax, "\n")
cat("Alpha (AQY):", alpha, "\n")
cat("Rd:", Rd, "\n")
cat("Ik:", Ik, "\n")
cat("Pgross:", Pgross, "\n")

```

```{r}
all_int <- read.csv("../output/all_interval_summaries.csv")
ggplot(all_int, aes(x = Light_Value, y = Slope)) +
  geom_point() +
  geom_line() +
  labs(
    x = "Light Level (PAR)",
    y = "Mean O2 (μmol/L)",
    title = "PI Curve") +
  theme_minimal()
all_int$Colony_ID <- as.character(all_int$Colony_ID)

# filter sample metadata
oct_2024 <- sample_meta %>%
  filter(grepl("^2024-10", Date))
oct_2024$Surface_Area_cm2 <- as.numeric(oct_2024$Surface_Area_cm2)

# merge with sample data info 
all_int <- all_int %>%
  inner_join(oct_2024, by = "Colony_ID") %>%
  filter(!is.na(Phenotype))
length(unique(all_int$Colony_ID))

# normalize 
all_int <- all_int %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Slope * Chamber_Volume_L / Surface_Area_cm2) * 60
  )
write.csv(all_int, "../output/pi_curves/Oct2024/all_interval_summaries_normalized_Oct2024.csv")
```

```{r}
# Nest by Colony
pi_nested <- all_int %>%
  select(Colony_ID, Light_Value, Slope_norm) %>%  # or whatever your P column is
  rename(I = Light_Value, P = Slope_norm) %>%
  group_by(Colony_ID) %>%
  nest()

# Define a function to fit the model
fit_pi_model <- function(df) {
  tryCatch({
    fit <- nlsLM(P ~ Pmax * tanh(alpha * I / Pmax) + Rd,
                 data = df,
                 start = list(Pmax = max(df$P), alpha = 0.01, Rd = -0.1))
    pars <- coef(fit)
    tibble(Pmax = pars["Pmax"],
           alpha = pars["alpha"],
           Rd = pars["Rd"],
           Ik = pars["Pmax"]/pars["alpha"],
           Pgross = pars["Pmax"] + pars["Rd"])
  }, error = function(e) tibble(Pmax = NA, alpha = NA, Rd = NA, Ik = NA, Pgross = NA))
}

# Apply the model to each colony
pi_results <- pi_nested %>%
  mutate(params = map(data, fit_pi_model)) %>%
  select(-data) %>%
  unnest(cols = params)

# merge with sample data info 
pi_results <- pi_results %>%
  inner_join(oct_2024, by = "Colony_ID")
write.csv(pi_results, "../output/pi_curves/Oct2024/pi_curve_metrics_Oct2024.csv")

# Pivot parameters into long format
pi_long <- pi_results %>%
  pivot_longer(cols = c(Pmax, alpha, Rd, Ik, Pgross),
               names_to = "Parameter",
               values_to = "Value")

# Boxplot
ggplot(pi_long, aes(x = Phenotype, y = Value, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, size = 1.5, alpha = 0.8) +
  facet_grid(Parameter ~ Species, scales = "free_y") +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw(base_size = 14) +
  labs(x = "Phenotype", y = "Value") +
  theme(legend.position = "none")
```

Stats
```{r}
pmax.lme <- lme(
  Pgross ~ Species * Phenotype,
  random = ~1 | Colony_ID,  # random intercept for each colony
  data = pi_results,
  na.action = na.exclude
)
summary(pmax.lme)
car::Anova(pmax.lme, type = 3)
em_date <- emmeans(pmax.lme, pairwise ~ Phenotype, adjust = "tukey")
summary(em_date)
```

```{r}
library(dplyr)
library(ggplot2)

# 1. Raw data plot
ggplot(all_int, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype")

# 2. Summarize by mean ± SEM
summary_df <- all_int %>%
  group_by(Species, Phenotype, Light_Value) %>%
  summarise(
    mean_slope = mean(Slope_norm, na.rm = TRUE),
    sem_slope = sd(Slope_norm, na.rm = TRUE)/sqrt(n()),
    .groups = "drop"
  )

# Plot mean ± SEM
ggplot(summary_df, aes(x = Light_Value, y = mean_slope, color = Phenotype)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = mean_slope - sem_slope, ymax = mean_slope + sem_slope, fill = Phenotype),
              alpha = 0.2, color = NA) +
  facet_wrap(~Species, scales = "free") +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m⁻² s⁻¹)", y = "Slope_norm (mean ± SEM)", color = "Phenotype", fill = "Phenotype")

ggplot() +
  # Raw data points and lines per colony
  geom_point(data = all_int, aes(x = Light_Value, y = Slope_norm, color = Phenotype), alpha = 0.5) +
  geom_line(data = all_int, aes(x = Light_Value, y = Slope_norm, group = Colony_ID, color = Phenotype), alpha = 0.3) +
  # Mean ± SEM ribbons and lines
  geom_ribbon(data = summary_df, aes(x = Light_Value, ymin = mean_slope - sem_slope, ymax = mean_slope + sem_slope, fill = Phenotype), alpha = 0.2) +
  geom_line(data = summary_df, aes(x = Light_Value, y = mean_slope, color = Phenotype), size = 1.2) +
  facet_wrap(~Species, scales = "free") +
  theme_bw(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype", fill = "Phenotype") +
  theme(legend.position = "top")
```









## Try with Jan 2025 

```{r}
getwd()
# Load the metadata ----
pi_curve_run_metadata <- read_csv("../data/pi_curves/pi_curve_metadata_run.csv")

run_meta <- read_csv("../data/pi_curves/pi_curve_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
sample_meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))

# List all raw data files ----

files <- list.files(path = "../data/pi_curves/raw_data/Jan2025/", pattern = "*.csv", full.names = TRUE)


# Read, process & join each file ----

all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "Ch\\d+") %>% str_remove("Ch") %>% as.numeric()

  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )

  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )

  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )

  return(raw_joined)
})


# 4️⃣ Inspect and save ----

all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

#write_csv(all_data, "all_joined_data.csv")

ggplot(all_data, aes(x = DateTime, y = Oxygen)) +
  geom_point() + 
  facet_wrap(~Colony_ID)
```