---
title: "PI curves TCB"
author: "Jill Ashey"
date: "2025-09-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("devtools")
#devtools::install_github("colin-olito/LoLinR")

library(tidyverse)
library(nls.multstart)
library(broom)
library(LoLinR)
library(readr)
library(lubridate)
library(fuzzyjoin)
library(stringr)
library(purrr)
library(future)
library(furrr)
library(minpack.lm)
library(lme4)
library(lmerTest)
library(nlme)
library(emmeans)
```

## TCB - Sept 2023 

### Prepare data 

Load run metadata 
```{r}
run_meta <- read_csv("../data/pi_curves/pi_curve_metadata_TCB_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
```

Load sample metadata
```{r}
sample_meta <- read_csv("../data/pi_curves/pi_curve_metadata_TCB_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```

List all raw files for Sept 2023 
```{r}
files <- list.files(path = "../data/pi_curves/raw_data/Sept2023_TCB/", pattern = "*.csv", full.names = TRUE)
```

Join O2 data with metadata files based on start and stop times 
```{r}
all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "channel\\d+") %>% str_remove("channel") %>% as.numeric()
  
  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )
  
  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )
  
  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )
  
  return(raw_joined)
})
```

Look at joined data and save
```{r}
all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

write_csv(all_data, "../output/pi_curves/Sept2023_TCB/all_joined_data_Sept2023_TCB.csv")
```

### Run localized linear regressions 

The interval fitting code will take about an hour to run and is computationally intense. If needed, run as a background job or on a remote server.

Make list of all Colony_ID / Run combinations (including blanks)
```{r}
colony_runs <- all_data %>%
  filter(!is.na(Colony_ID)) %>%
  distinct(Colony_ID, Run)
```

For this timepoint, only two out of the five runs had blanks. Therefore, I am going to calculate the blank intervals, average them, and then apply them to the regression for each colony. 

Calculate interval slopes for the blanks
```{r}
blank_data <- all_data %>%
  filter(Colony_ID == "blank")

blank_intervals <- blank_data %>%
  group_by(Run, Light_Value) %>%
  filter(!is.na(Colony_ID)) %>%
  group_split()
```

Fit regressions and extract slopes for each blank interval
```{r}
fit_reg <- function(df) {
  rankLocReg(
    xall = as.numeric(df$`Delta T [min]`),
    yall = df$Oxygen,
    alpha = 0.2,
    method = "pc",
    verbose = FALSE
  )
}

blank_fits <- map(blank_intervals, fit_reg)
blank_slopes <- map_dbl(blank_fits, ~ pluck(., "allRegs", "b1", 1))
blank_summary <- map2_dfr(
  blank_intervals,
  blank_slopes,
  ~ tibble(
    Light_Value = unique(.x$Light_Value),
    Blank_Slope = .y
  )
)
```

Average blank slopes per light value, across all blanks/runs
```{r}
avg_blank_slopes <- blank_summary %>%
  group_by(Light_Value) %>%
  summarise(Avg_Blank_Slope = mean(Blank_Slope, na.rm = TRUE), .groups = "drop")
```

Run interval fittings with blank correction
```{r}
# Function: blank correction + slope extraction + save/return interval summary
fit_colony_intervals_avgblank <- function(colony_id, run_num) {
  # Subset sample data
  sample_data <- all_data %>%
    filter(Colony_ID == colony_id, Run == run_num)
  if (nrow(sample_data) == 0) return(NULL)
  
  # Split data by light step
  intervals <- sample_data %>%
    group_by(Light_Value) %>%
    group_split()
  
  # Fit regressions for each interval
  fits <- map(intervals, fit_reg)
  micromol.L.s <- map_dbl(fits, ~ pluck(., "allRegs", "b1", 1))
  
  # Detailed interval summary
  interval_df <- map2_dfr(
    intervals,
    seq_along(intervals),
    ~ tibble(
      Colony_ID   = colony_id,
      Run         = run_num,
      Interval    = .y,
      Light_Value = unique(.x$Light_Value),
      N_points    = nrow(.x),
      O2_start    = head(.x$Oxygen, 1),
      O2_end      = tail(.x$Oxygen, 1),
      DeltaT_min  = min(.x$`Delta T [min]`),
      DeltaT_max  = max(.x$`Delta T [min]`),
      O2_min      = min(.x$Oxygen, na.rm = TRUE),
      O2_max      = max(.x$Oxygen, na.rm = TRUE),
      Slope       = micromol.L.s[.y]
    )
  ) %>%
    left_join(avg_blank_slopes, by = "Light_Value") %>%
    mutate(
      Blank_Corrected_Slope = Slope - Avg_Blank_Slope
    )
  
  # Write file (outside of the pipeline!)
  fn <- paste0("../output/pi_curves/Sept2023_TCB/intervals_", colony_id, "_run", run_num, "_Sept2023_TCB.csv")
  write_csv(interval_df, fn)
  
  interval_df
}

# Run for all samples and save big CSV
plan(multisession)

all_interval_summaries <- future_pmap_dfr(
  list(colony_runs$Colony_ID, colony_runs$Run),
  fit_colony_intervals_avgblank
)

write_csv(all_interval_summaries, "../output/pi_curves/Sept2023_TCB/all_interval_summaries_Sept2023_TCB.csv")
```

Read in interval data again so the above chunk does not need to be run
```{r}
all_int <- read.csv("../output/pi_curves/Sept2023_TCB/all_interval_summaries_Sept2023_TCB.csv")
all_int$Colony_ID <- as.character(all_int$Colony_ID)
```

Make sure surface area data is numeric 
```{r}
sample_meta$Surface_Area_cm2 <- as.numeric(sample_meta$Surface_Area_cm2)
```

Merge sample metadata with interval data 
```{r}
all_int <- all_int %>%
  inner_join(sample_meta, by = "Colony_ID") %>%
  filter(!is.na(Phenotype))
length(unique(all_int$Colony_ID))
```

Normalize interval data by chamber volume and surface area and save
```{r}
# normalize 
all_int <- all_int %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Blank_Corrected_Slope * Chamber_Volume_L / Surface_Area_cm2) * 60) # units are umol O2 cm-2 h-1
write.csv(all_int, "../output/pi_curves/Sept2023_TCB/all_interval_summaries_normalized_Sept2023_TCB.csv")
```

Plot PI curves across light values 
```{r}
ggplot(all_int, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype")
```

### Fit Platt model to determine PI curve metrics 

- Pmax 
- alpha 
- Ik 
- LEDR
- Pgross 

Nest by colony 
```{r}
pi_nested <- all_int %>%
  select(Colony_ID, Light_Value, Slope_norm) %>%  # or whatever your P column is
  rename(I = Light_Value, P = Slope_norm) %>%
  group_by(Colony_ID) %>%
  nest()
```

Define the Platt model and fit for each colony 
```{r}
# Define a function to fit the model
fit_pi_model <- function(df) {
  tryCatch({
    fit <- nlsLM(P ~ Pmax * tanh(alpha * I / Pmax) + Rd,
                 data = df,
                 start = list(Pmax = max(df$P), alpha = 0.01, Rd = -0.1))
    pars <- coef(fit)
    tibble(Pmax = pars["Pmax"],
           alpha = pars["alpha"],
           Rd = pars["Rd"],
           Ik = pars["Pmax"]/pars["alpha"],
           Pgross = pars["Pmax"] + pars["Rd"])
  }, error = function(e) tibble(Pmax = NA, alpha = NA, Rd = NA, Ik = NA, Pgross = NA))
}

# Apply the model to each colony
pi_results <- pi_nested %>%
  mutate(params = map(data, fit_pi_model)) %>%
  select(-data) %>%
  unnest(cols = params)
```

Merge pi curve metric data with sample metadata and save 
```{r}
pi_results <- pi_results %>%
  inner_join(sample_meta, by = "Colony_ID")
write.csv(pi_results, "../output/pi_curves/Sept2023_TCB/pi_curve_metrics_Sept2023_TCB.csv")
```

Pivot pi data to long format for plotting 
```{r}
pi_long <- pi_results %>%
  pivot_longer(cols = c(Pmax, alpha, Rd, Ik, Pgross),
               names_to = "Parameter",
               values_to = "Value")
```

Plot metrics 
```{r}
ggplot(pi_long, aes(x = Phenotype, y = Value, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, size = 1.5, alpha = 0.8) +
  facet_grid(Parameter ~ Species, scales = "free_y") +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw(base_size = 14) +
  labs(x = "Phenotype", y = "Value") +
  theme(legend.position = "none")
```

## Field vs TCB corals 

Read in normalized interval data for field and TCB corals from Sept 2023
```{r}
int_field<- read.csv("../output/pi_curves/Sept2023/all_interval_summaries_normalized_Sept2023.csv")
int_field$Project <- "Field"
int_field <- int_field %>%
  select(Colony_ID, Run.x, Light_Value, Date, Species, Phenotype, Slope_norm, Project)

int_tcb <- read.csv("../output/pi_curves/Sept2023_TCB/all_interval_summaries_normalized_Sept2023_TCB.csv")
int_tcb <- int_tcb %>%
  select(Colony_ID, Run.x, Light_Value, Date, Species, Phenotype, Slope_norm, Project)
```

Bind data together 
```{r}
all_int <- rbind(int_field, int_tcb)
```

Remove 237 from field -- major outlier 
```{r}
all_int_filt <- all_int %>%
  filter(!(Colony_ID == "237" & Project == "Field"))
```

Plot PI curves across light values 
```{r}
ggplot(all_int_filt, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  #facet_wrap(~Species+Timepoint, scales = "fixed") +
  facet_grid(rows = vars(Species), cols = vars(Project)) +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype")

ggplot(all_int_filt, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1.5, position = position_jitter(width = 10)) +
  geom_smooth(aes(fill = Phenotype), method = "loess", se = TRUE, linewidth = 1.2) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  #facet_wrap(~Species + Timepoint, scales = "fixed") +
  facet_grid(rows = vars(Species), cols = vars(Project)) +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype", fill = "Phenotype")
```

### PI curve metrics 

Read in calculated pi curve metrics 
```{r}
pi_metrics_field <- read.csv("../output/pi_curves/Sept2023/pi_curve_metrics_Sept2023.csv")

pi_metrics_tcb <- read.csv("../output/pi_curves/Sept2023_TCB/pi_curve_metrics_Sept2023_TCB.csv")
```

Bind data together 
```{r}
all_pi_metrics <- rbind(pi_metrics_field, pi_metrics_tcb)
```

#### Stats 

##### Pmax 

Assess normality of data 
```{r}
hist(all_pi_metrics$Pmax)

# Q-Q Plot
qqnorm(all_pi_metrics$Pmax)
qqline(all_pi_metrics$Pmax, col = "red")

# Shapiro-Wilk Test
shapiro.test(all_pi_metrics$Pmax)
```

Data is normal. 

Perform stats
```{r}
mod_pmax <- lmer(Pmax ~ Species * Phenotype * Project + (1 | Colony_ID),
                 data = all_pi_metrics,
                 REML = FALSE)
anova(mod_pmax, type = 3)   # type III sums of squares
emmeans(mod_pmax, pairwise ~ Species)
emmeans(mod_pmax, pairwise ~ Project)
```

Species and Project are significant. 

##### Alpha 

Assess normality of data 
```{r}
hist(all_pi_metrics$alpha)

# Q-Q Plot
qqnorm(all_pi_metrics$alpha)
qqline(all_pi_metrics$alpha, col = "red")

# Shapiro-Wilk Test
shapiro.test(all_pi_metrics$alpha)
```

Data is not normal, perform data transformation
```{r}
# Log10
all_pi_metrics$alpha_log <- log10(all_pi_metrics$alpha +1) # add 1 to avoid log10(0)
hist(all_pi_metrics$alpha_log)
qqnorm(all_pi_metrics$alpha_log)
qqline(all_pi_metrics$alpha_log, col = "red")
shapiro.test(all_pi_metrics$alpha_log)

# Sqrt
all_pi_metrics$alpha_sqrt <- sqrt(all_pi_metrics$alpha)
hist(all_pi_metrics$alpha_sqrt)
qqnorm(all_pi_metrics$alpha_sqrt)
qqline(all_pi_metrics$alpha_sqrt, col = "red")
shapiro.test(all_pi_metrics$alpha_sqrt)

# Inverse 
all_pi_metrics$alpha_inv <- 1 / (all_pi_metrics$alpha +1) # add 1 to avoid division by zero
hist(all_pi_metrics$alpha_inv)
qqnorm(all_pi_metrics$alpha_inv)
qqline(all_pi_metrics$alpha_inv, col = "red")
shapiro.test(all_pi_metrics$alpha_inv)
```

Only sqrt is normal. Proceed with sqrt 

Perform stats
```{r}
mod_alpha <- lmer(alpha_sqrt ~ Species * Phenotype * Project + (1 | Colony_ID),
                 data = all_pi_metrics,
                 REML = FALSE)
anova(mod_alpha, type = 3)   # type III sums of squares
emmeans(mod_alpha, pairwise ~ Phenotype)
emmeans(mod_alpha, pairwise ~ Project)
emmeans(mod_alpha, pairwise ~ Phenotype:Project)
emmeans(mod_alpha, pairwise ~ Phenotype:Project:Species)
```

Phenotype, Project, Phenotype:Project, and Species:Phenotype:Project are significant

##### Ik 

Assess normality of data 
```{r}
hist(all_pi_metrics$Ik)

# Q-Q Plot
qqnorm(all_pi_metrics$Ik)
qqline(all_pi_metrics$Ik, col = "red")

# Shapiro-Wilk Test
shapiro.test(all_pi_metrics$Ik)
```

Data is not normal, perform data transformation
```{r}
# Log10
all_pi_metrics$Ik_log <- log10(all_pi_metrics$Ik +1) # add 1 to avoid log10(0)
hist(all_pi_metrics$Ik_log)
qqnorm(all_pi_metrics$Ik_log)
qqline(all_pi_metrics$Ik_log, col = "red")
shapiro.test(all_pi_metrics$Ik_log)

# Sqrt
all_pi_metrics$Ik_sqrt <- sqrt(all_pi_metrics$Ik)
hist(all_pi_metrics$Ik_sqrt)
qqnorm(all_pi_metrics$Ik_sqrt)
qqline(all_pi_metrics$Ik_sqrt, col = "red")
shapiro.test(all_pi_metrics$Ik_sqrt)

# Inverse 
all_pi_metrics$Ik_inv <- 1 / (all_pi_metrics$Ik +1) # add 1 to avoid division by zero
hist(all_pi_metrics$Ik_inv)
qqnorm(all_pi_metrics$Ik_inv)
qqline(all_pi_metrics$Ik_inv, col = "red")
shapiro.test(all_pi_metrics$Ik_inv)
```

Both log and sqrt are normal. Proceed with log. 

Perform stats 
```{r}
mod_ik <- lmer(Ik_log ~ Species * Phenotype * Project + (1 | Colony_ID),
                 data = all_pi_metrics,
                 REML = FALSE)
anova(mod_ik, type = 3)   # type III sums of squares
emmeans(mod_alpha, pairwise ~ Species)
emmeans(mod_alpha, pairwise ~ Project)
```

Species and Project (sort of) are significant. 

##### Rd

Assess normality of data 
```{r}
hist(all_pi_metrics$Rd)

# Q-Q Plot
qqnorm(all_pi_metrics$Rd)
qqline(all_pi_metrics$Rd, col = "red")

# Shapiro-Wilk Test
shapiro.test(all_pi_metrics$Rd)
```

Data is normal. 

Perform stats 
```{r}
mod_rd <- lmer(Rd ~ Species * Phenotype * Project + (1 | Colony_ID),
                 data = all_pi_metrics,
                 REML = FALSE)
anova(mod_rd, type = 3)   # type III sums of squares
emmeans(mod_rd, pairwise ~ Species)
emmeans(mod_rd, pairwise ~ Phenotype)
emmeans(mod_rd, pairwise ~ Species:Phenotype:Project)
```

Species, Phenotype (sort of), and Species:Phenotype:Project are significant. 

##### Pgross

Assess normality of data 
```{r}
hist(all_pi_metrics$Pgross)

# Q-Q Plot
qqnorm(all_pi_metrics$Pgross)
qqline(all_pi_metrics$Pgross, col = "red")

# Shapiro-Wilk Test
shapiro.test(all_pi_metrics$Pgross)
```

Data is not normal, perform data transformation
```{r}
# Log10
all_pi_metrics$Pgross_log <- log10(all_pi_metrics$Pgross +1) # add 1 to avoid log10(0)
hist(all_pi_metrics$Pgross_log)
qqnorm(all_pi_metrics$Pgross_log)
qqline(all_pi_metrics$Pgross_log, col = "red")
shapiro.test(all_pi_metrics$Pgross_log)

# Sqrt
all_pi_metrics$Pgross_sqrt <- sqrt(all_pi_metrics$Pgross)
hist(all_pi_metrics$Pgross_sqrt)
qqnorm(all_pi_metrics$Pgross_sqrt)
qqline(all_pi_metrics$Pgross_sqrt, col = "red")
shapiro.test(all_pi_metrics$Pgross_sqrt)

# Inverse 
all_pi_metrics$Pgross_inv <- 1 / (all_pi_metrics$Pgross +1) # add 1 to avoid division by zero
hist(all_pi_metrics$Pgross_inv)
qqnorm(all_pi_metrics$Pgross_inv)
qqline(all_pi_metrics$Pgross_inv, col = "red")
shapiro.test(all_pi_metrics$Pgross_inv)
```

All produced normality. Proceed with log 

Perform stats 
```{r}
mod_pgross <- lmer(Pgross_log ~ Species * Phenotype * Project + (1 | Colony_ID),
                 data = all_pi_metrics,
                 REML = FALSE)
anova(mod_pgross, type = 3)   # type III sums of squares
emmeans(mod_pgross, pairwise ~ Species)
emmeans(mod_pgross, pairwise ~ Project)
emmeans(mod_pgross, pairwise ~ Species:Phenotype:Project)
```

Species and Project are significant. 

### Plot
```{r}
# Subset 'CAREER - Phys' to 'Field'
all_pi_metrics_sub <- all_pi_metrics %>%
  mutate(Project = ifelse(Project == "CAREER - Phys", "Field", Project))

# Select and reshape the variables for plotting
plot_data <- all_pi_metrics_sub %>%
  select(Project, Species, Phenotype, Pmax, alpha, Ik, Rd, Pgross) %>%
  pivot_longer(cols = c(Pmax, alpha, Ik, Rd, Pgross),
               names_to = "Metric", values_to = "Value")

# Plot boxplot with jitter, colored by Phenotype
ggplot(plot_data, aes(x = Project, y = Value, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6, position = position_dodge(width = 0.8)) +
  geom_jitter(aes(color = Phenotype), position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), 
              size = 1.5, alpha = 0.8, show.legend = FALSE) +
  facet_grid(Metric ~ Species, scales = "free_y") +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw(base_size = 14) +
  labs(x = "Project", y = "Value") +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust = 1))

```

