---
title: "PI curves KBay pairs 2023-2025"
author: "Jill Ashey"
date: "2025-08-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("devtools")
#devtools::install_github("colin-olito/LoLinR")

library(tidyverse)
library(nls.multstart)
library(broom)
library(LoLinR)
library(readr)
library(lubridate)
library(fuzzyjoin)
library(stringr)
library(purrr)
library(future)
library(furrr)
library(minpack.lm)
library(lme4)
library(lmerTest)
library(nlme)
library(emmeans)
```

### Look at run and sample metadata first 

Run metadata
```{r}
run_meta <- read_csv("../data/pi_curves/pi_curve_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))

# Summarize light values 
light_summary <- run_meta %>%
  group_by(Light_Level_Percent) %>%
  summarise(
    mean_light = mean(Light_Value, na.rm = TRUE),
    sd_light = sd(Light_Value, na.rm = TRUE),
    sem_light = sd(Light_Value, na.rm = TRUE) / sqrt(sum(!is.na(Light_Value))),
    .groups = "drop"
  )
```

Load sample metadata
```{r}
sample_meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```


## Sept 2023 

### Prepare data 

Load run metadata 
```{r}
run_meta <- read_csv("../data/pi_curves/pi_curve_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
```

Load sample metadata
```{r}
sample_meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```

List all raw files for Sept 2023 
```{r}
files <- list.files(path = "../data/pi_curves/raw_data/Sept2023/", pattern = "*.csv", full.names = TRUE)
```

Join O2 data with metadata files based on start and stop times 
```{r}
all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "channel\\d+") %>% str_remove("channel") %>% as.numeric()
  
  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )
  
  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )
  
  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )
  
  return(raw_joined)
})
```

Look at joined data and save
```{r}
all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

write_csv(all_data, "../output/pi_curves/Sept2023/all_joined_data_Sept2023.csv")
```

### Run localized linear regressions 

The interval fitting code will take about an hour to run and is computationally intense. If needed, run as a background job or on a remote server.

Make list of all Colony_ID / Run combinations (including blanks)
```{r}
colony_runs <- all_data %>%
  filter(!is.na(Colony_ID)) %>%
  distinct(Colony_ID, Run)
```

For this timepoint, only two out of the five runs had blanks. Therefore, I am going to calculate the blank intervals, average them, and then apply them to the regression for each colony. 

Calculate interval slopes for the blanks
```{r}
blank_data <- all_data %>%
  filter(Colony_ID == "blank")

blank_intervals <- blank_data %>%
  group_by(Run, Light_Value) %>%
  filter(!is.na(Colony_ID)) %>%
  group_split()
```

Fit regressions and extract slopes for each blank interval
```{r}
fit_reg <- function(df) {
  rankLocReg(
    xall = as.numeric(df$`Delta T [min]`),
    yall = df$Oxygen,
    alpha = 0.2,
    method = "pc",
    verbose = FALSE
  )
}

blank_fits <- map(blank_intervals, fit_reg)
blank_slopes <- map_dbl(blank_fits, ~ pluck(., "allRegs", "b1", 1))
blank_summary <- map2_dfr(
  blank_intervals,
  blank_slopes,
  ~ tibble(
    Light_Value = unique(.x$Light_Value),
    Blank_Slope = .y
  )
)
```

Average blank slopes per light value, across all blanks/runs
```{r}
avg_blank_slopes <- blank_summary %>%
  group_by(Light_Value) %>%
  summarise(Avg_Blank_Slope = mean(Blank_Slope, na.rm = TRUE), .groups = "drop")
```

Run interval fittings with blank correction
```{r}
# Function: blank correction + slope extraction + save/return interval summary
fit_colony_intervals_avgblank <- function(colony_id, run_num) {
  # Subset sample data
  sample_data <- all_data %>%
    filter(Colony_ID == colony_id, Run == run_num)
  if (nrow(sample_data) == 0) return(NULL)
  
  # Split data by light step
  intervals <- sample_data %>%
    group_by(Light_Value) %>%
    group_split()
  
  # Fit regressions for each interval
  fits <- map(intervals, fit_reg)
  micromol.L.s <- map_dbl(fits, ~ pluck(., "allRegs", "b1", 1))
  
  # Detailed interval summary
  interval_df <- map2_dfr(
    intervals,
    seq_along(intervals),
    ~ tibble(
      Colony_ID   = colony_id,
      Run         = run_num,
      Interval    = .y,
      Light_Value = unique(.x$Light_Value),
      N_points    = nrow(.x),
      O2_start    = head(.x$Oxygen, 1),
      O2_end      = tail(.x$Oxygen, 1),
      DeltaT_min  = min(.x$`Delta T [min]`),
      DeltaT_max  = max(.x$`Delta T [min]`),
      O2_min      = min(.x$Oxygen, na.rm = TRUE),
      O2_max      = max(.x$Oxygen, na.rm = TRUE),
      Slope       = micromol.L.s[.y]
    )
  ) %>%
    left_join(avg_blank_slopes, by = "Light_Value") %>%
    mutate(
      Blank_Corrected_Slope = Slope - Avg_Blank_Slope
    )
  
  # Write file (outside of the pipeline!)
  fn <- paste0("../output/pi_curves/Sept2023/intervals_", colony_id, "_run", run_num, "_Sept2023.csv")
  write_csv(interval_df, fn)
  
  interval_df
}

# Run for all samples and save big CSV
plan(multisession)

all_interval_summaries <- future_pmap_dfr(
  list(colony_runs$Colony_ID, colony_runs$Run),
  fit_colony_intervals_avgblank
)

write_csv(all_interval_summaries, "../output/pi_curves/Sept2023/all_interval_summaries_Sept2023.csv")
```

Read in interval data again so the above chunk does not need to be run
```{r}
all_int <- read.csv("../output/pi_curves/Sept2023/all_interval_summaries_Sept2023.csv")
all_int$Colony_ID <- as.character(all_int$Colony_ID)
```

Filter sample metadata for Sept 2023
```{r}
sept_2023 <- sample_meta %>%
  filter(grepl("^2023-09", Date))
sept_2023$Surface_Area_cm2 <- as.numeric(sept_2023$Surface_Area_cm2)
```

Merge sample metadata with interval data 
```{r}
all_int <- all_int %>%
  inner_join(sept_2023, by = "Colony_ID") %>%
  filter(!is.na(Phenotype))
length(unique(all_int$Colony_ID))
```

Normalize interval data by chamber volume and surface area and save
```{r}
# normalize 
all_int <- all_int %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Blank_Corrected_Slope * Chamber_Volume_L / Surface_Area_cm2) * 60) # units are umol O2 cm-2 h-1
write.csv(all_int, "../output/pi_curves/Sept2023/all_interval_summaries_normalized_Sept2023.csv")
```

Plot PI curves across light values 
```{r}
ggplot(all_int, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype")
```

Something is off with 237 at the 150 light level. The slope is -19. Remove for further analysis 
```{r}
all_int_sub <- all_int %>%
  filter(!Colony_ID == 237)
```

### Fit Platt model to determine PI curve metrics 

- Pmax 
- alpha 
- Ik 
- LEDR
- Pgross 

Nest by colony 
```{r}
pi_nested <- all_int %>%
  select(Colony_ID, Light_Value, Slope_norm) %>%  # or whatever your P column is
  rename(I = Light_Value, P = Slope_norm) %>%
  group_by(Colony_ID) %>%
  nest()
```

Define the Platt model and fit for each colony 
```{r}
# Define a function to fit the model
fit_pi_model <- function(df) {
  tryCatch({
    fit <- nlsLM(P ~ Pmax * tanh(alpha * I / Pmax) + Rd,
                 data = df,
                 start = list(Pmax = max(df$P), alpha = 0.01, Rd = -0.1))
    pars <- coef(fit)
    tibble(Pmax = pars["Pmax"],
           alpha = pars["alpha"],
           Rd = pars["Rd"],
           Ik = pars["Pmax"]/pars["alpha"],
           Pgross = pars["Pmax"] + pars["Rd"])
  }, error = function(e) tibble(Pmax = NA, alpha = NA, Rd = NA, Ik = NA, Pgross = NA))
}

# Apply the model to each colony
pi_results <- pi_nested %>%
  mutate(params = map(data, fit_pi_model)) %>%
  select(-data) %>%
  unnest(cols = params)
```

Merge pi curve metric data with sample metadata and save 
```{r}
pi_results <- pi_results %>%
  inner_join(sept_2023, by = "Colony_ID")
write.csv(pi_results, "../output/pi_curves/Sept2023/pi_curve_metrics_Sept2023.csv")
```

Pivot pi data to long format for plotting 
```{r}
pi_long <- pi_results %>%
  pivot_longer(cols = c(Pmax, alpha, Rd, Ik, Pgross),
               names_to = "Parameter",
               values_to = "Value")
```

Plot metrics 
```{r}
ggplot(pi_long, aes(x = Phenotype, y = Value, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, size = 1.5, alpha = 0.8) +
  facet_grid(Parameter ~ Species, scales = "free_y") +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw(base_size = 14) +
  labs(x = "Phenotype", y = "Value") +
  theme(legend.position = "none")
```

## Jan 2024 

### Prepare data 

Load run metadata 
```{r}
run_meta <- read_csv("../data/pi_curves/pi_curve_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
```

Load sample metadata
```{r}
sample_meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```

List all raw files for Jan 2024 
```{r}
files <- list.files(path = "../data/pi_curves/raw_data/Jan2024", pattern = "*.csv", full.names = TRUE)
```

Join O2 data with metadata files based on start and stop times 
```{r}
all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "channel\\d+") %>% str_remove("channel") %>% as.numeric()
  
  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )
  
  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )
  
  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )
  
  return(raw_joined)
})
```

Look at the joined data and save 
```{r}
all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

write_csv(all_data, "../output/pi_curves/Jan2024/all_joined_data_Jan2024.csv")
```

### Run localized linear regressions 

The interval fitting code will take about an hour to run and is computationally intense. If needed, run as a background job or on a remote server.

Make list of all Colony_ID / Run combinations (including blanks)
```{r}
colony_runs <- all_data %>%
  filter(!is.na(Colony_ID)) %>%
  distinct(Colony_ID, Run)
```

For this timepoint, only two out of the five runs had blanks. Therefore, I am going to calculate the blank intervals, average them, and then apply them to the regression for each colony. 

Calculate interval slopes for the blanks 
```{r}
blank_data <- all_data %>%
  filter(Colony_ID == "blank")

blank_intervals <- blank_data %>%
  group_by(Run, Light_Value) %>%
  group_split()
```

Fit regressions and extract slopes for each blank interval
```{r}
fit_reg <- function(df) {
  rankLocReg(
    xall = as.numeric(df$`Delta T [min]`),
    yall = df$Oxygen,
    alpha = 0.2,
    method = "pc",
    verbose = FALSE
  )
}

blank_fits <- map(blank_intervals, fit_reg)
blank_slopes <- map_dbl(blank_fits, ~ pluck(., "allRegs", "b1", 1))
blank_summary <- map2_dfr(
  blank_intervals,
  blank_slopes,
  ~ tibble(
    Light_Value = unique(.x$Light_Value),
    Blank_Slope = .y
  )
)
```

Average blank slopes per light value, across all blanks/runs
```{r}
avg_blank_slopes <- blank_summary %>%
  group_by(Light_Value) %>%
  summarise(Avg_Blank_Slope = mean(Blank_Slope, na.rm = TRUE), .groups = "drop")
```

Now run interval fittings with blank correction
```{r}
# Function: blank correction + slope extraction + save/return interval summary
fit_colony_intervals_avgblank <- function(colony_id, run_num) {
  # Sample data for this colony/run
  sample_data <- all_data %>%
    filter(Colony_ID == colony_id, Run == run_num)
  if (nrow(sample_data) == 0) return(NULL)
  
  # Split into intervals by light step
  intervals <- sample_data %>%
    group_by(Light_Value) %>%
    group_split()
  
  # Fit local regression for each light step
  fits <- map(intervals, fit_reg)
  micromol.L.s <- map_dbl(fits, ~ pluck(., "allRegs", "b1", 1))
  light_levels <- map_dbl(intervals, ~ unique(.x$Light_Value))
  
  # Build interval summary table
  interval_df <- tibble(
    Colony_ID   = colony_id,
    Run         = run_num,
    Light_Value = light_levels,
    Raw_Slope   = micromol.L.s
  ) %>%
    left_join(avg_blank_slopes, by = "Light_Value") %>%
    mutate(Blank_Corrected_Slope = Raw_Slope - Avg_Blank_Slope)
  
  # Save per colony/run (optional)
  fn <- paste0("../output/pi_curves/Jan2024/intervals_", colony_id, "_run", run_num, "_Jan2024.csv")
  write_csv(interval_df, fn)
  
  return(interval_df)
}

# Process all samples and save combined output
plan(multisession)

all_interval_summaries <- future_pmap_dfr(
  list(colony_runs$Colony_ID, colony_runs$Run),
  fit_colony_intervals_avgblank
)

write_csv(all_interval_summaries, "../output/pi_curves/Jan2024/all_interval_summaries_Jan2024.csv")

```

Read in interval data again so the above chunk does not need to be run
```{r}
all_int <- read.csv("../output/pi_curves/Jan2024/all_interval_summaries_Jan2024.csv")
all_int$Colony_ID <- as.character(all_int$Colony_ID)
```

Filter sample metadata for Jan 2024
```{r}
jan_2024 <- sample_meta %>%
  filter(grepl("^2024-01", Date))
jan_2024$Surface_Area_cm2 <- as.numeric(jan_2024$Surface_Area_cm2)
```

Merge sample metadata with interval data 
```{r}
all_int <- all_int %>%
  inner_join(jan_2024, by = "Colony_ID") %>%
  filter(!is.na(Phenotype))
length(unique(all_int$Colony_ID))
```

Normalize interval data by chamber volume and surface area and save
```{r}
# normalize 
all_int <- all_int %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Blank_Corrected_Slope * Chamber_Volume_L / Surface_Area_cm2) * 60) # units are umol O2 cm-2 h-1
write.csv(all_int, "../output/pi_curves/Jan2024/all_interval_summaries_normalized_Jan2024.csv")
```

Plot PI curves across light values 
```{r}
ggplot(all_int, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype")
```

### Fit Platt model to determine PI curve metrics 

- Pmax 
- alpha 
- Ik 
- LEDR
- Pgross 

Nest by colony 
```{r}
pi_nested <- all_int %>%
  select(Colony_ID, Light_Value, Slope_norm) %>%  # or whatever your P column is
  rename(I = Light_Value, P = Slope_norm) %>%
  group_by(Colony_ID) %>%
  nest()
```

Define the Platt model and fit for each colony 
```{r}
# Define a function to fit the model
fit_pi_model <- function(df) {
  tryCatch({
    fit <- nlsLM(P ~ Pmax * tanh(alpha * I / Pmax) + Rd,
                 data = df,
                 start = list(Pmax = max(df$P), alpha = 0.01, Rd = -0.1))
    pars <- coef(fit)
    tibble(Pmax = pars["Pmax"],
           alpha = pars["alpha"],
           Rd = pars["Rd"],
           Ik = pars["Pmax"]/pars["alpha"],
           Pgross = pars["Pmax"] + pars["Rd"])
  }, error = function(e) tibble(Pmax = NA, alpha = NA, Rd = NA, Ik = NA, Pgross = NA))
}

# Apply the model to each colony
pi_results <- pi_nested %>%
  mutate(params = map(data, fit_pi_model)) %>%
  select(-data) %>%
  unnest(cols = params)
```

Merge pi curve metric data with sample metadata and save 
```{r}
pi_results <- pi_results %>%
  inner_join(jan_2024, by = "Colony_ID")
write.csv(pi_results, "../output/pi_curves/Jan2024/pi_curve_metrics_Jan2024.csv")
```

Pivot pi data to long format for plotting 
```{r}
pi_long <- pi_results %>%
  pivot_longer(cols = c(Pmax, alpha, Rd, Ik, Pgross),
               names_to = "Parameter",
               values_to = "Value")
```

Plot metrics 
```{r}
ggplot(pi_long, aes(x = Phenotype, y = Value, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, size = 1.5, alpha = 0.8) +
  facet_grid(Parameter ~ Species, scales = "free_y") +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw(base_size = 14) +
  labs(x = "Phenotype", y = "Value") +
  theme(legend.position = "none")
```















## May 2024 

### Prepare data 

Load run metadata 
```{r}
run_meta <- read_csv("../data/pi_curves/pi_curve_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
```

Load sample metadata
```{r}
sample_meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```

List all raw files for Jan 2024 
```{r}
files <- list.files(path = "../data/pi_curves/raw_data/May2024", pattern = "*.csv", full.names = TRUE)
```

Join O2 data with metadata files based on start and stop times 
```{r}
all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "channel\\d+") %>% str_remove("channel") %>% as.numeric()
  
  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )
  
  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )
  
  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )
  
  return(raw_joined)
})
```

Look at the joined data and save 
```{r}
all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

write_csv(all_data, "../output/pi_curves/May2024/all_joined_data_May2024.csv")
```

### Run localized linear regressions 

The interval fitting code will take about an hour to run and is computationally intense. If needed, run as a background job or on a remote server.

Make list of all Colony_ID / Run combinations (including blanks)
```{r}
colony_runs <- all_data %>%
  filter(!is.na(Colony_ID)) %>%
  distinct(Colony_ID, Run)
```

Run interval fittings with blank correction
```{r}
# Function: blank correction + slope extraction + save/return interval summary
fit_colony_intervals <- function(colony_id, run_num) {
  
  # Sample data for this colony/run
  sample_data <- all_data %>%
    filter(Colony_ID == colony_id, Run == run_num)
  
  # Skip if no data
  if (nrow(sample_data) == 0) return(NULL)
  
  # Blank data for same run
  blank_data <- all_data %>%
    filter(Colony_ID == "blank", Run == run_num)
  
  # Blank correction if possible
  if (nrow(blank_data) > 0) {
    blank_means <- blank_data %>%
      group_by(Light_Value) %>%
      summarise(blank_O2 = mean(Oxygen, na.rm = TRUE), .groups = "drop")
    
    sample_data <- sample_data %>%
      left_join(blank_means, by = "Light_Value") %>%
      mutate(Oxygen = Oxygen - blank_O2)
  }
  
  # Split into intervals by light step
  intervals <- sample_data %>%
    group_by(Light_Value) %>%
    group_split()
  
  # Local regression
  fit_reg <- function(df) {
    rankLocReg(
      xall = as.numeric(df$`Delta T [min]`),
      yall = df$Oxygen,
      alpha = 0.2,
      method = "pc",
      verbose = FALSE
    )
  }
  
  fits <- map(intervals, fit_reg)
  
  # Slopes per interval
  micromol.L.s <- map_dbl(fits, ~ pluck(., "allRegs", "b1", 1))
  
  # Interval summary table
  interval_df <- map2_dfr(
    intervals,
    seq_along(intervals),
    ~ tibble(
      Colony_ID   = colony_id,
      Run         = run_num,
      Interval    = .y,
      Light_Value = unique(.x$Light_Value),
      N_points    = nrow(.x),
      O2_start    = head(.x$Oxygen, 1),
      O2_end      = tail(.x$Oxygen, 1),
      DeltaT_min  = min(.x$`Delta T [min]`),
      DeltaT_max  = max(.x$`Delta T [min]`),
      O2_min      = min(.x$Oxygen, na.rm = TRUE),
      O2_max      = max(.x$Oxygen, na.rm = TRUE),
      Slope       = micromol.L.s[.y]
    )
  )
  
  # Save file for this colony/run
  fn <- paste0("../output/pi_curves/May2024/intervals_", colony_id, "_run", run_num, "_May2024.csv")
  write_csv(interval_df, fn)
  
  return(interval_df)
}

# Run for all colony/run combos in parallel
plan(multisession)

all_interval_summaries <- future_pmap_dfr(
  list(colony_runs$Colony_ID, colony_runs$Run),
  fit_colony_intervals
)

# Save combined CSV
write_csv(all_interval_summaries, "../output/pi_curves/May2024/all_interval_summaries_May2024.csv")
```

Read in interval data again so the above chunk does not need to be run
```{r}
all_int <- read.csv("../output/pi_curves/May2024/all_interval_summaries_May2024.csv")
all_int$Colony_ID <- as.character(all_int$Colony_ID)
```

Filter sample metadata for May 2024
```{r}
may_2024 <- sample_meta %>%
  filter(grepl("^2024-05", Date))
may_2024$Surface_Area_cm2 <- as.numeric(may_2024$Surface_Area_cm2)
```

Merge sample metadata with interval data 
```{r}
all_int <- all_int %>%
  inner_join(may_2024, by = "Colony_ID") %>%
  filter(!is.na(Phenotype))
length(unique(all_int$Colony_ID))
```

Normalize interval data by chamber volume and surface area and save
```{r}
# normalize 
all_int <- all_int %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Slope * Chamber_Volume_L / Surface_Area_cm2) * 60) # units are umol O2 cm-2 h-1
write.csv(all_int, "../output/pi_curves/May2024/all_interval_summaries_normalized_May2024.csv")
```

Plot PI curves across light values 
```{r}
ggplot(all_int, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype")
```

### Fit Platt model to determine PI curve metrics 

- Pmax 
- alpha 
- Ik 
- LEDR
- Pgross 

Nest by colony 
```{r}
pi_nested <- all_int %>%
  select(Colony_ID, Light_Value, Slope_norm) %>%  # or whatever your P column is
  rename(I = Light_Value, P = Slope_norm) %>%
  group_by(Colony_ID) %>%
  nest()
```

Define the Platt model and fit for each colony 
```{r}
# Define a function to fit the model
fit_pi_model <- function(df) {
  tryCatch({
    fit <- nlsLM(P ~ Pmax * tanh(alpha * I / Pmax) + Rd,
                 data = df,
                 start = list(Pmax = max(df$P), alpha = 0.01, Rd = -0.1))
    pars <- coef(fit)
    tibble(Pmax = pars["Pmax"],
           alpha = pars["alpha"],
           Rd = pars["Rd"],
           Ik = pars["Pmax"]/pars["alpha"],
           Pgross = pars["Pmax"] + pars["Rd"])
  }, error = function(e) tibble(Pmax = NA, alpha = NA, Rd = NA, Ik = NA, Pgross = NA))
}

# Apply the model to each colony
pi_results <- pi_nested %>%
  mutate(params = map(data, fit_pi_model)) %>%
  select(-data) %>%
  unnest(cols = params)
```

Merge pi curve metric data with sample metadata and save 
```{r}
pi_results <- pi_results %>%
  inner_join(may_2024, by = "Colony_ID")
write.csv(pi_results, "../output/pi_curves/May2024/pi_curve_metrics_May2024.csv")
```

Pivot pi data to long format for plotting 
```{r}
pi_long <- pi_results %>%
  pivot_longer(cols = c(Pmax, alpha, Rd, Ik, Pgross),
               names_to = "Parameter",
               values_to = "Value")
```

Plot metrics 
```{r}
ggplot(pi_long, aes(x = Phenotype, y = Value, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, size = 1.5, alpha = 0.8) +
  facet_grid(Parameter ~ Species, scales = "free_y") +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw(base_size = 14) +
  labs(x = "Phenotype", y = "Value") +
  theme(legend.position = "none")
```

## Oct 2024 

### Prepare data 

Load run metadata 
```{r}
run_meta <- read_csv("../data/pi_curves/pi_curve_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
```

Load sample metadata
```{r}
sample_meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```

List all raw files for Oct 2024 
```{r}
files <- list.files(path = "../data/pi_curves/raw_data/Oct2024", pattern = "*.csv", full.names = TRUE)
```

Join O2 data with metadata files based on start and stop times 
```{r}
all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "Ch\\d+") %>% str_remove("Ch") %>% as.numeric()
  
  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )
  
  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )
  
  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )
  
  return(raw_joined)
})
```

Look at the joined data and save 
```{r}
all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

write_csv(all_data, "../output/pi_curves/Oct2024/all_joined_data_Oct2024.csv")
```

### Run localized linear regressions 

The interval fitting code will take about an hour to run and is computationally intense. If needed, run as a background job or on a remote server.

Make list of all Colony_ID / Run combinations (including blanks)
```{r}
colony_runs <- all_data %>%
  filter(!is.na(Colony_ID)) %>%
  distinct(Colony_ID, Run)
```

Run interval fittings with blank correction
```{r}
# Function: blank correction + slope extraction + save/return interval summary
fit_colony_intervals <- function(colony_id, run_num) {
  
  # Sample data for this colony/run
  sample_data <- all_data %>%
    filter(Colony_ID == colony_id, Run == run_num)
  
  # Skip if no data
  if (nrow(sample_data) == 0) return(NULL)
  
  # Blank data for same run
  blank_data <- all_data %>%
    filter(Colony_ID == "blank", Run == run_num)
  
  # Blank correction if possible
  if (nrow(blank_data) > 0) {
    blank_means <- blank_data %>%
      group_by(Light_Value) %>%
      summarise(blank_O2 = mean(Oxygen, na.rm = TRUE), .groups = "drop")
    
    sample_data <- sample_data %>%
      left_join(blank_means, by = "Light_Value") %>%
      mutate(Oxygen = Oxygen - blank_O2)
  }
  
  # Split into intervals by light step
  intervals <- sample_data %>%
    group_by(Light_Value) %>%
    group_split()
  
  # Local regression
  fit_reg <- function(df) {
    rankLocReg(
      xall = as.numeric(df$`Delta T [min]`),
      yall = df$Oxygen,
      alpha = 0.2,
      method = "pc",
      verbose = FALSE
    )
  }
  
  fits <- map(intervals, fit_reg)
  
  # Slopes per interval
  micromol.L.s <- map_dbl(fits, ~ pluck(., "allRegs", "b1", 1))
  
  # Interval summary table
  interval_df <- map2_dfr(
    intervals,
    seq_along(intervals),
    ~ tibble(
      Colony_ID   = colony_id,
      Run         = run_num,
      Interval    = .y,
      Light_Value = unique(.x$Light_Value),
      N_points    = nrow(.x),
      O2_start    = head(.x$Oxygen, 1),
      O2_end      = tail(.x$Oxygen, 1),
      DeltaT_min  = min(.x$`Delta T [min]`),
      DeltaT_max  = max(.x$`Delta T [min]`),
      O2_min      = min(.x$Oxygen, na.rm = TRUE),
      O2_max      = max(.x$Oxygen, na.rm = TRUE),
      Slope       = micromol.L.s[.y]
    )
  )
  
  # Save file for this colony/run
  fn <- paste0("../output/pi_curves/Oct2024/intervals_", colony_id, "_run", run_num, "_Oct2024.csv")
  write_csv(interval_df, fn)
  
  return(interval_df)
}

# 3️⃣ Run for all colony/run combos in parallel
plan(multisession)

all_interval_summaries <- future_pmap_dfr(
  list(colony_runs$Colony_ID, colony_runs$Run),
  fit_colony_intervals
)

# Save combined CSV
write_csv(all_interval_summaries, "../output/pi_curves/Oct2024/all_interval_summaries_Oct2024.csv")
```

Read in interval data again so the above chunk does not need to be run
```{r}
all_int <- read.csv("../output/pi_curves/Oct2024/all_interval_summaries_Oct2024.csv")
all_int$Colony_ID <- as.character(all_int$Colony_ID)
```

Filter sample metadata for Oct 2024
```{r}
oct_2024 <- sample_meta %>%
  filter(grepl("^2024-10", Date))
oct_2024$Surface_Area_cm2 <- as.numeric(oct_2024$Surface_Area_cm2)
```

Merge sample metadata with interval data 
```{r}
all_int <- all_int %>%
  inner_join(oct_2024, by = "Colony_ID") %>%
  filter(!is.na(Phenotype))
length(unique(all_int$Colony_ID))
```

Normalize interval data by chamber volume and surface area and save
```{r}
# normalize 
all_int <- all_int %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Slope * Chamber_Volume_L / Surface_Area_cm2) * 60) # units are umol O2 cm-2 h-1
write.csv(all_int, "../output/pi_curves/Oct2024/all_interval_summaries_normalized_Oct2024.csv")
```

Plot PI curves across light values 
```{r}
ggplot(all_int, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype")
```

### Fit Platt model to determine PI curve metrics 

- Pmax 
- alpha 
- Ik 
- LEDR
- Pgross 

Nest by colony 
```{r}
pi_nested <- all_int %>%
  select(Colony_ID, Light_Value, Slope_norm) %>%  # or whatever your P column is
  rename(I = Light_Value, P = Slope_norm) %>%
  group_by(Colony_ID) %>%
  nest()
```

Define the Platt model and fit for each colony 
```{r}
# Define a function to fit the model
fit_pi_model <- function(df) {
  tryCatch({
    fit <- nlsLM(P ~ Pmax * tanh(alpha * I / Pmax) + Rd,
                 data = df,
                 start = list(Pmax = max(df$P), alpha = 0.01, Rd = -0.1))
    pars <- coef(fit)
    tibble(Pmax = pars["Pmax"],
           alpha = pars["alpha"],
           Rd = pars["Rd"],
           Ik = pars["Pmax"]/pars["alpha"],
           Pgross = pars["Pmax"] + pars["Rd"])
  }, error = function(e) tibble(Pmax = NA, alpha = NA, Rd = NA, Ik = NA, Pgross = NA))
}

# Apply the model to each colony
pi_results <- pi_nested %>%
  mutate(params = map(data, fit_pi_model)) %>%
  select(-data) %>%
  unnest(cols = params)
```

Merge pi curve metric data with sample metadata and save 
```{r}
pi_results <- pi_results %>%
  inner_join(oct_2024, by = "Colony_ID")
write.csv(pi_results, "../output/pi_curves/Oct2024/pi_curve_metrics_Oct2024.csv")
```

Pivot pi data to long format for plotting 
```{r}
pi_long <- pi_results %>%
  pivot_longer(cols = c(Pmax, alpha, Rd, Ik, Pgross),
               names_to = "Parameter",
               values_to = "Value")
```

Plot metrics 
```{r}
ggplot(pi_long, aes(x = Phenotype, y = Value, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, size = 1.5, alpha = 0.8) +
  facet_grid(Parameter ~ Species, scales = "free_y") +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw(base_size = 14) +
  labs(x = "Phenotype", y = "Value") +
  theme(legend.position = "none")
```

## Jan 2025

### Prepare data 

Load run metadata 
```{r}
run_meta <- read_csv("../data/pi_curves/pi_curve_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
```

Load sample metadata
```{r}
sample_meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```

List all raw files for Jan 2025 
```{r}
files <- list.files(path = "../data/pi_curves/raw_data/Jan2025", pattern = "*.csv", full.names = TRUE)
```

Join O2 data with metadata files based on start and stop times 
```{r}
all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "Ch\\d+") %>% str_remove("Ch") %>% as.numeric()
  
  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )
  
  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )
  
  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )
  
  return(raw_joined)
})
```

Look at the joined data and save 
```{r}
all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

write_csv(all_data, "../output/pi_curves/Jan2025/all_joined_data_Jan2025.csv")
```

### Run localized linear regressions 

The interval fitting code will take about an hour to run and is computationally intense. If needed, run as a background job or on a remote server.

Make list of all Colony_ID / Run combinations (including blanks)
```{r}
colony_runs <- all_data %>%
  filter(!is.na(Colony_ID)) %>%
  distinct(Colony_ID, Run)
```

Run interval fittings with blank correction
```{r}
# Function: blank correction + slope extraction + save/return interval summary
fit_colony_intervals <- function(colony_id, run_num) {
  
  # Sample data for this colony/run
  sample_data <- all_data %>%
    filter(Colony_ID == colony_id, Run == run_num)
  
  # Skip if no data
  if (nrow(sample_data) == 0) return(NULL)
  
  # Blank data for same run
  blank_data <- all_data %>%
    filter(Colony_ID == "blank", Run == run_num)
  
  # Blank correction if possible
  if (nrow(blank_data) > 0) {
    blank_means <- blank_data %>%
      group_by(Light_Value) %>%
      summarise(blank_O2 = mean(Oxygen, na.rm = TRUE), .groups = "drop")
    
    sample_data <- sample_data %>%
      left_join(blank_means, by = "Light_Value") %>%
      mutate(Oxygen = Oxygen - blank_O2)
  }
  
  # Split into intervals by light step
  intervals <- sample_data %>%
    group_by(Light_Value) %>%
    group_split()
  
  # Local regression
  fit_reg <- function(df) {
    rankLocReg(
      xall = as.numeric(df$`Delta T [min]`),
      yall = df$Oxygen,
      alpha = 0.2,
      method = "pc",
      verbose = FALSE
    )
  }
  
  fits <- map(intervals, fit_reg)
  
  # Slopes per interval
  micromol.L.s <- map_dbl(fits, ~ pluck(., "allRegs", "b1", 1))
  
  # Interval summary table
  interval_df <- map2_dfr(
    intervals,
    seq_along(intervals),
    ~ tibble(
      Colony_ID   = colony_id,
      Run         = run_num,
      Interval    = .y,
      Light_Value = unique(.x$Light_Value),
      N_points    = nrow(.x),
      O2_start    = head(.x$Oxygen, 1),
      O2_end      = tail(.x$Oxygen, 1),
      DeltaT_min  = min(.x$`Delta T [min]`),
      DeltaT_max  = max(.x$`Delta T [min]`),
      O2_min      = min(.x$Oxygen, na.rm = TRUE),
      O2_max      = max(.x$Oxygen, na.rm = TRUE),
      Slope       = micromol.L.s[.y]
    )
  )
  
  # Save file for this colony/run
  fn <- paste0("../output/pi_curves/Jan2025/intervals_", colony_id, "_run", run_num, "_Jan2025.csv")
  write_csv(interval_df, fn)
  
  return(interval_df)
}

# 3️⃣ Run for all colony/run combos in parallel
plan(multisession)

all_interval_summaries <- future_pmap_dfr(
  list(colony_runs$Colony_ID, colony_runs$Run),
  fit_colony_intervals
)

# Save combined CSV
write_csv(all_interval_summaries, "../output/pi_curves/Jan2025/all_interval_summaries_Jan2025.csv")
```

Read in interval data again so the above chunk does not need to be run
```{r}
all_int <- read.csv("../output/pi_curves/Jan2025/all_interval_summaries_Jan2025.csv")
all_int$Colony_ID <- as.character(all_int$Colony_ID)
```

Filter sample metadata for Jan 2025
```{r}
jan_2025 <- sample_meta %>%
  filter(grepl("^2025-01", Date))
jan_2025$Surface_Area_cm2 <- as.numeric(jan_2025$Surface_Area_cm2)
```

Merge sample metadata with interval data 
```{r}
all_int <- all_int %>%
  inner_join(jan_2025, by = "Colony_ID") %>%
  filter(!is.na(Phenotype))
length(unique(all_int$Colony_ID))
```

Set surface area to 20 for all colonies. I think I messed up the SA measurements because they are all quite low so I need to redo them
```{r}
all_int <- all_int %>%
  mutate(Surface_Area_cm2 = if_else(
      !is.na(Surface_Area_cm2), 20, NA_real_))
```


Normalize interval data by chamber volume and surface area and save
```{r}
# normalize 
all_int <- all_int %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Slope * Chamber_Volume_L / Surface_Area_cm2) * 60) # units are umol O2 cm-2 h-1
write.csv(all_int, "../output/pi_curves/Jan2025/all_interval_summaries_normalized_Jan2025.csv")
```

Plot PI curves across light values 
```{r}
ggplot(all_int, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype")
```

### Fit Platt model to determine PI curve metrics 

- Pmax 
- alpha 
- Ik 
- LEDR
- Pgross 

Nest by colony 
```{r}
pi_nested <- all_int %>%
  select(Colony_ID, Light_Value, Slope_norm) %>%  # or whatever your P column is
  rename(I = Light_Value, P = Slope_norm) %>%
  group_by(Colony_ID) %>%
  nest()
```

Define the Platt model and fit for each colony 
```{r}
# Define a function to fit the model
fit_pi_model <- function(df) {
  tryCatch({
    fit <- nlsLM(P ~ Pmax * tanh(alpha * I / Pmax) + Rd,
                 data = df,
                 start = list(Pmax = max(df$P), alpha = 0.01, Rd = -0.1))
    pars <- coef(fit)
    tibble(Pmax = pars["Pmax"],
           alpha = pars["alpha"],
           Rd = pars["Rd"],
           Ik = pars["Pmax"]/pars["alpha"],
           Pgross = pars["Pmax"] + pars["Rd"])
  }, error = function(e) tibble(Pmax = NA, alpha = NA, Rd = NA, Ik = NA, Pgross = NA))
}

# Apply the model to each colony
pi_results <- pi_nested %>%
  mutate(params = map(data, fit_pi_model)) %>%
  select(-data) %>%
  unnest(cols = params)
```

Merge pi curve metric data with sample metadata and save 
```{r}
pi_results <- pi_results %>%
  inner_join(jan_2025, by = "Colony_ID")
write.csv(pi_results, "../output/pi_curves/Jan2025/pi_curve_metrics_Jan2025.csv")
```

Pivot pi data to long format for plotting 
```{r}
pi_long <- pi_results %>%
  pivot_longer(cols = c(Pmax, alpha, Rd, Ik, Pgross),
               names_to = "Parameter",
               values_to = "Value")
```

Plot metrics 
```{r}
ggplot(pi_long, aes(x = Phenotype, y = Value, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, size = 1.5, alpha = 0.8) +
  facet_grid(Parameter ~ Species, scales = "free_y") +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw(base_size = 14) +
  labs(x = "Phenotype", y = "Value") +
  theme(legend.position = "none")
```

## May 2025

### Prepare data 

Load run metadata 
```{r}
run_meta <- read_csv("../data/pi_curves/pi_curve_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
```

Load sample metadata
```{r}
sample_meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```

List all raw files for May 2025 
```{r}
files <- list.files(path = "../data/pi_curves/raw_data/May2025", pattern = "*.csv", full.names = TRUE)
```

Join O2 data with metadata files based on start and stop times 
```{r}
all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "Ch\\d+") %>% str_remove("Ch") %>% as.numeric()
  
  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )
  
  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )
  
  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )
  
  return(raw_joined)
})
```

Look at the joined data and save 
```{r}
all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

write_csv(all_data, "../output/pi_curves/May2025/all_joined_data_May2025.csv")
```

### Run localized linear regressions 

The interval fitting code will take about an hour to run and is computationally intense. If needed, run as a background job or on a remote server.

Make list of all Colony_ID / Run combinations (including blanks)
```{r}
colony_runs <- all_data %>%
  filter(!is.na(Colony_ID)) %>%
  distinct(Colony_ID, Run)
```

Run interval fittings with blank correction
```{r}
# Function: blank correction + slope extraction + save/return interval summary
fit_colony_intervals <- function(colony_id, run_num) {
  
  # Sample data for this colony/run
  sample_data <- all_data %>%
    filter(Colony_ID == colony_id, Run == run_num)
  
  # Skip if no data
  if (nrow(sample_data) == 0) return(NULL)
  
  # Blank data for same run
  blank_data <- all_data %>%
    filter(Colony_ID == "blank", Run == run_num)
  
  # Blank correction if possible
  if (nrow(blank_data) > 0) {
    blank_means <- blank_data %>%
      group_by(Light_Value) %>%
      summarise(blank_O2 = mean(Oxygen, na.rm = TRUE), .groups = "drop")
    
    sample_data <- sample_data %>%
      left_join(blank_means, by = "Light_Value") %>%
      mutate(Oxygen = Oxygen - blank_O2)
  }
  
  # Split into intervals by light step
  intervals <- sample_data %>%
    group_by(Light_Value) %>%
    group_split()
  
  # Local regression
  fit_reg <- function(df) {
    rankLocReg(
      xall = as.numeric(df$`Delta T [min]`),
      yall = df$Oxygen,
      alpha = 0.2,
      method = "pc",
      verbose = FALSE
    )
  }
  
  fits <- map(intervals, fit_reg)
  
  # Slopes per interval
  micromol.L.s <- map_dbl(fits, ~ pluck(., "allRegs", "b1", 1))
  
  # Interval summary table
  interval_df <- map2_dfr(
    intervals,
    seq_along(intervals),
    ~ tibble(
      Colony_ID   = colony_id,
      Run         = run_num,
      Interval    = .y,
      Light_Value = unique(.x$Light_Value),
      N_points    = nrow(.x),
      O2_start    = head(.x$Oxygen, 1),
      O2_end      = tail(.x$Oxygen, 1),
      DeltaT_min  = min(.x$`Delta T [min]`),
      DeltaT_max  = max(.x$`Delta T [min]`),
      O2_min      = min(.x$Oxygen, na.rm = TRUE),
      O2_max      = max(.x$Oxygen, na.rm = TRUE),
      Slope       = micromol.L.s[.y]
    )
  )
  
  # Save file for this colony/run
  fn <- paste0("../output/pi_curves/May2025/intervals_", colony_id, "_run", run_num, "_May2025.csv")
  write_csv(interval_df, fn)
  
  return(interval_df)
}

# Run for all colony/run combos in parallel
plan(multisession)

all_interval_summaries <- future_pmap_dfr(
  list(colony_runs$Colony_ID, colony_runs$Run),
  fit_colony_intervals
)

# Save combined CSV
write_csv(all_interval_summaries, "../output/pi_curves/May2025/all_interval_summaries_May2025.csv")
```

Read in interval data again so the above chunk does not need to be run
```{r}
all_int <- read.csv("../output/pi_curves/May2025/all_interval_summaries_May2025.csv")
all_int$Colony_ID <- as.character(all_int$Colony_ID)
```

Filter sample metadata for May 2025
```{r}
may_2025 <- sample_meta %>%
  filter(grepl("^2025-05", Date))
may_2025$Surface_Area_cm2 <- as.numeric(may_2025$Surface_Area_cm2)
```

Merge sample metadata with interval data 
```{r}
all_int <- all_int %>%
  inner_join(may_2025, by = "Colony_ID") %>%
  filter(!is.na(Phenotype))
length(unique(all_int$Colony_ID))
```

Normalize interval data by chamber volume and surface area and save
```{r}
# normalize 
all_int <- all_int %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Slope * Chamber_Volume_L / Surface_Area_cm2) * 60) # units are umol O2 cm-2 h-1
write.csv(all_int, "../output/pi_curves/May2025/all_interval_summaries_normalized_May2025.csv")
```

Plot PI curves across light values 
```{r}
ggplot(all_int, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype")
```

### Fit Platt model to determine PI curve metrics 

- Pmax 
- alpha 
- Ik 
- LEDR
- Pgross 

Nest by colony 
```{r}
pi_nested <- all_int %>%
  select(Colony_ID, Light_Value, Slope_norm) %>%  # or whatever your P column is
  rename(I = Light_Value, P = Slope_norm) %>%
  group_by(Colony_ID) %>%
  nest()
```

Define the Platt model and fit for each colony 
```{r}
# Define a function to fit the model
fit_pi_model <- function(df) {
  tryCatch({
    fit <- nlsLM(P ~ Pmax * tanh(alpha * I / Pmax) + Rd,
                 data = df,
                 start = list(Pmax = max(df$P), alpha = 0.01, Rd = -0.1))
    pars <- coef(fit)
    tibble(Pmax = pars["Pmax"],
           alpha = pars["alpha"],
           Rd = pars["Rd"],
           Ik = pars["Pmax"]/pars["alpha"],
           Pgross = pars["Pmax"] + pars["Rd"])
  }, error = function(e) tibble(Pmax = NA, alpha = NA, Rd = NA, Ik = NA, Pgross = NA))
}

# Apply the model to each colony
pi_results <- pi_nested %>%
  mutate(params = map(data, fit_pi_model)) %>%
  select(-data) %>%
  unnest(cols = params)
```

Merge pi curve metric data with sample metadata and save 
```{r}
pi_results <- pi_results %>%
  inner_join(may_2025, by = "Colony_ID")
write.csv(pi_results, "../output/pi_curves/May2025/pi_curve_metrics_May2025.csv")
```

Pivot pi data to long format for plotting 
```{r}
pi_long <- pi_results %>%
  pivot_longer(cols = c(Pmax, alpha, Rd, Ik, Pgross),
               names_to = "Parameter",
               values_to = "Value")
```

Plot metrics 
```{r}
ggplot(pi_long, aes(x = Phenotype, y = Value, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, size = 1.5, alpha = 0.8) +
  facet_grid(Parameter ~ Species, scales = "free_y") +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw(base_size = 14) +
  labs(x = "Phenotype", y = "Value") +
  theme(legend.position = "none")
```

## October 2025 

### Prepare data 

Load run metadata 
```{r}
run_meta <- read_csv("../data/pi_curves/pi_curve_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
```

Load sample metadata
```{r}
sample_meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```

List all raw files for October 2025 
```{r}
files <- list.files(path = "../data/pi_curves/raw_data/Oct2025", pattern = "*.csv", full.names = TRUE)
```

Join O2 data with metadata files based on start and stop times 
```{r}
all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "Ch\\d+") %>% str_remove("Ch") %>% as.numeric()
  
  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )
  
  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )
  
  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )
  
  return(raw_joined)
})
```

Look at the joined data and save 
```{r}
all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

write_csv(all_data, "../output/pi_curves/Oct2025/all_joined_data_Oct2025.csv")
```

### Run localized linear regressions 

The interval fitting code will take about an hour to run and is computationally intense. If needed, run as a background job or on a remote server.

Make list of all Colony_ID / Run combinations (including blanks)
```{r}
colony_runs <- all_data %>%
  filter(!is.na(Colony_ID)) %>%
  distinct(Colony_ID, Run)
```

Run interval fittings with blank correction
```{r}
# Function: blank correction + slope extraction + save/return interval summary
fit_colony_intervals <- function(colony_id, run_num) {
  
  # Sample data for this colony/run
  sample_data <- all_data %>%
    filter(Colony_ID == colony_id, Run == run_num)
  
  # Skip if no data
  if (nrow(sample_data) == 0) return(NULL)
  
  # Blank data for same run
  blank_data <- all_data %>%
    filter(Colony_ID == "blank", Run == run_num)
  
  # Blank correction if possible
  if (nrow(blank_data) > 0) {
    blank_means <- blank_data %>%
      group_by(Light_Value) %>%
      summarise(blank_O2 = mean(Oxygen, na.rm = TRUE), .groups = "drop")
    
    sample_data <- sample_data %>%
      left_join(blank_means, by = "Light_Value") %>%
      mutate(Oxygen = Oxygen - blank_O2)
  }
  
  # Split into intervals by light step
  intervals <- sample_data %>%
    group_by(Light_Value) %>%
    group_split()
  
  # Local regression
  fit_reg <- function(df) {
    rankLocReg(
      xall = as.numeric(df$`Delta T [min]`),
      yall = df$Oxygen,
      alpha = 0.2,
      method = "pc",
      verbose = FALSE
    )
  }
  
  fits <- map(intervals, fit_reg)
  
  # Slopes per interval
  micromol.L.s <- map_dbl(fits, ~ pluck(., "allRegs", "b1", 1))
  
  # Interval summary table
  interval_df <- map2_dfr(
    intervals,
    seq_along(intervals),
    ~ tibble(
      Colony_ID   = colony_id,
      Run         = run_num,
      Interval    = .y,
      Light_Value = unique(.x$Light_Value),
      N_points    = nrow(.x),
      O2_start    = head(.x$Oxygen, 1),
      O2_end      = tail(.x$Oxygen, 1),
      DeltaT_min  = min(.x$`Delta T [min]`),
      DeltaT_max  = max(.x$`Delta T [min]`),
      O2_min      = min(.x$Oxygen, na.rm = TRUE),
      O2_max      = max(.x$Oxygen, na.rm = TRUE),
      Slope       = micromol.L.s[.y]
    )
  )
  
  # Save file for this colony/run
  fn <- paste0("../output/pi_curves/Oct2025/intervals_", colony_id, "_run", run_num, "_Oct2025.csv")
  write_csv(interval_df, fn)
  
  return(interval_df)
}

# Run for all colony/run combos in parallel
plan(multisession)

all_interval_summaries <- future_pmap_dfr(
  list(colony_runs$Colony_ID, colony_runs$Run),
  fit_colony_intervals
)

# Save combined CSV
write_csv(all_interval_summaries, "../output/pi_curves/Oct2025/all_interval_summaries_Oct2025.csv")
```

Read in interval data again so the above chunk does not need to be run
```{r}
all_int <- read.csv("../output/pi_curves/Oct2025/all_interval_summaries_Oct2025.csv")
all_int$Colony_ID <- as.character(all_int$Colony_ID)
```

Filter sample metadata for May 2025
```{r}
oct_2025 <- sample_meta %>%
  filter(grepl("^2025-10", Date))
oct_2025$Surface_Area_cm2 <- as.numeric(oct_2025$Surface_Area_cm2)
```

NOTE FOR 10/2/25 - will need to set the SA as dummy values of 20 for now until we can do the fraggameter for SA

Merge sample metadata with interval data 
```{r}
all_int <- all_int %>%
  inner_join(oct_2025, by = "Colony_ID") %>%
  filter(!is.na(Phenotype))
length(unique(all_int$Colony_ID))
```

Normalize interval data by chamber volume and surface area and save
```{r}
# normalize 
all_int <- all_int %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Slope * Chamber_Volume_L / Surface_Area_cm2) * 60) # units are umol O2 cm-2 h-1
write.csv(all_int, "../output/pi_curves/Oct2025/all_interval_summaries_normalized_Oct2025.csv")
```

Plot PI curves across light values 
```{r}
ggplot(all_int, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype")
```

### Fit Platt model to determine PI curve metrics 

- Pmax 
- alpha 
- Ik 
- LEDR
- Pgross 

Nest by colony 
```{r}
pi_nested <- all_int %>%
  select(Colony_ID, Light_Value, Slope_norm) %>%  # or whatever your P column is
  rename(I = Light_Value, P = Slope_norm) %>%
  group_by(Colony_ID) %>%
  nest()
```

Define the Platt model and fit for each colony 
```{r}
# Define a function to fit the model
fit_pi_model <- function(df) {
  tryCatch({
    fit <- nlsLM(P ~ Pmax * tanh(alpha * I / Pmax) + Rd,
                 data = df,
                 start = list(Pmax = max(df$P), alpha = 0.01, Rd = -0.1))
    pars <- coef(fit)
    tibble(Pmax = pars["Pmax"],
           alpha = pars["alpha"],
           Rd = pars["Rd"],
           Ik = pars["Pmax"]/pars["alpha"],
           Pgross = pars["Pmax"] + pars["Rd"])
  }, error = function(e) tibble(Pmax = NA, alpha = NA, Rd = NA, Ik = NA, Pgross = NA))
}

# Apply the model to each colony
pi_results <- pi_nested %>%
  mutate(params = map(data, fit_pi_model)) %>%
  select(-data) %>%
  unnest(cols = params)
```

Merge pi curve metric data with sample metadata and save 
```{r}
pi_results <- pi_results %>%
  inner_join(oct_2025, by = "Colony_ID")
write.csv(pi_results, "../output/pi_curves/Oct2025/pi_curve_metrics_Oct2025.csv")
```

Pivot pi data to long format for plotting 
```{r}
pi_long <- pi_results %>%
  pivot_longer(cols = c(Pmax, alpha, Rd, Ik, Pgross),
               names_to = "Parameter",
               values_to = "Value")
```

Plot metrics 
```{r}
ggplot(pi_long, aes(x = Phenotype, y = Value, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, size = 1.5, alpha = 0.8) +
  facet_grid(Parameter ~ Species, scales = "free_y") +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  theme_bw(base_size = 14) +
  labs(x = "Phenotype", y = "Value") +
  theme(legend.position = "none")
```









## All time points 

### PI curves 

Read in normalized interval data 
```{r}
int_sept2023 <- read.csv("../output/pi_curves/Sept2023/all_interval_summaries_normalized_Sept2023.csv")
int_sept2023$Timepoint <- "Sept23"
int_sept2023 <- int_sept2023 %>%
  select(Colony_ID, Run.x, Light_Value, Date, Species, Phenotype, Slope_norm, Timepoint)

int_jan2024 <- read.csv("../output/pi_curves/Jan2024/all_interval_summaries_normalized_Jan2024.csv")
int_jan2024$Timepoint <- "Jan24"
int_jan2024 <- int_jan2024 %>%
  select(Colony_ID, Run.x, Light_Value, Date, Species, Phenotype, Slope_norm, Timepoint)

int_may2024 <- read.csv("../output/pi_curves/May2024/all_interval_summaries_normalized_May2024.csv")
int_may2024$Timepoint <- "May24"
int_may2024 <- int_may2024 %>%
  select(Colony_ID, Run.x, Light_Value, Date, Species, Phenotype, Slope_norm, Timepoint)

int_oct2024 <- read.csv("../output/pi_curves/Oct2024/all_interval_summaries_normalized_Oct2024.csv")
int_oct2024$Timepoint <- "Oct24"
int_oct2024 <- int_oct2024 %>%
  select(Colony_ID, Run.x, Light_Value, Date, Species, Phenotype, Slope_norm, Timepoint)

int_jan2025 <- read.csv("../output/pi_curves/Jan2025/all_interval_summaries_normalized_Jan2025.csv")
int_jan2025$Timepoint <- "Jan25"
int_jan2025 <- int_jan2025 %>%
  select(Colony_ID, Run.x, Light_Value, Date, Species, Phenotype, Slope_norm, Timepoint)

int_may2025 <- read.csv("../output/pi_curves/May2025/all_interval_summaries_normalized_May2025.csv")
int_may2025$Timepoint <- "May25"
int_may2025 <- int_may2025 %>%
  select(Colony_ID, Run.x, Light_Value, Date, Species, Phenotype, Slope_norm, Timepoint)

int_oct2025 <- read.csv("../output/pi_curves/Oct2025/all_interval_summaries_normalized_Oct2025.csv")
int_oct2025$Timepoint <- "Oct25"
int_oct2025 <- int_oct2025 %>%
  select(Colony_ID, Run.x, Light_Value, Date, Species, Phenotype, Slope_norm, Timepoint)
```

Bind data together 
```{r}
all_time_int <- rbind(int_sept2023, int_jan2024, int_may2024, int_oct2024, int_jan2025, int_may2025, int_oct2025)
```

Plot PI curves across light values 
```{r}
# Set levels 
all_time_int$Timepoint <- factor(all_time_int$Timepoint, levels = c("Sept23", "Jan24", "May24", "Oct24", "Jan25", "May25", "Oct25"))

ggplot(all_time_int, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  #facet_wrap(~Species+Timepoint, scales = "fixed") +
  facet_grid(rows = vars(Species), cols = vars(Timepoint)) +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype")

ggplot(all_time_int, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1.5, position = position_jitter(width = 10)) +
  geom_smooth(aes(fill = Phenotype), method = "loess", se = TRUE, linewidth = 1.2) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  #facet_wrap(~Species + Timepoint, scales = "fixed") +
  facet_grid(rows = vars(Species), cols = vars(Timepoint)) +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype", fill = "Phenotype")
```

Remove Colony 237 from Sept23, Colony 214 from Jan24 and Colony 28 from May25 and re-plot. Also remove Jan25 because the surface area measurements are messed up
```{r}
all_time_int_filt <- all_time_int %>%
  filter(!(Colony_ID == "237" & Timepoint == "Sept23")) %>%
  filter(!(Colony_ID == "214" & Timepoint == "Jan24")) %>%
  filter(!(Colony_ID == "28" & Timepoint == "May25")) %>%
  filter(!Timepoint == "Jan25")

ggplot(all_time_int_filt, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  #facet_wrap(~Species+Timepoint, scales = "fixed") +
  facet_grid(rows = vars(Species), cols = vars(Timepoint)) +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype")

ggplot(all_time_int_filt, aes(x = Light_Value, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1.5, position = position_jitter(width = 10)) +
  geom_smooth(aes(fill = Phenotype), method = "loess", se = TRUE, linewidth = 1.2) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  #facet_wrap(~Species + Timepoint, scales = "fixed") +
  facet_grid(rows = vars(Species), cols = vars(Timepoint)) +
  theme_classic(base_size = 14) +
  labs(x = "Light (µmol photons m-2 s-1)", y = "umol O2 cm-2 h-1", color = "Phenotype", fill = "Phenotype") + 
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text.y = element_text(face = "italic", size = 13),
    #legend.position = c(0.15, 0.83),  # adjust x, y for placement
    legend.position = "right",
    #legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 14.5)
  )
```




### PI curve metrics 

Read in calculated pi curve metrics 
```{r}
pi_metrics_sept2023 <- read.csv("../output/pi_curves/Sept2023/pi_curve_metrics_Sept2023.csv")
pi_metrics_sept2023$Timepoint <- "Sept23"

pi_metrics_jan2024 <- read.csv("../output/pi_curves/Jan2024/pi_curve_metrics_Jan2024.csv")
pi_metrics_jan2024$Timepoint <- "Jan24"

pi_metrics_may2024 <- read.csv("../output/pi_curves/May2024/pi_curve_metrics_May2024.csv")
pi_metrics_may2024$Timepoint <- "May24"

pi_metrics_oct2024 <- read.csv("../output/pi_curves/Oct2024/pi_curve_metrics_Oct2024.csv")
pi_metrics_oct2024$Timepoint <- "Oct24"

pi_metrics_jan2025 <- read.csv("../output/pi_curves/Jan2025/pi_curve_metrics_Jan2025.csv")
pi_metrics_jan2025$Timepoint <- "Jan25"

pi_metrics_may2025 <- read.csv("../output/pi_curves/May2025/pi_curve_metrics_May2025.csv")
pi_metrics_may2025$Timepoint <- "May25"

pi_metrics_oct2025 <- read.csv("../output/pi_curves/Oct2025/pi_curve_metrics_Oct2025.csv")
pi_metrics_oct2025$Timepoint <- "Oct25"
```

Bind data together 
```{r}
all_time_pi_metrics <- rbind(pi_metrics_sept2023, pi_metrics_jan2024, pi_metrics_may2024, pi_metrics_oct2024, pi_metrics_jan2025, pi_metrics_may2025, pi_metrics_oct2025)
```

Remove Jan25 for now (still need SA)
```{r}
all_time_pi_metrics <- all_time_pi_metrics %>%
  filter(!Timepoint == "Jan25")
```


#### Stats 

NEED TO DO

##### Pmax 

Assess normality of data 
```{r}
hist(all_time_pi_metrics$Pmax)

# Q-Q Plot
qqnorm(all_time_pi_metrics$Pmax)
qqline(all_time_pi_metrics$Pmax, col = "red")

# Shapiro-Wilk Test
shapiro.test(all_time_pi_metrics$Pmax)
```

Data is not normal, perform data transformation
```{r}
# Log10
all_time_pi_metrics$Pmax_log <- log10(all_time_pi_metrics$Pmax +1) # add 1 to avoid log10(0)
hist(all_time_pi_metrics$Pmax_log)
qqnorm(all_time_pi_metrics$Pmax_log)
qqline(all_time_pi_metrics$Pmax_log, col = "red")
shapiro.test(all_time_pi_metrics$Pmax_log)

# Sqrt
all_time_pi_metrics$Pmax_sqrt <- sqrt(all_time_pi_metrics$Pmax)
hist(all_time_pi_metrics$Pmax_sqrt)
qqnorm(all_time_pi_metrics$Pmax_sqrt)
qqline(all_time_pi_metrics$Pmax_sqrt, col = "red")
shapiro.test(all_time_pi_metrics$Pmax_sqrt)

# Inverse 
all_time_pi_metrics$Pmax_inv <- 1 / (all_time_pi_metrics$Pmax +1) # add 1 to avoid division by zero
hist(all_time_pi_metrics$Pmax_inv)
qqnorm(all_time_pi_metrics$Pmax_inv)
qqline(all_time_pi_metrics$Pmax_inv, col = "red")
shapiro.test(all_time_pi_metrics$Pmax_inv)
```

Both log and sqrt are normal. Proceed with log. 

Perform stats 
```{r}
mod_pmax <- lmer(Pmax_log ~ Species * Phenotype * Timepoint + (1 | Colony_ID),
                 data = all_time_pi_metrics,
                 REML = FALSE)
car::Anova(mod_pmax, type = 3)
```

Pmax is significant by species (Pcomp Pmax greater than Mcap Pmax) and Timepoint. ADD INFO ONCE I REDO JAN25 SURFACE AREA

Look at posthoc comparisons 
```{r}
em_spp <- emmeans(mod_pmax, pairwise ~ Species, adjust = "tukey")
summary(em_spp)

em_time <- emmeans(mod_pmax, pairwise ~ Timepoint, adjust = "tukey")
summary(em_time)
```

##### Alpha 

Assess normality of data 
```{r}
hist(all_time_pi_metrics$alpha)

# Q-Q Plot
qqnorm(all_time_pi_metrics$alpha)
qqline(all_time_pi_metrics$alpha, col = "red")

# Shapiro-Wilk Test
shapiro.test(all_time_pi_metrics$alpha)
```

Data is not normal, perform data transformation
```{r}
# Log10
all_time_pi_metrics$alpha_log <- log10(all_time_pi_metrics$alpha +1) # add 1 to avoid log10(0)
hist(all_time_pi_metrics$alpha_log)
qqnorm(all_time_pi_metrics$alpha_log)
qqline(all_time_pi_metrics$alpha_log, col = "red")
shapiro.test(all_time_pi_metrics$alpha_log)

# Sqrt
all_time_pi_metrics$alpha_sqrt <- sqrt(all_time_pi_metrics$alpha)
hist(all_time_pi_metrics$alpha_sqrt)
qqnorm(all_time_pi_metrics$alpha_sqrt)
qqline(all_time_pi_metrics$alpha_sqrt, col = "red")
shapiro.test(all_time_pi_metrics$alpha_sqrt)

# Inverse 
all_time_pi_metrics$alpha_inv <- 1 / (all_time_pi_metrics$alpha +1) # add 1 to avoid division by zero
hist(all_time_pi_metrics$alpha_inv)
qqnorm(all_time_pi_metrics$alpha_inv)
qqline(all_time_pi_metrics$alpha_inv, col = "red")
shapiro.test(all_time_pi_metrics$alpha_inv)
```

Only sqrt is normal. Proceed with sqrt 

Perform stats 
```{r}
mod_alpha <- lmer(alpha_sqrt ~ Species * Phenotype * Timepoint + (1 | Colony_ID),
                 data = all_time_pi_metrics,
                 REML = FALSE)
car::Anova(mod_alpha, type = 3)
```

Alpha is significant by Timepoint. ADD INFO ONCE I REDO JAN25 SURFACE AREA

Look at posthoc comparisons 
```{r}
em_time_alpha <- emmeans(mod_alpha, pairwise ~ Timepoint, adjust = "tukey")
summary(em_time_alpha)
```


##### Ik 

Assess normality of data 
```{r}
hist(all_time_pi_metrics$Ik)

# Q-Q Plot
qqnorm(all_time_pi_metrics$Ik)
qqline(all_time_pi_metrics$Ik, col = "red")

# Shapiro-Wilk Test
shapiro.test(all_time_pi_metrics$Ik)
```

Data is not normal, perform data transformation
```{r}
# Log10
all_time_pi_metrics$Ik_log <- log10(all_time_pi_metrics$Ik +1) # add 1 to avoid log10(0)
hist(all_time_pi_metrics$Ik_log)
qqnorm(all_time_pi_metrics$Ik_log)
qqline(all_time_pi_metrics$Ik_log, col = "red")
shapiro.test(all_time_pi_metrics$Ik_log)

# Sqrt
all_time_pi_metrics$Ik_sqrt <- sqrt(all_time_pi_metrics$Ik)
hist(all_time_pi_metrics$Ik_sqrt)
qqnorm(all_time_pi_metrics$Ik_sqrt)
qqline(all_time_pi_metrics$Ik_sqrt, col = "red")
shapiro.test(all_time_pi_metrics$Ik_sqrt)

# Inverse 
all_time_pi_metrics$Ik_inv <- 1 / (all_time_pi_metrics$Ik +1) # add 1 to avoid division by zero
hist(all_time_pi_metrics$Ik_inv)
qqnorm(all_time_pi_metrics$Ik_inv)
qqline(all_time_pi_metrics$Ik_inv, col = "red")
shapiro.test(all_time_pi_metrics$Ik_inv)
```

Both log and sqrt are normal. Proceed with log. 

Perform stats 
```{r}
mod_ik <- lmer(Ik_log ~ Species * Phenotype * Timepoint + (1 | Colony_ID),
                 data = all_time_pi_metrics,
                 REML = FALSE)
car::Anova(mod_ik, type = 3)
```

Ik is significant by species (Pcomp Ik greater than Mcap Ik) and Timepoint. ADD INFO ONCE I REDO JAN25 SURFACE AREA

Look at posthoc comparisons 
```{r}
em_time_ik <- emmeans(mod_ik, pairwise ~ Timepoint, adjust = "tukey")
summary(em_time_ik)

em_spp_ik <- emmeans(mod_ik, pairwise ~ Species, adjust = "tukey")
summary(em_spp_ik)
```

##### Rd 

Assess normality of data 
```{r}
hist(all_time_pi_metrics$Rd)

# Q-Q Plot
qqnorm(all_time_pi_metrics$Rd)
qqline(all_time_pi_metrics$Rd, col = "red")

# Shapiro-Wilk Test
shapiro.test(all_time_pi_metrics$Rd)
```

Data is not normal, perform data transformation
```{r}
# Log10
all_time_pi_metrics$Rd_log <- log10(all_time_pi_metrics$Rd +1) # add 1 to avoid log10(0)
hist(all_time_pi_metrics$Rd_log)
qqnorm(all_time_pi_metrics$Rd_log)
qqline(all_time_pi_metrics$Rd_log, col = "red")
shapiro.test(all_time_pi_metrics$Rd_log)

# Sqrt
all_time_pi_metrics$Rd_sqrt <- sqrt(all_time_pi_metrics$Rd)
hist(all_time_pi_metrics$Rd_sqrt)
qqnorm(all_time_pi_metrics$Rd_sqrt)
qqline(all_time_pi_metrics$Rd_sqrt, col = "red")
shapiro.test(all_time_pi_metrics$Rd_sqrt)

# Inverse 
all_time_pi_metrics$Rd_inv <- 1 / (all_time_pi_metrics$Rd +1) # add 1 to avoid division by zero
hist(all_time_pi_metrics$Rd_inv)
qqnorm(all_time_pi_metrics$Rd_inv)
qqline(all_time_pi_metrics$Rd_inv, col = "red")
shapiro.test(all_time_pi_metrics$Rd_inv)
```

No normality produced...I think there are issues because Rd is negative. 

Perform stats 
```{r}
mod_rd <- lmer(Rd ~ Species * Phenotype * Timepoint + (1 | Colony_ID),
                 data = all_time_pi_metrics,
                 REML = FALSE)
car::Anova(mod_rd, type = 3)
```

Rd is significant by species (Pcomp Rd more negative than Mcap Rd) and Timepoint. ADD INFO ONCE I REDO JAN25 SURFACE AREA

Look at posthoc comparisons 
```{r}
em_time_rd <- emmeans(mod_rd, pairwise ~ Timepoint, adjust = "tukey")
summary(em_time_rd)

em_spp_rd <- emmeans(mod_rd, pairwise ~ Species, adjust = "tukey")
summary(em_spp_rd)
```

##### Pgross

Assess normality of data 
```{r}
hist(all_time_pi_metrics$Pgross)

# Q-Q Plot
qqnorm(all_time_pi_metrics$Pgross)
qqline(all_time_pi_metrics$Pgross, col = "red")

# Shapiro-Wilk Test
shapiro.test(all_time_pi_metrics$Pgross)
```

Data is not normal, perform data transformation
```{r}
# Log10
all_time_pi_metrics$Pgross_log <- log10(all_time_pi_metrics$Pgross +1) # add 1 to avoid log10(0)
hist(all_time_pi_metrics$Pgross_log)
qqnorm(all_time_pi_metrics$Pgross_log)
qqline(all_time_pi_metrics$Pgross_log, col = "red")
shapiro.test(all_time_pi_metrics$Pgross_log)

# Sqrt
all_time_pi_metrics$Pgross_sqrt <- sqrt(all_time_pi_metrics$Pgross)
hist(all_time_pi_metrics$Pgross_sqrt)
qqnorm(all_time_pi_metrics$Pgross_sqrt)
qqline(all_time_pi_metrics$Pgross_sqrt, col = "red")
shapiro.test(all_time_pi_metrics$Pgross_sqrt)

# Inverse 
all_time_pi_metrics$Pgross_inv <- 1 / (all_time_pi_metrics$Pgross +1) # add 1 to avoid division by zero
hist(all_time_pi_metrics$Pgross_inv)
qqnorm(all_time_pi_metrics$Pgross_inv)
qqline(all_time_pi_metrics$Pgross_inv, col = "red")
shapiro.test(all_time_pi_metrics$Pgross_inv)
```

Both log and sqrt are normal. Proceed with log. 

Perform stats 
```{r}
mod_pgross <- lmer(Pgross_log ~ Species * Phenotype * Timepoint + (1 | Colony_ID),
                 data = all_time_pi_metrics,
                 REML = FALSE)
car::Anova(mod_pgross, type = 3)
```

Pgross is significant by species (Pcomp Pgross greater than Mcap Pgross), Timepoint and Species x Timepoint. ADD INFO ONCE I REDO JAN25 SURFACE AREA

Look at posthoc comparisons 
```{r}
em_time_pgross <- emmeans(mod_pgross, pairwise ~ Timepoint, adjust = "tukey")
summary(em_time_pgross)

em_spp_pgross <- emmeans(mod_pgross, pairwise ~ Species, adjust = "tukey")
summary(em_spp_pgross)

em_time_spp_pgross <- emmeans(mod_pgross, pairwise ~ Species*Timepoint, adjust = "tukey")
summary(em_time_spp_pgross)
```

#### Plot 

Calculate mean, sem, and sd 
```{r}
pi_metrics_summary <- all_time_pi_metrics %>%
  group_by(Timepoint, Phenotype, Species) %>%
  summarise(
    mean_Pmax = mean(Pmax, na.rm = TRUE),
    sd_Pmax = sd(Pmax, na.rm = T),
    se_Pmax = sd(Pmax, na.rm = T) / sqrt(length(Pmax)),
    sem_Pmax = sd(Pmax, na.rm = TRUE) / sqrt(sum(!is.na(Pmax))),
    mean_alpha = mean(alpha, na.rm = TRUE),
    sem_alpha = sd(alpha, na.rm = TRUE) / sqrt(sum(!is.na(alpha))),
    mean_Rd = mean(Rd, na.rm = TRUE),
    sem_Rd = sd(Rd, na.rm = TRUE) / sqrt(sum(!is.na(Rd))),
    mean_Ik = mean(Ik, na.rm = TRUE),
    sem_Ik = sd(Ik, na.rm = TRUE) / sqrt(sum(!is.na(Ik))),
    mean_Pgross = mean(Pgross, na.rm = TRUE),
    sem_Pgross = sd(Pgross, na.rm = TRUE) / sqrt(sum(!is.na(Pgross))),
    .groups = "drop")
```

Plot! 
```{r}
# Set levels 
pi_metrics_summary$Timepoint <- factor(pi_metrics_summary$Timepoint, levels = c("Sept23", "Jan24", "May24", "Oct24", "Jan25", "May25", "Oct25"))

# Pmax
ggplot(pi_metrics_summary, aes(x = Timepoint, y = mean_Pmax, color = Phenotype, group = Phenotype)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_Pmax - sem_Pmax, ymax = mean_Pmax + sem_Pmax, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Pmax",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )

# Alpha
ggplot(pi_metrics_summary, aes(x = Timepoint, y = mean_alpha, color = Phenotype, group = Phenotype)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_alpha - sem_alpha, ymax = mean_alpha + sem_alpha, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Alpha",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )

# Rd
ggplot(pi_metrics_summary, aes(x = Timepoint, y = mean_Rd, color = Phenotype, group = Phenotype)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_Rd - sem_Rd, ymax = mean_Rd + sem_Rd, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "LEDR",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )

# Ik
ggplot(pi_metrics_summary, aes(x = Timepoint, y = mean_Ik, color = Phenotype, group = Phenotype)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_Ik - sem_Ik, ymax = mean_Ik + sem_Ik, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Ik",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )

# Pgross
ggplot(pi_metrics_summary, aes(x = Timepoint, y = mean_Pgross, color = Phenotype, group = Phenotype)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_Pgross - sem_Pgross, ymax = mean_Pgross + sem_Pgross, color = Phenotype),
                width = 0.3, linewidth = 0.8) +  # color matches points/lines
  facet_wrap(~ Species) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")
  ) +
  theme_classic(base_size = 16) +  # bigger base size
  labs(
    x = "",
    y = "Pgross",
    color = "Phenotype"
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1),
    strip.text = element_text(face = "italic", size = 16),
    #legend.position = c(0.85, 0.83),  # adjust x, y for placement
    legend.position = "right",    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_blank(),    
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    axis.title.y = element_text(size = 16)
  )
```

The stats indicate no significant phenotype differences. Is this beneficial or negative acclimatization? I guess we wont know until the next stressor (ie the heatwave in the fall 2025). 
