---
title: "Porites ITS2"
author: "Jill Ashey"
date: "2026-03-31"
output: html_document
---

Porites compressa ITS2 analysis. Code from Ariana Huffmyer and Sarah Roderick.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

if (!require("tidyverse")) install.packages("tidyverse")
if (!require("readxl")) install.packages("readxl")
if (!require("phyloseq")) BiocManager::install("phyloseq")
if (!require("janitor")) install.packages("janitor")
if (!require("writexl")) install.packages("writexl")
if (!require("microViz")) BiocManager::install("microViz")

library("readxl")
library("phyloseq")
library("janitor")
library("writexl")
library("ggplot2")
library('cowplot')
library("janitor")
library("car")
library("phyloseq")
library("stringr")
library("ape")
library("DESeq2")
library("microViz")
library("vegan")
library("ComplexHeatmap")
library("tidyverse")
library("dplyr")
library(emmeans)
library(readxl)
library(patchwork)
```

Load metadata.  
```{r}
#create metadata and load
metadata <- read.csv("../data/ITS2/SymPortal_datasheet_Martynek.csv", header = TRUE, sep = ",")%>%dplyr::select(sample_name , Phenotype)

keep<-c("26", "27", "43", "44", "233", "234", "235", "236", "237", "238", "239", "240", "243", "244", "247", "248")

metadata <- metadata %>%
  filter(sample_name %in% keep)

rownames(metadata) <- metadata$sample_name

metadata <- as.matrix(metadata)
metadata <- sample_data(data.frame(metadata))
```

Load the data for strain level information.  
```{r}
#gather the taxonomy names  
taxnames <- read_tsv(
  file  = "../data/ITS2/20251021T032609_martyma1/post_med_seqs/638_20251021T051401_DBV_20251021T231108.seqs.absolute.abund_only.txt",
  n_max = 0) %>%
  dplyr::select(-1) %>%
  names(.)


#extract clade letter  
tax0 <- tibble(
  DIV = taxnames,
  clade = str_extract(DIV, "[A-Z]")
)

tax1 <- as.matrix(tax0)
rownames(tax1) <- tax0$DIV
tax <- tax_table(tax1)

otu0 <- readr::read_tsv(
  file  = "../data/ITS2/20251021T032609_martyma1/post_med_seqs/638_20251021T051401_DBV_20251021T231108.seqs.absolute.abund_and_meta.txt"
) %>%
  dplyr::select(-1) %>%
  filter(sample_name %in% keep)

otu1 <- as.matrix(otu0[, 39:289])
rownames(otu1) <- otu0$sample_name
otu <- otu_table(otu1, taxa_are_rows = FALSE)

coralDIV <- phyloseq(otu, tax, metadata)
coralDIV
```

Build DIV Data

First, view the number of sequence counts in our dataset.   
```{r}
samps <- readr::read_tsv(
  "../data/ITS2/20251021T032609_martyma1/post_med_seqs/638_20251021T051401_DBV_20251021T231108.seqs.absolute.abund_and_meta.txt"
) %>%
  dplyr::select(post_taxa_id_absolute_symbiodiniaceae_seqs, sample_name) %>%
  filter(sample_name %in% keep)

hist(samps$post_taxa_id_absolute_symbiodiniaceae_seqs, breaks = 20)
min(samps$post_taxa_id_absolute_symbiodiniaceae_seqs)
max(samps$post_taxa_id_absolute_symbiodiniaceae_seqs)
mean(samps$post_taxa_id_absolute_symbiodiniaceae_seqs)
```

We have >100,000 sequences for all samples. 

Plot tree.  
```{r}
random_tree = rtree(ntaxa(coralDIV), rooted=TRUE, tip.label=taxa_names(coralDIV))
plot(random_tree)
```

Merge phyloseq data. 
```{r}
phylo_coral_DIV = merge_phyloseq(coralDIV, metadata, random_tree)
phylo_coral_DIV
```

Filter out samples with low numbers in groups (n=1 per group) from phyloseq object.  
```{r}
phylo_coral_DIV<-ps_filter(phylo_coral_DIV)
phylo_coral_DIV<-ps_arrange(phylo_coral_DIV, Phenotype)
```

Now build a tree with the combined data.  
```{r}
plot_tree(phylo_coral_DIV, color="Phenotype", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
```

Plot a heatmap.  

```{r}
plot_heatmap(phylo_coral_DIV, "NMDS", "bray", "Phenotype", low="#66CCFF", high="#000033", na.value="white")
```

Assign object to a shorter name. 
```{r}
GP1 = phylo_coral_DIV
GP1
```

Display smaple sequencing depth. 
```{r}
min(sample_sums(GP1))
```

Transform to relative abundance and subsample to even sampling depth to minimum count in dataset (89462).  
```{r}
GP1_rare<-rarefy_even_depth(GP1, sample.size = min(sample_sums(GP1)),
  rngseed = TRUE, replace = TRUE, trimOTUs = TRUE, verbose = TRUE)
```

View sample sequencing depth again after subsampling. 
```{r}
sample_sums(GP1_rare)
GP1_rare
```

Samples are now all at 89,462 sequences. No OTUs were removed. 

Now transform counts to relative abundance expressed as a proportion.    
```{r}
GP1_rare = transform_sample_counts(GP1_rare, function(x) x/sum(x))
```

View tree and heatmap after subsampling.   
```{r}
plot_tree(GP1_rare, color="Phenotype", label.tips="taxa_names", ladderize="left", plot.margin=0.3) 
plot_heatmap(GP1_rare, "NMDS", "bray", "Phenotype", low="#66CCFF", high="#000033", na.value="white")
```

View NMDS ordination by OTU
```{r}
GP.ord <- ordinate(GP1_rare, "NMDS", "bray")
ordinate(GP1_rare, "NMDS", "bray")
p1 = plot_ordination(GP1_rare, GP.ord, type="taxa", title="taxa")
p1
```

Plot with samples now.  
```{r}
p2 = plot_ordination(GP1_rare, GP.ord, type="samples", color="Phenotype") 
p2_geom<- p2 + geom_polygon(aes(fill=Phenotype)) + geom_point(size=4) + ggtitle("samples")
p2_geom
```

View PCoA by OTU.  
```{r}
GP.ordb.pcoa <- ordinate(GP1_rare, "PCoA", "bray")
p1b = plot_ordination(GP1_rare, GP.ordb.pcoa, type="taxa", title="taxa")
p1b
```
Plot with samples now.  
```{r}
p2b = plot_ordination(GP1_rare, GP.ord, type="samples", color="parent"); p2b 
```
Output NMDS plot for visualization.  
```{r}
phen_names <- c("Resistant" = "Resistant", "Susceptible" = "Susceptible")
stress_label <- paste("Stress =", round(GP.ord$stress, 3))

p2b_viz<-p2  + 
  geom_point(aes(colour=Phenotype), size=4) + 
  #ggtitle("NMDS")+
  scale_colour_manual(name="Phenotype", values=c("#3b244d", "#f2916b"), labels=phen_names) +
  theme_classic() + 
  theme(
    panel.background=element_blank()
  ); p2b_viz

ggsave(plot=p2b_viz, "../output/ITS2/nmds_its2_div.png", h=5, w=6, dpi = 300)
```

Conduct a permanova in the vegan package.  
```{r}
metadata <- as(sample_data(GP1_rare), "data.frame")

perm1<-adonis2(phyloseq::distance(GP1_rare, method="bray") ~ Phenotype,
       data = metadata)

perm1
```

Permanova suggests there is NO significant difference by phenotype in ITS2 profiles (p=0.455).  

adonis2(formula = phyloseq::distance(GP1_rare, method = "bray") ~ Phenotype, data = metadata)
         Df SumOfSqs      R2     F Pr(>F)
Model     1 0.017195 0.05958 0.887  0.455
Residual 14 0.271413 0.94042             
Total    15 0.288608 1.00000 

Plot heatmap of ITS2 DIVs 

Filter out taxa that comprise <1% mean relative abundance across samples. 
```{r}
GP1_rare

GP1_prune = filter_taxa(GP1_rare, function(x) mean(x) > 0.01, TRUE)
GP1_prune 

sample_order<-c("26", "44", "234", "236", "238", "240", "244", "248", "27", "43", "233", "235", "237", "239", "243", "247")

# p<-plot_heatmap(GP1_prune, sample.label="parent", low="#66CCFF", high="#000033", na.value="white", taxa.label="DIV", sample.order = sample_order)
# p$scales$scales[[1]]$name <- "Phenotype"
# 
# print(p)

#ggplot2::ggsave("Mcap2020/Figures/ITS2/div_heatmap.jpg", p, width = 6, height = 6)
```

There were 73 OTUs before filtering and 14 after filtering. This indicates there were many rare taxa.   

Run the NMDS visualization again with the dataset with rare samples removed. 
```{r}
GP.ordb.filt <- ordinate(GP1_prune, "NMDS", "bray")

p3b = plot_ordination(GP1_prune, GP.ordb.filt, type="samples", color="parent"); p3b 

p3b_viz_filt<-p3b  + 
  geom_point(aes(colour=Phenotype), size=4) + 
  ggtitle("NMDS")+
  scale_colour_manual(name="Phenotype", values=c("#3b244d", "#f2916b")) +
  theme_classic() + 
  theme(
    panel.background=element_blank()
  ); p3b_viz_filt

ggsave(plot=p3b_viz_filt, "../output/ITS2/nmds_its2_div_filtered.png", h=5, w=6, dpi = 300)

# Get stress value
GP.ordb.filt$stress
```

Run PERMANOVA on filtered data set. 
```{r}
metadata_filt <- as(sample_data(GP1_prune), "data.frame")

perm2_filt<-adonis2(phyloseq::distance(GP1_prune, method="bray") ~ Phenotype,
       data = metadata_filt)
perm2_filt
```

There is no significant variation of dominant taxa by phenotype (p=0.587).

adonis2(formula = phyloseq::distance(GP1_prune, method = "bray") ~ Phenotype, data = metadata_filt)
         Df SumOfSqs      R2      F Pr(>F)
Model     1 0.006446 0.05016 0.7394  0.587
Residual 14 0.122042 0.94984              
Total    15 0.128487 1.00000

Draw by parent with Complex Heatmap.  
```{r}
library(dendsort)
library(WGCNA)
library(circlize)

table<-as.matrix(t(otu_table(GP1_prune)))
                 
row_dend = dendsort(hclust(dist(table)))
col_dend = dendsort(hclust(dist(t(table))))

#col_fun = colorRamp2(c(0, 50, 100), c("white", "darkgray", "black"))
col_fun = colorRamp2(c(0, .1, .25, .5), hcl_palette = "Blue-Red 3")

pdf(file = "../output/ITS2/DIV-ComplexHeatmap.pdf", height = 7, width = 7)
ComplexHeatmap::Heatmap(table, name = "Relative \nAbundance", row_title = "DIV", column_title = "ITS2 DIV Relative Abundance (>1%)", 
        col = col_fun, 
        row_names_side = "left", 
        row_dend_side = "left",
        width = unit(4, "in"), 
        height = unit(4, "in"), 
        column_dend_height = unit(.5, "in"),
        row_dend_width = unit(.5, "in"),
        column_dend_reorder = FALSE,
        row_dend_reorder = TRUE,
        row_gap = unit(2.5, "mm"),
        clustering_distance_rows = "euclidean",
        clustering_method_rows = "complete",
        border = TRUE, 
        column_names_gp =  gpar(fontsize = 12, border=FALSE),
        column_names_rot = 35,
        cluster_rows = row_dend, 
        cluster_columns = col_dend,
        column_order = sample_order, 
        row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE))
dev.off()
```

Calculate alpha diversity of those that have >1 sequence
```{r}
GP1_rare_untransformed<-rarefy_even_depth(GP1, sample.size = min(sample_sums(GP1)),
  rngseed = TRUE, replace = TRUE, trimOTUs = TRUE, verbose = TRUE)

GP2 <- prune_taxa(taxa_sums(GP1_rare_untransformed) > 1, GP1_rare_untransformed)
```

Plot all richness metrics.  
```{r}
plot_richness(GP2)
```

Plot the two most commonly used - Shannon and Inverse Simpson.  
```{r}
alpha1<-plot_richness(GP2, measures=c("InvSimpson", "Shannon"), x="Phenotype", color="Phenotype", nrow=1)

alpha1_plot <- alpha1 + geom_boxplot() + geom_point(size=3) + theme_classic() +   scale_colour_manual(name="Phenotype", values=c("#3b244d", "#f2916b"), labels=phen_names) +
  theme(axis.text.x=element_text(angle=45, hjust=1, color="black")) + 
  theme(axis.title.x=element_blank())+
  theme(legend.position="none");alpha1_plot

ggsave(plot=alpha1_plot, file="../output/ITS2/alpha_rare.png", width=7, height=5)
```

Test for differences in alpha diversity.  
```{r}
metadata <- as(sample_data(GP2), "data.frame")
metadata$sample<-rownames(metadata)

alpha_rare<-estimate_richness(GP2, measures=c("InvSimpson", "Shannon", "Chao1"))
alpha_rare$sample<-rownames(alpha_rare)

alpha_rare<-full_join(metadata, alpha_rare, by="sample")

summary(aov(Shannon~Phenotype, data=alpha_rare))
summary(aov(InvSimpson~Phenotype, data=alpha_rare))
summary(aov(Chao1~Phenotype, data=alpha_rare))
```

No differences in alpha diversity 

Load profile level data 
```{r}
#create metadata and load
metadata <- read.csv("../data/ITS2/SymPortal_datasheet_Martynek.csv", header = TRUE, sep = ",")%>%dplyr::select(sample_name , Phenotype)

keep<-c("26", "27", "43", "44", "233", "234", "235", "236", "237", "238", "239", "240", "243", "244", "247", "248")

metadata <- metadata %>%
  filter(sample_name %in% keep)

rownames(metadata) <- metadata$sample_name

metadata <- as.matrix(metadata)
metadata <- sample_data(data.frame(metadata))

# load data for taxonomy
tax0 <- read_tsv(
  file  = "../data/ITS2/20251021T032609_martyma1/its2_type_profiles/638_20251021T051401_DBV_20251021T231108.profiles.absolute.abund_and_meta.txt", n_max = 6) %>%
  dplyr::select(-2) %>% 
  gather(uid, value, -1) %>% 
  spread(1, value) %>%
  clean_names()

tax1 <- as.matrix(tax0[, -1], dimnames = list(tax0$uid, colnames(tax0[-1])))
rownames(tax1) <- tax0$uid
tax <- tax_table(tax1)

#load data for count matrix 
otu0 <- read_tsv(
  file  = "../data/ITS2/20251021T032609_martyma1/its2_type_profiles/638_20251021T051401_DBV_20251021T231108.profiles.absolute.abund_and_meta.txt", col_names = TRUE) 

#list<-c("UID", "sample", "797700", "799603", "798782", "798780", "798578", "798776")
#colnames(otu0)<-list

otu0<-otu0%>%
  filter(...2 %in% keep) %>%
  #rename("sample" = "...2") %>%
  dplyr::select(-1) %>%
  mutate(across(-...2, ~as.numeric(as.character(.)))) # Ensure conversion from factors/chars

otu1 <- as.matrix(otu0[, -1])
class(otu1) <- "numeric"
rownames(otu1) <- otu0$...2
otu <- otu_table(otu1, taxa_are_rows = FALSE)

taxa_names(otu)
taxa_names(tax)

#combine to phyloseq object
coral <- phyloseq(otu, tax, metadata)
coral
```

Create tree.  
```{r}
random_tree_profile = rtree(ntaxa(coral), rooted=TRUE, tip.label=taxa_names(coral))
```

Merge phyloseq data. 
```{r}
phylo_coral = merge_phyloseq(coral, metadata, random_tree_profile)
phylo_coral
```

Filter out samples with low numbers in groups from phyloseq object.  
```{r}
phylo_coral<-ps_filter(phylo_coral)

phylo_coral<-ps_arrange(phylo_coral, Phenotype)
```

Now build a tree with the combined data.  
```{r}
plot_tree(phylo_coral, color="Phenotype", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
```

Plot a heatmap.  
```{r}
plot_heatmap(phylo_coral, "NMDS", "bray", "Phenotype", low="#66CCFF", high="#000033", na.value="white")
```

Plot ordinations.  
```{r}
GP1_profile = phylo_coral
GP1_profile
```

Subsample to minimum sampling depth  
```{r}
min(sample_sums(GP1_profile))

GP1_rare_profile<-rarefy_even_depth(GP1_profile, sample.size = min(sample_sums(GP1_profile)),
  rngseed = TRUE, replace = TRUE, trimOTUs = TRUE, verbose = TRUE)

sample_sums(GP1_rare_profile)
```

All samples now have 65353 sequences

Transform to relative abundance as a proportion. 
```{r}
GP1_rare_profile = transform_sample_counts(GP1_rare_profile, function(x) x/sum(x))
```

View tree and heatmap after subsampling  
```{r}
plot_tree(GP1_rare_profile, color="Phenotype", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
plot_heatmap(GP1_rare_profile, "NMDS", "bray", "Phenotype", low="#66CCFF", high="#000033", na.value="white")
```

Plot abundance of each profile  
```{r}
t<-plot_bar(GP1_rare_profile, fill="its2_type_profile")
t$data$Sample <- factor(t$data$Sample, levels = sample_order)

# Define custom colors
# C1: Oranges/Yellows | C15: Deep Reds/Maroons
# my_colors <- c(
#   "#FFB347", "#FB8C00",
#   "#E31A1C", "#CC1517", "#B51012", "#9E0B0D", "#870608", "#700103")
my_colors <- c(
  "#41AB5D", "#74C476", # C1 (Warm Copper)
  "#9DBAD5", "#7393B3", "#5D7B93", "#465F73", "#2F4353", "#182733" # C15 (Slate Gradient)
)

# Plot
t_plot <- t + 
  scale_fill_manual(name = "ITS2 Profile", values = my_colors) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", linewidth = 1) + 
  # facet_grid remains the same
  facet_grid(~Phenotype, scales = "free", space = "free") + 
  # Force bars to sit flush on the x-axis
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format(accuracy = 1)) +
  ylab("Relative Abundance (%)") +
  theme_classic() +
  theme(
    # Angle x-axis text
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    # Color the facet headers
    strip.text = element_text(face = "bold", size = 12),
    # Note: strip.text colors are usually uniform; 
    # to color them individually, we target the strip background or use ggh4x
    strip.background = element_blank() 
  ); t_plot

rel_abund_data<-t$data

ggsave("../output/ITS2/profile_plot.png", t_plot, width=11, height=6, dpi = 300)
```

Calculate the average percent abundance for each taxa in each phenotype  
```{r}
means <- rel_abund_data %>%
  group_by(its2_type_profile, Phenotype) %>%
  # Add abundances within each profile and lifestage
  summarize_at("Abundance", mean)%>%
  spread(its2_type_profile, Abundance);means
```

Plot relative abundance mean values  
```{r}
rel_abund_plot<-rel_abund_data%>%
  dplyr::select(its2_type_profile, Abundance, Phenotype, Sample)%>%
  group_by(its2_type_profile, Phenotype)%>%
  dplyr::summarise(mean=mean(Abundance, na.rm=TRUE), se=sd(Abundance, na.rm=TRUE)/sqrt(length(Abundance)))%>%
  
  ggplot(aes(x=Phenotype, y=mean, color=its2_type_profile))+
  facet_wrap(~its2_type_profile, nrow=2)+
  geom_point(position=position_dodge(0.2), size=6)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1, position=position_dodge(0.2), linewidth=2)+
  ylab("Mean Relative Abundance")+
  ylim(-0.1,1.1)+
  xlab("Phenotype")+
  scale_fill_manual(name = "ITS2 Profile", values = my_colors) +
  #scale_x_discrete(labels = parent_names) + 
  #ggtitle("C")+
  theme_classic()+
  theme(
    legend.position="none",
    text=element_text(color="black", size=12), 
    axis.text=element_text(color="black", size=11), 
    axis.text.x=element_text(angle=45, hjust=1, vjust=1)
  );rel_abund_plot

ggsave("../output/ITS2/profile_plot_means.png", plot=rel_abund_plot, width=10, height=6)
```

Make plots publication ready 
```{r}
## NMDS
p2b_viz<-p2  + 
  geom_point(aes(colour=Phenotype), size=8) + 
  #ggtitle("NMDS")+
  scale_colour_manual(name="Phenotype", values=c("#3b244d", "#f2916b"), labels=phen_names) +
  theme_classic() + 
  theme(
    panel.background=element_blank(),
    axis.text.y   = element_text(size = 24, color = "black"), 
    axis.text.x   = element_text(size = 24, color = "black", face = "plain"),
    axis.title  = element_text(size = 27, face = "bold"),
    legend.position = "right", # or "bottom" depending on your preference
    legend.title = element_text(size = 20, face = "bold"),
    legend.text = element_text(size = 18), 
    legend.key.size = unit(1.5, "cm"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)); p2b_viz

## Relative abundances 
t_plot <- t + 
  scale_fill_manual(name = "ITS2 Profile", values = my_colors) +
  #geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", linewidth = 1) + 
  # facet_grid remains the same
  facet_grid(~Phenotype, scales = "free", space = "free") + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format(accuracy = 1)) +
  ylab("Relative Abundance (%)") +
  xlab("Colony ID") +
  theme_classic() +
    theme(
    panel.background=element_blank(),
    axis.text.y   = element_text(size = 24, color = "black"), 
    axis.text.x   = element_text(size = 14, color = "black", face = "plain", angle = 45, vjust = 1, hjust = 1),
    axis.title  = element_text(size = 27, face = "bold"),
    legend.position = "right", # or "bottom" depending on your preference
    legend.title = element_text(size = 20, face = "bold"),
    legend.text = element_text(size = 18), 
    legend.key.size = unit(1.5, "cm"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    strip.text = element_text(face = "bold", size = 22),
    strip.background = element_blank() 
  ); t_plot

# Combine the plots
combined_plot <- (p2b_viz | t_plot) + 
  plot_annotation(tag_levels = 'A') +
  plot_layout(widths = c(1.15, 2)) & # Give the bar plot more horizontal room
  theme(plot.tag = element_text(size = 30, face = "bold"))

# Save 
ggsave("../output/ITS2/NMDS_RelativeAbundance.pdf", combined_plot, width = 25, height = 8, units = "in", dpi = 300)
ggsave("../output/ITS2/NMDS_RelativeAbundance.png", combined_plot, width = 25, height = 8, units = "in", dpi = 300)
```

Plot relative abundance of ITS2 sequences and profiles. 
```{r}
# Melt the phyloseq object to a long-format dataframe
df_long <- psmelt(GP1_rare)

# Apply the 5% threshold logic per sample
df_final <- df_long %>%
  group_by(Sample) %>%
  mutate(DIV = ifelse(Abundance < 0.05, "Other", DIV)) %>%
  group_by(Sample, DIV, Phenotype) %>%
  dplyr::summarise(Abundance = sum(Abundance), .groups = "drop")

# Re-order samples to match your desired resistant/susceptible layout
df_final$Sample <- factor(df_final$Sample, levels = sample_order)

unique_divs <- unique(df_final$DIV)
c15_main <- "C15" # Assuming the main sequence is labeled 'C15'
other_variants <- setdiff(unique_divs, c(c15_main, "Other"))
blue_palette <- c("#2171b5", "#4292c6", "#6baed6", "#9ecae1", "#c6dbef")
my_pal_seq <- c(
  setNames("#2171b5", c15_main),                          # Darkest Blue for C15
  setNames(blue_palette[2:(length(other_variants)+1)], other_variants), # Lighter blues for others
  setNames("#D3D3D3", "Other"))

df_final$DIV <- factor(df_final$DIV, levels = c(other_variants, c15_main, "Other"))

# Standardize text size for plots
text_size_axis_title <- 20
text_size_axis_text  <- 14
text_size_legend     <- 16
text_size_strip      <- 18

t_seq <- ggplot(df_final, aes(x = Sample, y = Abundance, fill = DIV)) +
  geom_bar(stat = "identity", width = 0.9, color = "black", linewidth = 0.4) + 
  scale_fill_manual(name = "ITS2 Sequence", values = my_pal_seq) +
  facet_grid(~Phenotype, scales = "free", space = "free") +
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0.05))) +
  theme_classic() +
  theme(
    # Center axis text: hjust = 0.5 for centered, or 1 for angled alignment
    axis.text.x  = element_text(angle = 45, hjust = 1, size = text_size_axis_text, color = "black"),
    axis.text.y  = element_text(size = text_size_axis_text, color = "black"),
    axis.title   = element_text(face = "bold", size = text_size_axis_title),
    strip.text   = element_text(face = "bold", size = text_size_strip),
    strip.background = element_blank(),
    legend.title = element_text(face = "bold", size = text_size_legend),
    legend.text  = element_text(size = text_size_legend - 2),
    plot.margin  = margin(10, 10, 2, 10) # Minimal bottom margin to connect A and B
  ) +
  ylab("Relative Abundance (%)") +
  xlab(NULL); t_seq

my_colors_prof <- c("#41AB5D", "#74C476", "#9DBAD5", "#7393B3", "#5D7B93", "#465F73", "#2F4353", "#182733")
t_prof <- t + 
  #geom_bar(stat = "identity", width = bar_width_set, color = bar_outline_color, linewidth = bar_line_width) +
  scale_fill_manual(name = "ITS2 Profile", values = my_colors_prof) +
  facet_grid(~Phenotype, scales = "free", space = "free") + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format(accuracy = 1)) +
  theme_classic() +
  theme(
    axis.text.y   = element_text(size = text_size_axis_text, color = "black"), 
    axis.text.x   = element_text(size = text_size_axis_text, color = "black", angle = 45, hjust = 1),
    axis.title    = element_text(size = text_size_axis_title, face = "bold"),
    strip.text    = element_text(face = "bold", size = text_size_strip),
    strip.background = element_blank(),
    legend.title  = element_text(size = text_size_legend, face = "bold"),
    legend.text   = element_text(size = text_size_legend - 2),
    plot.margin   = margin(2, 10, 10, 10) # Minimal top margin
  ) +
  ylab("Relative Abundance (%)") +
  xlab("Colony ID"); t_prof

# This stacks them vertically as shown in your reference image
combined_its2_plot <- (t_seq / t_prof) + 
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(face = "bold", size = 24))

# 4. Save the combined figure
ggsave("../output/ITS2/combined_seq_profile_plot.png", combined_its2_plot, width=16, height=12, dpi=300)
ggsave("../output/ITS2/combined_seq_profile_plot.pdf", combined_its2_plot, width=16, height=12, dpi=300)
```

