---
title: "Surface area calculations"
author: "Jill Ashey"
date: "2026-02-26"
output: html_document
---

This script will calculate the surface area of the TPC 2025 fragments via the wax dipping method. Kristen calculated the TPC 2023 fragments via wax dipping. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

Load wax data and calculate change in weight 
```{r}
wax_data <- read.csv("../data/Surface_Area/Wax_weights_TPC2025.csv")
wax_data <- wax_data %>%
  mutate(delta.mass.g = weight2.g - weight1.g)
```

Subset standards and calculate the surface area 
```{r}
wax_data_std <- wax_data %>%
  filter(Sample == "Standard") %>%
  mutate(radius = Diameter/2,
         surface.area.cm2 = (pi * Diameter * Height) + (pi * radius^2))
```

Calculate curve coefficients for slope and intercept based on standards 
```{r}
std.curve <- lm(surface.area.cm2 ~ delta.mass.g + 0, data = wax_data_std)
std.curve$coefficients
summary(std.curve)$r.squared
plot(surface.area.cm2 ~ delta.mass.g, data = wax_data_std, 
     main="Standard Curve (Forced through Origin)", 
     xlab="Wax Weight Gain (g)", ylab="Surface Area (cm2)")
abline(std.curve, col="red")
```

Calculate the surface area of samples using the standard curve 
```{r}
wax_data_samp <- wax_data %>%
  filter(Sample == "Coral") %>%
  #mutate(surface.area.cm2 = std.curve$coefficients[2] * delta.mass.g + std.curve$coefficients[1]) %>%
  mutate(surface.area.cm2 = predict(std.curve, newdata = .)) # predict automatically uses the slope from std curve
```

Make sure samples fall within range of stds 
```{r}
range(wax_data_samp$surface.area.cm2, na.rm = T)
range(wax_data_std$surface.area.cm2)

mean(wax_data_samp$surface.area.cm2, na.rm = T)
```

Keep only colony and surface area info
```{r}
wax_data_samp <- wax_data_samp %>%
  dplyr::select(Colony_ID, weight1.g, weight2.g, delta.mass.g, surface.area.cm2)
```

Save as csv 
```{r}
write.csv(wax_data_samp, "../data/Surface_Area/Surface_Area_wax_TPC2025.csv")
```
