---
title: "Temperature"
author: "Jill Ashey"
date: "2025-08-14"
output: html_document
---

Plot temperature from Patch Reef 13 in KBay

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(readr)
library(tidyverse)
library(lubridate)
library(purrr)
```

Read in data from 2013-2023 (data obtained from Brown et al. 2023 - https://www.bco-dmo.org/dataset/905047)
```{r}
temps <- read.csv("../data/environmental/905047_v1_infield_temp.csv", header = T)
```

```{r}
# Make sure your datetime column is actually a POSIXct time object
temps$ISO_DateTime_UTC <- as.POSIXct(temps$ISO_DateTime_UTC,
                                          format = "%Y-%m-%dT%H:%MZ",
                                          tz = "UTC")

# Plot
ggplot(temps, aes(x = ISO_DateTime_UTC, y = Temperature, color = Site)) +
  geom_point() +
  labs(x = "Time (UTC)",y = "Temperature (°C)") +
  theme_minimal()
```

Read in data from hobo loggers from NSF career project. These are all in different formats so they need to be read in and processed differently. 
```{r}
# Sept 2022 - April 2023 
df1 <- read.csv("../data/environmental/2022-Sept_2023-April_Reef13_1m_21152005.csv", header = T)
df2 <- read.csv("../data/environmental/2022-Sept_2023-April_Reef13_1m_21152007.csv", header = T)
df3 <- read.csv("../data/environmental/2022-Sept_2023-April_Reef13_3m_21152009.csv", header = T)

# April 2024 - Sept 2024 
df4 <- read.csv("../data/environmental/2024-09-30_Reef13_21152005.csv", skip = 1, header = T)
df5 <- read.csv("../data/environmental/2024-09-30_Reef13_21152007.csv", skip = 1, header = T)
df6 <- read.csv("../data/environmental/2024-09-30_Reef13_21152009.csv", skip = 1, header = T)

# Jan 2025 - May 2025
df7 <- read.csv("../data/environmental/21152005.csv", header = T)
df8 <- read.csv("../data/environmental/21152007.csv", header = T)
df9 <- read.csv("../data/environmental/21152009.csv", header = T)

# Sept 2023 - Jan 2024 
df10 <- read.csv("../data/environmental/Sept2023_Jan2024_Reef13_insitu_temp_21152005_3m.csv", skip = 1, header = FALSE)
df11 <- read.csv("../data/environmental/Sept2023_Jan2024_Reef13_insitu_temp_21152007_1m.csv", skip = 1, header = FALSE)
df12 <- read.csv("../data/environmental/Sept2023_Jan2024_Reef13_insitu_temp_21152009_1m.csv", skip = 1, header = FALSE)
```

```{r}
## --- Group 1: like df1 (separate Date + Time) ---
df_group1 <- list(df1, df2, df3)

df_group1_clean <- map2(df_group1, paste0("df", 1:3), ~
  .x %>%
    mutate(
      DateTime = dmy_hms(paste(Date, Time..GMT.10.00)),
      Temp_C = Temp...C,
      Source = .y
    ) %>%
    select(DateTime, Temp_C, Source)
)

## --- Group 2: like df4 (combined DateTime, long header) ---
df_group2 <- list(df4, df5, df6)

df_group2_clean <- map2(df_group2, paste0("df", 4:6), ~
  .x %>%
    mutate(
      DateTime = mdy_hms(Date.Time..GMT.10.00),
      Temp_C = as.numeric(
        .x[[grep("Temp", names(.x), value = TRUE)[1]]]
      ),
      Source = .y
    ) %>%
    select(DateTime, Temp_C, Source)
)

## --- Group 3: like df7 (separate Date + Time again, but US format) ---
df_group3 <- list(df7, df8, df9)

df_group3_clean <- map2(df_group3, paste0("df", 7:9), ~
  .x %>%
    mutate(
      DateTime = mdy_hms(paste(Date, Time..GMT.10.00)),
      Temp_C = Temp...C,
      Source = .y
    ) %>%
    select(DateTime, Temp_C, Source)
)

## --- Group 4: df10–df12 (Date, Time, Temp_C) --- 
df_group4 <- list(df10, df11, df12)

df_group4_clean <- map2(df_group4, paste0("df", 10:12), ~
  .x %>%
    mutate(
      DateTime = dmy_hms(paste(V1, V2)),   # Date in dmy format (19-09-23)
      Temp_C   = as.numeric(V3),
      Source   = .y
    ) %>%
    select(DateTime, Temp_C, Source)
)

## --- Combine all ---
all_data <- bind_rows(df_group1_clean, df_group2_clean, df_group3_clean, df_group4_clean)

head(all_data)

ggplot(all_data, aes(x = DateTime, y = Temp_C)) +
  geom_point() +
  scale_x_datetime( date_breaks = "1 month", date_labels = "%b") + 

  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time")
```

Combine data 
```{r}
# Clean temps
temps_clean <- temps %>%
  mutate(
    DateTime = ymd_hms(ISO_DateTime_UTC, tz = "UTC"),  # handles yyyy-mm-ddTHH:MMZ
    Temp_C   = Temperature,
    Source   = Site
  ) %>%
  select(DateTime, Temp_C, Source)

# Make sure all_data has the same structure
all_data_clean <- all_data %>%
  select(DateTime, Temp_C, Source)

# Combine
combined_data <- bind_rows(all_data_clean, temps_clean)
```

```{r}
ggplot(combined_data, aes(x = DateTime, y = Temp_C)) + 
  geom_point() + 
  geom_hline(yintercept = 28, color = "grey", size = 1, linetype = "dashed") +
  geom_hline(yintercept = 29, color = "grey", size = 1, linetype = "solid") +
  theme_classic() + 
  labs(y = "Temperature (°C)", x = "") + 
  scale_x_datetime( date_breaks = "1 year", date_labels = "%Y") + 
  geom_vline(xintercept = as.POSIXct("2014-10-01 01:17:00"), color = "red", size = 1, linetype = "dashed") +
  geom_vline(xintercept = as.POSIXct("2015-10-01 00:00:00"), color = "red", size = 1, linetype = "dashed") +
  geom_vline(xintercept = as.POSIXct("2019-10-01 00:00:00"), color = "red", size = 1, linetype = "dashed") +
  geom_vline(xintercept = as.POSIXct("2023-09-14 00:00:00"), color = "blue", size = 1, linetype = "dotdash") +
  geom_vline(xintercept = as.POSIXct("2024-01-20 00:00:00"), color = "blue", size = 1, linetype = "dotdash") + 
  geom_vline(xintercept = as.POSIXct("2024-05-09 00:00:00"), color = "blue", size = 1, linetype = "dotdash") +
  geom_vline(xintercept = as.POSIXct("2024-10-01 00:00:00"), color = "blue", size = 1, linetype = "dotdash") +
  geom_vline(xintercept = as.POSIXct("2025-01-09 00:00:00"), color = "blue", size = 1, linetype = "dotdash") +
  geom_vline(xintercept = as.POSIXct("2025-05-08 00:00:00"), color = "blue", size = 1, linetype = "dotdash") +
  theme(axis.text.x = element_text(size = 30,
                                   color = "black")) +
  theme(axis.text.y = element_text(size = 30,
                                   color = "black"),
        axis.title.y = element_text(size = 35,
                                    face = "bold",
                                    margin = margin(t = 0, r = 20, b = 0, l = 0)))


```


Summarize 
```{r}
daily_summary <- combined_data %>%
  mutate(Date = as.Date(DateTime)) %>%
  group_by(Date, Source) %>%   # keep Source if you want separate curves per Source
  summarise(
    mean_temp = mean(Temp_C, na.rm = TRUE),
    sd_temp   = sd(Temp_C, na.rm = TRUE),
    .groups = "drop") %>%
  mutate(
    lower = mean_temp - sd_temp,
    upper = mean_temp + sd_temp)
```



Plot
```{r}
ggplot(daily_summary, aes(x = Date, y = mean_temp, colour = Source)) +
  geom_ribbon(
    aes(ymin = lower, ymax = upper, fill = Source),
    alpha = 0.3, colour = NA, show.legend = FALSE
  ) +
  geom_smooth(
    method = "loess", span = 0.1, se = FALSE,
    aes(colour = Source), alpha = 1, show.legend = TRUE
  ) +
  theme_minimal() +
  labs(y = "Daily Mean Temperature (°C)", x = "Time") +
  scale_x_date(
    date_breaks = "1 year", 
    date_labels = "%Y"
  ) +
  geom_vline(xintercept = as.Date("2014-10-01"), color = "red", linetype = "dashed") +
  geom_vline(xintercept = as.Date("2015-10-01"), color = "red", linetype = "dashed") +
  geom_vline(xintercept = as.Date("2019-10-01"), color = "red", linetype = "dashed")
```

Other plot options 

Option 1
```{r}
daily_summary <- combined_data %>%
  mutate(Date = as.Date(DateTime)) %>%
  group_by(Date) %>%
  summarise(mean_temp = mean(Temp_C, na.rm = TRUE),
            sd_temp   = sd(Temp_C, na.rm = TRUE)) %>%
  mutate(lower = mean_temp - sd_temp,
         upper = mean_temp + sd_temp)

ggplot(daily_summary, aes(x = Date, y = mean_temp)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "lightblue", alpha = 0.3) +
  geom_line(color = "blue") +
  theme_minimal() +
  labs(y = "Daily Mean Temperature (°C)", x = "Date") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  geom_vline(xintercept = as.Date(c("2014-10-01", "2015-10-01", "2019-10-01")),
             color = "red", linetype = "dashed")

```

Option 2
```{r}
ggplot(combined_data, aes(x = DateTime, y = Temp_C)) +
  geom_point(alpha = 0.05, size = 0.5, color = "blue") +
  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time") +
  scale_x_datetime(date_breaks = "1 year", date_labels = "%Y") +
  geom_vline(xintercept = as.POSIXct(c("2014-10-01 01:17:00",
                                       "2015-10-01 00:00:00",
                                       "2019-10-01 00:00:00")),
             color = "red", linetype = "dashed")

```

Option 3
```{r}
ggplot(combined_data, aes(x = DateTime, y = Temp_C)) +
  geom_point(alpha = 0.5, size = 0.5, color = "black") +
  geom_smooth(method = "loess", span = 0.05, se = FALSE, color = "black") +
  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time") +
  scale_x_datetime(date_breaks = "1 year", date_labels = "%Y") +
  geom_vline(xintercept = as.POSIXct(c("2014-10-01 01:17:00",
                                       "2015-10-01 00:00:00",
                                       "2019-10-01 00:00:00")),
             color = "red", linetype = "dashed")

```


Calculate MMM
```{r}
library(dplyr)
library(lubridate)
library(zoo)

# Add year and month columns
combined_data <- combined_data %>%
  mutate(Year = year(DateTime), Month = month(DateTime))

# Calculate mean temperature for each month of each year
monthly_means <- combined_data %>%
  group_by(Year, Month) %>%
  summarise(mean_temp = mean(Temp_C, na.rm = TRUE), .groups = 'drop')

# Find the maximum monthly mean for each year
annual_max_monthly_means <- monthly_means %>%
  group_by(Year) %>%
  summarise(max_monthly_mean = max(mean_temp), .groups = 'drop')

# Calculate MMM (mean of the annual maxima)
MMM_value <- mean(annual_max_monthly_means$max_monthly_mean, na.rm = TRUE)

# 1. Set your MMM
MMM_value <- MMM_value # change this to your known MMM

# 2. Calculate anomaly (DHW anomaly)
combined_data <- combined_data %>%
  mutate(anomaly = pmax(0, Temp_C - (MMM_value + 1)))

# 3. Convert half-hourly anomalies to degree-weeks (sum over 12-week rolling window)
# Compute how many measurements in a week:
samples_per_week <- 7 * 24 * 2 # 2 samples per hour

# 4. Create a rolling sum window (12 weeks = 12 * samples_per_week)
window_length <- 12 * samples_per_week

combined_data <- combined_data %>%
  arrange(DateTime) %>%
  mutate(
    DHW = rollapply(
      anomaly,
      width = window_length,
      FUN = function(x) sum(x) / samples_per_week, # degree-weeks
      align = "right",
      fill = NA
    )
  )
```

```{r}
library(ggplot2)
library(scales)

# Ensure your DHW is numeric
combined_data$DHW <- as.numeric(combined_data$DHW)

ggplot(combined_data, aes(x = DateTime)) +
  geom_line(aes(y = Temp_C, color = "Temperature (°C)")) +
  geom_line(aes(y = DHW, color = "DHW")) +
  scale_y_continuous(
    name = "Temperature (°C)",
    sec.axis = sec_axis(~ ., name = "Degree Heating Weeks (DHW)")
  ) +
  scale_color_manual(
    values = c("Temperature (°C)" = "steelblue", "DHW" = "firebrick")
  ) +
  labs(
    x = "Date",
    color = "",
    title = "Sea Temperature and Degree Heating Weeks Time Series"
  ) +
  theme_minimal() +
  theme(
    axis.title.y.left = element_text(color = "steelblue"),
    axis.title.y.right = element_text(color = "firebrick"),
    legend.position = "top"
  )

## scale right y axis: 
scale_factor <- max(combined_data$Temp_C, na.rm=TRUE) / max(combined_data$DHW, na.rm=TRUE)
ggplot(combined_data, aes(x = DateTime)) +
  geom_line(aes(y = Temp_C, color = "Temperature (°C)")) +
  geom_line(aes(y = DHW * scale_factor, color = "DHW (scaled)")) + # scale DHW for display
  scale_y_continuous(
    name = "Temperature (°C)",
    sec.axis = sec_axis(~ . / scale_factor, name = "Degree Heating Weeks (DHW)")
  ) +
  scale_color_manual(
    values = c("Temperature (°C)" = "steelblue", "DHW (scaled)" = "firebrick")
  ) +
  labs(
    x = "Date",
    color = "",
    title = "Sea Temperature and Degree Heating Weeks Time Series"
  ) +
  theme_minimal() +
  theme(
    axis.title.y.left = element_text(color = "steelblue"),
    axis.title.y.right = element_text(color = "firebrick"),
    legend.position = "top"
  )
```

other code to calculate DHW
```{r}
library(dplyr)
library(lubridate)
library(zoo)

# 1. Daily mean temps
daily_data <- combined_data %>%
  mutate(Date = as.Date(DateTime)) %>%
  group_by(Date) %>%
  summarise(Temp_C = mean(Temp_C, na.rm = TRUE)) %>%
  ungroup()

# 2. Estimate Maximum Monthly Mean (MMM)
# If you don’t have climatology, take the mean of the warmest month in your dataset
daily_data <- daily_data %>%
  mutate(Month = month(Date)) 

mmm <- daily_data %>%
  group_by(Month) %>%
  summarise(MonthlyMean = mean(Temp_C, na.rm = TRUE)) %>%
  arrange(desc(MonthlyMean)) %>%
  slice(1) %>%
  pull(MonthlyMean)

cat("MMM =", mmm, "\n")

# 3. Calculate HotSpot (only values >1°C above MMM)
daily_data <- daily_data %>%
  mutate(HotSpot = ifelse(Temp_C > (mmm + 1), Temp_C - mmm, 0))

# 4. Calculate Degree Heating Weeks (DHW)
# 84-day (12-week) rolling sum of HotSpots, converted to °C-weeks
daily_data <- daily_data %>%
  mutate(DHW = rollapply(HotSpot, width = 84, FUN = sum, align = "right", fill = NA) / 7)

head(daily_data)

```

Read in data 
```{r}
dhw <- read.csv("../data/environmental/Reef13 temp hotspot through 2023.csv")
mean(dhw$DHW)
max(dhw$DHW)

```

Compare brown v NOAA
```{r}
library(dplyr)
library(zoo)

# --- Parameters ---
MMM <- 27.3                # climatological MMM (Heron Reef, Weeks et al. 2008)
threshold <- MMM + 1       # bleaching threshold = 28.3 °C

# --- Daily data already computed ---
# daily_data has columns: Date, Temp_C

daily_data <- daily_data %>%
  arrange(Date) %>%
  mutate(
    # NOAA CRW method -------------------------
    HS_NOAA = ifelse(Temp_C > threshold, Temp_C - threshold, 0),
    DHW_NOAA = rollapply(HS_NOAA, width = 84, FUN = sum,
                         align = "right", fill = NA) / 7,

    # Brown et al. (2023) method --------------
    HS_Brown = ifelse(Temp_C >= threshold, Temp_C - MMM, 0),
    DHW_Brown = rollapply(HS_Brown, width = 90, FUN = sum,
                          align = "right", fill = NA) / 7
  )

# Preview
daily_data %>%
  select(Date, Temp_C, HS_NOAA, DHW_NOAA, HS_Brown, DHW_Brown) %>%
  head(20)

library(ggplot2)

ggplot(daily_data, aes(x = Date)) +
  # Temp on left axis
  geom_line(aes(y = Temp_C), color = "steelblue") +
  geom_hline(yintercept = 28.3, linetype = "dashed", color = "red") +  # threshold
  
  # Brown DHW on right axis
  geom_line(aes(y = DHW_Brown * 3), color = "darkorange") +  
  # multiply by ~3 so scales are similar (tweak to match your data range)
  
  scale_y_continuous(
    name = "Daily Mean Temp (°C)",
    sec.axis = sec_axis(~./3, name = "Brown DHW (°C-weeks)")  # inverse of *3
  ) +
  theme_minimal() +
  labs(title = "Temperature & DHW (Brown et al. 2023)", x = "Date")

ggplot(daily_data, aes(x = Date)) +
  # DHW bars (Brown)
  geom_col(aes(y = DHW_Brown), fill = "firebrick", alpha = 0.6) +
  
  # Temperature line
  geom_line(aes(y = Temp_C), color = "steelblue") +
  geom_hline(yintercept = 28.3, linetype = "dashed", color = "red") +
  
  scale_y_continuous(
    name = "Temp (°C) / DHW (°C-weeks)"   # shared axis label
  ) +
  theme_minimal() +
  labs(title = "Temperature (line) & DHW (bars)", x = "Date")

# Define scaling factor to map DHW (0-10) onto temperature axis (20-30)
scale_factor <- 10 / (30 - 20)  # 10/10 = 1, but let's keep the example general

ggplot(daily_data, aes(x = Date)) +
  geom_col(aes(y = DHW_Brown / scale_factor), fill = "firebrick", alpha = 0.6) +  # Scale DHW down for plotting
  geom_line(aes(y = Temp_C), color = "steelblue") +
  geom_hline(yintercept = 28.3, linetype = "dashed", color = "red") +
  scale_y_continuous(
    name = "Temp (°C)",
    limits = c(20, 30),
    sec.axis = sec_axis(~ . * scale_factor, name = "DHW (°C-weeks)")  # Right axis scales back up
  ) +
  theme_minimal() +
  labs(title = "Temperature (line) & DHW (bars)", x = "Date")

```


```{r}
# Transformation functions
# From DHW to temperature scale for plotting
transform_dhw_to_temp <- function(x) 20 + (x / 10) * 10

# Inverse transformation for secondary axis labels
transform_temp_to_dhw <- function(x) (x - 20) / 10 * 10

ggplot(daily_data, aes(x = Date)) +
  geom_col(aes(y = transform_dhw_to_temp(DHW_Brown)), fill = "firebrick", alpha = 0.6) +
  geom_line(aes(y = Temp_C), color = "steelblue") +
  geom_hline(yintercept = 28.3, linetype = "dashed", color = "red") +
  scale_y_continuous(
    name = "Temp (°C)",
    limits = c(20, 30),
    sec.axis = sec_axis(~ transform_temp_to_dhw(.), name = "DHW (°C-weeks)")
  ) +
  theme_minimal() +
  labs(title = "Temperature (line) & DHW (bars)", x = "Date")


```

```{r}
transform_dhw_to_temp <- function(x) 20 + (x / 10) * 10
transform_temp_to_dhw <- function(x) (x - 20) / 10 * 10

ggplot(daily_data, aes(x = Date)) +
  geom_col(aes(y = transform_dhw_to_temp(DHW_Brown)), fill = "firebrick", alpha = 0.6) +
  geom_line(aes(y = Temp_C), color = "steelblue") +
  geom_hline(yintercept = 28.3, linetype = "dashed", color = "red") +
  scale_y_continuous(
    name = "Temp (°C)",
    sec.axis = sec_axis(~ transform_temp_to_dhw(.), name = "DHW (°C-weeks)")
  ) +
  coord_cartesian(ylim = c(20, 30)) +  # Zoom y-axis without dropping data
  theme_minimal() +
  labs(title = "Temperature (line) & DHW (bars)", x = "Date")



####
# Adjust DHW y values relative to baseline at 20
dhw_offset <- function(x) (x / 10) * 10  # scale DHW from 0-10 to 0-10 (height only)

ggplot(daily_data, aes(x = Date)) +
  geom_col(aes(y = dhw_offset(DHW_Brown)), fill = "firebrick", alpha = 0.6) +
  geom_line(aes(y = Temp_C - 20), color = "steelblue") +  # shift Temp for sync with DHW offset
  geom_hline(yintercept = 8.3, linetype = "dashed", color = "red") +  # 28.3 - 20 = 8.3 adjusted
  scale_y_continuous(
    name = "DHW (°C-weeks)",
    limits = c(0, 10),
    sec.axis = sec_axis(~ . + 20, name = "Temp (°C)")
  ) +
  theme_minimal() +
  labs(title = "Temperature (line) & DHW (bars)", x = "Date")


#####
scale_factor <- 10 / (30 - 20)  # scale DHW range (0-10) to temp range (20-30)

ggplot(daily_data, aes(x = Date)) +
  geom_line(aes(y = Temp_C), color = "steelblue") +
  geom_line(aes(y = DHW_Brown / scale_factor), color = "firebrick", alpha = 0.6) +
  geom_hline(yintercept = 28.3, linetype = "dashed", color = "red") +
  scale_y_continuous(
    name = "Temp (°C)",
    limits = c(20, 30),
    sec.axis = sec_axis(~ . * scale_factor, name = "DHW (°C-weeks)")
  ) +
  theme_minimal() +
  labs(title = "Temperature & DHW (lines)", x = "Date")

```

```{r}
library(ggplot2)
library(dplyr)

# Calculate scaling factor
scale_factor <- 10 / (30 - 20)  # 10 / 10 = 1

# Prepare your data for geom_rect: 
# For bars we need xmin, xmax, ymin, ymax
# Approximate bar width from Date diff:
bar_width <- diff(range(daily_data$Date)) / nrow(daily_data) * 0.9

daily_data <- daily_data %>%
  mutate(
    xmin = Date - bar_width / 2,
    xmax = Date + bar_width / 2,
    ymin = 20,  # baseline temperature
    ymax = 20 + DHW_Brown / scale_factor  # scaled DHW on temp scale
  )

ggplot() +
  geom_rect(
    data = daily_data_filtered,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = "firebrick", alpha = 0.6
  ) +
  geom_line(data = daily_data, aes(x = Date, y = Temp_C), color = "steelblue") +
  geom_hline(yintercept = 28.3, linetype = "dashed", color = "red") +
  scale_y_continuous(
    name = "Temp (°C)",
    limits = c(20, 30),
    sec.axis = sec_axis(~ . - 20, name = "DHW (°C-weeks)")
  ) +
  theme_minimal() +
  labs(title = "Temperature (line) & DHW (bars)", x = "Date")
```

Discuss w/ Katie. 

