---
title: "Temperature"
author: "Jill Ashey"
date: "2025-08-14"
output: html_document
---

Plot temperature from Patch Reef 13 in KBay

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(readr)
library(tidyverse)
library(lubridate)
library(purrr)
library(zoo)
library(scales)
```

Read in data from 2013-2023 (data obtained from Brown et al. 2023 - https://www.bco-dmo.org/dataset/905047)
```{r}
temps <- read.csv("../data/environmental/905047_v1_infield_temp.csv", header = T)
```

Format date and time for data from 2013-2023
```{r}
# Create a date column for plotting
temps <- temps %>%
  mutate(date = as.Date(paste(year, month, day, sep = "-")))

# Plot
ggplot(temps, aes(x = date, y = Temperature, color = Site)) +
  geom_point() +
  labs(x = "Time (UTC)",y = "Temperature (°C)") +
  theme_minimal()
```

Calculate DHW
```{r}
MMM <- 27.3
threshold <- MMM + 1 # 28.3

daily_mean <- temps %>%
  group_by(year, month, day) %>%
  summarize(daily_temp_mean = mean(Temperature, na.rm = T)) %>%
  ungroup() %>%
  arrange(year, month, day) %>%
  mutate(
    date = as.Date(paste(year, month, day, sep = "-")),
    HS = ifelse(daily_temp_mean >= threshold, daily_temp_mean - MMM, 0)
  )

# Check if dates are continuous:
date_seq <- seq(min(daily_mean$date), max(daily_mean$date), by = "day")
missing_dates <- setdiff(date_seq, daily_mean$date)

if(length(missing_dates) > 0) {
  # Fill missing dates with HS = 0, daily_temp_mean = NA or interpolated
  missing_df <- data.frame(
    year = as.integer(format(missing_dates, "%Y")),
    month = as.integer(format(missing_dates, "%m")),
    day = as.integer(format(missing_dates, "%d")),
    daily_temp_mean = NA,
    date = missing_dates,
    HS = 0
  )
  daily_mean <- bind_rows(daily_mean, missing_df) %>%
    arrange(date)
}

# Now compute rolling DHW (90-day window)
daily_mean <- daily_mean %>%
  mutate(DHW = rollapply(HS / 7, width = 90, sum, align = "right", fill = NA))

```

Plot! 
```{r}
# Define your MMM and bleaching thresholds
MMM <- 27.3
bleaching_threshold <- MMM + 1 # 28.3

# Create a date column for plotting
daily_mean <- daily_mean %>%
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(DHW = ifelse(is.na(DHW), 0, DHW))

# scale constants (temp 20–30, DHW 0–10)
temp_min <- 20
temp_max <- 30
dhw_min  <- 0
dhw_max  <- 10

# scaling factor
scale_factor <- (temp_max - temp_min) / (dhw_max - dhw_min)

ggplot(daily_mean, aes(x = date)) +
  geom_line(aes(y = daily_temp_mean, color = "Temperature"), linewidth = 0.8) +
  geom_line(aes(y = DHW * scale_factor + temp_min, color = "DHW"), linewidth = 0.8, alpha = 0.8) +
  scale_y_continuous(
    name = "Daily Temperature (°C)",
    limits = c(temp_min, temp_max),
    sec.axis = sec_axis(
      ~ (. - temp_min) / scale_factor,
      name = "Degree Heating Weeks (°C-weeks)",
      breaks = seq(0, 10, 2)
    )
  ) +
  scale_color_manual(
    name = NULL,
    values = c("Temperature" = "steelblue", "DHW" = "firebrick")
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.title.y.left  = element_text(color = "steelblue"),
    axis.title.y.right = element_text(color = "firebrick"),
    legend.position = "top"
  )
```

Plot with the points from temps
```{r}
MMM <- 27.3
threshold <- MMM + 1

# Prepare daily_mean with date
daily_mean <- daily_mean %>% 
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(datetime = as.POSIXct(date, format = "%Y-%m-%d", tz = "UTC"))

# Prepare hourly temps with POSIXct datetime
temps <- temps %>%
  mutate(
    datetime = make_datetime(year = year, month = month, day = day, hour = hour, min = 0, sec = 0, tz = "UTC"))

# Plot
ggplot() +
  # Hourly temperatures as light blue semi-transparent dots
  geom_point(data = temps, aes(x = datetime, y = Temperature), color = "grey60", alpha = 0.3, size = 0.5) +
  
  # Daily mean temperature as black line
  geom_line(data = daily_mean, aes(x = datetime, y = daily_temp_mean, color = "Temperature"), color = "black", linewidth = 0.8) +
  
  # DHW as gray line scaled to temperature y-axis limits (20-30)
  geom_line(data = daily_mean, aes(x = datetime, 
                                   y = DHW * (30 - 20) / (10 - 0) + 20), 
            color = "firebrick", size = 1, alpha = 0.7) +
  
  # Horizontal lines for MMM and Bleaching Threshold
  geom_hline(yintercept = MMM, linetype = "dashed", color = "black") +
  geom_hline(yintercept = threshold, linetype = "solid", color = "black") +
  
  # Primary y-axis limits and secondary axis that maps 20-30 temp to 0-10 DHW
  scale_y_continuous(
    name = "Temperature (°C)",
    limits = c(20, 30),
    sec.axis = sec_axis(
      trans = ~ (. - 20) * 10 / (30 - 20),
      name = "Degree Heating Weeks (°C wk⁻¹)",
      breaks = seq(0, 10, 1)
    )
  ) +
  theme_minimal()

```




















Read in data from hobo loggers from NSF CAREER project. These are all in different formats so they need to be read in and processed differently. 
```{r}
# Sept 2022 - April 2023 
df1 <- read.csv("../data/environmental/2022-Sept_2023-April_Reef13_1m_21152005.csv", header = T)
df2 <- read.csv("../data/environmental/2022-Sept_2023-April_Reef13_1m_21152007.csv", header = T)
df3 <- read.csv("../data/environmental/2022-Sept_2023-April_Reef13_3m_21152009.csv", header = T)

# April 2024 - Sept 2024 
df4 <- read.csv("../data/environmental/2024-09-30_Reef13_21152005.csv", skip = 1, header = T)
df5 <- read.csv("../data/environmental/2024-09-30_Reef13_21152007.csv", skip = 1, header = T)
df6 <- read.csv("../data/environmental/2024-09-30_Reef13_21152009.csv", skip = 1, header = T)

# Jan 2025 - May 2025
df7 <- read.csv("../data/environmental/21152005.csv", header = T)
df8 <- read.csv("../data/environmental/21152007.csv", header = T)
df9 <- read.csv("../data/environmental/21152009.csv", header = T)

# Sept 2023 - Jan 2024 
df10 <- read.csv("../data/environmental/Sept2023_Jan2024_Reef13_insitu_temp_21152005_3m.csv", skip = 1, header = FALSE)
df11 <- read.csv("../data/environmental/Sept2023_Jan2024_Reef13_insitu_temp_21152007_1m.csv", skip = 1, header = FALSE)
df12 <- read.csv("../data/environmental/Sept2023_Jan2024_Reef13_insitu_temp_21152009_1m.csv", skip = 1, header = FALSE)

# May 2023 - Sept 2023
df13 <- read.csv("../data/environmental/2023-04-14_to_2023-09-19_21152007_Reef13_3m.csv", header = T)
df14 <- read.csv("../data/environmental/2023-04-14_to_2023-09-19_21152007_Reef13_1m.csv", header = F, skip = 1)

# Jan 2024 - May 2024
df15 <- read.csv("../data/environmental/2024-05-08_Kbay_21152007.csv", header = T)
df16 <- read.csv("../data/environmental/2024-05-09_Kbay_21152005.csv", header = T)
df17 <- read.csv("../data/environmental/2024-05-09_Kbay_21152009.csv", header = T)

# October 2024 - Jan 2025
df18 <- read.csv("../data/environmental/Jan2025_KBay_21152005.csv", header = T)
df19 <- read.csv("../data/environmental/Jan2025_KBay_21152007.csv", header = T)
df20 <- read.csv("../data/environmental/Jan2025_KBay_21152009.csv", header = T)

```

Process data 
```{r}
## --- Group 1: like df1 (separate Date + Time) ---
df_group1 <- list(df1, df2, df3)

df_group1_clean <- map2(df_group1, paste0("df", 1:3), ~
  .x %>%
    mutate(
      DateTime = dmy_hms(paste(Date, Time..GMT.10.00)),
      Temp_C = Temp...C,
      Source = .y
    ) %>%
    select(DateTime, Temp_C, Source)
)

## --- Group 2: like df4 (combined DateTime, long header) ---
df_group2 <- list(df4, df5, df6)

df_group2_clean <- map2(df_group2, paste0("df", 4:6), ~
  .x %>%
    mutate(
      DateTime = mdy_hms(Date.Time..GMT.10.00),
      Temp_C = as.numeric(
        .x[[grep("Temp", names(.x), value = TRUE)[1]]]
      ),
      Source = .y
    ) %>%
    select(DateTime, Temp_C, Source)
)

## --- Group 3: like df7 (separate Date + Time again, but US format) ---
df_group3 <- list(df7, df8, df9)

df_group3_clean <- map2(df_group3, paste0("df", 7:9), ~
  .x %>%
    mutate(
      DateTime = mdy_hms(paste(Date, Time..GMT.10.00)),
      Temp_C = Temp...C,
      Source = .y
    ) %>%
    select(DateTime, Temp_C, Source)
)

## --- Group 4: df10–df12 (Date, Time, Temp_C) --- 
df_group4 <- list(df10, df11, df12)

df_group4_clean <- map2(df_group4, paste0("df", 10:12), ~
  .x %>%
    mutate(
      DateTime = dmy_hms(paste(V1, V2)),   # Date in dmy format (19-09-23)
      Temp_C   = as.numeric(V3),
      Source   = .y
    ) %>%
    select(DateTime, Temp_C, Source)
)

## --- Group 5: like df13 (separate Date + Time again, but US format) ---
df_group5 <- list(df13)

df_group5_clean <- map2(df_group5, paste0("df", 13), ~
  .x %>%
    mutate(
      DateTime = ymd_hms(paste(Date, Time..GMT.10.00)),
      Temp_C = Temp...C,
      Source = .y
    ) %>%
    select(DateTime, Temp_C, Source)
)

## --- Group 6: like df10 (Date, Time, Temp_C) --- 
df_group6 <- list(df14)

df_group6_clean <- map2(df_group6, paste0("df", 14), ~
  .x %>%
    mutate(
      DateTime = ymd_hms(paste(V1, V2)),   # Date in dmy format (19-09-23)
      Temp_C   = as.numeric(V3),
      Source   = .y
    ) %>%
    select(DateTime, Temp_C, Source)
)

## --- Group 7: df15, df16, df17 (DateTime, Temp_C) --- 
df_group7 <- list(df15, df16, df17)

df_group7_clean <- map2(df_group7, paste0("df", 15:17), ~
  .x %>%
    mutate(
      DateTime = dmy_hm(Date.Time..GMT.10.00),
      Temp_C = as.numeric(
        .x[[grep("Temp", names(.x), value = TRUE)[1]]]
      ),
      Source = .y
    ) %>%
    select(DateTime, Temp_C, Source)
)

## --- Group 8: df18, df19, df20 (DateTime, Temp_C) --- 
df_group8 <- list(df18, df19, df20)

df_group8_clean <- map2(df_group8, paste0("df", 18:20), ~
  .x %>%
    mutate(
      DateTime = mdy_hms(Date.Time..GMT.07.00),
      Temp_C = as.numeric(
        .x[[grep("Temp", names(.x), value = TRUE)[1]]]
      ),
      Source = .y
    ) %>%
    select(DateTime, Temp_C, Source)
)

## --- Combine all ---
all_data <- bind_rows(df_group1_clean, df_group2_clean, df_group3_clean, df_group4_clean, df_group5_clean, df_group6_clean, df_group7_clean, df_group8_clean)

head(all_data)

ggplot(all_data, aes(x = DateTime, y = Temp_C, color = Source)) +
  geom_point() +
  scale_x_datetime( date_breaks = "1 month", date_labels = "%b") + 

  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time")
```

Something weird is going on with the temperature data...some not parsing...find a better way to do this. 








Separate out year, month and day 
```{r}
all_data <- all_data %>%
  mutate(
    year = year(DateTime),
    month = month(DateTime),
    day = day(DateTime),
    hour = hour(DateTime))
```

Calculate DHW
```{r}
daily_mean_nsf <- all_data %>%
  group_by(year, month, day) %>%
  summarize(daily_temp_mean = mean(Temp_C, na.rm = T)) %>%
  ungroup() %>%
  arrange(year, month, day) %>%
  mutate(
    date = as.Date(paste(year, month, day, sep = "-")),
    HS = ifelse(daily_temp_mean >= threshold, daily_temp_mean - MMM, 0)
  )

# Check if dates are continuous:
date_seq <- seq(min(daily_mean_nsf$date), max(daily_mean_nsf$date), by = "day")
missing_dates <- setdiff(date_seq, daily_mean_nsf$date)

if(length(missing_dates) > 0) {
  # Fill missing dates with HS = 0, daily_temp_mean = NA or interpolated
  missing_df <- data.frame(
    year = as.integer(format(missing_dates, "%Y")),
    month = as.integer(format(missing_dates, "%m")),
    day = as.integer(format(missing_dates, "%d")),
    daily_temp_mean = NA,
    date = missing_dates,
    HS = 0
  )
  daily_mean_nsf <- bind_rows(daily_mean_nsf, missing_df) %>%
    arrange(date)
}

# Now compute rolling DHW (90-day window)
daily_mean_nsf <- daily_mean_nsf %>%
  mutate(DHW = rollapply(HS / 7, width = 90, sum, align = "right", fill = NA))
```

Plot! 
```{r}
# Define your MMM and bleaching thresholds
MMM <- 27.3
threshold <- MMM + 1 # 28.3

# Create a date column for plotting
daily_mean_nsf <- daily_mean_nsf %>%
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(DHW = ifelse(is.na(DHW), 0, DHW))

# scale constants (temp 20–30, DHW 0–10)
temp_min <- 20
temp_max <- 30
dhw_min  <- 0
dhw_max  <- 10

# scaling factor
scale_factor <- (temp_max - temp_min) / (dhw_max - dhw_min)

ggplot(daily_mean_nsf, aes(x = date)) +
  geom_line(aes(y = daily_temp_mean, color = "Temperature"), linewidth = 0.8) +
  geom_line(aes(y = DHW * scale_factor + temp_min, color = "DHW"), linewidth = 0.8, alpha = 0.8) +
  scale_y_continuous(
    name = "Daily Temperature (°C)",
    limits = c(temp_min, temp_max),
    sec.axis = sec_axis(
      ~ (. - temp_min) / scale_factor,
      name = "Degree Heating Weeks (°C-weeks)",
      breaks = seq(0, 10, 2)
    )
  ) +
  scale_color_manual(
    name = NULL,
    values = c("Temperature" = "steelblue", "DHW" = "firebrick")
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.title.y.left  = element_text(color = "steelblue"),
    axis.title.y.right = element_text(color = "firebrick"),
    legend.position = "top"
  )
```

Plot with the points from hourly data
```{r}
MMM <- 27.3
threshold <- MMM + 1

# Prepare daily_mean with date
daily_mean_nsf <- daily_mean_nsf %>% 
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(datetime = as.POSIXct(date, format = "%Y-%m-%d", tz = "UTC"))

# Prepare hourly temps with POSIXct datetime
all_data$year <- as.integer(all_data$year)
all_data$month <- as.integer(all_data$month)
all_data <- all_data %>%
  mutate(
    datetime = make_datetime(year = year, month = month, day = day, hour = hour, tz = "UTC"))

# Plot
ggplot() +
  # Hourly temperatures as light blue semi-transparent dots
  geom_point(data = all_data, aes(x = datetime, y = Temp_C), color = "grey60", alpha = 0.3, size = 0.5) +
  
  # Daily mean temperature as black line
  geom_line(data = daily_mean_nsf, aes(x = datetime, y = daily_temp_mean, color = "Temperature"), color = "black", linewidth = 0.8) +
  
  # DHW as gray line scaled to temperature y-axis limits (20-30)
  geom_line(data = daily_mean_nsf, aes(x = datetime, 
                                   y = DHW * (30 - 20) / (10 - 0) + 20), 
            color = "firebrick", size = 1, alpha = 0.7) +
  
  # Horizontal lines for MMM and Bleaching Threshold
  geom_hline(yintercept = MMM, linetype = "dashed", color = "black") +
  geom_hline(yintercept = threshold, linetype = "solid", color = "black") +
  
  # Primary y-axis limits and secondary axis that maps 20-30 temp to 0-10 DHW
  scale_y_continuous(
    name = "Temperature (°C)",
    limits = c(20, 30),
    sec.axis = sec_axis(
      trans = ~ (. - 20) * 10 / (30 - 20),
      name = "Degree Heating Weeks (°C wk⁻¹)",
      breaks = seq(0, 10, 1)
    )
  ) +
  theme_minimal()

```


### Combine 

Select and rename specific cols 
```{r}
temps_sub <- temps %>%
  select(Temperature, hour, year, day, month, datetime) %>%
  rename(Temperature = Temperature,  # already named correctly, just for clarity
         hour = hour,
         year = year,
         month = month,
         day = day,
         datetime = datetime)

all_data_sub <- all_data %>%
  select(Temp_C, year, month, day, hour, datetime) %>%
  rename(Temperature = Temp_C) %>%
  select(Temperature, hour, year, month, day, datetime)

# Reorder columns in temps_sub to match all_data_sub
temps_sub <- temps_sub %>%
  select(Temperature, hour, year, month, day, datetime)

# Bind
combined_data <- bind_rows(temps_sub, all_data_sub)
```

Calculate DHW
```{r}
daily_mean <- combined_data %>%
  group_by(year, month, day) %>%
  summarize(daily_temp_mean = mean(Temperature, na.rm = T)) %>%
  ungroup() %>%
  arrange(year, month, day) %>%
  mutate(
    date = as.Date(paste(year, month, day, sep = "-")),
    HS = ifelse(daily_temp_mean >= threshold, daily_temp_mean - MMM, 0)
  )

# Check if dates are continuous:
date_seq <- seq(min(daily_mean$date), max(daily_mean$date), by = "day")
missing_dates <- setdiff(date_seq, daily_mean$date)

if(length(missing_dates) > 0) {
  # Fill missing dates with HS = 0, daily_temp_mean = NA or interpolated
  missing_df <- data.frame(
    year = as.integer(format(missing_dates, "%Y")),
    month = as.integer(format(missing_dates, "%m")),
    day = as.integer(format(missing_dates, "%d")),
    daily_temp_mean = NA,
    date = missing_dates,
    HS = 0
  )
  daily_mean <- bind_rows(daily_mean, missing_df) %>%
    arrange(date)
}

# Now compute rolling DHW (90-day window)
daily_mean <- daily_mean %>%
  mutate(DHW = rollapply(HS / 7, width = 90, sum, align = "right", fill = NA))
```

Plot! 
```{r}
# Define your MMM and bleaching thresholds
MMM <- 27.3
threshold <- MMM + 1 # 28.3

# Create a date column for plotting
daily_mean <- daily_mean %>%
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(DHW = ifelse(is.na(DHW), 0, DHW))

# scale constants (temp 20–30, DHW 0–10)
temp_min <- 20
temp_max <- 30
dhw_min  <- 0
dhw_max  <- 10

# scaling factor
scale_factor <- (temp_max - temp_min) / (dhw_max - dhw_min)

ggplot(daily_mean, aes(x = date)) +
  geom_line(aes(y = daily_temp_mean, color = "Temperature"), linewidth = 0.8) +
  geom_line(aes(y = DHW * scale_factor + temp_min, color = "DHW"), linewidth = 0.8, alpha = 0.8) +
  scale_y_continuous(
    name = "Daily Temperature (°C)",
    limits = c(temp_min, temp_max),
    sec.axis = sec_axis(
      ~ (. - temp_min) / scale_factor,
      name = "Degree Heating Weeks (°C-weeks)",
      breaks = seq(0, 10, 2)
    )
  ) +
  scale_color_manual(
    name = NULL,
    values = c("Temperature" = "steelblue", "DHW" = "firebrick")
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.title.y.left  = element_text(color = "steelblue"),
    axis.title.y.right = element_text(color = "firebrick"),
    legend.position = "top"
  )
```

Plot with the points from temps
```{r}
MMM <- 27.3
threshold <- MMM + 1

# Prepare daily_mean with date
daily_mean <- daily_mean %>% 
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(datetime = as.POSIXct(date, format = "%Y-%m-%d", tz = "UTC"))

# Prepare hourly temps with POSIXct datetime
combined_data <- combined_data %>%
  mutate(
    datetime = make_datetime(year = year, month = month, day = day, hour = hour, min = 0, sec = 0, tz = "UTC"))

# Make df of sampling dates to mark on plot 
sample_dates <- data.frame(
  datetime = as.POSIXct(c("2023-09-14 00:00", "2024-01-20 00:00", "2024-05-09 00:00", "2024-10-01 00:00", "2025-01-09 00:00", "2025-05-08 00:00")),
  sample_id = c("Sept2023", "Jan2024", "May2024", "Oct2024", "Jan2025", "May2025"))

# Plot
whole_plot <- ggplot() +
  # Hourly temperatures as light semi-transparent dots
  geom_point(data = combined_data, aes(x = datetime, y = Temperature), color = "grey60", alpha = 0.3, size = 2) +
  
  # Daily mean temperature as black line
  geom_line(data = daily_mean, aes(x = datetime, y = daily_temp_mean, color = "Temperature"), color = "black", linewidth = 3) +
  
  # DHW as gray line scaled to temperature y-axis limits (20-30)
  geom_line(data = daily_mean, aes(x = datetime, 
                                   y = DHW * (30 - 20) / (10 - 0) + 20), 
            color = "firebrick", size = 2.5, alpha = 0.6) +
  
  # Horizontal lines for MMM and Bleaching Threshold
  geom_hline(yintercept = MMM, linetype = "dashed", color = "black", size = 2.5) +
  geom_hline(yintercept = threshold, linetype = "solid", color = "black", size = 2.5) +
  geom_point(data = sample_dates, aes(x = datetime, y = 29), shape = 23, fill = "skyblue", color = "black", size = 8) +   # add stars at y = 29°C on the temperature axis

  # Primary y-axis limits and secondary axis that maps 20-30 temp to 0-10 DHW
  scale_y_continuous(
    name = "Temperature (°C)",
    limits = c(20, 30),
    sec.axis = sec_axis(
      trans = ~ (. - 20) * 10 / (30 - 20),
      name = "Degree Heating Weeks (°C wk-1)",
      breaks = seq(0, 10, 2))) +
  scale_x_datetime(date_breaks = "1 year", date_labels = "%Y") +
  xlab("") +
  theme_classic() +
  theme(
    axis.title.y.left  = element_text(color = "black"),
    axis.title.y.right = element_text(color = "firebrick")) +
  theme(axis.ticks = element_line(size = 1, color = "black"),
        axis.ticks.length = unit(0.3, "cm"),
        axis.text.x = element_text(size = 24,
                                   color = "black"), 
        axis.text.y = element_text(size = 24,
                                   color = "black"),
        axis.title.y = element_text(size = 28,
                                    face = "bold"))

ggsave("../output/environmental/KBay_temps_DHW_2013-2025.png", plot = whole_plot, width=18, height=7)
ggsave("../output/environmental/KBay_temps_DHW_2013-2025.pdf", plot = whole_plot, width=18, height=7)
```

Plot with only the sampling data 
```{r}
MMM <- 27.3
threshold <- MMM + 1

# Prepare daily_mean with date
daily_mean_nsf <- daily_mean_nsf %>% 
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  filter(date >= as.Date("2022-09-02")) %>%
  mutate(datetime = as.POSIXct(date, format = "%Y-%m-%d", tz = "UTC"))

# Prepare hourly temps with POSIXct datetime
all_data$year <- as.integer(all_data$year)
all_data$month <- as.integer(all_data$month)
all_data <- all_data %>%
  mutate(
    datetime = make_datetime(year = year, month = month, day = day, hour = hour, tz = "UTC"))

# Plot
nsf_plot <- ggplot() +
  # Hourly temperatures as light blue semi-transparent dots
  geom_point(
    data = all_data,
    aes(x = datetime, y = Temp_C),
    color = "grey60", alpha = 0.3, size = 2) +
  # Daily mean temperature as black line
  geom_line(
    data = daily_mean_nsf,
    aes(x = datetime, y = daily_temp_mean, color = "Temperature"),
    color = "black", linewidth = 3) +
  # Horizontal lines for MMM and Bleaching Threshold
  geom_hline(yintercept = MMM, linetype = "dashed", color = "black", size = 2.5) +
  geom_hline(yintercept = threshold, linetype = "solid", color = "black", size = 2.5) +
  geom_point(data = sample_dates, aes(x = datetime, y = 29), shape = 23, fill = "skyblue", color = "black", size = 8) +   # add stars at y = 29°C on the temperature axis
  scale_x_datetime(date_breaks = "3 months", date_labels = "%b %Y") +
  xlab("") +
  scale_y_continuous(name = "Temperature (°C)", limits = c(20, 30)) +
  theme_classic() +
  theme(axis.ticks = element_line(size = 1, color = "black"),
        axis.ticks.length = unit(0.3, "cm"),
        axis.text.x = element_text(size = 24,
                                   color = "black",
                                   angle = 45,
                                   hjust = 1), 
        axis.text.y = element_text(size = 24,
                                   color = "black"),
        axis.title.y = element_text(size = 28,
                                    face = "bold"))

ggsave("../output/environmental/KBay_temps_2022-2025.png", plot = nsf_plot, width=18, height=7)
ggsave("../output/environmental/KBay_temps_2022-2025.pdf", plot = nsf_plot, width=18, height=7)
```

Combine both plots into one figure and save
```{r}
all_temp_plot <- nsf_plot + whole_plot + 
  plot_layout(ncol = 1) + 
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(size = 28, face = "bold"))

ggsave("../output/environmental/KBay_temps.pdf", all_temp_plot, width = 30, height = 18, units = "in")
ggsave("../output/environmental/KBay_temps.png", all_temp_plot, width = 30, height = 18, units = "in", dpi = 300)
```

Calculate the avg temp for the week and day prior to sampling dates 
```{r}
# make sure datetime columns are POSIXct
all_data <- all_data %>%
  mutate(datetime = ymd_hms(datetime))

sample_dates <- sample_dates %>%
  mutate(datetime = ymd(datetime))

# function to calculate mean + sem + n
calc_stats <- function(data, start_date, end_date, sample_id, window) {
  df <- data %>%
    filter(datetime >= start_date & datetime < end_date)
  
  if(nrow(df) > 0) {
    mean_val <- mean(df$Temp_C, na.rm = TRUE)
    sem_val  <- sd(df$Temp_C, na.rm = TRUE) / sqrt(sum(!is.na(df$Temp_C)))
    n_val    <- sum(!is.na(df$Temp_C))
  } else {
    mean_val <- NA
    sem_val  <- NA
    n_val    <- 0
  }
  
  tibble(
    sample_id = sample_id,
    window = window,
    mean_temp = mean_val,
    sem_temp = sem_val,
    n = n_val
  )
}

# loop over sample_dates and calculate stats for week, day, month
results <- purrr::map_dfr(1:nrow(sample_dates), function(i) {
  samp_date <- sample_dates$datetime[i]
  samp_id   <- sample_dates$sample_id[i]
  
  week_prior_start  <- samp_date - days(7)
  day_prior_start   <- samp_date - days(2)
  month_prior_start <- samp_date - days(30)
  
  bind_rows(
    calc_stats(all_data, week_prior_start, samp_date, samp_id, "week_prior"),
    calc_stats(all_data, day_prior_start, samp_date, samp_id, "day_prior"),
    calc_stats(all_data, month_prior_start, samp_date, samp_id, "month_prior")
  )
})

results

```



