---
title: "Temperature"
author: "Jill Ashey"
date: "2025-08-14"
output: html_document
---

Plot temperature from Patch Reef 13 in KBay

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(readr)
library(tidyverse)
library(lubridate)
library(naniar)
library(purrr)
library(zoo)
library(scales)
library(patchwork)
library(ggridges)
library(ggtext)
```

Read in data from 2013-2023 (data obtained from Brown et al. 2023 - https://www.bco-dmo.org/dataset/905047)
```{r}
temps <- read.csv("../data/environmental/905047_v1_infield_temp.csv", header = T)
```

Format date and time for data from 2013-2023
```{r}
# Create a date column for plotting
temps <- temps %>%
  mutate(date = as.Date(paste(year, month, day, sep = "-")))

# Plot
ggplot(temps, aes(x = date, y = Temperature, color = Site)) +
  geom_point() +
  labs(x = "Time (UTC)",y = "Temperature (°C)") +
  theme_minimal()
```

Calculate DHW
```{r}
MMM <- 27.3
threshold <- MMM + 1 # 28.3

daily_mean <- temps %>%
  group_by(year, month, day) %>%
  dplyr::summarize(daily_temp_mean = mean(Temperature, na.rm = T)) %>%
  ungroup() %>%
  dplyr::arrange(year, month, day) %>%
  mutate(
    date = as.Date(paste(year, month, day, sep = "-")),
    HS = ifelse(daily_temp_mean >= threshold, daily_temp_mean - MMM, 0)
  )

# Check if dates are continuous:
date_seq <- seq(min(daily_mean$date), max(daily_mean$date), by = "day")
missing_dates <- setdiff(date_seq, daily_mean$date)

if(length(missing_dates) > 0) {
  # Fill missing dates with HS = 0, daily_temp_mean = NA or interpolated
  missing_df <- data.frame(
    year = as.integer(format(missing_dates, "%Y")),
    month = as.integer(format(missing_dates, "%m")),
    day = as.integer(format(missing_dates, "%d")),
    daily_temp_mean = NA,
    date = missing_dates,
    HS = 0
  )
  daily_mean <- bind_rows(daily_mean, missing_df) %>%
    arrange(date)
}

# Now compute rolling DHW (90-day window)
daily_mean <- daily_mean %>%
  mutate(DHW = rollapply(HS / 7, width = 90, sum, align = "right", fill = NA))

```

Plot! 
```{r}
# Define your MMM and bleaching thresholds
MMM <- 27.3
bleaching_threshold <- MMM + 1 # 28.3

# Create a date column for plotting
daily_mean <- daily_mean %>%
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(DHW = ifelse(is.na(DHW), 0, DHW))

# scale constants (temp 20–30, DHW 0–10)
temp_min <- 20
temp_max <- 30
dhw_min  <- 0
dhw_max  <- 10

# scaling factor
scale_factor <- (temp_max - temp_min) / (dhw_max - dhw_min)

ggplot(daily_mean, aes(x = date)) +
  geom_line(aes(y = daily_temp_mean, color = "Temperature"), linewidth = 0.8) +
  geom_line(aes(y = DHW * scale_factor + temp_min, color = "DHW"), linewidth = 0.8, alpha = 0.8) +
  scale_y_continuous(
    name = "Daily Temperature (°C)",
    limits = c(temp_min, temp_max),
    sec.axis = sec_axis(
      ~ (. - temp_min) / scale_factor,
      name = "Degree Heating Weeks (°C-weeks)",
      breaks = seq(0, 10, 2)
    )
  ) +
  scale_color_manual(
    name = NULL,
    values = c("Temperature" = "steelblue", "DHW" = "firebrick")
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.title.y.left  = element_text(color = "steelblue"),
    axis.title.y.right = element_text(color = "firebrick"),
    legend.position = "top"
  )
```

Plot with the points from temps
```{r}
MMM <- 27.3
threshold <- MMM + 1

# Prepare daily_mean with date
daily_mean <- daily_mean %>% 
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(datetime = as.POSIXct(date, format = "%Y-%m-%d", tz = "UTC"))

# Prepare hourly temps with POSIXct datetime
temps <- temps %>%
  mutate(
    datetime = make_datetime(year = year, month = month, day = day, hour = hour, min = 0, sec = 0, tz = "UTC"))

# Plot
ggplot() +
  # Hourly temperatures as light blue semi-transparent dots
  geom_point(data = temps, aes(x = datetime, y = Temperature), color = "grey60", alpha = 0.3, size = 0.5) +
  
  # Daily mean temperature as black line
  geom_line(data = daily_mean, aes(x = datetime, y = daily_temp_mean, color = "Temperature"), color = "black", linewidth = 0.8) +
  
  # DHW as gray line scaled to temperature y-axis limits (20-30)
  geom_line(data = daily_mean, aes(x = datetime, 
                                   y = DHW * (30 - 20) / (10 - 0) + 20), 
            color = "firebrick", size = 1, alpha = 0.7) +
  
  # Horizontal lines for MMM and Bleaching Threshold
  geom_hline(yintercept = MMM, linetype = "dashed", color = "black") +
  geom_hline(yintercept = threshold, linetype = "solid", color = "black") +
  
  # Primary y-axis limits and secondary axis that maps 20-30 temp to 0-10 DHW
  scale_y_continuous(
    name = "Temperature (°C)",
    limits = c(20, 30),
    sec.axis = sec_axis(
      trans = ~ (. - 20) * 10 / (30 - 20),
      name = "Degree Heating Weeks (°C wk⁻¹)",
      breaks = seq(0, 10, 1)
    )
  ) +
  theme_minimal()

```




















Read in data from hobo loggers from NSF CAREER project. These are all in different formats so they need to be read in and processed differently. 
```{r}
# Sept 2022 - April 2023 
sept22_apr23_21152005 <- read.csv("../data/environmental/2022-Sept_2023-April_Reef13_1m_21152005.csv")
sept22_apr23_21152005 <- sept22_apr23_21152005 %>%
  mutate(DateTime = dmy_hms(paste(Date, Time..GMT.10.00)),
         Temp_C = Temp...C) %>%
  dplyr::select(DateTime, Temp_C)

sept22_apr23_21152007 <- read.csv("../data/environmental/2022-Sept_2023-April_Reef13_1m_21152007.csv")
sept22_apr23_21152007 <- sept22_apr23_21152007 %>%
  mutate(DateTime = dmy_hms(paste(Date, Time..GMT.10.00)),
         Temp_C = Temp...C) %>%
  dplyr::select(DateTime, Temp_C)

sept22_apr23_21152009 <- read.csv("../data/environmental/2022-Sept_2023-April_Reef13_3m_21152009.csv")
sept22_apr23_21152009 <- sept22_apr23_21152009 %>%
  mutate(DateTime = dmy_hms(paste(Date, Time..GMT.10.00)),
         Temp_C = Temp...C) %>%
  dplyr::select(DateTime, Temp_C)

sept22_apr23 <- rbind(sept22_apr23_21152005, sept22_apr23_21152007, sept22_apr23_21152009)
ggplot(sept22_apr23, aes(x = DateTime, y = Temp_C)) +
  geom_point() +
  scale_x_datetime( date_breaks = "1 month", date_labels = "%b") + 
  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time")

# Apr 2023 - Sept 2023
apr23_sept_23_21152007_3m <- read.csv("../data/environmental/2023-04-14_to_2023-09-19_21152007_Reef13_3m.csv", header = T)
apr23_sept_23_21152007_3m <- apr23_sept_23_21152007_3m %>%
  mutate(DateTime = ymd_hms(paste(Date, Time..GMT.10.00)),
         Temp_C = Temp...C) %>%
  dplyr::select(DateTime, Temp_C)

apr23_sept_23_21152007_1m <- read.csv("../data/environmental/2023-04-14_to_2023-09-19_21152007_Reef13_1m.csv", header = F, skip = 1)
apr23_sept_23_21152007_1m <- apr23_sept_23_21152007_1m %>%
  mutate(DateTime = ymd_hms(paste(V1, V2)),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

apr23_sept23 <- rbind(apr23_sept_23_21152007_3m, apr23_sept_23_21152007_1m)
ggplot(apr23_sept23, aes(x = DateTime, y = Temp_C)) +
  geom_point() +
  scale_x_datetime( date_breaks = "1 month", date_labels = "%b") + 
  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time")

# Sept 2023 - Jan 2024
sept23_jan24_21152005 <- read.csv("../data/environmental/Sept2023_Jan2024_Reef13_insitu_temp_21152005_3m.csv", header = F, skip = 1) %>%
  na.omit()
sept23_jan24_21152005 <- sept23_jan24_21152005 %>%
  mutate(DateTime = dmy_hms(paste(V1, V2)),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

sept23_jan24_21152007 <- read.csv("../data/environmental/Sept2023_Jan2024_Reef13_insitu_temp_21152007_1m.csv", header = F, skip = 1) %>%
  na.omit()
sept23_jan24_21152007 <- sept23_jan24_21152007 %>%
  mutate(DateTime = dmy_hms(paste(V1, V2)),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

sept23_jan24_21152009 <- read.csv("../data/environmental/Sept2023_Jan2024_Reef13_insitu_temp_21152009_1m.csv", header = F, skip = 1) %>%
  na.omit()
sept23_jan24_21152009 <- sept23_jan24_21152009 %>%
  mutate(DateTime = dmy_hms(paste(V1, V2)),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

sept23_jan24 <- rbind(sept23_jan24_21152005, sept23_jan24_21152007, sept23_jan24_21152009)
ggplot(sept23_jan24, aes(x = DateTime, y = Temp_C)) +
  geom_point() +
  scale_x_datetime( date_breaks = "1 month", date_labels = "%b") + 
  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time")

# Jan 2024 - May 2024
jan24_may24_21152005 <- read.csv("../data/environmental/2024-05-09_Kbay_21152005.csv", header = F, skip = 1) %>%
  na.omit()
jan24_may24_21152005 <- jan24_may24_21152005 %>%
  mutate(DateTime = mdy_hm(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

jan24_may24_21152007 <- read.csv("../data/environmental/2024-05-08_Kbay_21152007.csv", header = F, skip = 1) %>%
  na.omit()
jan24_may24_21152007 <- jan24_may24_21152007 %>%
  mutate(DateTime = mdy_hm(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

jan24_may24_21152009 <- read.csv("../data/environmental/2024-05-09_Kbay_21152009.csv", header = F, skip = 1) %>%
  na.omit()
jan24_may24_21152009 <- jan24_may24_21152009 %>%
  mutate(DateTime = mdy_hm(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

jan24_may24 <- rbind(jan24_may24_21152005, jan24_may24_21152007, jan24_may24_21152009)
ggplot(jan24_may24, aes(x = DateTime, y = Temp_C)) +
  geom_point() +
  scale_x_datetime( date_breaks = "1 month", date_labels = "%b") + 
  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time")

# May 2024 - Sept 2024
may24_sept24_21152005 <- read.csv("../data/environmental/2024-09-30_Reef13_21152005.csv", skip = 2, header = F)
may24_sept24_21152005 <- may24_sept24_21152005 %>%
  mutate(DateTime = mdy_hms(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

may24_sept24_21152007 <- read.csv("../data/environmental/2024-09-30_Reef13_21152007.csv", skip = 2, header = F)
may24_sept24_21152007 <- may24_sept24_21152007 %>%
  mutate(DateTime = mdy_hms(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

may24_sept24_21152009 <- read.csv("../data/environmental/2024-09-30_Reef13_21152009.csv", skip = 2, header = F)
may24_sept24_21152009 <- may24_sept24_21152009 %>%
  mutate(DateTime = mdy_hms(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

may24_sept24 <- rbind(may24_sept24_21152005, may24_sept24_21152007, may24_sept24_21152009)
ggplot(may24_sept24, aes(x = DateTime, y = Temp_C)) +
  geom_point() +
  scale_x_datetime( date_breaks = "1 month", date_labels = "%b") + 
  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time")

# Oct 2024 - Jan 2025
oct24_jan25_21152005 <- read.csv("../data/environmental/Jan2025_KBay_21152005.csv", header = F, skip = 1)
oct24_jan25_21152005 <- oct24_jan25_21152005 %>%
  mutate(DateTime = mdy_hm(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

oct24_jan25_21152007 <- read.csv("../data/environmental/Jan2025_KBay_21152007.csv", header = F, skip = 1)
oct24_jan25_21152007 <- oct24_jan25_21152007 %>%
  mutate(DateTime = mdy_hm(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

oct24_jan25_21152009 <- read.csv("../data/environmental/Jan2025_KBay_21152009.csv", header = F, skip = 1)
oct24_jan25_21152009 <- oct24_jan25_21152009 %>%
  mutate(DateTime = mdy_hms(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

oct24_jan25 <- rbind(oct24_jan25_21152005, oct24_jan25_21152007, oct24_jan25_21152009)
ggplot(oct24_jan25, aes(x = DateTime, y = Temp_C)) +
  geom_point() +
  scale_x_datetime( date_breaks = "1 month", date_labels = "%b") + 
  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time")

# Jan 2025 - May 2025
jan25_may25_21152005 <- read.csv("../data/environmental/21152005.csv", header = T)
jan25_may25_21152005 <- jan25_may25_21152005 %>%
  mutate(DateTime = mdy_hms(paste(Date, Time..GMT.10.00)),
         Temp_C = Temp...C) %>%
  dplyr::select(DateTime, Temp_C)

jan25_may25_21152007 <- read.csv("../data/environmental/21152007.csv", header = T)
jan25_may25_21152007 <- jan25_may25_21152007 %>%
  mutate(DateTime = mdy_hms(paste(Date, Time..GMT.10.00)),
         Temp_C = Temp...C) %>%
  dplyr::select(DateTime, Temp_C)

jan25_may25_21152009 <- read.csv("../data/environmental/21152009.csv", header = T)
jan25_may25_21152009 <- jan25_may25_21152009 %>%
  mutate(DateTime = mdy_hms(paste(Date, Time..GMT.10.00)),
         Temp_C = Temp...C) %>%
  dplyr::select(DateTime, Temp_C)

jan25_may25 <- rbind(jan25_may25_21152005, jan25_may25_21152007, jan25_may25_21152009)
ggplot(oct24_jan25, aes(x = DateTime, y = Temp_C)) +
  geom_point() +
  scale_x_datetime( date_breaks = "1 month", date_labels = "%b") + 
  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time")

# May 2025 - Oct 2025 - lost one of the U22 loggers on the reef (21152005)
may25_oct25_21152007 <- read.csv("../data/environmental/Oct2025_21152007.csv", skip = 2, header = F)
may25_oct25_21152007 <- may25_oct25_21152007 %>%
  mutate(DateTime = mdy_hms(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

may25_oct25_21152009 <- read.csv("../data/environmental/Oct2025_21152009.csv", skip = 2, header = F)
may25_oct25_21152009 <- may25_oct25_21152009 %>%
  mutate(DateTime = mdy_hms(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

may25_oct25 <- rbind(may25_oct25_21152007, may25_oct25_21152009)
may25_oct25 <- may25_oct25 %>% # remove times when logger was out of water 
  mutate(DateTime = ymd_hms(DateTime)) %>%
  filter(DateTime >= ymd("2025-05-13"),
         DateTime <= ymd_hms("2025-10-01 08:30:00"))
ggplot(may25_oct25, aes(x = DateTime, y = Temp_C)) +
  geom_point() +
  scale_x_datetime( date_breaks = "1 month", date_labels = "%b") + 
  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time")

# Oct 2025 - Jan 2026
oct25_jan26_21152007 <- read.csv("../data/environmental/Jan2026_21152007.csv", skip = 2, header = F)
oct25_jan26_21152007 <- oct25_jan26_21152007 %>%
  mutate(DateTime = mdy_hms(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

oct25_jan26_21152009 <- read.csv("../data/environmental/Jan2026_21152009.csv", skip = 2, header = F)
oct25_jan26_21152009 <- oct25_jan26_21152009 %>%
  mutate(DateTime = mdy_hms(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

oct25_jan26_20947172 <- read.csv("../data/environmental/Jan2026_20947172.csv", skip = 2, header = F)
oct25_jan26_20947172 <- oct25_jan26_20947172 %>%
  mutate(DateTime = mdy_hms(V2),
         Temp_C = V3) %>%
  dplyr::select(DateTime, Temp_C)

oct25_jan26 <- rbind(oct25_jan26_21152007, oct25_jan26_21152009, oct25_jan26_20947172)
oct25_jan26 <- oct25_jan26 %>% # remove times when logger was out of water 
  mutate(DateTime = ymd_hms(DateTime)) %>%
  filter(DateTime >= ymd("2025-10-14"),
         DateTime <= ymd_hms("2026-01-20 08:30:00"))
ggplot(oct25_jan26, aes(x = DateTime, y = Temp_C)) +
  geom_point() +
  scale_x_datetime( date_breaks = "1 month", date_labels = "%b") + 
  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time")
```

Bind all data together, save and plot 
```{r}
all_data <- rbind(sept22_apr23, apr23_sept23, sept23_jan24, jan24_may24, may24_sept24, oct24_jan25, jan25_may25, may25_oct25, oct25_jan26)

# Write csv
write.csv(all_data, "../output/environmental/NSF_CAREER_KBay_PR13_Temps.csv")

ggplot(all_data, aes(x = DateTime, y = Temp_C)) +
  geom_point() +
  scale_x_datetime( date_breaks = "1 month", date_labels = "%b") + 
  theme_minimal() +
  labs(y = "Temperature (°C)", x = "Time")
```

Separate out year, month and day 
```{r}
all_data <- all_data %>%
  mutate(
    year = year(DateTime),
    month = month(DateTime),
    day = day(DateTime),
    hour = hour(DateTime))
```

Calculate DHW
```{r}
daily_mean_nsf <- all_data %>%
  dplyr::group_by(year, month, day) %>%
  dplyr::summarize(daily_temp_mean = mean(Temp_C, na.rm = TRUE), .groups = "drop") %>% 
  dplyr::arrange(year, month, day) %>%
  dplyr::mutate(
    date = as.Date(paste(year, month, day, sep = "-")),
    HS = ifelse(daily_temp_mean >= threshold, daily_temp_mean - MMM, 0)
  )

# Check if dates are continuous:
date_seq <- seq(min(daily_mean_nsf$date), max(daily_mean_nsf$date), by = "day")
missing_dates <- setdiff(date_seq, daily_mean_nsf$date)

if(length(missing_dates) > 0) {
  # Fill missing dates with HS = 0, daily_temp_mean = NA or interpolated
  missing_df <- data.frame(
    year = as.integer(format(missing_dates, "%Y")),
    month = as.integer(format(missing_dates, "%m")),
    day = as.integer(format(missing_dates, "%d")),
    daily_temp_mean = NA,
    date = missing_dates,
    HS = 0
  )
  daily_mean_nsf <- bind_rows(daily_mean_nsf, missing_df) %>%
    arrange(date)
}

# Now compute rolling DHW (90-day window)
daily_mean_nsf <- daily_mean_nsf %>%
  mutate(DHW = rollapply(HS / 7, width = 90, sum, align = "right", fill = NA))
```

Plot! 
```{r}
# Define your MMM and bleaching thresholds
MMM <- 27.3
threshold <- MMM + 1 # 28.3

# Create a date column for plotting
daily_mean_nsf <- daily_mean_nsf %>%
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(DHW = ifelse(is.na(DHW), 0, DHW))

# scale constants (temp 20–30, DHW 0–10)
temp_min <- 20
temp_max <- 30
dhw_min  <- 0
dhw_max  <- 10

# scaling factor
scale_factor <- (temp_max - temp_min) / (dhw_max - dhw_min)

ggplot(daily_mean_nsf, aes(x = date)) +
  geom_line(aes(y = daily_temp_mean, color = "Temperature"), linewidth = 0.8) +
  geom_line(aes(y = DHW * scale_factor + temp_min, color = "DHW"), linewidth = 0.8, alpha = 0.8) +
  scale_y_continuous(
    name = "Daily Temperature (°C)",
    limits = c(temp_min, temp_max),
    sec.axis = sec_axis(
      ~ (. - temp_min) / scale_factor,
      name = "Degree Heating Weeks (°C-weeks)",
      breaks = seq(0, 10, 2)
    )
  ) +
  scale_color_manual(
    name = NULL,
    values = c("Temperature" = "steelblue", "DHW" = "firebrick")
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.title.y.left  = element_text(color = "steelblue"),
    axis.title.y.right = element_text(color = "firebrick"),
    legend.position = "top"
  )
```

Plot with the points from hourly data
```{r}
MMM <- 27.3
threshold <- MMM + 1

# Prepare daily_mean with date
daily_mean_nsf <- daily_mean_nsf %>% 
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(datetime = as.POSIXct(date, format = "%Y-%m-%d", tz = "UTC"))

# Prepare hourly temps with POSIXct datetime
all_data$year <- as.integer(all_data$year)
all_data$month <- as.integer(all_data$month)
all_data <- all_data %>%
  mutate(
    datetime = make_datetime(year = year, month = month, day = day, hour = hour, tz = "UTC"))

# Plot
ggplot() +
  # Hourly temperatures as light blue semi-transparent dots
  geom_point(data = all_data, aes(x = datetime, y = Temp_C), color = "grey60", alpha = 0.3, size = 0.5) +
  
  # Daily mean temperature as black line
  geom_line(data = daily_mean_nsf, aes(x = datetime, y = daily_temp_mean, color = "Temperature"), color = "black", linewidth = 0.8) +
  
  # DHW as gray line scaled to temperature y-axis limits (20-30)
  geom_line(data = daily_mean_nsf, aes(x = datetime, 
                                   y = DHW * (30 - 20) / (10 - 0) + 20), 
            color = "firebrick", size = 1, alpha = 0.7) +
  
  # Horizontal lines for MMM and Bleaching Threshold
  geom_hline(yintercept = MMM, linetype = "dashed", color = "black") +
  geom_hline(yintercept = threshold, linetype = "solid", color = "black") +
  
  # Primary y-axis limits and secondary axis that maps 20-30 temp to 0-10 DHW
  scale_y_continuous(
    name = "Temperature (°C)",
    limits = c(20, 30),
    sec.axis = sec_axis(
      trans = ~ (. - 20) * 10 / (30 - 20),
      name = "Degree Heating Weeks (°C/week)",
      breaks = seq(0, 10, 1)
    )
  ) +
  theme_minimal()
```


### Combine 

Select and rename specific cols 
```{r}
temps_sub <- temps %>%
  dplyr::select(Temperature, hour, year, day, month, datetime) %>%
  dplyr::rename(Temperature = Temperature,  # already named correctly, just for clarity
         hour = hour,
         year = year,
         month = month,
         day = day,
         datetime = datetime)

all_data_sub <- all_data %>%
  dplyr::rename(Temperature = Temp_C) %>%
  dplyr::select(Temperature, hour, year, month, day, datetime)

# Reorder columns in temps_sub to match all_data_sub
temps_sub <- temps_sub %>%
  dplyr::select(Temperature, hour, year, month, day, datetime)

# Bind
combined_data <- bind_rows(temps_sub, all_data_sub)
```

Calculate DHW
```{r}
daily_mean <- combined_data %>%
  group_by(year, month, day) %>%
  dplyr::summarize(daily_temp_mean = mean(Temperature, na.rm = T)) %>%
  ungroup() %>%
  dplyr::arrange(year, month, day) %>%
  mutate(
    date = as.Date(paste(year, month, day, sep = "-")),
    HS = ifelse(daily_temp_mean >= threshold, daily_temp_mean - MMM, 0)
  )

# Check if dates are continuous:
date_seq <- seq(min(daily_mean$date), max(daily_mean$date), by = "day")
missing_dates <- setdiff(date_seq, daily_mean$date)

if(length(missing_dates) > 0) {
  # Fill missing dates with HS = 0, daily_temp_mean = NA or interpolated
  missing_df <- data.frame(
    year = as.integer(format(missing_dates, "%Y")),
    month = as.integer(format(missing_dates, "%m")),
    day = as.integer(format(missing_dates, "%d")),
    daily_temp_mean = NA,
    date = missing_dates,
    HS = 0
  )
  daily_mean <- bind_rows(daily_mean, missing_df) %>%
    arrange(date)
}

# Now compute rolling DHW (90-day window)
daily_mean <- daily_mean %>%
  mutate(DHW = rollapply(HS / 7, width = 90, sum, align = "right", fill = NA))
```

Plot! 
```{r}
# Define your MMM and bleaching thresholds
MMM <- 27.3
threshold <- MMM + 1 # 28.3

# Create a date column for plotting
daily_mean <- daily_mean %>%
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(DHW = ifelse(is.na(DHW), 0, DHW))

# scale constants (temp 20–30, DHW 0–10)
temp_min <- 20
temp_max <- 30
dhw_min  <- 0
dhw_max  <- 10

# scaling factor
scale_factor <- (temp_max - temp_min) / (dhw_max - dhw_min)

ggplot(daily_mean, aes(x = date)) +
  geom_line(aes(y = daily_temp_mean, color = "Temperature"), linewidth = 0.8) +
  geom_line(aes(y = DHW * scale_factor + temp_min, color = "DHW"), linewidth = 0.8, alpha = 0.8) +
  scale_y_continuous(
    name = "Daily Temperature (°C)",
    limits = c(temp_min, temp_max),
    sec.axis = sec_axis(
      ~ (. - temp_min) / scale_factor,
      name = "Degree Heating Weeks (°C-weeks)",
      breaks = seq(0, 10, 2)
    )
  ) +
  scale_color_manual(
    name = NULL,
    values = c("Temperature" = "steelblue", "DHW" = "firebrick")
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.title.y.left  = element_text(color = "steelblue"),
    axis.title.y.right = element_text(color = "firebrick"),
    legend.position = "top"
  )
```

Plot with the points from temps
```{r}
MMM <- 27.3
threshold <- MMM + 1

# Prepare daily_mean with date
daily_mean <- daily_mean %>% 
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(datetime = as.POSIXct(date, format = "%Y-%m-%d", tz = "UTC"))

# Prepare hourly temps with POSIXct datetime
combined_data <- combined_data %>%
  mutate(
    datetime = make_datetime(year = year, month = month, day = day, hour = hour, min = 0, sec = 0, tz = "UTC"))

# Make df of sampling dates to mark on plot 
sample_dates <- data.frame(
  datetime = as.POSIXct(c("2023-09-14 00:00", "2025-10-01 00:00")),
  sample_id = c("Sept2023", "Oct2025"))

# Plot
whole_plot <- ggplot() +
  # Hourly temperatures as light semi-transparent dots
  geom_point(data = combined_data, aes(x = datetime, y = Temperature), color = "grey60", alpha = 0.3, size = 2) +
  
  # Daily mean temperature as black line
  geom_line(data = daily_mean, aes(x = datetime, y = daily_temp_mean, color = "Temperature"), color = "black", linewidth = 3) +
  
  # DHW as gray line scaled to temperature y-axis limits (20-30)
  geom_line(data = daily_mean, aes(x = datetime, 
                                   y = DHW * (30 - 20) / (10 - 0) + 20), 
            color = "firebrick", size = 2.5, alpha = 0.6) +
  
  # Horizontal lines for MMM and Bleaching Threshold
  geom_hline(yintercept = MMM, linetype = "dashed", color = "black", size = 2.5) +
  geom_hline(yintercept = threshold, linetype = "solid", color = "black", size = 2.5) +
  geom_point(data = sample_dates, aes(x = datetime, y = 29), shape = 23, fill = "blue", color = "black", size = 8) +   # add stars at y = 29°C on the temperature axis

  # Primary y-axis limits and secondary axis that maps 20-30 temp to 0-10 DHW
  scale_y_continuous(
    name = "Temperature (°C)",
    limits = c(20, 30),
    sec.axis = sec_axis(
      trans = ~ (. - 20) * 10 / (30 - 20),
      name = "Degree Heating Weeks (°C wk-1)",
      breaks = seq(0, 10, 2))) +
  scale_x_datetime(date_breaks = "1 year", date_labels = "%Y") +
  xlab("") +
  theme_classic() +
  theme(
    axis.title.y.left  = element_text(color = "black"),
    axis.title.y.right = element_text(color = "firebrick")) +
  theme(axis.ticks = element_line(size = 1, color = "black"),
        axis.ticks.length = unit(0.3, "cm"),
        axis.text.x = element_text(size = 24,
                                   color = "black"), 
        axis.text.y = element_text(size = 24,
                                   color = "black"),
        axis.title.y = element_text(size = 28,
                                    face = "bold"))

ggsave("../output/environmental/KBay_temps_DHW_2013-2026.png", plot = whole_plot, width=18, height=7)
ggsave("../output/environmental/KBay_temps_DHW_2013-2026.pdf", plot = whole_plot, width=18, height=7)
```

Plot with only the sampling data 
```{r}
MMM <- 27.3
threshold <- MMM + 1

# Prepare daily_mean with date
daily_mean_nsf <- daily_mean_nsf %>% 
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  mutate(datetime = as.POSIXct(date, format = "%Y-%m-%d", tz = "UTC"))

# Prepare hourly temps with POSIXct datetime
all_data$year <- as.integer(all_data$year)
all_data$month <- as.integer(all_data$month)
all_data <- all_data %>%
  mutate(
    datetime = make_datetime(year = year, month = month, day = day, hour = hour, tz = "UTC"))

# Plot
nsf_plot <- ggplot() +
  # Hourly temperatures as light blue semi-transparent dots
  geom_point(data = all_data, aes(x = datetime, y = Temp_C), color = "grey60", alpha = 0.3, size = 2) +
  # Daily mean temperature as black line
  geom_line(data = daily_mean_nsf, aes(x = datetime, y = daily_temp_mean, color = "Temperature"), color = "black", linewidth = 3) +
  
  # DHW as gray line scaled to temperature y-axis limits (20-30)
  geom_line(data = daily_mean_nsf, aes(x = datetime, 
                                   y = DHW * (30 - 20) / (10 - 0) + 20), 
            color = "firebrick", size = 2.5, alpha = 0.6) +
  
  # Horizontal lines for MMM and Bleaching Threshold
  geom_hline(yintercept = MMM, linetype = "dashed", color = "black", size = 2.5) +
  geom_hline(yintercept = threshold, linetype = "solid", color = "black", size = 2.5) +
    geom_point(data = sample_dates, aes(x = datetime, y = 29), shape = 23, fill = "skyblue", color = "black", size = 8) +   # add stars at y = 29°C on the temperature axis

  # Primary y-axis limits and secondary axis that maps 20-30 temp to 0-10 DHW
  scale_y_continuous(
    name = "Temperature (°C)",
    limits = c(20, 30),
    sec.axis = sec_axis(
      trans = ~ (. - 20) * 10 / (30 - 20),
      name = "Degree Heating Weeks (°C wk-1)",
      breaks = seq(0, 10, 2))) +
  scale_x_datetime(date_breaks = "3 months", date_labels = "%b %Y") +
  xlab("") +
  theme_classic() +
  theme(
    axis.title.y.left  = element_text(color = "black"),
    axis.title.y.right = element_text(color = "firebrick")) +
  theme(axis.ticks = element_line(size = 1, color = "black"),
        axis.ticks.length = unit(0.3, "cm"),
        axis.text.x = element_text(size = 24,
                                   color = "black",
                                   angle = 45,
                                   hjust = 1),
        axis.text.y = element_text(size = 24,
                                   color = "black"),
        axis.title.y = element_text(size = 28,
                                    face = "bold"))

ggsave("../output/environmental/KBay_temps_2022-2026.png", plot = nsf_plot, width=18, height=7)
ggsave("../output/environmental/KBay_temps_2022-2026.pdf", plot = nsf_plot, width=18, height=7)
```

Combine both plots into one figure and save
```{r}
all_temp_plot <- nsf_plot + whole_plot + 
  plot_layout(ncol = 1) + 
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(size = 28, face = "bold"))

ggsave("../output/environmental/KBay_temps.pdf", all_temp_plot, width = 30, height = 18, units = "in")
ggsave("../output/environmental/KBay_temps.png", all_temp_plot, width = 30, height = 18, units = "in", dpi = 300)
```

Calculate the avg temp for the week and day prior to sampling dates 
```{r}
# make sure datetime columns are POSIXct
all_data <- all_data %>%
  mutate(datetime = ymd_hms(datetime))

sample_dates <- sample_dates %>%
  mutate(datetime = ymd(datetime))

# function to calculate mean + sem + n
calc_stats <- function(data, start_date, end_date, sample_id, window) {
  df <- data %>%
    filter(datetime >= start_date & datetime < end_date)
  
  if(nrow(df) > 0) {
    mean_val <- mean(df$Temp_C, na.rm = TRUE)
    sem_val  <- sd(df$Temp_C, na.rm = TRUE) / sqrt(sum(!is.na(df$Temp_C)))
    n_val    <- sum(!is.na(df$Temp_C))
  } else {
    mean_val <- NA
    sem_val  <- NA
    n_val    <- 0
  }
  
  tibble(
    sample_id = sample_id,
    window = window,
    mean_temp = mean_val,
    sem_temp = sem_val,
    n = n_val
  )
}

# loop over sample_dates and calculate stats for week, day, month
results <- purrr::map_dfr(1:nrow(sample_dates), function(i) {
  samp_date <- sample_dates$datetime[i]
  samp_id   <- sample_dates$sample_id[i]
  
  week_prior_start  <- samp_date - days(7)
  day_prior_start   <- samp_date - days(2)
  month_prior_start <- samp_date - days(30)
  
  bind_rows(
    calc_stats(all_data, week_prior_start, samp_date, samp_id, "week_prior"),
    calc_stats(all_data, day_prior_start, samp_date, samp_id, "day_prior"),
    calc_stats(all_data, month_prior_start, samp_date, samp_id, "month_prior")
  )
})

results

```


## TPC env plotting 

Look at the six months prior to TPC sampling specifically 
```{r}
# 2023
temps_2023 <- all_data %>%
  filter(DateTime <= "2023-09-23 01:00:00") %>%
  filter(DateTime >= "2023-03-23 01:00:00")
temps_2023_high <- temps_2023 %>%
  dplyr::filter(Temp_C > 27.3)
hours_above_28_2023 <- temps_2023 %>%
  filter(Temp_C > 27.3) %>%
  dplyr::summarise(hours = n() * 0.5) %>%
  pull(hours)

## How many hrs were above 28 in the month before collection?
collection_time <- as.POSIXct("2023-09-23 00:00:00", tz = "UTC")
start_window <- collection_time %m-% months(2)  # 2 month before
end_window   <- collection_time                 # up to collection
temps_month_before <- temps_2023_high %>%
  filter(DateTime >= start_window,
         DateTime <  end_window)
hours_in_month_before <- nrow(temps_month_before) * 0.5
hours_in_month_before

# 2025
temps_2025 <- all_data %>%
  filter(DateTime <= "2025-10-01 01:00:00") %>%
  filter(DateTime >= "2025-04-01 01:00:00")
temps_2025_high <- temps_2025 %>%
  dplyr::filter(Temp_C > 27.3)
hours_above_28_2025 <- temps_2025 %>%
  filter(Temp_C > 27.3) %>%
  dplyr::summarise(hours = n() * 0.5) %>%
  pull(hours)

## How many hrs were above 28 in the month before collection?
collection_time <- as.POSIXct("2025-10-01 00:00:00", tz = "UTC")
start_window <- collection_time %m-% months(2)  # 2 month before
end_window   <- collection_time                 # up to collection
temps_month_before <- temps_2025_high %>%
  filter(DateTime >= start_window,
         DateTime <  end_window)
hours_in_month_before <- nrow(temps_month_before) * 0.5
hours_in_month_before
```

A: Compare 2019, 2023, and 2025 on same plot -- different colored lines. Add MMM as 27.3 and MMM+1 as horizontal lines. B: add background shading for MMM









Plot 2023 and 2025
```{r}
# 1. Setup Collection Dates and Combine Data
temps_2023$Collection_Date <- as.POSIXct("2023-09-23")
temps_2025$Collection_Date <- as.POSIXct("2025-10-01")

all_temps <- bind_rows(
  temps_2023 %>% mutate(Sampling_Year = "2023"),
  temps_2025 %>% mutate(Sampling_Year = "2025")
) %>%
  mutate(
    DateTime = as.POSIXct(DateTime),
    Days_Before = as.numeric(difftime(DateTime, Collection_Date, units = "days"))
  ) %>%
  filter(Days_Before >= -365 & Days_Before <= 0)

# Define New Colors (Teal and Gold)
env_colors <- c("2023" = "#008080", "2025" = "#D4AF37")

# PANEL A 
p_time <- ggplot(all_temps, aes(x = Days_Before, y = Temp_C, color = Sampling_Year)) +
   geom_hline(yintercept = 28, linetype = "dashed", color = "grey40", alpha = 0.5) +
   geom_line(alpha = 0.3, linewidth = 0.3) +
   geom_smooth(method = "gam", linewidth = 1.4, se = FALSE) +
   scale_color_manual(values = env_colors) +
   
   # ADD na.rm = TRUE here to handle missing data
   scale_y_continuous(breaks = seq(floor(min(all_temps$Temp_C, na.rm = TRUE)), 
                                   ceiling(max(all_temps$Temp_C, na.rm = TRUE)), by = 2)) +
   
   theme_classic() +
   theme(
     axis.text = element_text(size = 15, color = "black"),
     axis.title = element_text(size = 16, face = "plain"),
     legend.position = "none" 
   ) +
   labs(x = "Days Before Collection", y = "Temperature (°C)")

# PANEL B: Frequency as Percentage (Right)
freq_data <- all_temps %>%
  mutate(temp_bin = round(Temp_C)) %>%
  group_by(Sampling_Year, temp_bin) %>%
  dplyr::summarise(count = n(), .groups = "drop_last") %>%
  mutate(percent = (count / sum(count)) * 100)

p_freq <- ggplot(freq_data, aes(x = temp_bin, y = percent, fill = Sampling_Year)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.85, color = "white") +
  scale_fill_manual(values = env_colors) +
  scale_x_continuous(breaks = seq(22, 32, 1)) +
  theme_classic() +
  theme(
    axis.text = element_text(size = 15, color = "black"),
    axis.title = element_text(size = 16, face = "plain"),
    legend.position = "right", 
    legend.background = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  labs(x = "Temperature Bin (°C)", 
       y = "Percentage of Total Time (%)", 
       fill = "Year")

# 3. Assemble Side-by-Side and Label with Patchwork
# plot_annotation(tag_levels = 'A') automatically labels the plots
final_env_plot <- (p_time + p_freq) + 
  plot_layout(widths = c(1.3, 1)) + 
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(face = "bold", size = 14))

# Save 
ggsave("../output/environmental/Temperatures_2023_and_2025.pdf", 
       plot = final_env_plot, 
       width = 18, 
       height = 6, 
       units = "in", 
       device = "pdf",
       dpi = 300)
ggsave("../output/environmental/Temperatures_2023_and_2025.png", 
       plot = final_env_plot, 
       width = 18, 
       height = 6, 
       units = "in", 
       dpi = 300)
```

## Plotting for PI TPC paper 

Subset by March-October for 2019, 2021, 2023, and 2025 
```{r}
combined_data$datetime <- as.POSIXct(combined_data$datetime)

subset_data <- combined_data %>%
  filter(year %in% c(2019, 2021, 2023, 2025),
         month >= 3 & month <= 10)
```

We have some data missing from 2019 and 2021. I am going to supplement this data with water temp data from the NOAA HIMB station (Station MOKH1 - 1612480). Read in data from 2019 and 2021 
```{r}
temp_2019 <- read.table("../data/environmental/mokh1h2019.txt", sep = "", skip = 0, header = F, na.strings = "NA")[,c(1:4, 15)]
colnames(temp_2019) <- c("year", "month", "day", "hour", "Temperature")
temp_2021 <- read.table("../data/environmental/mokh1h2021.txt", sep = "", skip = 0, header = F, na.strings = "NA")[,c(1:4, 15)]
colnames(temp_2021) <- c("year", "month", "day", "hour", "Temperature")
temp_2025 <- read.table("../data/environmental/mokh1h2025.txt", sep = "", skip = 0, header = F, na.strings = "NA")[,c(1:4, 15)]
colnames(temp_2025) <- c("year", "month", "day", "hour", "Temperature")

# Bind data together and subset for temps between March - October 
mokh_temps <- bind_rows(temp_2019, temp_2021, temp_2025) %>%
  replace_with_na(replace = list(Temperature = c("MM", "9   9.", ".0 999", "999", "0 999."))) %>%
  filter(month >= 3 & month <= 10)
```

Identify missing rows 
```{r}
missing_rows <- mokh_temps %>%
     anti_join(subset_data,
               by = c("year", "month", "day", "hour"))
```

Append the subset data
```{r}
subset_filled <- subset_data %>%
  bind_rows(
    missing_rows %>%
      transmute(
        Temperature = Temperature,
        hour        = hour,
        year        = year,
        month       = month,
        day         = day,
        datetime    = as.POSIXct(
          sprintf("%d-%02d-%02d %02d:00:00", year, month, day, hour),
          tz = "UTC"))) %>%
  arrange(year, month, day, hour)

```

Plot 2019, 2021, 2023, and 2025
```{r}
subset_filled$datetime <- as.POSIXct(subset_filled$datetime)

# Keep only one unique row per timestamp
subset_cleaned <- subset_filled %>%
  distinct(datetime, .keep_all = TRUE)

#### PANEL A: Time Series
plot_summary <- subset_cleaned %>%
  filter(year %in% c(2019, 2021, 2023, 2025),
         month >= 3 & month <= 10) %>%
  mutate(dummy_date = as.Date(paste0("2024-", month, "-", day))) %>%
  group_by(year, dummy_date) %>%
  dplyr::summarise(
    # Use a custom check to prevent Inf if all values are NA
    mean_temp = if(all(is.na(Temperature))) NA else mean(Temperature, na.rm = TRUE),
    min_temp  = if(all(is.na(Temperature))) NA else min(Temperature, na.rm = TRUE),
    max_temp  = if(all(is.na(Temperature))) NA else max(Temperature, na.rm = TRUE),
    .groups = "drop")

label_text <- paste0(
  "<span style='color:#e41a1c;'>**2019:** 5.3 DHWs</span><br>",
  "<span style='color:#377eb8;'>**2021:** 0 DHWs</span><br>",
  "<span style='color:#4daf4a;'>**2023:** 0 DHWs</span><br>",
  "<span style='color:#984ea3;'>**2025:** 0.19 DHWs</span>"
)

p_time <- ggplot(plot_summary, aes(x = dummy_date, color = as.factor(year), fill = as.factor(year))) +
  # Shading for Daily Min/Max (geom_ribbon handles NAs by leaving gaps)
  geom_ribbon(aes(ymin = min_temp, ymax = max_temp), alpha = 0.2, color = NA) +
  # All lines are now solid
  geom_line(aes(y = mean_temp), linewidth = 1.5, linetype = "solid") +
  geom_hline(yintercept = 27.3, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 28.3, linetype = "dotted", color = "black") +
  scale_x_date(date_labels = "%b", 
               date_breaks = "1 month", 
               limits = c(as.Date("2024-03-01"), as.Date("2024-11-01")),
               expand = c(0,0)) +
  scale_y_continuous(breaks = seq(20, 32, by = 1)) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  # annotate(
  #   "richtext",
  #   x = as.Date("2024-09-30"), # Positions box at the far right
  #   y = 22,                     # Positions box at the bottom
  #   label = label_text,
  #   fill = "white",             # Background color
  #   color = "black",            # Border color
  #   label.size = 0.8,
  #   label.padding = unit(c(0.5, 0.5, 0.5, 0.5), "lines"),
  #   label.r = unit(0, "lines"), # Square corners
  #   hjust = 1, vjust = 0,       # Align bottom-right
  #   size = 7) +                    # Adjust text size to match your base_size
  theme_classic(base_size = 26) + # Massive base size
  theme(
    axis.text = element_text(color = "black", size = 22),
    axis.title = element_text(face = "bold", size = 28),
    legend.title = element_text(face = "bold", size = 24),
    legend.text = element_text(size = 22),
    legend.position = "right",
    plot.margin = margin(10, 10, 10, 10)
  ) +
  labs(x = "", y = "Temperature (°C)", color = "Year", fill = "Year"); p_time

#### PANEL B: Frequency
# freq_data_clean <- subset_filled %>%
#   filter(year %in% c(2019, 2021, 2023, 2025), month >= 3, month <= 10) %>%
#   # Create 0.5 bins: multiplying by 2, rounding, then dividing by 2
#   mutate(temp_bin = round(Temperature * 2) / 2) %>% 
#   filter(!is.na(temp_bin), temp_bin >= 22, temp_bin <= 30.5) %>%
#   group_by(year, temp_bin) %>%
#   # 'count' here represents total hours since each row is an hour
#   dplyr::summarise(total_hours = n(), .groups = "drop")
# p_hours <- ggplot(freq_data_clean, aes(x = temp_bin, y = total_hours, fill = as.factor(year))) +
#   # Keep your background shading
#   annotate("rect", xmin = 26.5, xmax = 27.5, ymin = 0, ymax = Inf, fill = "grey70", alpha = 0.5) +
#   annotate("rect", xmin = 27.5, xmax = Inf, ymin = 0, ymax = Inf, fill = "grey40", alpha = 0.5) +
#   # Use position_dodge with a specified width so bars don't overlap strangely
#   geom_bar(stat = "identity", position = position_dodge(width = 0.4), color = "black", linewidth = 0.4) +
#   scale_y_continuous(expand = c(0, 0)) +
#   # Update X axis to show half-degree breaks
#   scale_x_continuous(breaks = seq(22, 30, 0.5), limits = c(21.5, 30.5)) +
#   scale_fill_brewer(palette = "Set1") +
#   theme_classic(base_size = 22) +
#   theme(
#     axis.text.x = element_text(angle = 45, hjust = 1), # Angle labels if they get crowded
#     axis.text = element_text(color = "black"),
#     axis.title = element_text(face = "bold"),
#     legend.title = element_blank()
#   ) +
#   labs(x = "Temperature Bin (°C)", y = "Total Time (Hours)"); p_hours
subset_cleaned$year <- factor(subset_cleaned$year, levels = c("2025", "2023", "2021", "2019"))
year_colors <- c("2019" = "#e41a1c", "2021" = "#377eb8", "2023" = "#4daf4a", "2025" = "#984ea3")
dhw_labels <- data.frame(
  year = factor(c("2019", "2021", "2023", "2025"), 
                levels = c("2025", "2023", "2021", "2019")),
  dhw_text = c("5.3 DHWs", "0 DHWs", "0 DHWs", "0.19 DHWs")
)

ridge_freq <- ggplot(subset_cleaned, aes(x = Temperature, y = as.factor(year), 
                          # Use cut to create 3 shading zones
                          fill = after_stat(cut(x, 
                                                breaks = c(-Inf, 27.3, 28.3, Inf), 
                                                labels = c("Normal", "Warning", "Stress"))), 
                          color = as.factor(year))) + 
  geom_density_ridges_gradient(
    scale = 0.9, 
    rel_min_height = 0.01,
    linewidth = 1.5) +
  geom_text(data = dhw_labels, 
            aes(x = 21.5, y = year, label = dhw_text, color = year), 
            inherit.aes = FALSE,
            hjust = 0, 
            vjust = -1.6, # Adjust this to move text up/down relative to the ridge
            size = 7, 
            fontface = "bold") +
  # Manual colors: Gray for normal, Light Red for warning, Dark Red for stress
  scale_fill_manual(
    values = c("Normal" = "gray80", "Warning" = "#fbb4ae", "Stress" = "#e41a1c"), 
    guide = "none"
  ) +
  scale_color_manual(values = year_colors, guide = "none") +
  #scale_color_brewer(palette = "Set1", guide = "none") + 
  scale_x_continuous(breaks = seq(22, 32, 1), expand = c(0.01, 0)) +
  scale_y_discrete(expand = expansion(mult = c(0.01, 0.4))) +
  theme_classic(base_size = 26) + 
  theme(
    axis.text = element_text(color = "black", size = 22),
    axis.title = element_text(face = "bold", size = 28), 
    legend.position = "none",
    plot.margin = margin(10, 10, 10, 10)) +
  labs(x = "Temperature (°C)", y = "Year"); ridge_freq

## Combine and Save
final_plot <- (p_time + ridge_freq) + 
  plot_layout(widths = c(1.8, 1), guides = "collect") + 
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 32, face = "bold"))

# Save 
ggsave("../output/environmental/Temperatures_Mar-Oct_2019_2021_2023_2025.pdf", 
       plot = final_plot, 
       width = 20, 
       height = 8, 
       units = "in", 
       device = "pdf",
       dpi = 300)
ggsave("../output/environmental/Temperatures_Mar-Oct_2019_2021_2023_2025.png", 
       plot = final_plot, 
       width = 20, 
       height = 8, 
       units = "in", 
       dpi = 300)
```

Calculate percentage of time "in the red" (ie above 28.3C) and above MMM (27.3)
```{r}
coral_bleaching_summary <- subset_cleaned %>%
  filter(month >= 3 & month <= 10) %>%
  group_by(year) %>%
  dplyr::summarize(
    total_hours = n(),
    hours_above_28.3 = sum(Temperature > 28.3, na.rm = TRUE),
    percent_stress = (hours_above_28.3 / total_hours) * 100)
mmm_summary <- subset_cleaned %>%
  filter(month >= 3 & month <= 10) %>%
  group_by(year) %>%
  dplyr::summarize(
    total_hours = n(),
    hours_above_27.3 = sum(Temperature > 27.3, na.rm = TRUE),
    percent_stress = (hours_above_27.3 / total_hours) * 100)
```



