---
title: "Net calcification TPC"
author: "Jill Ashey"
date: "2025-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(car)
library(tidyverse)
library(broom)
library(readr)
library(lubridate)
library(purrr)
library(furrr)
library(fuzzyjoin)
library(future)
library(rTPC)
library(nls.multstart)
library(emmeans)
library(lme4)
library(nlme)
library(Rmisc)
library(boot)
library(nlme)
library(propagate)

n_cores <- detectCores() - 1
```

# Functions for analysis 

```{r}
sharpe_schoolfield <- function(lnc, E, Eh, Th, temp, Tc) {
  Tc <- 273.15 + Tc
  k <- 8.62e-5
  boltzmann.term <- lnc + log(exp(E/k*(1/Tc - 1/temp))) # units are eV/K, electrovolts/Kelvin
  inactivation.term <- log(1/(1 + exp(Eh/k*(1/Th - 1/temp))))
  return(boltzmann.term + inactivation.term)}

get_topt <- function(E, Th, Eh){
  k <- 8.62e-05
  return((Eh * Th) / (Eh + (k * Th * log((Eh / E) - 1))))}

get_tpc_ci <- function(model, new_data) {
  # predictNLS calculates the intervals based on the Taylor expansion (Delta Method)
  pred <- predictNLS(model, newdata = new_data, interval = "confidence", alpha = 0.05)
  return(as.data.frame(pred$summary))}
```

# 2023

Read in data 
```{r}
calc_2023 <- read.csv("../data/TPC_calcification/TPC_calcification_all_standardized_without_blanks_2023_KB.csv")
```

## Fit TPC curve 

Set target temperature to numeric, convert temperature to Kelvin and log transform the calcification data 
```{r}
calc_2023 <- calc_2023 %>%
  mutate(Target.Temperature = as.numeric(Target.Temperature),
         Target.Temperature.K = Target.Temperature + 273.15,
         log.rate = log(umol.CaCO3h.1.cm.2 + 1))
```

Fit curves for all colonies using `sharpe_schoolfield` function defined above
```{r}
fits_2023 <- calc_2023 %>%
  group_by(Phenotype,Species,Colony.ID) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(log.rate ~ sharpe_schoolfield(lnc, E, Eh, Th, temp = Target.Temperature.K, Tc = 27),
                                                data = .x,
                                                iter = 1000,
                                                start_lower = c(lnc = -10, E = 0.1, Eh = 0.2, Th = 285),
                                                start_upper = c(lnc = 10, E = 2, Eh = 5, Th = 330),
                                                supp_errors = 'Y',
                                                na.action = na.omit,
                                                lower = c(lnc = -10, E = 0, Eh = 0, Th = 0))))
```

Get TPC parameters for each colony 
```{r}
params_2023 <- fits_2023 %>%
  unnest_legacy(fit %>% purrr::map(tidy))

# Remove Colonies 202 and 242 
params_2023 <- params_2023 %>%
  filter(!Colony.ID == "202") %>%
  filter(!Colony.ID == "242")
```

Calculate Topt and Rmax for each colony 
```{r}
Topt_Rmax_2023 <- params_2023 %>%
  dplyr::select(Colony.ID, Phenotype, Species, term, estimate) %>%
  tidyr::pivot_wider(names_from = term, values_from = estimate) %>%
  mutate(
    Topt_K = get_topt(E, Th, Eh),
    Topt_C = Topt_K - 273.15,
    Rmax_log = sharpe_schoolfield(lnc, E, Eh, Th, temp = Topt_K, Tc = 27),
    Rmax_linear = exp(Rmax_log) - 1)
```

Pivot data so that each trait is a row 
```{r}
traits_long_2023 <- Topt_Rmax_2023 %>%
  select(Colony.ID, Phenotype, Species, lnc, E, Eh, Topt_C, Rmax_log) %>%
  pivot_longer(cols = c(lnc, E, Eh, Topt_C, Rmax_log), 
               names_to = "trait", 
               values_to = "estimate")
  
```

## lnc (r_tref) - 2023

Subset by lnc data
```{r}
lnc_long_2023 <- traits_long_2023 %>%
  filter(trait == "lnc")
```

Check for normality
```{r}
hist(lnc_long_2023$estimate)

# Q-Q Plot
qqnorm(lnc_long_2023$estimate)
qqline(lnc_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(lnc_long_2023$estimate)
```

Data is normal! 

Run linear mixed effects model 
```{r}
lnc.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony.ID, data=lnc_long_2023, na.action=na.exclude)
summary(lnc.lme)
anova(lnc.lme)
```

Species is significant. Species x Phenotype is moderately significant (p=0.054)

## E - 2023

Subset by E data
```{r}
e_long_2023 <- traits_long_2023 %>%
  filter(trait == "E")
```

Check for normality
```{r}
hist(e_long_2023$estimate)

# Q-Q Plot
qqnorm(e_long_2023$estimate)
qqline(e_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(e_long_2023$estimate)
```

Data is normal! 

Run linear mixed effects model 
```{r}
e.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony.ID, data=e_long_2023, na.action=na.exclude)
summary(e.lme)
anova(e.lme)
```

Species is significant. 

## Eh - 2023

Subset by Eh data
```{r}
eh_long_2023 <- traits_long_2023 %>%
  filter(trait == "Eh")
```

Check for normality
```{r}
hist(eh_long_2023$estimate)

# Q-Q Plot
qqnorm(eh_long_2023$estimate)
qqline(eh_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(eh_long_2023$estimate)
```

Data is not normal, apply transformations. 
```{r}
# Log10
eh_long_2023$eh_log <- log10(eh_long_2023$estimate +1) # add 1 to avoid log10(0)
hist(eh_long_2023$eh_log)
qqnorm(eh_long_2023$eh_log)
qqline(eh_long_2023$eh_log, col = "red")
shapiro.test(eh_long_2023$eh_log)
```

Run linear mixed effects model 
```{r}
eh.lme <- lme(eh_log~Species*Phenotype, random = ~1|Colony.ID, data=eh_long_2023, na.action=na.exclude)
summary(eh.lme)
anova(eh.lme)
```

No significant differences. 

## Topt - 2023

Subset by Topt data
```{r}
topt_long_2023 <- traits_long_2023 %>%
  filter(trait == "Topt_C")
```

Check for normality
```{r}
hist(topt_long_2023$estimate)

# Q-Q Plot
qqnorm(topt_long_2023$estimate)
qqline(topt_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(topt_long_2023$estimate)
```

Data is normal! 

Run linear mixed effects model 
```{r}
topt.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony.ID, data=topt_long_2023, na.action=na.exclude)
summary(topt.lme)
anova(topt.lme)
```

No significant differences. 

## Rmax_log - 2023

Subset by Rmax_log data
```{r}
rmax_long_2023 <- traits_long_2023 %>%
  filter(trait == "Rmax_log")
```

Check for normality
```{r}
hist(rmax_long_2023$estimate)

# Q-Q Plot
qqnorm(rmax_long_2023$estimate)
qqline(rmax_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(rmax_long_2023$estimate)
```

Data is normal! 

Run linear mixed effects model 
```{r}
rmax.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony.ID, data=rmax_long_2023, na.action=na.exclude)
summary(rmax.lme)
anova(rmax.lme)
```

Species is significant. 

## Visualization for 2023

Plot
```{r}
# Create the point plot with Mean and SE
ggplot(traits_long_2023, aes(x = Phenotype, y = estimate, color = Phenotype)) +
  # Add raw colony data as small, transparent points to show the spread
  geom_jitter(width = 0.1, alpha = 0.3, size = 1.5) +
  stat_summary(fun = mean, geom = "point", size = 4) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 1) +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "2023 Thermal Performance Traits",
    y = "Estimate (Mean ± SE)",
    x = "Phenotype"
  )

## Box plot 
ggplot(traits_long_2023, aes(x = Phenotype, y = estimate, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.6, color = "black") +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.2, color = "black") +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.major.x = element_blank() # Clean up vertical lines
  ) +
  labs(
    y = "Estimated Parameter Value",
    x = "Phenotype"
  )
```

Multidimensional visualization via PCA
```{r}
# Select only numeric traits for PCA
pca_prep <- Topt_Rmax_2023 %>%
  ungroup() %>%
  select(lnc, E, Eh, Topt_C, Rmax_log)

# Run PCA (scale = TRUE is essential here)
pca_fit <- prcomp(pca_prep, scale. = TRUE)

# Plot
autoplot(pca_fit, data = Topt_Rmax_2023, colour = 'Phenotype', 
         shape = 'Species', frame = TRUE, frame.type = 'norm',
         loadings = TRUE, loadings.label = TRUE, 
         loadings.label.size = 3, loadings.colour = "black", loadings.label.colour = "black") +
  scale_color_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  scale_fill_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  theme_bw() +
  labs(title = "Multivariate Thermal Phenotype PCA",
       caption = "Arrows indicate the direction of trait influence.")
```

# 2025

## Calculate calcification rates 

Read in data 
```{r}
calc_2025 <- read.csv("../data/TPC_calcification/Calcification_data_TPC.csv", header = T)
```

Calculate incubation time 
```{r}
calc_2025 <- calc_2025 %>%
  mutate(
    start = hm(Time_start),
    end   = hm(Time_end),
    elapsed_mins  = as.numeric(end - start, units = "mins"),
    elapsed_hours = elapsed_mins / 60)
```

Calculate delta alkalinity 
```{r}
calc_2025 <- calc_2025 %>%
  mutate(Delta_TA_raw = Starting_alk_value - Alk_avg)
```

Calculate blanks 
```{r}
blank_delta_ta <- calc_2025 %>%
  filter(ColonyID == "blank" | Notes == "blank") %>%
  group_by(Date, Target_Temperature) %>%
  summarise(mean_blank_delta_ta = mean(Delta_TA_raw, na.rm = TRUE)) %>%
  pull(mean_blank_delta_ta)
```

Subtract blank and calculate net calcification
```{r}
calc_2025 <- calc_2025 %>%
  mutate(
    Delta_TA_corr = Delta_TA_raw - blank_delta_ta,
    # Assuming seawater density 1.028 g/mL and volume in mL to L, and no coral weight (thus calc mg/L/hr)
    Net_Calcification = (Delta_TA_corr / 2) * (Sample_Volume_mL / 1000) * 1.028 / elapsed_hours) %>%
  filter(!is.na(Net_Calcification) & !is.nan(Net_Calcification))
```

Normalize by SA
```{r}
calc_2025 <- calc_2025 %>%
  mutate(Net_Calcification_cm2 = Net_Calcification / Surface_Area,
         Net_Calcification_cm2 = log(Net_Calcification_cm2 + 1))
```

Read in meta
```{r}
meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")
```

Join with calc 
```{r}
meta_unique <- meta %>%
  group_by(Colony_ID) %>%
  slice(1) %>%   # or use slice_max(Date) if you want the most recent entry
  ungroup()

calc_2025 <- calc_2025 %>%
  left_join(meta_unique, by = c("ColonyID" = "Colony_ID")) %>%
  select(ColonyID, Target_Temperature, Net_Calcification_cm2, Species, Phenotype) %>%
  mutate(log.rate = log(Net_Calcification_cm2 + 1))
```

Plot 
```{r}
# Filter out blanks and missing phenotype/species
calc_plot <- calc_2025 %>%
  filter(!is.na(Phenotype),
         !is.na(Species),
         ColonyID != "blank")

ggplot(calc_plot, aes(
  x = Target_Temperature,
  y = log.rate,
  color = Phenotype,
  fill = Phenotype
)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "loess", se = TRUE, span = 0.6, alpha = 0.25) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  labs(
    x = "Target Temperature (°C)",
    y = expression("Log(Net Calcification ("*cm^-2*"))"),
    color = "Phenotype",
    fill = "Phenotype")
```

## Fit TPC curve

Set target temperature to numeric and convert temperature to Kelvin
```{r}
calc_2025 <- calc_2025 %>%
  mutate(Target_Temperature = as.numeric(Target_Temperature),
         Target.Temperature.K = Target_Temperature + 273.15) %>%
  filter(!is.na(Phenotype),
         !is.na(Species),
         ColonyID != "blank")
```

Fit curves for all colonies using `sharpe_schoolfield` function defined above
```{r}
fits_2025 <- calc_2025 %>%
  group_by(Phenotype,Species,ColonyID) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(log.rate ~ sharpe_schoolfield(lnc, E, Eh, Th, temp = Target.Temperature.K, Tc = 27),
                                                data = .x,
                                                iter = 1000,
                                                start_lower = c(lnc = -10, E = 0.1, Eh = 0.2, Th = 285),
                                                start_upper = c(lnc = 10, E = 2, Eh = 5, Th = 330),
                                                supp_errors = 'Y',
                                                na.action = na.omit,
                                                lower = c(lnc = -10, E = 0, Eh = 0, Th = 0))))
```

Get TPC parameters for each colony 
```{r}
params_2025 <- fits_2025 %>%
  unnest_legacy(fit %>% purrr::map(tidy))

# Remove Colonies 3, 221, and 240
params_2025 <- params_2025 %>%
  filter(!ColonyID == "240") %>%
  filter(!ColonyID == "221") %>%
  filter(!ColonyID == "3") 
```

Calculate Topt and Rmax for each colony 
```{r}
Topt_Rmax_2025 <- params_2025 %>%
  dplyr::select(ColonyID, Phenotype, Species, term, estimate) %>%
  tidyr::pivot_wider(names_from = term, values_from = estimate) %>%
  mutate(
    Topt_K = get_topt(E, Th, Eh),
    Topt_C = Topt_K - 273.15,
    Rmax_log = sharpe_schoolfield(lnc, E, Eh, Th, temp = Topt_K, Tc = 27),
    Rmax_linear = exp(Rmax_log) - 1)
```

Pivot data so that each trait is a row 
```{r}
traits_long_2025 <- Topt_Rmax_2025 %>%
  select(ColonyID, Phenotype, Species, lnc, E, Eh, Topt_C, Rmax_log) %>%
  pivot_longer(cols = c(lnc, E, Eh, Topt_C, Rmax_log), 
               names_to = "trait", 
               values_to = "estimate")
```

## lnc (r_tref) - 2025

Subset by lnc data
```{r}
lnc_long_2025 <- traits_long_2025 %>%
  filter(trait == "lnc")
```

Check for normality
```{r}
hist(lnc_long_2025$estimate)

# Q-Q Plot
qqnorm(lnc_long_2025$estimate)
qqline(lnc_long_2025$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(lnc_long_2025$estimate)
```

Data is not normal, perform data transformation
```{r}
# Log10
lnc_long_2025$lnc_log <- log10(lnc_long_2025$estimate + 1) # add 1 to avoid log10(0)
hist(lnc_long_2025$lnc_log)
qqnorm(lnc_long_2025$lnc_log)
qqline(lnc_long_2025$lnc_log, col = "red")
shapiro.test(lnc_long_2025$lnc_log)

# Sqrt
lnc_long_2025$calc_sqrt <- sqrt(lnc_long_2025$estimate)
hist(lnc_long_2025$calc_sqrt)
qqnorm(lnc_long_2025$calc_sqrt)
qqline(lnc_long_2025$calc_sqrt, col = "red")
shapiro.test(lnc_long_2025$calc_sqrt)

# Inverse 
lnc_long_2025$lnc_inv <- 1 / (lnc_long_2025$estimate +1) # add 1 to avoid division by zero
hist(lnc_long_2025$lnc_inv)
qqnorm(lnc_long_2025$lnc_inv)
qqline(lnc_long_2025$lnc_inv, col = "red")
shapiro.test(lnc_long_2025$lnc_inv)
```

Normality not produced. Come back to this. 

Run linear mixed effects model 
```{r}
lnc.lme <- lme(estimate~Species*Phenotype, random = ~1|ColonyID, data=lnc_long_2025, na.action=na.exclude)
summary(lnc.lme)
anova(lnc.lme)
```

No significant differences.

## E - 2025

Subset by E data
```{r}
e_long_2025 <- traits_long_2025 %>%
  filter(trait == "E")
```

Check for normality
```{r}
hist(e_long_2025$estimate)

# Q-Q Plot
qqnorm(e_long_2025$estimate)
qqline(e_long_2025$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(e_long_2025$estimate)
```

Data is not normal, perform data transformation
```{r}
# Log10
e_long_2025$e_log <- log10(e_long_2025$estimate + 1) # add 1 to avoid log10(0)
hist(e_long_2025$e_log)
qqnorm(e_long_2025$e_log)
qqline(e_long_2025$e_log, col = "red")
shapiro.test(e_long_2025$e_log)

# Sqrt
e_long_2025$e_sqrt <- sqrt(e_long_2025$estimate)
hist(e_long_2025$e_sqrt)
qqnorm(e_long_2025$e_sqrt)
qqline(e_long_2025$e_sqrt, col = "red")
shapiro.test(e_long_2025$e_sqrt)

# Inverse 
e_long_2025$e_inv <- 1 / (e_long_2025$estimate +1) # add 1 to avoid division by zero
hist(e_long_2025$e_inv)
qqnorm(e_long_2025$e_inv)
qqline(e_long_2025$e_inv, col = "red")
shapiro.test(e_long_2025$e_inv)
```

Normality not produced. Come back to this. 

Run linear mixed effects model 
```{r}
e.lme <- lme(estimate~Species*Phenotype, random = ~1|ColonyID, data=e_long_2025, na.action=na.exclude)
summary(e.lme)
anova(e.lme)
```

No significant differences. 

## Eh - 2025

Subset by Eh data
```{r}
eh_long_2025 <- traits_long_2025 %>%
  filter(trait == "Eh")
```

Check for normality
```{r}
hist(eh_long_2025$estimate)

# Q-Q Plot
qqnorm(eh_long_2025$estimate)
qqline(eh_long_2025$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(eh_long_2025$estimate)
```

Data is not normal, perform data transformation
```{r}
# Log10
eh_long_2025$eh_log <- log10(eh_long_2025$estimate + 1) # add 1 to avoid log10(0)
hist(eh_long_2025$eh_log)
qqnorm(eh_long_2025$eh_log)
qqline(eh_long_2025$eh_log, col = "red")
shapiro.test(eh_long_2025$eh_log)

# Sqrt
eh_long_2025$eh_sqrt <- sqrt(eh_long_2025$estimate)
hist(eh_long_2025$eh_sqrt)
qqnorm(eh_long_2025$eh_sqrt)
qqline(eh_long_2025$eh_sqrt, col = "red")
shapiro.test(eh_long_2025$eh_sqrt)

# Inverse 
eh_long_2025$eh_inv <- 1 / (eh_long_2025$estimate +1) # add 1 to avoid division by zero
hist(eh_long_2025$eh_inv)
qqnorm(eh_long_2025$eh_inv)
qqline(eh_long_2025$eh_inv, col = "red")
shapiro.test(eh_long_2025$eh_inv)
```

All produced normality. Move forward with log

Run linear mixed effects model 
```{r}
eh.lme <- lme(eh_log~Species*Phenotype, random = ~1|ColonyID, data=eh_long_2025, na.action=na.exclude)
summary(eh.lme)
anova(eh.lme)
```

Phenotype signficant. 

## Topt - 2025

Subset by Topt data
```{r}
topt_long_2025 <- traits_long_2025 %>%
  filter(trait == "Topt_C")
```

Check for normality
```{r}
hist(topt_long_2025$estimate)

# Q-Q Plot
qqnorm(topt_long_2025$estimate)
qqline(topt_long_2025$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(topt_long_2025$estimate)
```

Data is normal! 

Run linear mixed effects model 
```{r}
topt.lme <- lme(estimate~Species*Phenotype, random = ~1|ColonyID, data=topt_long_2025, na.action=na.exclude)
summary(topt.lme)
anova(topt.lme)
```

No significant differences. 

## Rmax_log - 2025

Subset by Rmax_log data
```{r}
rmax_long_2025 <- traits_long_2025 %>%
  filter(trait == "Rmax_log")
```

Check for normality
```{r}
hist(rmax_long_2025$estimate)

# Q-Q Plot
qqnorm(rmax_long_2025$estimate)
qqline(rmax_long_2025$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(rmax_long_2025$estimate)
```

Data is normal!

Run linear mixed effects model 
```{r}
rmax.lme <- lme(estimate~Species*Phenotype, random = ~1|ColonyID, data=rmax_long_2025, na.action=na.exclude)
summary(rmax.lme)
anova(rmax.lme)
```

Species is significant. 

## Visualization for 2025

Plot
```{r}
# Create the point plot with Mean and SE
ggplot(traits_long_2025, aes(x = Phenotype, y = estimate, color = Phenotype)) +
  # Add raw colony data as small, transparent points to show the spread
  geom_jitter(width = 0.1, alpha = 0.3, size = 1.5) +
  stat_summary(fun = mean, geom = "point", size = 4) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 1) +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "2025 Thermal Performance Traits",
    y = "Estimate (Mean ± SE)",
    x = "Phenotype"
  )

## Box plot 
ggplot(traits_long_2025, aes(x = Phenotype, y = estimate, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.6, color = "black") +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.2, color = "black") +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.major.x = element_blank() # Clean up vertical lines
  ) +
  labs(
    y = "Estimated Parameter Value",
    x = "Phenotype"
  )
```

Multidimensional visualization via PCA
```{r}
# Select only numeric traits for PCA
pca_prep <- Topt_Rmax_2025 %>%
  ungroup() %>%
  select(lnc, E, Eh, Topt_C, Rmax_log)

# Run PCA (scale = TRUE is essential here)
pca_fit <- prcomp(pca_prep, scale. = TRUE)

# Plot
autoplot(pca_fit, data = Topt_Rmax_2025, colour = 'Phenotype', 
         shape = 'Species', frame = TRUE, frame.type = 'norm',
         loadings = TRUE, loadings.label = TRUE, 
         loadings.label.size = 3, loadings.colour = "black", loadings.label.colour = "black") +
  scale_color_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  scale_fill_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  theme_bw() +
  labs(title = "Multivariate Thermal Phenotype PCA",
       caption = "Arrows indicate the direction of trait influence.")
```

# Compare 2023 and 2025

Rename Colony id column from 2023 so it matches 2025 and add year as a column to both dfs
```{r}
traits_long_2023 <- traits_long_2023 %>%
  mutate(Year = "2023") %>%
  #dplyr::rename("ColonyID" = "Colony.ID") %>%
  mutate(ColonyID = as.character(ColonyID))

traits_long_2025 <- traits_long_2025 %>%
  mutate(Year = "2025")
```

Bind dfs together 
```{r}
traits_long_all <- rbind(traits_long_2023, traits_long_2025)
```

Visualize differences between 2023 and 2025
```{r}
ggplot(traits_long_all, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  # 1. Add Error Bars (Standard Error)
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, size = 1) +
  
  # 2. Add Lines to connect the means over time
  stat_summary(fun = mean, geom = "line", size = 1.2) +
  
  # 3. Add the Mean Points
  stat_summary(fun = mean, geom = "point", size = 4) +
  
  # 4. Facet by Trait and Species
  # Using 'free_y' because Topt (~30) and E (~0.5) have very different scales
  facet_grid(trait ~ Species, scales = "free_y") +
  
  # 5. Styling and Colors
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 10),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  labs(
    title = "Trait Evolution: 2023 vs 2025",
    y = "Estimated Value (Mean ± SE)",
    x = "Year"
  )
```


