---
title: "Net calcification TPC"
author: "Jill Ashey"
date: "2025-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(car)
library(tidyverse)
library(broom)
library(readr)
library(lubridate)
library(purrr)
library(furrr)
library(fuzzyjoin)
library(future)
library(rTPC)
library(nls.multstart)
library(emmeans)
library(lme4)
library(nlme)
library(Rmisc)
library(boot)
library(nlme)
library(propagate)
library(patchwork)
library(lmerTest)
library(performance)
library(insight)
library(ggtext)

#n_cores <- detectCores() - 1
```

# Functions for analysis 

```{r}
sharpe_schoolfield <- function(lnc, E, Eh, Th, temp, Tc) {
  Tc <- 273.15 + Tc
  k <- 8.62e-5
  boltzmann.term <- lnc + log(exp(E/k*(1/Tc - 1/temp))) # units are eV/K, electrovolts/Kelvin
  inactivation.term <- log(1/(1 + exp(Eh/k*(1/Th - 1/temp))))
  return(boltzmann.term + inactivation.term)}

get_topt <- function(E, Th, Eh){
  k <- 8.62e-05
  return((Eh * Th) / (Eh + (k * Th * log((Eh / E) - 1))))}

get_tpc_ci <- function(model, new_data) {
  # predictNLS calculates the intervals based on the Taylor expansion (Delta Method)
  pred <- predictNLS(model, newdata = new_data, interval = "confidence", alpha = 0.05)
  return(as.data.frame(pred$summary))}
```

Run this before running any ANOVAs
```{r}
options(contrasts = c("contr.sum", "contr.poly"))
```

# 2023

Read in data 
```{r}
calc_2023 <- read.csv("../data/TPC_calcification/TPC_calcification_all_standardized_without_blanks_2023_KB.csv")
```

## Fit TPC curve 

Set target temperature to numeric, convert temperature to Kelvin and log transform the calcification data 
```{r}
calc_2023 <- calc_2023 %>%
  mutate(Target.Temperature = as.numeric(Target.Temperature),
         Target.Temperature.K = Target.Temperature + 273.15,
         log.rate = log(umol.CaCO3h.1.cm.2 + 1))
```

Fit curves for all colonies using `sharpe_schoolfield` function defined above
```{r}
fits_2023 <- calc_2023 %>%
  group_by(Phenotype,Species,Colony.ID) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(log.rate ~ sharpe_schoolfield(lnc, E, Eh, Th, temp = Target.Temperature.K, Tc = 27),
                                                data = .x,
                                                iter = 1000,
                                                start_lower = c(lnc = -10, E = 0.1, Eh = 0.2, Th = 285),
                                                start_upper = c(lnc = 10, E = 2, Eh = 5, Th = 330),
                                                supp_errors = 'Y',
                                                na.action = na.omit,
                                                lower = c(lnc = -10, E = 0, Eh = 0, Th = 0))))
```

Get TPC parameters for each colony 
```{r}
params_2023 <- fits_2023 %>%
  unnest_legacy(fit %>% purrr::map(tidy))

# Remove Colonies 202 and 242 
params_2023 <- params_2023 %>%
  filter(!Colony.ID == "202") %>%
  filter(!Colony.ID == "242")
```

Calculate Topt and Rmax for each colony 
```{r}
Topt_Rmax_2023 <- params_2023 %>%
  dplyr::select(Colony.ID, Phenotype, Species, term, estimate) %>%
  tidyr::pivot_wider(names_from = term, values_from = estimate) %>%
  mutate(
    Topt_K = get_topt(E, Th, Eh),
    Topt_C = Topt_K - 273.15,
    Rmax_log = sharpe_schoolfield(lnc, E, Eh, Th, temp = Topt_K, Tc = 27),
    Rmax_linear = exp(Rmax_log) - 1)
```

Pivot data so that each trait is a row 
```{r}
traits_long_2023 <- Topt_Rmax_2023 %>%
  dplyr::select(Colony.ID, Phenotype, Species, lnc, E, Eh, Topt_C, Rmax_log) %>%
  pivot_longer(cols = c(lnc, E, Eh, Topt_C, Rmax_log), 
               names_to = "trait", 
               values_to = "estimate")
  
```

## Visualization for 2023

Plot
```{r}
# Create the point plot with Mean and SE
ggplot(traits_long_2023, aes(x = Phenotype, y = estimate, color = Phenotype)) +
  # Add raw colony data as small, transparent points to show the spread
  geom_jitter(width = 0.1, alpha = 0.3, size = 1.5) +
  stat_summary(fun = mean, geom = "point", size = 4) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 1) +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "2023 Thermal Performance Traits",
    y = "Estimate (Mean ± SE)",
    x = "Phenotype"
  )

## Box plot 
ggplot(traits_long_2023, aes(x = Phenotype, y = estimate, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.6, color = "black") +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.2, color = "black") +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.major.x = element_blank() # Clean up vertical lines
  ) +
  labs(
    y = "Estimated Parameter Value",
    x = "Phenotype"
  )
```

### Plot with confidence intervals 

Generate and aggregate predictions
```{r}
# 1. Define the temperature range for the smooth curves
new_data <- data.frame(
  Target.Temperature.K = seq(18 + 273.15 - 1,  # 17°C
                      38 + 273.15 + 1,  # 39°C
                      length.out = 200))

# 2. Generate predictions for every colony model
# This creates a "long" dataframe of all potential curves
preds_all_2023 <- fits_2023 %>%
  mutate(preds = map(fit, ~{
    p <- try(predict(.x, newdata = new_data), silent = TRUE)
    if(inherits(p, "try-error")) return(NULL)
    data.frame(Target.Temperature.K = new_data$Target.Temperature.K, 
               pred_val = as.numeric(p))
  })) %>%
  filter(!map_lgl(preds, is.null)) %>%
  unnest(preds)

# 3. Calculate the MEAN and CI of the curves per group
# At each temperature step, find the mean performance across all colonies
preds_mean_summary_2023 <- preds_all_2023 %>%
  group_by(Phenotype, Species, Target.Temperature.K) %>%
  dplyr::summarise(
    mean_val = mean(pred_val, na.rm = TRUE),
    sd_val = sd(pred_val, na.rm = TRUE),
    n = n(),
    se_val = sd_val / sqrt(n),
    # 95% Confidence Interval of the Mean
    lower = mean_val - (1.96 * se_val),
    upper = mean_val + (1.96 * se_val),
    .groups = "drop"
  )
```

Plot TPC with fitted curves 
```{r}
ggplot(preds_mean_summary_2023, aes(x = Target.Temperature.K - 273.15, y = mean_val, color = Phenotype, fill = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5, color = NA) +
  geom_line(linewidth = 1.5) +
  scale_x_continuous(
    breaks = seq(18, 40, by = 2), 
    limits = c(17, 40)) +
  geom_point(data = calc_2023, aes(x = Target.Temperature.K - 273.15, y = log.rate), 
             alpha = 0.2, size = 1, inherit.aes = TRUE) +
  facet_wrap(~ Species) +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 12),
    panel.grid = element_blank()
  ) +
  labs(
    x = "Temperature (°C)", 
    y = expression(bold("Net calcification (log(" *mu*"mol " *cm^-2 *hr^-1*"))")),
    title = "Calcification Fitted Thermal Performance Curves (2023)"
  )
```

### Plot with confidence intervals and the environmental data 

```{r}
temps <- read.csv("../output/environmental/NSF_CAREER_KBay_PR13_Temps.csv")
```

Subset for the year prior to Sept 23, 2023
```{r}
temps_sub <- temps %>%
  filter(DateTime <= "2023-09-23 01:00:00")
```

```{r}
temp_freq_2023 <- temps_sub %>%
  # Rounding to the nearest whole degree centers the bar on that number
  mutate(temp_bin = round(Temp_C, 0)) %>% 
  group_by(temp_bin) %>%
  dplyr::summarise(count = n(), .groups = "drop") %>%
  # Calculate the percentage of total time spent at that temperature
  mutate(percent = (count / sum(count)) * 100)
```

Plot with temp bars
```{r}
# Axis Parameters
p_min <- -0.5  # Hard floor of the plot
p_max <- 0.65   # Peak of the Y-axis
p_total <- p_max - p_min
s_max <- 30    # Max percent on the right axis
scale_factor <- p_total / s_max

ggplot() +
  geom_rect(data = temp_freq_2023, 
            aes(xmin = temp_bin - 0.45, xmax = temp_bin + 0.45, 
                ymin = p_min, ymax = p_min + (percent * scale_factor)), 
            fill = "#8DBBCC", color = "#727E8B", linewidth = 0.2) +
  geom_point(data = calc_2023, 
             aes(x = Target.Temperature.K - 273.15, y = log.rate, color = Phenotype), 
             alpha = 0.5, size = 1.2, position = position_jitter(width = 0.1)) +
  geom_ribbon(data = preds_mean_summary_2023, 
              aes(x = Target.Temperature.K - 273.15, ymin = lower, ymax = upper, fill = Phenotype), 
              alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, color = "grey", linewidth = 0.5, linetype = "dashed") +
  geom_line(data = preds_mean_summary_2023, 
            aes(x = Target.Temperature.K - 273.15, y = mean_val, color = Phenotype), 
            linewidth = 1.3) +
  scale_y_continuous(
    name = expression("Net calcification(log(" *mu*"mol " *cm^-2 *hr^-1*"))"),
    limits = c(p_min, p_max),
    expand = c(0, 0),
    sec.axis = sec_axis(~ (. - p_min) / scale_factor, name = "Percent days (%)", breaks = seq(0, 40, 5))
  ) +
  scale_x_continuous(breaks = seq(18, 40, by = 2), limits = c(17.5, 40.5), expand = c(0,0)) +
  facet_wrap(~ Species, nrow = 1) +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_classic() +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 13),
    axis.line = element_line(color = "black"),
    panel.spacing = unit(2, "lines")
  ) +
  labs(x = "Temperature (°C)")
```



# 2025

## Calculate calcification rates 

Read in data 
```{r}
calc_2025 <- read.csv("../data/TPC_calcification/Calcification_data_TPC.csv", header = T)
```

Calculate incubation time 
```{r}
calc_2025 <- calc_2025 %>%
  mutate(
    start = hm(Time_start),
    end   = hm(Time_end),
    elapsed_mins  = as.numeric(end - start, units = "mins"),
    elapsed_hours = elapsed_mins / 60)
```

Calculate delta alkalinity 
```{r}
calc_2025 <- calc_2025 %>%
  mutate(Delta_TA_raw = Starting_alk_value - Alk_avg)
```

Calculate blanks 
```{r}
blank_delta_ta <- calc_2025 %>%
  filter(ColonyID == "blank" | Notes == "blank") %>%
  group_by(Date, Target_Temperature) %>%
  summarise(mean_blank_delta_ta = mean(Delta_TA_raw, na.rm = TRUE)) %>%
  pull(mean_blank_delta_ta)
```

Subtract blank and calculate net calcification
```{r}
calc_2025 <- calc_2025 %>%
  mutate(
    Delta_TA_corr = Delta_TA_raw - blank_delta_ta,
    # Assuming seawater density 1.028 g/mL and volume in mL to L, and no coral weight (thus calc mg/L/hr)
    Net_Calcification = (Delta_TA_corr / 2) * (Sample_Volume_mL / 1000) * 1.028 / elapsed_hours) %>%
  filter(!is.na(Net_Calcification) & !is.nan(Net_Calcification))
```

Normalize by SA
```{r}
calc_2025 <- calc_2025 %>%
  mutate(Net_Calcification_cm2 = Net_Calcification / Surface_Area_cm2)
```

Read in meta
```{r}
meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")
```

Join with calc 
```{r}
meta_unique <- meta %>%
  group_by(Colony_ID) %>%
  slice(1) %>%   # or use slice_max(Date) if you want the most recent entry
  ungroup()

calc_2025 <- calc_2025 %>%
  left_join(meta_unique, by = c("ColonyID" = "Colony_ID")) %>%
  dplyr::select(ColonyID, Target_Temperature, Net_Calcification_cm2, Species, Phenotype) %>%
  mutate(log.rate = log(Net_Calcification_cm2 + 1))
```

Save calcification data as csv
```{r}
write.csv(calc_2025, "../output/TPC_Calcification/TPC_Calcification_normalized_2025.csv")
```

Plot 
```{r}
# Filter out blanks and missing phenotype/species
calc_plot <- calc_2025 %>%
  filter(!is.na(Phenotype),
         !is.na(Species),
         ColonyID != "blank",
         ColonyID != "213") 

ggplot(calc_plot, aes(
  x = Target_Temperature,
  y = log.rate,
  color = Phenotype,
  fill = Phenotype
)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "loess", se = TRUE, span = 0.6, alpha = 0.25) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  labs(
    x = "Target Temperature (°C)",
    y = expression("Log(Net Calcification ("*cm^-2*"))"),
    color = "Phenotype",
    fill = "Phenotype")
```

## Fit TPC curve

Set target temperature to numeric and convert temperature to Kelvin
```{r}
calc_2025 <- calc_2025 %>%
  mutate(Target_Temperature = as.numeric(Target_Temperature),
         Target.Temperature.K = Target_Temperature + 273.15) %>%
  filter(!is.na(Phenotype),
         !is.na(Species),
         ColonyID != "blank",
         ColonyID != "213") # Remove 213, outlier
```

Fit curves for all colonies using `sharpe_schoolfield` function defined above
```{r}
fits_2025 <- calc_2025 %>%
  group_by(Phenotype,Species,ColonyID) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(log.rate ~ sharpe_schoolfield(lnc, E, Eh, Th, temp = Target.Temperature.K, Tc = 27),
                                                data = .x,
                                                iter = 1000,
                                                start_lower = c(lnc = -10, E = 0.1, Eh = 0.2, Th = 285),
                                                start_upper = c(lnc = 10, E = 2, Eh = 5, Th = 330),
                                                supp_errors = 'Y',
                                                na.action = na.omit,
                                                lower = c(lnc = -10, E = 0, Eh = 0, Th = 0))))
```

Get TPC parameters for each colony 
```{r}
params_2025 <- fits_2025 %>%
  unnest_legacy(fit %>% purrr::map(tidy))

# Remove Colonies 3, 221, and 240 based on E values being 0 and 210 and 211 for E values being super high
params_2025 <- params_2025 %>%
  filter(!ColonyID == "240") %>%
  filter(!ColonyID == "221") %>%
  filter(!ColonyID == "3") %>%
  filter(!ColonyID == "210") %>%
  filter(!ColonyID == "211")
```

Calculate Topt and Rmax for each colony 
```{r}
Topt_Rmax_2025 <- params_2025 %>%
  dplyr::select(ColonyID, Phenotype, Species, term, estimate) %>%
  tidyr::pivot_wider(names_from = term, values_from = estimate) %>%
  mutate(
    Topt_K = get_topt(E, Th, Eh),
    Topt_C = Topt_K - 273.15,
    Rmax_log = sharpe_schoolfield(lnc, E, Eh, Th, temp = Topt_K, Tc = 27),
    Rmax_linear = exp(Rmax_log) - 1)
```

Pivot data so that each trait is a row 
```{r}
traits_long_2025 <- Topt_Rmax_2025 %>%
  dplyr::select(ColonyID, Phenotype, Species, lnc, E, Eh, Topt_C, Rmax_log) %>%
  pivot_longer(cols = c(lnc, E, Eh, Topt_C, Rmax_log), 
               names_to = "trait", 
               values_to = "estimate")
```

## Visualization for 2025

Plot
```{r}
# Create the point plot with Mean and SE
ggplot(traits_long_2025, aes(x = Phenotype, y = estimate, color = Phenotype)) +
  # Add raw colony data as small, transparent points to show the spread
  geom_jitter(width = 0.1, alpha = 0.3, size = 1.5) +
  stat_summary(fun = mean, geom = "point", size = 4) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 1) +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "2025 Thermal Performance Traits",
    y = "Estimate (Mean ± SE)",
    x = "Phenotype"
  )

## Box plot 
ggplot(traits_long_2025, aes(x = Phenotype, y = estimate, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.6, color = "black") +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.2, color = "black") +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.major.x = element_blank() # Clean up vertical lines
  ) +
  labs(
    y = "Estimated Parameter Value",
    x = "Phenotype"
  )
```

## Plot with confidence intervals 

Generate and aggregate predictions
```{r}
# 1. Define the temperature range for the smooth curves
new_data <- data.frame(
  Target.Temperature.K = seq(18 + 273.15 - 1,  # 17°C
                      38 + 273.15 + 1,  # 39°C
                      length.out = 200))

# 2. Generate predictions for every colony model
# This creates a "long" dataframe of all potential curves
preds_all_2025 <- fits_2025 %>%
  mutate(preds = map(fit, ~{
    # We use a safe predict call in case some models didn't converge
    p <- try(predict(.x, newdata = new_data), silent = TRUE)
    if(inherits(p, "try-error")) return(NULL)
    data.frame(Target.Temperature.K = new_data$Target.Temperature.K, pred_val = p)
  })) %>%
  filter(!map_lgl(preds, is.null)) %>% # Remove failed fits
  unnest(preds)

# 3. Calculate the MEAN and CI of the curves per group
# At each temperature step, find the mean performance across all colonies
preds_mean_summary_2025 <- preds_all_2025 %>%
  group_by(Phenotype, Species, Target.Temperature.K) %>%
  dplyr::summarise(
    mean_val = mean(pred_val, na.rm = TRUE),
    sd_val = sd(pred_val, na.rm = TRUE),
    n = n(),
    se_val = sd_val / sqrt(n),
    # 95% Confidence Interval of the Mean
    lower = mean_val - (1.96 * se_val),
    upper = mean_val + (1.96 * se_val),
    .groups = "drop"
  )
```

Plot TPC with fitted curves 
```{r}
ggplot(preds_mean_summary_2025, aes(x = Target.Temperature.K - 273.15, y = mean_val, color = Phenotype, fill = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5, color = NA) +
  geom_line(linewidth = 1.5) +
  scale_x_continuous(
    breaks = seq(18, 40, by = 2), 
    limits = c(17, 40)) +
  geom_point(data = calc_2025, aes(x = Target.Temperature.K - 273.15, y = log.rate), 
             alpha = 0.2, size = 1, inherit.aes = TRUE) +
  facet_wrap(~ Species) +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 12),
    panel.grid = element_blank()
  ) +
  labs(
    x = "Temperature (°C)", 
    y = expression(bold("Net calcification (log(" *mu*"mol " *cm^-2 *hr^-1*"))")),
    title = "Calcification Fitted Thermal Performance Curves (2025)"
  )
```

## Plot with confidence intervals and the environmental data 

```{r}
temps <- read.csv("../output/environmental/NSF_CAREER_KBay_PR13_Temps.csv")
```

Subset for the year prior to Oct 7, 2025
```{r}
temps_sub <- temps %>%
  filter(DateTime <= "2025-10-06 01:00:00")
```

```{r}
temp_freq_2025 <- temps_sub %>%
  # Rounding to the nearest whole degree centers the bar on that number
  mutate(temp_bin = round(Temp_C, 0)) %>% 
  group_by(temp_bin) %>%
  dplyr::summarise(count = n(), .groups = "drop") %>%
  # Calculate the percentage of total time spent at that temperature
  mutate(percent = (count / sum(count)) * 100)
```

Plot with temp bars
```{r}
# Axis Parameters
p_min <- -1.0  # Hard floor of the plot
p_max <- 0.75   # Peak of the Y-axis
p_total <- p_max - p_min
s_max <- 30    # Max percent on the right axis
scale_factor <- p_total / s_max

ggplot() +
  geom_rect(data = temp_freq_2025, 
            aes(xmin = temp_bin - 0.45, xmax = temp_bin + 0.45, 
                ymin = p_min, ymax = p_min + (percent * scale_factor)), 
            fill = "#8DBBCC", color = "#727E8B", linewidth = 0.2) +
  geom_point(data = calc_2025, 
             aes(x = Target.Temperature.K - 273.15, y = log.rate, color = Phenotype), 
             alpha = 0.5, size = 1.2, position = position_jitter(width = 0.1)) +
  geom_ribbon(data = preds_mean_summary_2025, 
              aes(x = Target.Temperature.K - 273.15, ymin = lower, ymax = upper, fill = Phenotype), 
              alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, color = "grey", linewidth = 0.5, linetype = "dashed") +
  geom_line(data = preds_mean_summary_2025, 
            aes(x = Target.Temperature.K - 273.15, y = mean_val, color = Phenotype), 
            linewidth = 1.3) +
  scale_y_continuous(
    name = expression("Net calcification (log(" *mu*"mol " *cm^-2 *hr^-1*"))"),
    limits = c(p_min, p_max),
    expand = c(0, 0),
    sec.axis = sec_axis(~ (. - p_min) / scale_factor, name = "Percent days (%)", breaks = seq(0, 40, 5))
  ) +
  scale_x_continuous(breaks = seq(18, 40, by = 2), limits = c(17.5, 40.5), expand = c(0,0)) +
  facet_wrap(~ Species, nrow = 1) +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_classic() +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 13),
    axis.line = element_line(color = "black"),
    panel.spacing = unit(2, "lines")
  ) +
  labs(x = "Temperature (°C)")
```

# Compare 2023 and 2025

Rename Colony id column from 2023 so it matches 2025 and add year as a column to both dfs
```{r}
traits_long_2023 <- traits_long_2023 %>%
  mutate(Year = "2023") %>%
  dplyr::rename("ColonyID" = "Colony.ID") %>%
  mutate(ColonyID = as.character(ColonyID))

traits_long_2025 <- traits_long_2025 %>%
  mutate(Year = "2025")
```

Bind dfs together 
```{r}
traits_long_all <- rbind(traits_long_2023, traits_long_2025)
```

## MCAP Statistics comparing 2023 v. 2025 TPC photosynthesis metrics 

Subset by Mcap only 
```{r}
traits_long_all_mcap <- traits_long_all %>%
  filter(Species == "Montipora capitata")
```

### lnc (r_tref)

Subset by lnc data
```{r}
lnc_long <- traits_long_all_mcap %>%
  filter(trait == "lnc")
```

Check for normality
```{r}
hist(lnc_long$estimate)

# Q-Q Plot
qqnorm(lnc_long$estimate)
qqline(lnc_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(lnc_long$estimate)
```

Data is not normal. Transform data to achieve normality 
```{r}
lnc_long$log <- log(lnc_long$estimate)
hist(lnc_long$log)
qqnorm(lnc_long$log)
qqline(lnc_long$log, col = "red")
shapiro.test(lnc_long$log)

lnc_long$sqrt <- sqrt(lnc_long$estimate)
hist(lnc_long$sqrt)
qqnorm(lnc_long$sqrt)
qqline(lnc_long$sqrt, col = "red")
shapiro.test(lnc_long$sqrt)
```

Log produced normality. 

Run linear mixed effects model 
```{r}
lnc.model <- lmer(log ~ Year * Phenotype + (1|ColonyID), data = lnc_long)
car::Anova(lnc.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(lnc.model, which = 1) 
lattice::qqmath(lnc.model)

# Post hoc tests (even though no sig diffs)
emmeans(lnc.model, pairwise ~ Phenotype | Year)
emmeans(lnc.model, pairwise ~ Year | Phenotype)

# Check plot to make sure results make sense
ggplot(lnc_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year is significant. 

### E

Subset by E data
```{r}
e_long <- traits_long_all_mcap %>%
  filter(trait == "E")
```

Check for normality
```{r}
hist(e_long$estimate)

# Q-Q Plot
qqnorm(e_long$estimate)
qqline(e_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(e_long$estimate)
```

Data is not normal. Transform data to achieve normality 
```{r}
e_long$log <- log(e_long$estimate)
hist(e_long$log)
qqnorm(e_long$log)
qqline(e_long$log, col = "red")
shapiro.test(e_long$log)

e_long$sqrt <- sqrt(e_long$estimate)
hist(e_long$sqrt)
qqnorm(e_long$sqrt)
qqline(e_long$sqrt, col = "red")
shapiro.test(e_long$sqrt)
```

Square root transformation produced normality. 

Run linear mixed effects model 
```{r}
e.model <- lmer(sqrt ~ Year * Phenotype + (1|ColonyID), data = e_long)
car::Anova(e.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(e.model, which = 1) 
lattice::qqmath(e.model)

# Post hoc tests
emmeans(e.model, pairwise ~ Year | Phenotype)
emmeans(e.model, pairwise ~ Phenotype | Year)

# Check plot to make sure results make sense
ggplot(e_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Phenotype x Year is marginally significant. 

### Eh

Subset by E data
```{r}
eh_long <- traits_long_all_mcap %>%
  filter(trait == "Eh")
```

Check for normality
```{r}
hist(eh_long$estimate)

# Q-Q Plot
qqnorm(eh_long$estimate)
qqline(eh_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(eh_long$estimate)
```

Data is not normal. Transform data to achieve normality 
```{r}
eh_long$log <- log(eh_long$estimate)
hist(eh_long$log)
qqnorm(eh_long$log)
qqline(eh_long$log, col = "red")
shapiro.test(eh_long$log)

eh_long$sqrt <- sqrt(eh_long$estimate)
hist(eh_long$sqrt)
qqnorm(eh_long$sqrt)
qqline(eh_long$sqrt, col = "red")
shapiro.test(eh_long$sqrt)
```

Log transformation produced normality. 

Run linear mixed effects model 
```{r}
eh.model <- lmer(log ~ Year * Phenotype + (1|ColonyID), data = eh_long)
car::Anova(eh.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(eh.model, which = 1) 
lattice::qqmath(eh.model)

# Check plot to make sure results make sense
ggplot(eh_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

No significant differences. 

### Topt

Subset by Topt data
```{r}
topt_long <- traits_long_all_mcap %>%
  filter(trait == "Topt_C")
```

Check for normality
```{r}
hist(topt_long$estimate)

# Q-Q Plot
qqnorm(topt_long$estimate)
qqline(topt_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(topt_long$estimate)
```

Data is normal.

Run linear mixed effects model 
```{r}
topt.model <- lmer(estimate ~ Year * Phenotype + (1|ColonyID), data = topt_long)
car::Anova(topt.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(topt.model, which = 1) 
lattice::qqmath(topt.model)

# Check plot to make sure results make sense
ggplot(topt_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

No significant differences. 

### Rmax

Subset by E data
```{r}
rmax_long <- traits_long_all_mcap %>%
  filter(trait == "Rmax_log")
```

Check for normality
```{r}
hist(rmax_long$estimate)

# Q-Q Plot
qqnorm(rmax_long$estimate)
qqline(rmax_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(rmax_long$estimate)
```

Data is normal. 

Run linear mixed effects model 
```{r}
rmax.model <- lmer(estimate ~ Year * Phenotype + (1|ColonyID), data = rmax_long)
car::Anova(rmax.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(rmax.model, which = 1) 
lattice::qqmath(rmax.model)

# Check plot to make sure results make sense
ggplot(rmax_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

No significant differences. 

## PCOMP Statistics comparing 2023 v. 2025 TPC photosynthesis metrics 

Subset by Pcomp only 
```{r}
traits_long_all_pcomp <- traits_long_all %>%
  filter(Species == "Porites compressa")
```

### lnc (r_tref)

Subset by lnc data
```{r}
lnc_long <- traits_long_all_pcomp %>%
  filter(trait == "lnc")
```

Check for normality
```{r}
hist(lnc_long$estimate)

# Q-Q Plot
qqnorm(lnc_long$estimate)
qqline(lnc_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(lnc_long$estimate)
```

Data is not normal. Transform data to achieve normality 
```{r}
lnc_long$log <- log(lnc_long$estimate)
hist(lnc_long$log)
qqnorm(lnc_long$log)
qqline(lnc_long$log, col = "red")
shapiro.test(lnc_long$log)

lnc_long$sqrt <- sqrt(lnc_long$estimate)
hist(lnc_long$sqrt)
qqnorm(lnc_long$sqrt)
qqline(lnc_long$sqrt, col = "red")
shapiro.test(lnc_long$sqrt)
```

Log produced normality. 

Run linear mixed effects model 
```{r}
lnc.model <- lmer(log ~ Year * Phenotype + (1|ColonyID), data = lnc_long)
car::Anova(lnc.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(lnc.model, which = 1) 
lattice::qqmath(lnc.model)

# Post hoc tests 
emmeans(lnc.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(lnc_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year is significant.

### E

Subset by E data
```{r}
e_long <- traits_long_all_pcomp %>%
  filter(trait == "E")
```

Check for normality
```{r}
hist(e_long$estimate)

# Q-Q Plot
qqnorm(e_long$estimate)
qqline(e_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(e_long$estimate)
```

Data is not normal. Transform data to achieve normality 
```{r}
e_long$log <- log(e_long$estimate)
hist(e_long$log)
qqnorm(e_long$log)
qqline(e_long$log, col = "red")
shapiro.test(e_long$log)

e_long$sqrt <- sqrt(e_long$estimate)
hist(e_long$sqrt)
qqnorm(e_long$sqrt)
qqline(e_long$sqrt, col = "red")
shapiro.test(e_long$sqrt)
```

Square root transformation produced normality. 

Run linear mixed effects model 
```{r}
e.model <- lmer(sqrt ~ Year * Phenotype + (1|ColonyID), data = e_long)
car::Anova(e.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(e.model, which = 1) 
lattice::qqmath(e.model)

# Post hoc tests 
emmeans(lnc.model, pairwise ~ Phenotype)

# Check plot to make sure results make sense
ggplot(e_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Phenotype on the edge of significance. 

### Eh

Subset by E data
```{r}
eh_long <- traits_long_all_pcomp %>%
  filter(trait == "Eh")
```

Check for normality
```{r}
hist(eh_long$estimate)

# Q-Q Plot
qqnorm(eh_long$estimate)
qqline(eh_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(eh_long$estimate)
```

Data is not normal. Transform data to achieve normality 
```{r}
eh_long$log <- log(eh_long$estimate)
hist(eh_long$log)
qqnorm(eh_long$log)
qqline(eh_long$log, col = "red")
shapiro.test(eh_long$log)

eh_long$sqrt <- sqrt(eh_long$estimate)
hist(eh_long$sqrt)
qqnorm(eh_long$sqrt)
qqline(eh_long$sqrt, col = "red")
shapiro.test(eh_long$sqrt)
```

Run linear mixed effects model 
```{r}
eh.model <- lmer(log ~ Year * Phenotype + (1|ColonyID), data = eh_long)
car::Anova(eh.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(eh.model, which = 1) 
lattice::qqmath(eh.model)

# Check plot to make sure results make sense
ggplot(eh_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

No significant differences. 

### Topt

Subset by Topt data
```{r}
topt_long <- traits_long_all_pcomp %>%
  filter(trait == "Topt_C")
```

Check for normality
```{r}
hist(topt_long$estimate)

# Q-Q Plot
qqnorm(topt_long$estimate)
qqline(topt_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(topt_long$estimate)
```

Data is normal.

Run linear mixed effects model 
```{r}
topt.model <- lmer(estimate ~ Year * Phenotype + (1|ColonyID), data = topt_long)
car::Anova(topt.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(topt.model, which = 1) 
lattice::qqmath(topt.model)

# Check plot to make sure results make sense
ggplot(topt_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

No significant differences. 

### Rmax

Subset by E data
```{r}
rmax_long <- traits_long_all_pcomp %>%
  filter(trait == "Rmax_log")
```

Check for normality
```{r}
hist(rmax_long$estimate)

# Q-Q Plot
qqnorm(rmax_long$estimate)
qqline(rmax_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(rmax_long$estimate)
```

Data is not normal. Transform data to achieve normality 
```{r}
rmax_long$log <- log(rmax_long$estimate)
hist(rmax_long$log)
qqnorm(rmax_long$log)
qqline(rmax_long$log, col = "red")
shapiro.test(rmax_long$log)

rmax_long$sqrt <- sqrt(rmax_long$estimate)
hist(rmax_long$sqrt)
qqnorm(rmax_long$sqrt)
qqline(rmax_long$sqrt, col = "red")
shapiro.test(rmax_long$sqrt)
```

Sqrt produces normality. 

Run linear mixed effects model 
```{r}
rmax.model <- lmer(estimate ~ Year * Phenotype + (1|ColonyID), data = rmax_long)
car::Anova(rmax.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(rmax.model, which = 1) 
lattice::qqmath(rmax.model)

# Post hoc tests 
emmeans(rmax.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(rmax_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year is significant. 

## Visualize 2023 v. 2025 TPC calcification 

Visualize differences between 2023 and 2025
```{r}
ggplot(traits_long_all, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  # 1. Add Error Bars (Standard Error)
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, size = 1) +
  
  # 2. Add Lines to connect the means over time
  stat_summary(fun = mean, geom = "line", size = 1.2) +
  
  # 3. Add the Mean Points
  stat_summary(fun = mean, geom = "point", size = 4) +
  
  # 4. Facet by Trait and Species
  # Using 'free_y' because Topt (~30) and E (~0.5) have very different scales
  facet_grid(trait ~ Species, scales = "free_y") +
  
  # 5. Styling and Colors
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 10),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  labs(
    title = "Trait Evolution: 2023 vs 2025",
    y = "Estimated Value (Mean ± SE)",
    x = "Year"
  )
```

Make the plot pretty
```{r}
# Define a mapping for the Y-axis labels
trait_labels <- c(
  "E" = "Activation~Energy~(eV)",
  "Eh" = "Deactivation~Energy~(eV)",
  "lnc" = "atop(Net~calcification~at~27*degree*C, log(mu*mol~hr^-1~cm^-2))",
  "Rmax_log" = "atop(Maximum~calcification, log(mu*mol~hr^-1~cm^-2))",
  "Topt_C" = "Thermal~optimum~(degree*C)"
)

# Define the traits vector in the order you want them plotted
traits <- c("E", "Eh", "lnc", "Topt_C", "Rmax_log")

## Spp in same plot 
create_combined_trait_plot <- function(data, tr, plot_letter, is_bottom_row = FALSE) {
  
  # Filter data for the specific trait but keep both species
  plot_data <- data %>% filter(trait == tr)
  
  p <- ggplot(plot_data, aes(x = Year, y = estimate, 
                             color = Phenotype, 
                             linetype = Species, 
                             group = interaction(Phenotype, Species))) +
    # Error bars (Species-specific linetypes can apply here too if desired)
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, size = 0.6, linetype = "solid") +
    stat_summary(fun = mean, geom = "line", size = 0.8) +
    stat_summary(fun = mean, geom = "point", size = 2.5) +
    # Custom Scales
    scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
    scale_linetype_manual(values = c("Montipora capitata" = "solid", 
                                     "Porites compressa" = "dashed")) +
    theme_classic() + 
    theme(
      axis.text = element_text(size = 9, color = "black"),
      axis.title.y = element_text(size = 9, face = "plain"),
      plot.title = element_text(size = 10, face = "bold", hjust = 0)
    ) +
    labs(title = plot_letter, x = NULL, y = trait_labels[[tr]])

  # Hide X-axis for all but the bottom plot
  if (!is_bottom_row) {
    p <- p + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  }
  
  return(p)
}

# Generate the single column of combined plots
combined_trait_plots <- lapply(1:5, function(i) {
  create_combined_trait_plot(traits_long_all, traits[i], 
                             LETTERS[i], is_bottom_row = (i == 5))
})

# Assemble the vertical stack
final_trait_figure <- wrap_plots(combined_trait_plots, ncol = 1) + 
  plot_layout(guides = "collect") +
  plot_annotation(
    theme = theme(plot.title = element_text(size = 14, face = "bold"))
  ) & 
  theme(legend.position = "right",
        legend.box = "vertical") # Stacks Phenotype and Species legends
final_trait_figure

# Save 
ggsave("../output/TPC_Calcification/Calcification_TPC_metrics_2023_2025.pdf", 
       plot = final_trait_figure, 
       width = 8, 
       height = 12, 
       units = "in", 
       device = "pdf",
       dpi = 300)
ggsave("../output/TPC_Calcification/Calcification_TPC_metrics_2023_2025.png", 
       plot = final_trait_figure, 
       width = 8, 
       height = 12, 
       units = "in", 
       dpi = 300)
```

Add significance data to plots
```{r}
# Add labels 
sig_labels <- c(
  "S***, PxY*",                 # for E
  "n.s.",       # for Eh
  "S***, Y*",         # for tref
  "n.s.", # for Topt_C
  "S***, Y***, SxY*"                     # for Rmax_log 
)

create_combined_trait_plot <- function(data, tr, plot_letter, sig_text = "", is_bottom_row = FALSE) {
  
  plot_data <- data %>% filter(trait == tr)
  p <- ggplot(plot_data, aes(x = Year, y = estimate, 
                               color = Phenotype, 
                               linetype = Species, 
                               group = interaction(Phenotype, Species))) +
     # Error bars - slightly thicker for publication
     stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, size = 0.8, linetype = "solid") +
     stat_summary(fun = mean, geom = "line", size = 1.0) +
     stat_summary(fun = mean, geom = "point", size = 3.5) +
     
     # SIGNIFICANCE LABELS - Positioned in bottom right
     annotate("text", x = Inf, y = -Inf, label = sig_text, 
              hjust = 1.1, vjust = -1.2, size = 4, color = "black", face = "plain") +
     
     scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
     scale_linetype_manual(values = c("Montipora capitata" = "solid", 
                                      "Porites compressa" = "dashed")) +
     theme_classic() + 
     theme(
       # Larger text for readability in print
       axis.text = element_text(size = 15, color = "black"),
       axis.title.y = element_text(size = 15, face = "plain", margin = margin(r = 10)),
       plot.title = element_text(size = 18, face = "bold", hjust = 0),
       # Ensure panels have enough breathing room
       plot.margin = margin(t = 5, r = 5, b = 5, l = 5)
     ) +
     labs(title = plot_letter, x = NULL, y = parse(text = trait_labels[[tr]]))

  if (!is_bottom_row) {
    p <- p + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  } else {
    # Bold the year labels on the bottom row
    p <- p + theme(axis.text.x = element_text(size = 18))
  }
  
  return(p)
}

# Define traits and your updated sig labels
traits <- c("E", "Eh", "lnc", "Topt_C", "Rmax_log")

# Apply the function across all traits
combined_trait_plots <- lapply(1:5, function(i) {
  create_combined_trait_plot(
    data = traits_long_all, 
    tr = traits[i], 
    plot_letter = LETTERS[i], 
    sig_text = sig_labels[i],
    is_bottom_row = (i == 5)
  )
})

# Assemble with Patchwork
final_trait_figure <- wrap_plots(combined_trait_plots, ncol = 1) + 
  plot_layout(guides = "collect") +
  plot_annotation(
    # Optional: add an overall title or leave NULL for publication
    title = NULL 
  ) & 
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11),
    legend.key.width = unit(1.5, "cm"), # Makes linetypes easier to distinguish
    legend.spacing.y = unit(0.5, "cm")
  )

# Save 
ggsave("../output/TPC_Calcification/Calcification_TPC_metrics_stats_2023_2025.pdf", 
       plot = final_trait_figure, 
       width = 10, 
       height = 16, 
       units = "in", 
       device = "pdf",
       dpi = 300)
ggsave("../output/TPC_Calcification/Calcification_TPC_metrics_stats_2023_2025.png", 
       plot = final_trait_figure, 
       width = 10, 
       height = 16, 
       units = "in", 
       dpi = 300)
```

Plot with species as separate columns
```{r}
# 1. Significance Labels
sig_data <- data.frame(
  trait = rep(c("E", "Eh", "lnc", "Topt_C", "Rmax_log"), each = 2),
  Species = rep(c("Montipora capitata", "Porites compressa"), 5),
  sig_label = c(
    "PxY*", "P*",      # E (A, B)
    "n.s.", "n.s.",    # Eh (C, D)
    "n.s.", "Y*",      # lnc (E, F)
    "n.s.", "n.s.",    # Topt_C (G, H)
    "n.s.", "Y*"       # Rmax_log (I, J)
  )
)

trait_labels <- c(
  "E" = "Activation~Energy~(eV)",
  "Eh" = "Deactivation~Energy~(eV)",
  "lnc" = "atop(Net~calcification~at~27*degree*C, log(mu*mol~hr^-1~cm^-2))",
  "Rmax_log" = "atop(Maximum~calcification, log(mu*mol~hr^-1~cm^-2))",
  "Topt_C" = "Thermal~optimum~(degree*C)"
)

# 2. Calculate Shared Y-Limits
trait_limits_df <- traits_long_all %>%
  as.data.frame() %>%
  dplyr::group_by(trait, Species, Year, Phenotype) %>%
  dplyr::summarise(
    m = mean(estimate, na.rm = TRUE),
    se = sd(estimate, na.rm = TRUE) / sqrt(n()),
    .groups = 'drop'
  ) %>%
  dplyr::group_by(trait) %>%
  dplyr::summarise(
    ymin = min(m - se, na.rm = TRUE),
    ymax = max(m + se, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  dplyr::mutate(ymax = ymax + (ymax - ymin) * 0.25, 
                ymin = ymin - (ymax - ymin) * 0.1)

# 3. Tag list
tag_list <- c('A','B','C','D','E','F','G','H','I','J')

# 4. Panel function
create_panel_plot <- function(data, tr, spp, ylimits, is_bottom_row = FALSE, panel_tag = NULL) {
  
  plot_data <- data %>% as.data.frame() %>% 
    dplyr::filter(trait == tr, Species == spp)
  sig_text <- sig_data %>% 
    dplyr::filter(trait == tr, Species == spp) %>% 
    dplyr::pull(sig_label)
  
  y_label <- parse(text = paste0("bold(", trait_labels[[tr]], ")"))

  p <- ggplot(plot_data, aes(x = factor(Year), y = estimate, 
                              color = Phenotype, group = Phenotype)) +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.15, linewidth = 1.2) +
    stat_summary(fun = mean, geom = "line", linewidth = 1.5) +
    stat_summary(fun = mean, geom = "point", size = 5) +
    annotate("text", x = Inf, y = -Inf, label = sig_text, 
             hjust = 1.1, vjust = -1.2, size = 7, color = "black", fontface = "plain") +
    scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
    theme_classic() + 
    theme(
      axis.text.y   = element_text(size = 24, color = "black"), 
      axis.text.x   = element_text(size = 24, color = "black", face = "bold"),
      axis.title.y  = element_text(size = 27, face = "bold"),
      plot.margin   = margin(t = 30, r = 15, b = 15, l = 15),
      plot.tag      = element_text(size = 30, face = "bold"),
      plot.tag.position = "topleft"
    ) +
    labs(x = NULL, y = y_label, tag = panel_tag) +
    coord_cartesian(ylim = ylimits)

  if (!is_bottom_row) {
    p <- p + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  }
  
  return(p)
}

# 5. Generate plots
# NOTE: outer loop = species, inner loop = traits (gives A-E top row, F-J bottom row)
all_plots <- list()
species_list <- c("Montipora capitata", "Porites compressa")
traits <- c("E", "Eh", "lnc", "Topt_C", "Rmax_log")
tag_list <- c('A','B','C','D','E','F','G','H','I','J') # 10 tags for 10 plots
counter <- 1

for (j in 1:length(species_list)) { 
  for (i in 1:length(traits)) {    
    current_limits <- trait_limits_df %>% 
      dplyr::filter(trait == traits[i]) %>% 
      dplyr::select(ymin, ymax) %>% 
      as.numeric()
      
    all_plots[[counter]] <- create_panel_plot(
      data = traits_long_all, 
      tr = traits[i], 
      spp = species_list[j],
      ylimits = current_limits, 
      is_bottom_row = (j == 2),
      panel_tag = tag_list[counter]
    )
    counter <- counter + 1
  }
}

# 6. Species labels 
mcap_header <- wrap_elements(full = grid::textGrob("Montipora capitata", 
                               x = unit(0.01, "npc"), 
                               hjust = 0,
                               gp = gpar(fontsize = 30, fontface = "bold.italic")))
                             
pcomp_header <- wrap_elements(full = grid::textGrob("Porites compressa", 
                                x = unit(0.01, "npc"), 
                                hjust = 0,
                                gp = gpar(fontsize = 30, fontface = "bold.italic")))

# 7. Assemble
final_trait_figure <- (
  mcap_header / 
  (all_plots[[1]] | all_plots[[2]] | all_plots[[3]] | all_plots[[4]] | all_plots[[5]]) /
  pcomp_header /
  (all_plots[[6]] | all_plots[[7]] | all_plots[[8]] | all_plots[[9]] | all_plots[[10]])
) + 
plot_layout(heights = c(0.12, 1, 0.12, 1), guides = "collect") & 
theme(
  legend.position = "right", # or "bottom" depending on your preference
  legend.title = element_text(size = 20, face = "bold"),
  legend.text = element_text(size = 18), 
  legend.key.size = unit(1.5, "cm"),
  plot.margin = margin(t = 10, r = 10, b = 10, l = 10) 
)

# Save
ggsave("../output/TPC_Calcification/Calcification_TPC_metrics_stats_separate_spp_2023_2025.pdf", 
       plot = final_trait_figure, width = 30, height = 14, units = "in", device = "pdf", dpi = 300)
ggsave("../output/TPC_Calcification/Calcification_TPC_metrics_stats_separate_spp_2023_2025.png", 
       plot = final_trait_figure, width = 30, height = 14, units = "in", dpi = 300)
```

Plot TPC calcification curves with temp bars 
```{r}
# Select specific cols from each df 
calc_2023 <- calc_2023 %>%
  dplyr::select(Colony.ID, Target.Temperature, umol.CaCO3h.1.cm.2, Species, Phenotype, log.rate, Target.Temperature.K) %>%
  dplyr::rename(Target_Temperature = Target.Temperature,
         Net_Calcification_cm2 = umol.CaCO3h.1.cm.2) %>%
  mutate(Colony.ID = as.character(Colony.ID))

# 1. Combine all data and create a unified "Header" for each plot
all_calc <- bind_rows(
  calc_2023 %>% mutate(Year = "2023"),
  calc_2025 %>% mutate(Year = "2025")
) %>% mutate(Plot_Header = paste(Species, Year))

all_preds <- bind_rows(
  preds_mean_summary_2023 %>% mutate(Year = "2023"),
  preds_mean_summary_2025 %>% mutate(Year = "2025")
) %>% mutate(Plot_Header = paste(Species, Year))

species_list <- c("Montipora capitata", "Porites compressa")

# Rebuild all_freq by duplicating the year data for each species
all_freq <- bind_rows(
  # For 2023
  map_dfr(species_list, ~temp_freq_2023 %>% mutate(Species = .x, Year = "2023")),
  # For 2025
  map_dfr(species_list, ~temp_freq_2025 %>% mutate(Year = "2025", Species = .x))
) %>% 
  mutate(Plot_Header = paste(Species, Year))

# Ensure all_photo and all_preds also have the Plot_Header
all_calc <- all_calc %>% mutate(Plot_Header = paste(Species, Year))
all_preds <- all_preds %>% mutate(Plot_Header = paste(Species, Year))

# 2. Axis Parameters
p_min <- -0.5; p_max <- 0.75; p_total <- p_max - p_min
s_max <- 30; scale_factor <- p_total / s_max

# 3. Create the Plot
final_tpc_plot <- ggplot() +
  # Environmental bars
  geom_rect(data = all_freq, 
            aes(xmin = temp_bin - 0.45, xmax = temp_bin + 0.45, 
                ymin = p_min, ymax = p_min + (percent * scale_factor)), 
            fill = "#8DBBCC", color = "#727E8B", linewidth = 0.2) +
  # Raw data points
  geom_point(data = all_calc, 
             aes(x = Target.Temperature.K - 273.15, y = log.rate, color = Phenotype), 
             alpha = 0.4, size = 1.5, position = position_jitter(width = 0.1)) +
  # Confidence ribbons
  geom_ribbon(data = all_preds, 
              aes(x = Target.Temperature.K - 273.15, ymin = lower, ymax = upper, fill = Phenotype), 
              alpha = 0.3, color = NA) +
  # Mean lines
  geom_line(data = all_preds, 
            aes(x = Target.Temperature.K - 273.15, y = mean_val, color = Phenotype), 
            linewidth = 1.2) +
  geom_hline(yintercept = 0, color = "grey", linewidth = 0.5, linetype = "dashed") +
  
  # Scaling and Dual Axes
  scale_y_continuous(
    name = expression("Net calcification (log(" *mu*"mol " *cm^-2 *hr^-1*"))"),
    limits = c(p_min, p_max), expand = c(0, 0),
    sec.axis = sec_axis(~ (. - p_min) / scale_factor, name = "Percent days (%)", breaks = seq(0, 30, 5))
  ) +
  scale_x_continuous(breaks = seq(18, 40, by = 2), limits = c(17.5, 40.5), expand = c(0,0)) +
  
  # FACET: This creates 4 distinct panels with your custom headers
  facet_wrap(~ Plot_Header, ncol = 2) +
  
  # Styling
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_classic() +
theme_classic() +
  theme(
    # --- INCREASED TEXT SIZES ---
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14),
    
    # Secondary axis styling (matches the blue bars)
    axis.title.y.right = element_text(color = "black", size = 13, margin = margin(l = 15)),
    axis.text.y.right = element_text(color = "black", size = 11),
    
    # Facet header styling
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 14, hjust = 0),
    
    # Legend styling
    legend.position = "top",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 13, face = "bold"),
    
    # Spacing
    panel.spacing = unit(2, "lines"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
  ) +
  labs(x = "Temperature (°C)")

final_labeled_tpc <- final_tpc_plot + 
  #plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(face = "bold", size = 16))

final_labeled_tpc

# Save 
ggsave("../output/TPC_Calcification/Calcification_TPC_curve_temp_bars_2023_2025.pdf", 
       plot = final_labeled_tpc, 
       width = 12, 
       height = 8, 
       units = "in", 
       device = "pdf",
       dpi = 300)
ggsave("../output/TPC_Calcification/Calcification_TPC_curve_temp_bars_2023_2025.png", 
       plot = final_labeled_tpc, 
       width = 12, 
       height = 8, 
       units = "in", 
       dpi = 300)
```

Plot curves w/o temp bars 
```{r}
# 1. Axis Parameters (Adjusted for single axis)
p_min <- -0.5; p_max <- 0.75

# 2. Create the Plot
final_tpc_plot <- ggplot() +
  # Raw data points
  geom_point(data = all_calc, 
             aes(x = Target.Temperature.K - 273.15, y = log.rate, color = Phenotype), 
             alpha = 0.4, size = 1.5, position = position_jitter(width = 0.1)) +
  # Confidence ribbons
  geom_ribbon(data = all_preds, 
              aes(x = Target.Temperature.K - 273.15, ymin = lower, ymax = upper, fill = Phenotype), 
              alpha = 0.3, color = NA) +
  # Mean lines
  geom_line(data = all_preds, 
            aes(x = Target.Temperature.K - 273.15, y = mean_val, color = Phenotype), 
            linewidth = 1.2) +
  geom_hline(yintercept = 0, color = "grey", linewidth = 0.5, linetype = "dashed") +
  
  # Simplified Single Axis
  scale_y_continuous(
    name = expression(bold("Net calcification (log(" *mu*"mol " *cm^-2 *hr^-1*"))")),
    limits = c(p_min, p_max), 
    expand = c(0, 0)
  ) +
  scale_x_continuous(breaks = seq(18, 40, by = 2), limits = c(17.5, 40.5), expand = c(0,0)) +
  
  # FACET: Using ggtext or a simple expression trick for italics is complex in facet_wrap.
  # The easiest way for publication is to italicize the whole strip or use labeller.
  facet_wrap(~ Plot_Header, ncol = 2) +
  
  # Styling
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_classic() +
  theme(
    axis.text = element_text(size = 16, color = "black"),
    axis.title = element_text(size = 18, face = "bold"),
    
    # Facet header styling - Italics applied here
    strip.background = element_blank(),
    strip.text = element_text(face = "italic", size = 18, hjust = 0),
    
    legend.position = "top",
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 15, face = "bold"),
    
    panel.spacing = unit(2, "lines"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
  ) +
  labs(x = "Temperature (°C)")

# If you ONLY want the Species name italicized and not the Year:
# Use element_markdown from library(ggtext) and wrap the species in "*"
library(ggtext)
 final_tpc_plot <- final_tpc_plot + 
   theme(strip.text = element_markdown())

final_labeled_tpc <- final_tpc_plot + 
  theme(plot.tag = element_text(face = "bold", size = 20))

final_labeled_tpc

# Save 
ggsave("../output/TPC_Calcification/Calcification_TPC_curve_2023_2025.pdf", 
       plot = final_labeled_tpc, 
       width = 12, 
       height = 8, 
       units = "in", 
       device = "pdf",
       dpi = 300)
ggsave("../output/TPC_Calcification/Calcification_TPC_curve_2023_2025.png", 
       plot = final_labeled_tpc, 
       width = 12, 
       height = 8, 
       units = "in", 
       dpi = 300)
```

