---
title: "TPCs"
author: "Jill Ashey"
date: "2025-09-15"
output: html_document
---

Interesting: https://www.nature.com/articles/s41467-024-53046-2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(car)
library(tidyverse)
library(broom)
library(readr)
library(lubridate)
library(purrr)
library(furrr)
library(fuzzyjoin)
library(future)
library(rTPC)
library(nls.multstart)
library(emmeans)
library(lme4)
library(nlme)
library(Rmisc)
library(boot)
library(nlme)
library(propagate)
library(patchwork)
library(lmerTest)
library(performance)
library(insight)
```

# Functions for analysis 

```{r}
sharpe_schoolfield <- function(lnc, E, Eh, Th, temp, Tc) {
  Tc <- 273.15 + Tc
  k <- 8.62e-5
  boltzmann.term <- lnc + log(exp(E/k*(1/Tc - 1/temp))) # units are eV/K, electrovolts/Kelvin
  inactivation.term <- log(1/(1 + exp(Eh/k*(1/Th - 1/temp))))
  return(boltzmann.term + inactivation.term)}

get_topt <- function(E, Th, Eh){
  k <- 8.62e-05
  return((Eh * Th) / (Eh + (k * Th * log((Eh / E) - 1))))}

get_tpc_ci <- function(model, new_data) {
  # predictNLS calculates the intervals based on the Taylor expansion (Delta Method)
  pred <- predictNLS(model, newdata = new_data, interval = "confidence", alpha = 0.05)
  return(as.data.frame(pred$summary))}
```

Run this before running any ANOVAs
```{r}
options(contrasts = c("contr.sum", "contr.poly"))
```

# Photosynthesis
## Sept 2023

### Prepare data 

Load run metadata 
```{r}
run_meta <- read_csv("../data/TPC/TPC_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
```

Load sample metadata 
```{r}
sample_meta <- read_csv("../data/TPC/TPC_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```

List all raw files for Sept 2023 
```{r}
files <- list.files(path = "../data/TPC_Photosynthesis/raw_data/Sept2023/", pattern = "*.csv", full.names = TRUE)
```

Join O2 data with metadata files based on start and stop times 
```{r}
all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "channel\\d+") %>% str_remove("channel") %>% as.numeric()
  
  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )
  
  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )
  
  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )
  
  return(raw_joined)
})
```

Look at joined data and save
```{r}
all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

write_csv(all_data, "../output/TPC_Photosynthesis/Sept2023/TPC_all_joined_data_Sept2023.csv")
```

### Run localized linear regressions 

The interval fitting code will take about an hour to run and is computationally intense. If needed, run as a background job or on a remote server.

Make list of all Colony_ID / Run combinations (including blanks)
```{r}
colony_runs <- all_data %>%
  filter(!is.na(Colony_ID)) %>%
  distinct(Colony_ID, Run)
```

For this timepoint, only two out of the five runs had blanks. Therefore, I am going to calculate the blank intervals, average them, and then apply them to the regression for each colony. 

Calculate interval slopes for the blanks
```{r}
blank_data <- all_data %>%
  filter(Colony_ID == "blank")

blank_intervals <- blank_data %>%
  group_by(Run, Temperature.y) %>%
  filter(!is.na(Colony_ID)) %>%
  group_split()
```

Fit regressions and extract slopes for each blank interval
```{r}
fit_reg <- function(df) {
  rankLocReg(
    xall = as.numeric(df$`Delta T [min]`),
    yall = df$Oxygen,
    alpha = 0.2,
    method = "pc",
    verbose = FALSE
  )
}

blank_fits <- map(blank_intervals, fit_reg)
blank_slopes <- map_dbl(blank_fits, ~ pluck(., "allRegs", "b1", 1))
blank_summary <- map2_dfr(
  blank_intervals,
  blank_slopes,
  ~ tibble(
    Temperature.y = unique(.x$Temperature.y),
    Blank_Slope = .y
  )
)
```

Average blanks
```{r}
avg_blank_slopes <- blank_summary %>%
  group_by(Temperature.y) %>%
  summarise(Avg_Blank_Slope = mean(Blank_Slope, na.rm = TRUE), .groups = "drop") %>%
  rename("Temperature" = "Temperature.y")
```

Fit regressions for each interval 
```{r}
fit_colony_intervals_avgblank <- function(colony_id, run_num) {
  # Subset sample data
  sample_data <- thinned_data %>%
    filter(Colony_ID == colony_id, Run == run_num)
  if (nrow(sample_data) == 0) return(NULL)
  
  # Split data by light step
  intervals <- sample_data %>%
    group_by(Temperature.y) %>%
    group_split()
  
  # Fit regressions for each interval
  fits <- map(intervals, fit_reg)
  micromol.L.s <- map_dbl(fits, ~ pluck(., "allRegs", "b1", 1))
  
  # Detailed interval summary
  interval_df <- map2_dfr(
    intervals,
    seq_along(intervals),
    ~ tibble(
      Colony_ID   = colony_id,
      Run         = run_num,
      Interval    = .y,
      Temperature = unique(.x$Temperature.y),
      N_points    = nrow(.x),
      O2_start    = head(.x$Oxygen, 1),
      O2_end      = tail(.x$Oxygen, 1),
      DeltaT_min  = min(.x$`Delta T [min]`),
      DeltaT_max  = max(.x$`Delta T [min]`),
      O2_min      = min(.x$Oxygen, na.rm = TRUE),
      O2_max      = max(.x$Oxygen, na.rm = TRUE),
      Slope       = micromol.L.s[.y]
    )
  ) %>%
    left_join(avg_blank_slopes, by = "Temperature") %>%
    mutate(
      Blank_Corrected_Slope = Slope - Avg_Blank_Slope
    )
  
  # Write file (outside of the pipeline!)
  fn <- paste0("../output/TPC_Photosynthesis/Sept2023/intervals_", colony_id, "_run", run_num, "_Sept2023_TPC.csv")
  write_csv(interval_df, fn)
  
  interval_df
}

# Run for all samples and save big CSV
plan(multisession)

all_interval_summaries <- future_pmap_dfr(
  list(colony_runs$Colony_ID, colony_runs$Run),
  fit_colony_intervals_avgblank
)

write_csv(all_interval_summaries, "../output/TPC_Photosynthesis/Sept2023/all_interval_summaries_Sept2023_TPC.csv")
```

Read in interval data (so I don't have to run again)
```{r}
all_int_tpc <- read_csv("../output/TPC_Photosynthesis/Sept2023/all_interval_summaries_Sept2023_TPC.csv")
all_int_tpc$Colony_ID <- as.character(all_int_tpc$Colony_ID)
```

Filter sample metadata for Sept 2023
```{r}
sept_2023 <- sample_meta %>%
  filter(grepl("^2023-09", Date))
sept_2023$Surface_Area_cm2 <- as.numeric(sept_2023$Surface_Area_cm2)
```

Merge sample metadata with interval data 
```{r}
all_int_tpc <- all_int_tpc %>%
  inner_join(sept_2023, by = c("Colony_ID" = "Colony_ID", "Temperature" = "Target_Temperature")) %>%
  filter(!is.na(Phenotype))
length(unique(all_int_tpc$Colony_ID))
```

Normalize interval data by chamber volume and surface area and save
```{r}
# normalize 
all_int_tpc <- all_int_tpc %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Blank_Corrected_Slope * Chamber_Volume_L / Surface_Area_cm2) * 60) # units are umol O2 cm-2 h-1
write.csv(all_int_tpc, "../output/TPC_Photosynthesis/Sept2023/all_interval_summaries_normalized_Sept2023_TPC.csv")
```

Read data back in so I don't have to run above code again
```{r}
all_int_tpc <- read.csv("../output/TPC_Photosynthesis/Sept2023/all_interval_summaries_normalized_Sept2023_TPC.csv")
```

Plot curves across temperatures
```{r}
ggplot(all_int_tpc, aes(x = Temperature, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "umol O2 cm-2 h-1", color = "Phenotype")

ggplot(all_int_tpc, aes(x = Temperature, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(aes(fill = Phenotype), method = "loess", se = TRUE, linewidth = 1.2) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "umol O2 cm-2 h-1", color = "Phenotype")
```

Log and plot 
```{r}
all_int_tpc$log.rate <- log(all_int_tpc$Slope_norm + 1)  #logging and adding 0.1 because a log of zero does not exist

ggplot(all_int_tpc, aes(x = Temperature, y = log.rate, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "umol O2 cm-2 h-1", color = "Phenotype")

ggplot(all_int_tpc, aes(x = Temperature, y = log.rate, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(aes(fill = Phenotype), method = "loess", se = TRUE, linewidth = 1.2) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "log(umol O2 cm-2 h-1)", color = "Phenotype")
```

### Fit TPC curve

Set target temperature to numeric and convert temperature to Kelvin
```{r}
photo_2023 <- all_int_tpc %>%
  mutate(Temperature = as.numeric(Temperature),
         Temperature.K = Temperature + 273.15)
```

Fit curves for all colonies using `sharpe_schoolfield` function defined above
```{r}
fits_2023 <- photo_2023 %>%
  group_by(Phenotype,Species,Colony_ID) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(log.rate ~ sharpe_schoolfield(lnc, E, Eh, Th, temp = Temperature.K, Tc = 27),
                                                data = .x,
                                                iter = 1000,
                                                start_lower = c(lnc = -10, E = 0.1, Eh = 0.2, Th = 285),
                                                start_upper = c(lnc = 10, E = 2, Eh = 5, Th = 330),
                                                supp_errors = 'Y',
                                                na.action = na.omit,
                                                lower = c(lnc = -10, E = 0, Eh = 0, Th = 0))))
```

Get TPC parameters for each colony 
```{r}
params_2023 <- fits_2023 %>%
  unnest_legacy(fit %>% purrr::map(tidy)) %>%
  filter(!Colony_ID == "233")
```

Calculate Topt and Rmax for each colony 
```{r}
Topt_Rmax_2023 <- params_2023 %>%
  dplyr::select(Colony_ID, Phenotype, Species, term, estimate) %>%
  tidyr::pivot_wider(names_from = term, values_from = estimate) %>%
  mutate(
    Topt_K = get_topt(E, Th, Eh),
    Topt_C = Topt_K - 273.15,
    Rmax_log = sharpe_schoolfield(lnc, E, Eh, Th, temp = Topt_K, Tc = 27),
    Rmax_linear = exp(Rmax_log) - 1)
```

Pivot data so that each trait is a row 
```{r}
traits_long_2023 <- Topt_Rmax_2023 %>%
  dplyr::select(Colony_ID, Phenotype, Species, lnc, E, Eh, Topt_C, Rmax_log) %>%
  pivot_longer(cols = c(lnc, E, Eh, Topt_C, Rmax_log), 
               names_to = "trait", 
               values_to = "estimate")
```

### Visualization for 2023

Plot
```{r}
# Create the point plot with Mean and SE
ggplot(traits_long_2023, aes(x = Phenotype, y = estimate, color = Phenotype)) +
  # Add raw colony data as small, transparent points to show the spread
  geom_jitter(width = 0.1, alpha = 0.3, size = 1.5) +
  stat_summary(fun = mean, geom = "point", size = 4) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 1) +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "2023 Thermal Performance Traits",
    y = "Estimate (Mean ± SE)",
    x = "Phenotype"
  )

## Box plot 
ggplot(traits_long_2023, aes(x = Phenotype, y = estimate, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.6, color = "black") +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.2, color = "black") +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.major.x = element_blank() # Clean up vertical lines
  ) +
  labs(
    y = "Estimated Parameter Value",
    x = "Phenotype"
  )
```

### Plot with confidence intervals 

Generate and aggregate predictions
```{r}
# 1. Define the temperature range for the smooth curves
new_data <- data.frame(
  Temperature.K = seq(18 + 273.15 - 1,  # 17°C
                      38 + 273.15 + 1,  # 39°C
                      length.out = 200))

# 2. Generate predictions for every colony model
# This creates a "long" dataframe of all potential curves
preds_all_2023 <- fits_2023 %>%
  mutate(preds = map(fit, ~{
    # We use a safe predict call in case some models didn't converge
    p <- try(predict(.x, newdata = new_data), silent = TRUE)
    if(inherits(p, "try-error")) return(NULL)
    data.frame(Temperature.K = new_data$Temperature.K, pred_val = p)
  })) %>%
  filter(!map_lgl(preds, is.null)) %>% # Remove failed fits
  unnest(preds)

# 3. Calculate the MEAN and CI of the curves per group
# At each temperature step, find the mean performance across all colonies
preds_mean_summary_2023 <- preds_all_2023 %>%
  group_by(Phenotype, Species, Temperature.K) %>%
  dplyr::summarise(
    mean_val = mean(pred_val, na.rm = TRUE),
    sd_val = sd(pred_val, na.rm = TRUE),
    n = n(),
    se_val = sd_val / sqrt(n),
    # 95% Confidence Interval of the Mean
    lower = mean_val - (1.96 * se_val),
    upper = mean_val + (1.96 * se_val),
    .groups = "drop"
  )
```

Plot TPC with fitted curves 
```{r}
ggplot(preds_mean_summary_2023, aes(x = Temperature.K - 273.15, y = mean_val, color = Phenotype, fill = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5, color = NA) +
  geom_line(linewidth = 1.5) +
  scale_x_continuous(
    breaks = seq(18, 40, by = 2), 
    limits = c(17, 40)) +
  geom_point(data = photo_2023, aes(x = Temperature.K - 273.15, y = Slope_norm), 
             alpha = 0.2, size = 1, inherit.aes = TRUE) +
  facet_wrap(~ Species) +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 12),
    panel.grid = element_blank()
  ) +
  labs(
    x = "Temperature (°C)", 
    y = expression(bold("Net photosynthesis (log(" *mu*"mol " *cm^-2 *hr^-1*"))")),
    title = "Photosynthesis Fitted Thermal Performance Curves (2023)"
  )
```

### Plot with confidence intervals and the environmental data 

```{r}
temps <- read.csv("../output/environmental/NSF_CAREER_KBay_PR13_Temps.csv")
```

Subset for the year prior to Sept 23, 2023
```{r}
temps_sub <- temps %>%
  filter(DateTime <= "2023-09-23 01:00:00")
```

```{r}
temp_freq_2023 <- temps_sub %>%
  # Rounding to the nearest whole degree centers the bar on that number
  mutate(temp_bin = round(Temp_C, 0)) %>% 
  group_by(temp_bin) %>%
  dplyr::summarise(count = n(), .groups = "drop") %>%
  # Calculate the percentage of total time spent at that temperature
  mutate(percent = (count / sum(count)) * 100)
```

Plot with temp bars
```{r}
# Axis Parameters
p_min <- -0.5  # Hard floor of the plot
p_max <- 1.5   # Peak of the Y-axis
p_total <- p_max - p_min
s_max <- 30    # Max percent on the right axis
scale_factor <- p_total / s_max

ggplot() +
  geom_rect(data = temp_freq_2023, 
            aes(xmin = temp_bin - 0.45, xmax = temp_bin + 0.45, 
                ymin = p_min, ymax = p_min + (percent * scale_factor)), 
            fill = "#8DBBCC", color = "#727E8B", linewidth = 0.2) +
  geom_point(data = photo_2023, 
             aes(x = Temperature.K - 273.15, y = log.rate, color = Phenotype), 
             alpha = 0.5, size = 1.2, position = position_jitter(width = 0.1)) +
  geom_ribbon(data = preds_mean_summary_2023, 
              aes(x = Temperature.K - 273.15, ymin = lower, ymax = upper, fill = Phenotype), 
              alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, color = "grey", linewidth = 0.5, linetype = "dashed") +
  geom_line(data = preds_mean_summary_2023, 
            aes(x = Temperature.K - 273.15, y = mean_val, color = Phenotype), 
            linewidth = 1.3) +
  scale_y_continuous(
    name = expression("Net photosynthesis (log(" *mu*"mol " *cm^-2 *hr^-1*"))"),
    limits = c(p_min, p_max),
    expand = c(0, 0),
    sec.axis = sec_axis(~ (. - p_min) / scale_factor, name = "Percent days (%)", breaks = seq(0, 40, 5))
  ) +
  scale_x_continuous(breaks = seq(18, 40, by = 2), limits = c(17.5, 40.5), expand = c(0,0)) +
  facet_wrap(~ Species, nrow = 1) +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_classic() +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 13),
    axis.line = element_line(color = "black"),
    panel.spacing = unit(2, "lines")
  ) +
  labs(x = "Temperature (°C)")
```

## Oct 2025

### Prepare data 

Load run metadata 
```{r}
run_meta <- read_csv("../data/TPC/TPC_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
```

Load sample metadata 
```{r}
sample_meta <- read_csv("../data/TPC/TPC_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```

List all raw files for Oct 2025 
```{r}
files <- list.files(path = "../data/TPC_Photosynthesis/raw_data/Oct2025/", pattern = "*.csv", full.names = TRUE)
```

Join O2 data with metadata files based on start and stop times 
```{r}
all_data <- map_dfr(files, function(f) {
  
  # This looks for "Ch" and pulls the digits immediately following it
  chan_num <- as.numeric(str_extract(basename(f), "(?<=Ch)\\d+"))
  
  raw <- read_csv(f) %>%
    mutate(
      DateTime = dmy_hms(Date), 
      Date_join = as.Date(DateTime),
      Channel = chan_num
    )
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )
  # Ensure sample_meta$Date is class 'Date' before running this
  final <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date_join" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )
  
  return(final)
})
```

Look at joined data and save
```{r}
all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

write_csv(all_data, "../output/TPC_Photosynthesis/Oct2025/TPC_all_joined_data_Oct2025.csv")
```

### Run localized linear regressions 

The interval fitting code will take about an hour to run and is computationally intense. If needed, run as a background job or on a remote server.

Make list of all Colony_ID / Run combinations (including blanks)
```{r}
colony_runs <- all_data %>%
  filter(!is.na(Colony_ID)) %>%
  distinct(Colony_ID, Run)
```

Calculate interval slopes for blanks
```{r}
blank_data <- all_data %>%
  filter(Colony_ID == "blank")

blank_intervals <- blank_data %>%
  group_by(Run, Temperature.y) %>%
  filter(!is.na(Colony_ID)) %>%
  group_split()
```

Fit regressions and extract slopes for each blank interval
```{r}
fit_reg <- function(df) {
  rankLocReg(
    xall = as.numeric(df$`Delta T [min]`),
    yall = df$Oxygen,
    alpha = 0.2,
    method = "pc",
    verbose = FALSE
  )
}

blank_fits <- map(blank_intervals, fit_reg)
blank_slopes <- map_dbl(blank_fits, ~ pluck(., "allRegs", "b1", 1))
blank_summary <- map2_dfr(
  blank_intervals,
  blank_slopes,
  ~ tibble(
    Temperature.y = unique(.x$Temperature.y),
    Blank_Slope = .y
  )
)
```

Average blanks
```{r}
avg_blank_slopes <- blank_summary %>%
  group_by(Temperature.y) %>%
  summarise(Avg_Blank_Slope = mean(Blank_Slope, na.rm = TRUE), .groups = "drop") %>%
  rename("Temperature" = "Temperature.y")
```

Fit regressions for each interval
```{r}
fit_colony_intervals_avgblank <- function(colony_id, run_num) {
  # Subset sample data
  sample_data <- thinned_data %>%
    filter(Colony_ID == colony_id, Run == run_num)
  if (nrow(sample_data) == 0) return(NULL)
  
  # Split data by light step
  intervals <- sample_data %>%
    group_by(Temperature.y) %>%
    group_split()
  
  # Fit regressions for each interval
  fits <- map(intervals, fit_reg)
  micromol.L.s <- map_dbl(fits, ~ pluck(., "allRegs", "b1", 1))
  
  # Detailed interval summary
  interval_df <- map2_dfr(
    intervals,
    seq_along(intervals),
    ~ tibble(
      Colony_ID   = colony_id,
      Run         = run_num,
      Interval    = .y,
      Temperature = unique(.x$Temperature.y),
      N_points    = nrow(.x),
      O2_start    = head(.x$Oxygen, 1),
      O2_end      = tail(.x$Oxygen, 1),
      DeltaT_min  = min(.x$`Delta T [min]`),
      DeltaT_max  = max(.x$`Delta T [min]`),
      O2_min      = min(.x$Oxygen, na.rm = TRUE),
      O2_max      = max(.x$Oxygen, na.rm = TRUE),
      Slope       = micromol.L.s[.y]
    )
  ) %>%
    left_join(avg_blank_slopes, by = "Temperature") %>%
    mutate(
      Blank_Corrected_Slope = Slope - Avg_Blank_Slope
    )
  
  # Write file (outside of the pipeline!)
  fn <- paste0("../output/TPC_Photosynthesis/Oct2025/intervals_", colony_id, "_run", run_num, "_Oct2025_TPC.csv")
  write_csv(interval_df, fn)
  
  interval_df
}

# Run for all samples and save big CSV
plan(multisession)

all_interval_summaries <- future_pmap_dfr(
  list(colony_runs$Colony_ID, colony_runs$Run),
  fit_colony_intervals_avgblank
)

write_csv(all_interval_summaries, "../output/TPC_Photosynthesis/Oct2025/all_interval_summaries_Oct2025_TPC.csv")
```

Read in interval data (so I don't have to run again)
```{r}
all_int_tpc <- read_csv("../output/TPC_Photosynthesis/Oct2025/all_interval_summaries_Oct2025_TPC.csv")
all_int_tpc$Colony_ID <- as.character(all_int_tpc$Colony_ID)
```

Filter sample metadata for Oct 2025
```{r}
oct_2025 <- sample_meta %>%
  filter(grepl("^2025-10", Date))
oct_2025$Surface_Area_cm2 <- as.numeric(oct_2025$Surface_Area_cm2)
```

Merge sample metadata with interval data 
```{r}
all_int_tpc <- all_int_tpc %>%
  inner_join(oct_2025, by = c("Colony_ID" = "Colony_ID", "Temperature" = "Target_Temperature")) %>%
  filter(!is.na(Phenotype))
length(unique(all_int_tpc$Colony_ID))
```

Normalize interval data by chamber volume and surface area and save
```{r}
# normalize 
all_int_tpc <- all_int_tpc %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Blank_Corrected_Slope * Chamber_Volume_L / Surface_Area_cm2) * 60) # units are umol O2 cm-2 h-1
write.csv(all_int_tpc, "../output/TPC_Photosynthesis/Oct2025/all_interval_summaries_normalized_Oct2025_TPC.csv")
```

Read data back in so I don't have to run above code again
```{r}
all_int_tpc <- read.csv("../output/TPC_Photosynthesis/Oct2025/all_interval_summaries_normalized_Oct2025_TPC.csv")
```

Plot curves across temperatures
```{r}
ggplot(all_int_tpc, aes(x = Temperature, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "umol O2 cm-2 h-1", color = "Phenotype")

ggplot(all_int_tpc, aes(x = Temperature, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(aes(fill = Phenotype), method = "loess", se = TRUE, linewidth = 1.2) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "umol O2 cm-2 h-1", color = "Phenotype")
```

Log and plot 
```{r}
all_int_tpc$log.rate <- log(all_int_tpc$Slope_norm + 1)  #logging and adding 0.1 because a log of zero does not exist

ggplot(all_int_tpc, aes(x = Temperature, y = log.rate, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "umol O2 cm-2 h-1", color = "Phenotype")

ggplot(all_int_tpc, aes(x = Temperature, y = log.rate, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(aes(fill = Phenotype), method = "loess", se = TRUE, linewidth = 1.2) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "log(umol O2 cm-2 h-1)", color = "Phenotype")
```

## Fit TPC curve

Set target temperature to numeric and convert temperature to Kelvin
```{r}
photo_2025 <- all_int_tpc %>%
  mutate(Temperature = as.numeric(Temperature),
         Temperature.K = Temperature + 273.15)
```

Fit curves for all colonies using `sharpe_schoolfield` function defined above
```{r}
fits_2025 <- photo_2025 %>%
  group_by(Phenotype,Species,Colony_ID) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(log.rate ~ sharpe_schoolfield(lnc, E, Eh, Th, temp = Temperature.K, Tc = 27),
                                                data = .x,
                                                iter = 1000,
                                                start_lower = c(lnc = -10, E = 0.1, Eh = 0.2, Th = 285),
                                                start_upper = c(lnc = 10, E = 2, Eh = 5, Th = 330),
                                                supp_errors = 'Y',
                                                na.action = na.omit,
                                                lower = c(lnc = -10, E = 0, Eh = 0, Th = 0))))
```

Get TPC parameters for each colony 
```{r}
params_2025 <- fits_2025 %>%
  unnest_legacy(fit %>% purrr::map(tidy)) %>%
  filter(!Colony_ID == "206")
```

Calculate Topt and Rmax for each colony 
```{r}
Topt_Rmax_2025 <- params_2025 %>%
  dplyr::select(Colony_ID, Phenotype, Species, term, estimate) %>%
  tidyr::pivot_wider(names_from = term, values_from = estimate) %>%
  mutate(
    Topt_K = get_topt(E, Th, Eh),
    Topt_C = Topt_K - 273.15,
    Rmax_log = sharpe_schoolfield(lnc, E, Eh, Th, temp = Topt_K, Tc = 27),
    Rmax_linear = exp(Rmax_log) - 1)
```

Pivot data so that each trait is a row 
```{r}
traits_long_2025 <- Topt_Rmax_2025 %>%
  dplyr::select(Colony_ID, Phenotype, Species, lnc, E, Eh, Topt_C, Rmax_log) %>%
  pivot_longer(cols = c(lnc, E, Eh, Topt_C, Rmax_log), 
               names_to = "trait", 
               values_to = "estimate")
```

## Visualization for 2025

Plot
```{r}
# Create the point plot with Mean and SE
ggplot(traits_long_2025, aes(x = Phenotype, y = estimate, color = Phenotype)) +
  # Add raw colony data as small, transparent points to show the spread
  geom_jitter(width = 0.1, alpha = 0.3, size = 1.5) +
  stat_summary(fun = mean, geom = "point", size = 4) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 1) +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "2025 Thermal Performance Traits",
    y = "Estimate (Mean ± SE)",
    x = "Phenotype"
  )

## Box plot 
ggplot(traits_long_2025, aes(x = Phenotype, y = estimate, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.6, color = "black") +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.2, color = "black") +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.major.x = element_blank() # Clean up vertical lines
  ) +
  labs(
    y = "Estimated Parameter Value",
    x = "Phenotype"
  )
```

## Plot with confidence intervals 

Generate and aggregate predictions
```{r}
# 1. Define the temperature range for the smooth curves
new_data <- data.frame(
  Temperature.K = seq(18 + 273.15 - 1,  # 17°C
                      38 + 273.15 + 1,  # 39°C
                      length.out = 200))

# 2. Generate predictions for every colony model
# This creates a "long" dataframe of all potential curves
preds_all_2025 <- fits_2025 %>%
  mutate(preds = map(fit, ~{
    # We use a safe predict call in case some models didn't converge
    p <- try(predict(.x, newdata = new_data), silent = TRUE)
    if(inherits(p, "try-error")) return(NULL)
    data.frame(Temperature.K = new_data$Temperature.K, pred_val = p)
  })) %>%
  filter(!map_lgl(preds, is.null)) %>% # Remove failed fits
  unnest(preds)

# 3. Calculate the MEAN and CI of the curves per group
# At each temperature step, find the mean performance across all colonies
preds_mean_summary_2025 <- preds_all_2025 %>%
  group_by(Phenotype, Species, Temperature.K) %>%
  dplyr::summarise(
    mean_val = mean(pred_val, na.rm = TRUE),
    sd_val = sd(pred_val, na.rm = TRUE),
    n = n(),
    se_val = sd_val / sqrt(n),
    # 95% Confidence Interval of the Mean
    lower = mean_val - (1.96 * se_val),
    upper = mean_val + (1.96 * se_val),
    .groups = "drop"
  )
```

Plot TPC with fitted curves 
```{r}
ggplot(preds_mean_summary_2025, aes(x = Temperature.K - 273.15, y = mean_val, color = Phenotype, fill = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5, color = NA) +
  geom_line(linewidth = 1.5) +
  scale_x_continuous(
    breaks = seq(18, 40, by = 2), 
    limits = c(17, 40)) +
  geom_point(data = photo_2025, aes(x = Temperature.K - 273.15, y = Slope_norm), 
             alpha = 0.2, size = 1, inherit.aes = TRUE) +
  facet_wrap(~ Species) +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 12),
    panel.grid = element_blank()
  ) +
  labs(
    x = "Temperature (°C)", 
    y = expression(bold("Net photosynthesis (log(" *mu*"mol " *cm^-2 *hr^-1*"))")),
    title = "Photosynthesis Fitted Thermal Performance Curves (2025)"
  )
```

## Plot with confidence intervals and the environmental data 

```{r}
temps <- read.csv("../output/environmental/NSF_CAREER_KBay_PR13_Temps.csv")
```

Subset for the year prior to Oct 7, 2025
```{r}
temps_sub <- temps %>%
  filter(DateTime <= "2025-10-06 01:00:00")
```

```{r}
temp_freq_2025 <- temps_sub %>%
  # Rounding to the nearest whole degree centers the bar on that number
  mutate(temp_bin = round(Temp_C, 0)) %>% 
  group_by(temp_bin) %>%
  dplyr::summarise(count = n(), .groups = "drop") %>%
  # Calculate the percentage of total time spent at that temperature
  mutate(percent = (count / sum(count)) * 100)
```

Plot with temp bars
```{r}
# Axis Parameters
p_min <- -0.5  # Hard floor of the plot
p_max <- 1.5   # Peak of the Y-axis
p_total <- p_max - p_min
s_max <- 30    # Max percent on the right axis
scale_factor <- p_total / s_max

ggplot() +
  geom_rect(data = temp_freq_2025, 
            aes(xmin = temp_bin - 0.45, xmax = temp_bin + 0.45, 
                ymin = p_min, ymax = p_min + (percent * scale_factor)), 
            fill = "#8DBBCC", color = "#727E8B", linewidth = 0.2) +
  geom_point(data = photo_2025, 
             aes(x = Temperature.K - 273.15, y = log.rate, color = Phenotype), 
             alpha = 0.5, size = 1.2, position = position_jitter(width = 0.1)) +
  geom_ribbon(data = preds_mean_summary_2025, 
              aes(x = Temperature.K - 273.15, ymin = lower, ymax = upper, fill = Phenotype), 
              alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, color = "grey", linewidth = 0.5, linetype = "dashed") +
  geom_line(data = preds_mean_summary_2025, 
            aes(x = Temperature.K - 273.15, y = mean_val, color = Phenotype), 
            linewidth = 1.3) +
  scale_y_continuous(
    name = expression("Net photosynthesis (log(" *mu*"mol " *cm^-2 *hr^-1*"))"),
    limits = c(p_min, p_max),
    expand = c(0, 0),
    sec.axis = sec_axis(~ (. - p_min) / scale_factor, name = "Percent days (%)", breaks = seq(0, 40, 5))
  ) +
  scale_x_continuous(breaks = seq(18, 40, by = 2), limits = c(17.5, 40.5), expand = c(0,0)) +
  facet_wrap(~ Species, nrow = 1) +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_classic() +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 13),
    axis.line = element_line(color = "black"),
    panel.spacing = unit(2, "lines")
  ) +
  labs(x = "Temperature (°C)")
```

# Compare 2023 and 2025

Rename Colony id column from 2023 so it matches 2025 and add year as a column to both dfs
```{r}
traits_long_2023 <- traits_long_2023 %>%
  mutate(Year = "2023") %>%
  #dplyr::rename("ColonyID" = "Colony.ID") %>%
  mutate(Colony_ID = as.character(Colony_ID))

traits_long_2025 <- traits_long_2025 %>%
  mutate(Year = "2025") %>%
  mutate(Colony_ID = as.character(Colony_ID))
```

Bind dfs together 
```{r}
traits_long_all <- rbind(traits_long_2023, traits_long_2025)
unique(traits_long_all$trait)

# Save 
write.csv(traits_long_all, "../output/TPC_Photosynthesis/TPC_metrics.csv")
```

## MCAP Statistics comparing 2023 v. 2025 TPC photosynthesis metrics 

Read in data again if needed
```{r}
traits_long_all <- read.csv("../output/TPC_Photosynthesis/TPC_metrics.csv")
```

Subset by Mcap only 
```{r}
traits_long_all_mcap <- traits_long_all %>%
  filter(Species == "Montipora capitata")
```

### lnc (r_tref)

Subset by lnc data
```{r}
lnc_long <- traits_long_all_mcap %>%
  filter(trait == "lnc")
```

Check for normality
```{r}
hist(lnc_long$estimate)

# Q-Q Plot
qqnorm(lnc_long$estimate)
qqline(lnc_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(lnc_long$estimate)
```

Data is normal.

Run linear-mixed effects model
```{r}
lnc.model <- lmer(estimate ~ Year * Phenotype + (1|Colony_ID), data = lnc_long)
car::Anova(lnc.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(lnc.model, which = 1) 
lattice::qqmath(lnc.model)

# Post-hoc tests
emmeans(lnc.model, pairwise ~ Year)
emmeans(lnc.model, pairwise ~ Year | Phenotype)
emmeans(lnc.model, pairwise ~ Phenotype | Year)

# Check plot to make sure results make sense
ggplot(lnc_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year and Year x Phenotype is significant. 

### E

Subset by E data
```{r}
e_long <- traits_long_all_mcap %>%
  filter(trait == "E")
```

Check for normality
```{r}
hist(e_long$estimate)

# Q-Q Plot
qqnorm(e_long$estimate)
qqline(e_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(e_long$estimate)
```

Data is normal.

Run linear mixed effects model 
```{r}
e.model <- lmer(estimate ~ Year * Phenotype + (1|Colony_ID), data = e_long)
car::Anova(e.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(e.model, which = 1) 
lattice::qqmath(e.model)

# Post-hoc tests
emmeans(e.model, pairwise ~ Phenotype)
emmeans(e.model, pairwise ~ Phenotype)
emmeans(e.model, pairwise ~ Year | Phenotype)
emmeans(e.model, pairwise ~ Phenotype | Year)

# Check plot to make sure results make sense
ggplot(e_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year, Phenotype, Year x Phenotype are significant. 

### Eh

Subset by E data
```{r}
eh_long <- traits_long_all_mcap %>%
  filter(trait == "Eh")
```

Check for normality
```{r}
hist(eh_long$estimate)

# Q-Q Plot
qqnorm(eh_long$estimate)
qqline(eh_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(eh_long$estimate)
```

Data is normal.

Run linear mixed effects model 
```{r}
eh.model <- lmer(estimate ~ Year * Phenotype + (1|Colony_ID), data = eh_long)
car::Anova(eh.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(eh.model, which = 1) 
lattice::qqmath(eh.model)

# Post-hoc tests
emmeans(eh.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(eh_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year is significant. 

### Topt

Subset by E data
```{r}
topt_long <- traits_long_all_mcap %>%
  filter(trait == "Topt_C")
```

Check for normality
```{r}
hist(topt_long$estimate)

# Q-Q Plot
qqnorm(topt_long$estimate)
qqline(topt_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(topt_long$estimate)
```

Data is normal. 

Run linear mixed effects model 
```{r}
topt.model <- lmer(estimate ~ Year * Phenotype + (1|Colony_ID), data = topt_long)
car::Anova(topt.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(topt.model, which = 1) 
lattice::qqmath(topt.model)

# Post-hoc tests
emmeans(topt.model, pairwise ~ Phenotype)

# Check plot to make sure results make sense
ggplot(topt_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Phenotype is significant.

### Rmax

Subset by E data
```{r}
rmax_long <- traits_long_all_mcap %>%
  filter(trait == "Rmax_log")
```

Check for normality
```{r}
hist(rmax_long$estimate)

# Q-Q Plot
qqnorm(rmax_long$estimate)
qqline(rmax_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(rmax_long$estimate)
```

Data is normal.

Run linear mixed effects model 
```{r}
rmax.model <- lmer(estimate ~ Year * Phenotype + (1|Colony_ID), data = rmax_long)
car::Anova(rmax.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(rmax.model, which = 1) 
lattice::qqmath(rmax.model)

# Post-hoc tests
emmeans(rmax.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(rmax_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year is significant. 

## PCOMP Statistics comparing 2023 v. 2025 TPC photosynthesis metrics 

Subset by Pcomp only 
```{r}
traits_long_all_pcomp <- traits_long_all %>%
  filter(Species == "Porites compressa")
```

### lnc (r_tref)

Subset by lnc data
```{r}
lnc_long <- traits_long_all_pcomp %>%
  filter(trait == "lnc")
```

Check for normality
```{r}
hist(lnc_long$estimate)

# Q-Q Plot
qqnorm(lnc_long$estimate)
qqline(lnc_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(lnc_long$estimate)
```

Data is not normal. Transform data to achieve normality 
```{r}
lnc_long$log <- log(lnc_long$estimate)
hist(lnc_long$log)
qqnorm(lnc_long$log)
qqline(lnc_long$log, col = "red")
shapiro.test(lnc_long$log)

lnc_long$sqrt <- sqrt(lnc_long$estimate)
hist(lnc_long$sqrt)
qqnorm(lnc_long$sqrt)
qqline(lnc_long$sqrt, col = "red")
shapiro.test(lnc_long$sqrt)
```

Log produced normality. 

Run linear mixed effects model 
```{r}
lnc.model <- lmer(log ~ Year * Phenotype + (1|Colony_ID), data = lnc_long)
car::Anova(lnc.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(lnc.model, which = 1) 
lattice::qqmath(lnc.model)

# Post-hoc tests
emmeans(lnc.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(lnc_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year is significant. 

### E

Subset by E data
```{r}
e_long <- traits_long_all_pcomp %>%
  filter(trait == "E")
```

Check for normality
```{r}
hist(e_long$estimate)

# Q-Q Plot
qqnorm(e_long$estimate)
qqline(e_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(e_long$estimate)
```

Data is not normal. Transform data to achieve normality 
```{r}
e_long$log <- log(e_long$estimate)
hist(e_long$log)
qqnorm(e_long$log)
qqline(e_long$log, col = "red")
shapiro.test(e_long$log)

e_long$sqrt <- sqrt(e_long$estimate)
hist(e_long$sqrt)
qqnorm(e_long$sqrt)
qqline(e_long$sqrt, col = "red")
shapiro.test(e_long$sqrt)
```

Sqrt produced normality. 

Run linear mixed effects model 
```{r}
e.model <- lmer(sqrt ~ Year * Phenotype + (1|Colony_ID), data = e_long)
car::Anova(e.model, type = 3)

# Model normality 
par(mfrow = c(1, 2)) 
plot(e.model, which = 1) 
lattice::qqmath(e.model)

# Post-hoc tests
emmeans(e.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(e_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

### Eh

Subset by E data
```{r}
eh_long <- traits_long_all_pcomp %>%
  filter(trait == "Eh")
```

Check for normality
```{r}
hist(eh_long$estimate)

# Q-Q Plot
qqnorm(eh_long$estimate)
qqline(eh_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(eh_long$estimate)
```

Data is not normal. Transform data to achieve normality 
```{r}
eh_long$log <- log(eh_long$estimate)
hist(eh_long$log)
qqnorm(eh_long$log)
qqline(eh_long$log, col = "red")
shapiro.test(eh_long$log)

eh_long$sqrt <- sqrt(eh_long$estimate)
hist(eh_long$sqrt)
qqnorm(eh_long$sqrt)
qqline(eh_long$sqrt, col = "red")
shapiro.test(eh_long$sqrt)
```

Sqrt produces normality. 

Run linear mixed effects model 
```{r}
eh.model <- lmer(sqrt ~ Year * Phenotype + (1|Colony_ID), data = eh_long)
car::Anova(eh.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(eh.model, which = 1) 
lattice::qqmath(eh.model)

# Post-hoc tests
emmeans(eh.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(eh_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year is significant. 

### Topt

Subset by E data
```{r}
topt_long <- traits_long_all_pcomp %>%
  filter(trait == "Topt_C")
```

Check for normality
```{r}
hist(topt_long$estimate)

# Q-Q Plot
qqnorm(topt_long$estimate)
qqline(topt_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(topt_long$estimate)
```

Data is not normal. Transform data to achieve normality 
```{r}
topt_long$log <- log(topt_long$estimate)
hist(topt_long$log)
qqnorm(topt_long$log)
qqline(topt_long$log, col = "red")
shapiro.test(topt_long$log)

topt_long$sqrt <- sqrt(topt_long$estimate)
hist(topt_long$sqrt)
qqnorm(topt_long$sqrt)
qqline(topt_long$sqrt, col = "red")
shapiro.test(topt_long$sqrt)
```

Run linear mixed effects model 
```{r}
topt.model <- lmer(estimate ~ Year * Phenotype + (1|Colony_ID), data = topt_long)
car::Anova(topt.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(topt.model, which = 1) 
lattice::qqmath(topt.model)

# Post-hoc tests
emmeans(topt.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(topt_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

Year is significant. 

### Rmax

Subset by E data
```{r}
rmax_long <- traits_long_all_pcomp %>%
  filter(trait == "Rmax_log")
```

Check for normality
```{r}
hist(rmax_long$estimate)

# Q-Q Plot
qqnorm(rmax_long$estimate)
qqline(rmax_long$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(rmax_long$estimate)
```

Data is normal.

Run linear mixed effects model 
```{r}
rmax.model <- lmer(estimate ~ Year * Phenotype + (1|Colony_ID), data = rmax_long)
car::Anova(rmax.model, type = "III")

# Model normality 
par(mfrow = c(1, 2)) 
plot(rmax.model, which = 1) 
lattice::qqmath(rmax.model)

# Post-hoc tests
emmeans(rmax.model, pairwise ~ Year)

# Check plot to make sure results make sense
ggplot(rmax_long, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1) +
  facet_wrap(~Species) +
  theme_classic()
```

No significant differences. 

## Visualize 2023 v. 2025 TPC photosynthesis 

Visualize differences between 2023 and 2025
```{r}
ggplot(traits_long_all, aes(x = Year, y = estimate, color = Phenotype, group = Phenotype)) +
  # 1. Add Error Bars (Standard Error)
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, size = 1) +
  
  # 2. Add Lines to connect the means over time
  stat_summary(fun = mean, geom = "line", size = 1.2) +
  
  # 3. Add the Mean Points
  stat_summary(fun = mean, geom = "point", size = 4) +
  
  # 4. Facet by Trait and Species
  # Using 'free_y' because Topt (~30) and E (~0.5) have very different scales
  facet_grid(trait ~ Species, scales = "free_y") +
  
  # 5. Styling and Colors
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 10),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  labs(
    title = "Trait Evolution: 2023 vs 2025",
    y = "Estimated Value (Mean ± SE)",
    x = "Year"
  )
```

Make the plot pretty
```{r}
# Define a mapping for the Y-axis labels
trait_labels <- c(
  "E" = "Activation~Energy~(eV)",
  "Eh" = "Deactivation~Energy~(eV)",
  "lnc" = "atop(Net~photosynthesis~at~27*degree*C, log(mu*mol~hr^-1~cm^-2))",
  "Rmax_log" = "atop(Maximum~photosynthesis, log(mu*mol~hr^-1~cm^-2))",
  "Topt_C" = "Thermal~optimum~(degree*C)"
)

# Define the traits vector in the order you want them plotted
traits <- c("E", "Eh", "lnc", "Topt_C", "Rmax_log")

## Spp in same plot 
create_combined_trait_plot <- function(data, tr, plot_letter, is_bottom_row = FALSE) {
  
  # Filter data for the specific trait but keep both species
  plot_data <- data %>% filter(trait == tr)
  
  p <- ggplot(plot_data, aes(x = Year, y = estimate, 
                             color = Phenotype, 
                             linetype = Species, 
                             group = interaction(Phenotype, Species))) +
    # Error bars (Species-specific linetypes can apply here too if desired)
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, size = 0.6, linetype = "solid") +
    stat_summary(fun = mean, geom = "line", size = 0.8) +
    stat_summary(fun = mean, geom = "point", size = 2.5) +
    # Custom Scales
    scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
    scale_linetype_manual(values = c("Montipora capitata" = "solid", 
                                     "Porites compressa" = "dashed")) +
    theme_classic() + 
    theme(
      axis.text = element_text(size = 9, color = "black"),
      axis.title.y = element_text(size = 9, face = "plain"),
      plot.title = element_text(size = 10, face = "bold", hjust = 0)
    ) +
    labs(title = plot_letter, x = NULL, y = trait_labels[[tr]])

  # Hide X-axis for all but the bottom plot
  if (!is_bottom_row) {
    p <- p + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  }
  
  return(p)
}

# Generate the single column of combined plots
combined_trait_plots <- lapply(1:5, function(i) {
  create_combined_trait_plot(traits_long_all, traits[i], 
                             LETTERS[i], is_bottom_row = (i == 5))
})

# Assemble the vertical stack
final_trait_figure <- wrap_plots(combined_trait_plots, ncol = 1) + 
  plot_layout(guides = "collect") +
  plot_annotation(
    theme = theme(plot.title = element_text(size = 14, face = "bold"))
  ) & 
  theme(legend.position = "right",
        legend.box = "vertical") # Stacks Phenotype and Species legends
final_trait_figure

# Save 
ggsave("../output/TPC_Photosynthesis/Photosynthesis_TPC_metrics_2023_2025.pdf", 
       plot = final_trait_figure, 
       width = 8, 
       height = 12, 
       units = "in", 
       device = "pdf",
       dpi = 300)
ggsave("../output/TPC_Photosynthesis/Photosynthesis_TPC_metrics_2023_2025.png", 
       plot = final_trait_figure, 
       width = 8, 
       height = 12, 
       units = "in", 
       dpi = 300)
```

Add significance data to plots
```{r}
# Add labels 
sig_labels <- c(
  "S*, P*, Y***, SxP*, PxY*",                 # for E
  "Y***, SxY*",       # for Eh
  "S***, Y***, PxY*",         # for tref
  "P*, Y***, SxP*, SxY*", # for Topt_C
  "S***, Y*, PxY*"                     # for Rmax_log 
)

create_combined_trait_plot <- function(data, tr, plot_letter, sig_text = "", is_bottom_row = FALSE) {
  
  plot_data <- data %>% filter(trait == tr)
  p <- ggplot(plot_data, aes(x = Year, y = estimate, 
                               color = Phenotype, 
                               linetype = Species, 
                               group = interaction(Phenotype, Species))) +
     # Error bars - slightly thicker for publication
     stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, size = 0.8, linetype = "solid") +
     stat_summary(fun = mean, geom = "line", size = 1.0) +
     stat_summary(fun = mean, geom = "point", size = 3.5) +
     
     # SIGNIFICANCE LABELS - Positioned in bottom right
     annotate("text", x = Inf, y = -Inf, label = sig_text, 
              hjust = 1.1, vjust = -1.2, size = 4, color = "black") +
     
     scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
     scale_linetype_manual(values = c("Montipora capitata" = "solid", 
                                      "Porites compressa" = "dashed")) +
     theme_classic() + 
     theme(
       # Larger text for readability in print
       axis.text = element_text(size = 15, color = "black"),
       axis.title.y = element_text(size = 15, face = "plain", margin = margin(r = 10)),
       plot.title = element_text(size = 18, face = "bold", hjust = 0),
       # Ensure panels have enough breathing room
       plot.margin = margin(t = 5, r = 5, b = 5, l = 5)
     ) +
     labs(title = plot_letter, x = NULL, y = parse(text = trait_labels[[tr]]))

  if (!is_bottom_row) {
    p <- p + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  } else {
    # Bold the year labels on the bottom row
    p <- p + theme(axis.text.x = element_text(size = 18))
  }
  
  return(p)
}

# Define traits and your updated sig labels
traits <- c("E", "Eh", "lnc", "Topt_C", "Rmax_log")

# Apply the function across all traits
combined_trait_plots <- lapply(1:5, function(i) {
  create_combined_trait_plot(
    data = traits_long_all, 
    tr = traits[i], 
    plot_letter = LETTERS[i], 
    sig_text = sig_labels[i],
    is_bottom_row = (i == 5)
  )
})

# Assemble with Patchwork
final_trait_figure <- wrap_plots(combined_trait_plots, ncol = 1) + 
  plot_layout(guides = "collect") +
  plot_annotation(
    # Optional: add an overall title or leave NULL for publication
    title = NULL 
  ) & 
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11),
    legend.key.width = unit(1.5, "cm"), # Makes linetypes easier to distinguish
    legend.spacing.y = unit(0.5, "cm")
  )

# Save 
ggsave("../output/TPC_Photosynthesis/Photosynthesis_TPC_metrics_stats_2023_2025.pdf", 
       plot = final_trait_figure, 
       width = 10, 
       height = 16, 
       units = "in", 
       device = "pdf",
       dpi = 300)
ggsave("../output/TPC_Photosynthesis/Photosynthesis_TPC_metrics_stats_2023_2025.png", 
       plot = final_trait_figure, 
       width = 10, 
       height = 16, 
       units = "in", 
       dpi = 300)
```

Plot with species as separate columns
```{r}
# 1. Update/Verify Significance Labels (No changes needed here)
sig_data <- data.frame(
  trait = rep(c("E", "Eh", "lnc", "Topt_C", "Rmax_log"), each = 2),
  Species = rep(c("Montipora capitata", "Porites compressa"), 5),
  sig_label = c(
    "Y***, P*, PxY*", "Y***",      # E (A, B)
    "Y***", "Y***",    # Eh (C, D)
    "Y***, PxY*", "Y***",      # lnc (E, F)
    "P*", "Y***",    # Topt_C (G, H)
    "Y*", "n.s."       # Rmax_log (I, J)
  )
)

trait_labels <- c(
  "E" = "Activation~Energy~(eV)",
  "Eh" = "Deactivation~Energy~(eV)",
  "lnc" = "atop(Net~photosynthesis~at~bold(27)*degree*C, log(mu*mol~hr^-1~cm^-2))",
  "Rmax_log" = "atop(Maximum~photosynthesis, log(mu*mol~hr^-1~cm^-2))",
  "Topt_C" = "Thermal~optimum~(degree*C)"
)

# Calculate common Y-limits for each trait 
# This ensures that both species share the same scale for a given trait row.
# We calculate the mean +/- SE for every group to find the true extent of the data.
trait_limits_df <- traits_long_all %>%
  group_by(trait, Species, Year, Phenotype) %>%
  dplyr::summarise(
    m = mean(estimate, na.rm = TRUE),
    se = sd(estimate, na.rm = TRUE) / sqrt(n()),
    .groups = 'drop'
  ) %>%
  group_by(trait) %>%
  dplyr::summarise(
    ymin = min(m - se, na.rm = TRUE),
    ymax = max(m + se, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  # Add padding to the top for significance labels and visibility
  mutate(ymax = ymax + (ymax - ymin) * 0.05,
         ymin = ymin - (ymax - ymin) * 0.05)

# 2. Updated Function for individual panels
# Added 'ylimits' argument
create_panel_plot <- function(data, tr, spp, plot_letter, ylimits, is_bottom_row = FALSE, is_left_col = FALSE) {
  
  plot_data <- data %>% filter(trait == tr, Species == spp)
  sig_text <- sig_data %>% filter(trait == tr, Species == spp) %>% pull(sig_label)
  
  y_label <- if(is_left_col) {
    parse(text = paste0("bold(", trait_labels[[tr]], ")"))
  } else {
    NULL
  }

p <- ggplot(plot_data, aes(x = factor(Year), y = estimate, color = Phenotype, group = Phenotype)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.1, linewidth = 1.0) +
    stat_summary(fun = mean, geom = "line", linewidth = 1.2) +
    stat_summary(fun = mean, geom = "point", size = 4) +
    
    # Apply the consistent y-limits here
    coord_cartesian(ylim = ylimits) +
    
    annotate("text", x = Inf, y = -Inf, label = sig_text, 
             hjust = 1.1, vjust = -1.2, size = 6, color = "black", fontface = "plain") +
    
    scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
    theme_classic() + 
    theme(
      axis.text = element_text(size = 16, color = "black"), 
      axis.text.x = element_text(face = "bold"), 
      axis.title.y = element_text(size = 18), 
      plot.title = element_text(size = 22, face = "bold"),
      plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
    ) +
    labs(title = plot_letter, x = NULL, y = y_label)
if (tr == "Topt_C") {
    # Define the specific range and 0.5 increments you requested
    # We expand the limits slightly if the data exceeds 30 or goes below 27.5
    topt_min <- min(27.5, ylimits[1])
    topt_max <- max(30, ylimits[2])
    
    p <- p + scale_y_continuous(
      limits = c(topt_min, topt_max), 
      breaks = seq(27.5, 30, by = 0.5)
    )
  } else {
    # For all other traits, use the pre-calculated shared limits
    p <- p + coord_cartesian(ylim = ylimits)
  }

  if (!is_bottom_row) {
    p <- p + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  }
  
  return(p)
}

# 3. Create Species Header Labels
mcap_label <- ggplot() + 
  annotate("text", x = 0.5, y = 0.5, label = "Montipora capitata", fontface = "bold.italic", size = 8) + 
  theme_void()
pcomp_label <- ggplot() + 
  annotate("text", x = 0.5, y = 0.5, label = "Porites compressa", fontface = "bold.italic", size = 8) + 
  theme_void()

# 4. Generate the 10 panels
traits <- c("E", "Eh", "lnc", "Topt_C", "Rmax_log")
species_list <- c("Montipora capitata", "Porites compressa")
all_plots <- list()
letters_idx <- LETTERS[1:10]
counter <- 1

for (i in 1:length(traits)) {
  # Get limits for this specific trait
  current_limits <- trait_limits_df %>% 
    filter(trait == traits[i]) %>% 
    dplyr::select(ymin, ymax) %>% 
    as.numeric()

  for (j in 1:length(species_list)) {
    all_plots[[counter]] <- create_panel_plot(
      data = traits_long_all,
      tr = traits[i],
      spp = species_list[j],
      plot_letter = letters_idx[counter],
      ylimits = current_limits, # Pass the trait-specific common limits
      is_bottom_row = (i == 5),
      is_left_col = (j == 1)
    )
    counter <- counter + 1
  }
}

# 5. Assemble and update Legend sizes
final_trait_figure <- (mcap_label + pcomp_label) / 
  wrap_plots(all_plots, ncol = 2) + 
  plot_layout(heights = c(1, 15), guides = "collect") & 
  theme(
    legend.position = "top",
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 16),
    legend.margin = margin(b = 20)
  )

# Save
ggsave("../output/TPC_Photosynthesis/Photosynthesis_TPC_metrics_stats_separate_spp_2023_2025.png", 
       plot = final_trait_figure, 
       width = 11, 
       height = 20, 
       units = "in", 
       dpi = 300)

ggsave("../output/TPC_Photosynthesis/Photosynthesis_TPC_metrics_stats_separate_spp_2023_2025.pdf", 
       plot = final_trait_figure, 
       width = 11, 
       height = 20, 
       units = "in", 
       device = "pdf",
       dpi = 300)
```

Plot TPC photosynthesis curves with temp bars 
```{r}
# Select specific cols 
photo_2023 <- photo_2023 %>%
  dplyr::select(Colony_ID, Temperature, Slope_norm, Species, Phenotype, log.rate, Temperature.K) %>%
  mutate(Colony_ID = as.character(Colony_ID))

photo_2025 <- photo_2025 %>%
  mutate(Colony_ID = as.character(Colony_ID)) %>%
  dplyr::select(Colony_ID, Temperature, Slope_norm, Species, Phenotype, log.rate, Temperature.K)

# 1. Combine all data and create a unified "Header" for each plot
all_photo <- bind_rows(
  photo_2023 %>% mutate(Year = "2023"),
  photo_2025 %>% mutate(Year = "2025")
) %>% mutate(Plot_Header = paste(Species, Year))

all_preds <- bind_rows(
  preds_mean_summary_2023 %>% mutate(Year = "2023"),
  preds_mean_summary_2025 %>% mutate(Year = "2025")
) %>% mutate(Plot_Header = paste(Species, Year))

species_list <- c("Montipora capitata", "Porites compressa")

# Rebuild all_freq by duplicating the year data for each species
all_freq <- bind_rows(
  # For 2023
  map_dfr(species_list, ~temp_freq_2023 %>% mutate(Species = .x, Year = "2023")),
  # For 2025
  map_dfr(species_list, ~temp_freq_2025 %>% mutate(Year = "2025", Species = .x))
) %>% 
  mutate(Plot_Header = paste(Species, Year))

# Ensure all_photo and all_preds also have the Plot_Header
all_photo <- all_photo %>% mutate(Plot_Header = paste(Species, Year))
all_preds <- all_preds %>% mutate(Plot_Header = paste(Species, Year))

# 2. Axis Parameters
p_min <- -0.5; p_max <- 1.5; p_total <- p_max - p_min
s_max <- 30; scale_factor <- p_total / s_max

# 3. Create the Plot
final_tpc_plot <- ggplot() +
  # Environmental bars
  geom_rect(data = all_freq, 
            aes(xmin = temp_bin - 0.45, xmax = temp_bin + 0.45, 
                ymin = p_min, ymax = p_min + (percent * scale_factor)), 
            fill = "#8DBBCC", color = "#727E8B", linewidth = 0.2) +
  # Raw data points
  geom_point(data = all_photo, 
             aes(x = Temperature.K - 273.15, y = log.rate, color = Phenotype), 
             alpha = 0.4, size = 1.5, position = position_jitter(width = 0.1)) +
  # Confidence ribbons
  geom_ribbon(data = all_preds, 
              aes(x = Temperature.K - 273.15, ymin = lower, ymax = upper, fill = Phenotype), 
              alpha = 0.3, color = NA) +
  # Mean lines
  geom_line(data = all_preds, 
            aes(x = Temperature.K - 273.15, y = mean_val, color = Phenotype), 
            linewidth = 1.2) +
  geom_hline(yintercept = 0, color = "grey", linewidth = 0.5, linetype = "dashed") +
  
  # Scaling and Dual Axes
  scale_y_continuous(
    name = expression("Net photosynthesis (log(" *mu*"mol " *cm^-2 *hr^-1*"))"),
    limits = c(p_min, p_max), expand = c(0, 0),
    sec.axis = sec_axis(~ (. - p_min) / scale_factor, name = "Percent days (%)", breaks = seq(0, 30, 5))
  ) +
  scale_x_continuous(breaks = seq(18, 40, by = 2), limits = c(17.5, 40.5), expand = c(0,0)) +
  
  # FACET: This creates 4 distinct panels with your custom headers
  facet_wrap(~ Plot_Header, ncol = 2) +
  
  # Styling
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_classic() +
theme_classic() +
  theme(
    # --- INCREASED TEXT SIZES ---
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14),
    
    # Secondary axis styling (matches the blue bars)
    axis.title.y.right = element_text(color = "black", size = 13, margin = margin(l = 15)),
    axis.text.y.right = element_text(color = "black", size = 11),
    
    # Facet header styling
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 14, hjust = 0),
    
    # Legend styling
    legend.position = "top",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 13, face = "bold"),
    
    # Spacing
    panel.spacing = unit(2, "lines"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
  ) +
  labs(x = "Temperature (°C)")

final_labeled_tpc <- final_tpc_plot + 
  #plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(face = "bold", size = 16))

final_labeled_tpc

# Save 
ggsave("../output/TPC_Photosynthesis/Photosynthesis_TPC_curve_temp_bars_2023_2025.pdf", 
       plot = final_labeled_tpc, 
       width = 12, 
       height = 8, 
       units = "in", 
       device = "pdf",
       dpi = 300)
ggsave("../output/TPC_Photosynthesis/Photosynthesis_TPC_curve_temp_bars_2023_2025.png", 
       plot = final_labeled_tpc, 
       width = 12, 
       height = 8, 
       units = "in", 
       dpi = 300)
```

Plot curves w/o temp bars 
```{r}
# 1. Axis Parameters (Adjusted for single axis)
p_min <- -0.5; p_max <- 1.5

# 2. Create the Plot
final_tpc_plot <- ggplot() +
  # Raw data points
  geom_point(data = all_photo, 
             aes(x = Temperature.K - 273.15, y = log.rate, color = Phenotype), 
             alpha = 0.4, size = 1.5, position = position_jitter(width = 0.1)) +
  # Confidence ribbons
  geom_ribbon(data = all_preds, 
              aes(x = Temperature.K - 273.15, ymin = lower, ymax = upper, fill = Phenotype), 
              alpha = 0.3, color = NA) +
  # Mean lines
  geom_line(data = all_preds, 
            aes(x = Temperature.K - 273.15, y = mean_val, color = Phenotype), 
            linewidth = 1.2) +
  geom_hline(yintercept = 0, color = "grey", linewidth = 0.5, linetype = "dashed") +
  
  # Simplified Single Axis
  scale_y_continuous(
    name = expression(bold("Net photosynthesis (log(" *mu*"mol " *cm^-2 *hr^-1*"))")),
    limits = c(p_min, p_max), 
    expand = c(0, 0)
  ) +
  scale_x_continuous(breaks = seq(18, 40, by = 2), limits = c(17.5, 40.5), expand = c(0,0)) +
  
  # FACET: Using ggtext or a simple expression trick for italics is complex in facet_wrap.
  # The easiest way for publication is to italicize the whole strip or use labeller.
  facet_wrap(~ Plot_Header, ncol = 2) +
  
  # Styling
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_classic() +
  theme(
    axis.text = element_text(size = 16, color = "black"),
    axis.title = element_text(size = 18, face = "bold"),
    
    # Facet header styling - Italics applied here
    strip.background = element_blank(),
    strip.text = element_text(face = "italic", size = 18, hjust = 0),
    
    legend.position = "top",
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 15, face = "bold"),
    
    panel.spacing = unit(2, "lines"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
  ) +
  labs(x = "Temperature (°C)")

# If you ONLY want the Species name italicized and not the Year:
# Use element_markdown from library(ggtext) and wrap the species in "*"
library(ggtext)
final_tpc_plot <- final_tpc_plot + 
   theme(strip.text = element_markdown())

final_labeled_tpc <- final_tpc_plot + 
  theme(plot.tag = element_text(face = "bold", size = 20))

final_labeled_tpc

# Save 
ggsave("../output/TPC_Photosynthesis/Photosynthesis_TPC_curve_2023_2025.pdf", 
       plot = final_labeled_tpc, 
       width = 12, 
       height = 8, 
       units = "in", 
       device = "pdf",
       dpi = 300)
ggsave("../output/TPC_Photosynthesis/Photosynthesis_TPC_curve_2023_2025.png", 
       plot = final_labeled_tpc, 
       width = 12, 
       height = 8, 
       units = "in", 
       dpi = 300)
```


