---
title: "Net calcification TPC"
author: "Jill Ashey"
date: "2025-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rTPC)
library(broom)
library(nls.multstart)
library(boot)
library(patchwork) 
library(parallel)
library(MuMIn)
library(purrr)
library(ggfortify) 

n_cores <- detectCores() - 1
```

# 2023

Read in data 
```{r}
calc_2023 <- read.csv("../data/TPC_calcification/TPC_calcification_all_standardized_without_blanks_2023_KB.csv")
```

Kristen already manually calculated the calcification rates in excel. 

Plot raw data 
```{r}
calc_plot_2023 <- calc_2023 %>%
  filter(!is.na(Phenotype),
         !is.na(Species),
         Colony.ID != "blank")

ggplot(calc_plot_2023, aes(
  x = Target.Temperature,
  y = umol.CaCO3h.1.cm.2,
  color = Phenotype,
  fill = Phenotype
)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "loess", se = TRUE, span = 0.6, alpha = 0.25) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  labs(
    x = "Target Temperature (°C)",
    y = expression("Net Calcification ("*cm^-2*")"),
    color = "Phenotype",
    fill = "Phenotype")
```


## rTPC for 2023 

Because some values are negative, we need to 'shift' the data away from negative values while still maintaining the shape of the curve.
```{r}
# log 
calc_2023 <- calc_2023 %>%
  mutate(log_calc = log(umol.CaCO3h.1.cm.2 + 1))

calc_shifted_2023 <- calc_2023 %>%
  filter(!is.na(Species)) %>%
  mutate(
    rate_shifted = log_calc,
    temp = Target.Temperature)
```

Fit multiple models to find best fit for our data 
```{r}
safe_fit_aic_v2 <- function(d, mod_name) {
  tryCatch({
    sv <- get_start_vals(d$temp, d$rate_shifted, model_name = mod_name)
    
    # We use a formula approach for each model to ensure parameter names match
    if(mod_name == 'pawar_2018'){
       fit <- nls_multstart(rate_shifted ~ pawar_2018(temp, r_tref, e, eh, topt, tref = 25),
                            data = d, iter = 100, start_lower = sv - 10, start_upper = sv + 10, supp_errors = 'Y')
    } 
    else if(mod_name == 'sharpeschoolhigh_1981'){
       fit <- nls_multstart(rate_shifted ~ sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 25),
                            data = d, iter = 100, start_lower = sv - 10, start_upper = sv + 10, supp_errors = 'Y')
    } 
    else if(mod_name == 'gaussian_1987'){
       fit <- nls_multstart(rate_shifted ~ gaussian_1987(temp, rmax, topt, a),
                            data = d, iter = 100, start_lower = sv - 10, start_upper = sv + 10, supp_errors = 'Y')
    } 
    else if(mod_name == 'boatman_2017'){
       fit <- nls_multstart(rate_shifted ~ boatman_2017(temp, rmax, tmin, tmax, topt, a),
                            data = d, iter = 100, start_lower = sv - 10, start_upper = sv + 10, supp_errors = 'Y')
    }
    
    return(MuMIn::AICc(fit))
  }, error = function(e) return(NA))
}

# Run all models 
all_group_aic_2023 <- calc_shifted_2023 %>%
  group_by(Species, Phenotype) %>%
  nest() %>%
  crossing(model_name = c('sharpeschoolhigh_1981', 'pawar_2018', 'boatman_2017', 'gaussian_1987')) %>%
  mutate(aic_score = map2_dbl(data, model_name, ~safe_fit_aic_v2(.x, .y))) %>%
  group_by(Species, Phenotype) %>%
  mutate(delta_AICc = aic_score - min(aic_score, na.rm = TRUE)) %>%
  arrange(Species, Phenotype, delta_AICc)
```

Pawar_2018 model is the best fit for 3 out of 4 groups (Pcomp susceptible had Boatman_2017 as a better fit but for comparability across groups, we will stick with Pawar_2018). 

Fit the pawar 2018 function and run bootstrap across groups 
```{r}
# Define the Pawar-based bootstrap function
fit_pawar_boot <- function(x) {
  mod <- 'pawar_2018'
  
  # Wider search range (+/- 10) for stability
  fit <- nls_multstart(rate_shifted ~ pawar_2018(temp, r_tref, e, eh, topt, tref = 25),
                       data = x,
                       iter = 50, 
                       start_lower = get_start_vals(x$temp, x$rate_shifted, model_name = mod) - 10,
                       start_upper = get_start_vals(x$temp, x$rate_shifted, model_name = mod) + 10,
                       lower = get_lower_lims(x$temp, x$rate_shifted, model_name = mod),
                       upper = get_upper_lims(x$temp, x$rate_shifted, model_name = mod),
                       supp_errors = 'Y')
  
  # Calculate derived parameters
  extra_params <- calc_params(fit)
  
  # Get the raw r_tref estimate from the model
  r_tref_val <- broom::tidy(fit) %>% filter(term == "r_tref") %>% pull(estimate)
  
  # Return: rmax, topt, e, eh, r_tref_25
  all_stats <- c(as.numeric(extra_params$rmax), 
                 as.numeric(extra_params$topt), 
                 as.numeric(extra_params$e), 
                 as.numeric(extra_params$eh),
                 as.numeric(r_tref_val))
  
  names(all_stats) <- c("rmax", "topt", "e", "eh", "r_tref_25")
  return(all_stats)
}

# Run the bootstrap
boot_results_2023 <- calc_shifted_2023 %>%
  group_by(Species, Phenotype) %>%
  nest() %>%
  mutate(boot = map(data, ~boot(data = .x, 
                                 statistic = function(d, i) as.numeric(fit_pawar_boot(d[i,])), 
                                 R = 1000,
                                 parallel = "multicore", 
                                 ncpus = n_cores)))

# Tidy the results for interpretation
boot_stats_2023 <- boot_results_2023 %>%
  mutate(stats = map(boot, ~broom::tidy(.x, conf.int = TRUE, conf.method = 'perc'))) %>%
  unnest(stats) %>%
  ungroup() %>%
  mutate(term = rep(c("rmax", "topt", "e", "eh", "r_tref_25"), 4))
```

Clean up data and create 5 panel plot
```{r}
# Clean up the NAs by using the average of the confidence intervals where statistic is NA
boot_stats_cleaned_2023 <- boot_stats_final %>%
  mutate(statistic = ifelse(is.na(statistic), (conf.low + conf.high) / 2, statistic))

# Define the 5 traits we want to plot
traits <- c("r_tref_25", "rmax", "topt", "e", "eh")

# Create a list to store the plots
plot_list <- list()

for (i in 1:length(traits)) {
  trait_name <- traits[i]
  df_sub <- subset(boot_stats_cleaned_2023, term == trait_name)
  
  # Set specific labels for each trait
  y_label <- case_when(
    trait_name == "r_tref_25" ~ "Rate at 25°C (log)",
    trait_name == "rmax"      ~ "Rmax (log)",
    trait_name == "topt"      ~ "Topt (°C)",
    trait_name == "e"         ~ "Activation (e)",
    trait_name == "eh"        ~ "Deactivation (eh)"
  )
  
  # Build the plot using your specific style
  p <- ggplot(df_sub, aes(x = Phenotype, y = statistic, color = Phenotype)) +
    geom_errorbar(aes(ymin = statistic - std.error, ymax = statistic + std.error), 
                  width = 0.15, linewidth = 1) +
    geom_point(size = 4) +
    facet_grid(~Species) + # Species side-by-side in each row
    scale_color_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
    theme_bw(base_size = 12) +
    theme(
      legend.position = "none",
      strip.background = element_rect(fill = "gray95"),
      strip.text = element_text(face = "bold"),
      panel.grid.minor = element_blank()
    ) +
    labs(y = y_label, x = NULL)
  
  # Hide x-axis text for all except the bottom-most plot
  if (i < length(traits)) {
    p <- p + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  }
  
  plot_list[[i]] <- p
}

# Combine all 5 plots into one tall column
final_5_panel <- wrap_plots(plot_list, ncol = 1)
print(final_5_panel)
```

Plot log scale calcification rates across temps 
```{r}
calc_plot_2023 <- calc_2023 %>%
  filter(!is.na(Phenotype),
         !is.na(Species),
         Colony.ID != "blank")

ggplot(calc_plot_2023, aes(
  x = Target.Temperature,
  y = log_calc,
  color = Phenotype,
  fill = Phenotype
)) +
  geom_point(size = 3, alpha = 0.5) +
  geom_smooth(method = "loess", se = TRUE, span = 0.6, alpha = 0.25) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  labs(
    x = "Target Temperature (°C)",
    y = expression("log(Net Calcification ("*cm^-2*"))"),
    color = "Phenotype",
    fill = "Phenotype")
```

Plot as PCA
```{r}
# 1. Extract all 1000 replicates for all 5 traits from the 'boot' column
pca_prep_clean <- boot_results_final %>%
  mutate(reps = map(boot, ~as.data.frame(.x$t))) %>%
  unnest(reps) %>%
  ungroup() %>%
  # Rename based on your specific column order
  rename(rmax = V1, topt = V2, e = V3, eh = V4, r_tref_25 = V5) %>%
  # REMOVE THE NAs
  drop_na(rmax, topt, e, eh, r_tref_25)

# 2. Perform PCA (Scaling is vital because Topt and Rmax have different magnitudes)
pca_fit <- prcomp(pca_prep_clean[, c("rmax", "topt", "e", "eh", "r_tref_25")], scale. = TRUE)

# 3. Plot with Loadings (Arrows)
autoplot(pca_fit, data = pca_prep_clean, colour = 'Phenotype', 
         shape = 'Species', frame = TRUE, frame.type = 'norm',
         loadings = TRUE, loadings.label = TRUE, 
         loadings.label.size = 3, loadings.colour = "black", loadings.label.colour = "black") +
  scale_color_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  scale_fill_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  theme_bw() +
  labs(title = "Multivariate Thermal Phenotype PCA",
       subtitle = "Bootstrap cloud showing separation of coral groups",
       caption = "Arrows indicate the direction of trait influence.")
```

# 2025

## Calculate calcification rates 

Read in data 
```{r}
calc <- read.csv("../data/TPC_calcification/Calcification_data_TPC.csv", header = T)
```

Calculate incubation time 
```{r}
calc <- calc %>%
  mutate(
    start = hm(Time_start),
    end   = hm(Time_end),
    elapsed_mins  = as.numeric(end - start, units = "mins"),
    elapsed_hours = elapsed_mins / 60)
```

Calculate delta alkalinity 
```{r}
calc <- calc %>%
  mutate(Delta_TA_raw = Starting_alk_value - Alk_avg)
```

Calculate blanks 
```{r}
blank_delta_ta <- calc %>%
  filter(ColonyID == "blank" | Notes == "blank") %>%
  group_by(Date, Target_Temperature) %>%
  summarise(mean_blank_delta_ta = mean(Delta_TA_raw, na.rm = TRUE))# %>%
  #pull(mean_blank_delta_ta)
```

Subtract blank and calculate net calcification
```{r}
# Correct Delta_TA by subtracting blank
calc <- calc %>%
  left_join(blank_delta_ta, by = c("Date", "Target_Temperature")) 

calc <- calc %>%
  mutate(
    Delta_TA_corr = Delta_TA_raw - mean_blank_delta_ta,
    # Assuming seawater density 1.028 g/mL and volume in mL to L, and no coral weight (thus calc mg/L/hr)
    Net_Calcification = (Delta_TA_corr / 2) * (Sample_Volume_mL / 1000) * 1.028 / elapsed_hours) %>%
  filter(!is.na(Net_Calcification) & !is.nan(Net_Calcification))
```

Normalize by SA
```{r}
calc <- calc %>%
  mutate(Net_Calcification_cm2 = Net_Calcification / Surface_Area)
```

Read in meta
```{r}
meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")
```

Join with calc 
```{r}
meta_unique <- meta %>%
  group_by(Colony_ID) %>%
  slice(1) %>%   # or use slice_max(Date) if you want the most recent entry
  ungroup()

calc_meta <- calc %>%
  left_join(meta_unique, by = c("ColonyID" = "Colony_ID")) %>%
  select(ColonyID, Target_Temperature, Net_Calcification_cm2, Species, Phenotype)
```

Plot 
```{r}
# Filter out blanks and missing phenotype/species
calc_plot <- calc_meta %>%
  filter(!is.na(Phenotype),
         !is.na(Species),
         ColonyID != "blank")

ggplot(calc_plot, aes(
  x = Target_Temperature,
  y = Net_Calcification_cm2,
  color = Phenotype,
  fill = Phenotype
)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "loess", se = TRUE, span = 0.6, alpha = 0.25) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  labs(
    x = "Target Temperature (°C)",
    y = expression("Net Calcification ("*cm^-2*")"),
    color = "Phenotype",
    fill = "Phenotype")
```

## rTPC 

Because some values are negative, we need to 'shift' the data away from negative values while still maintaining the shape of the curve.
```{r}
min_rate <- min(calc_meta$Net_Calcification_cm2, na.rm = TRUE)
shift_val <- abs(min_rate) + 0.01 

calc_shifted <- calc_meta %>%
  filter(!is.na(Species)) %>%
  mutate(
    rate_shifted = Net_Calcification_cm2 + shift_val,
    temp = Target_Temperature)

# calc_clean <- calc_meta %>%
#   filter(!is.na(Species)) %>%
#   mutate(group = paste(Species, Phenotype, sep = "_")) %>%
#   rename(temp = Target_Temperature, rate = Net_Calcification_cm2)
```

Fitting curves for each group
```{r}
mod_pawar <- 'pawar_2018'

fit_pawar <- calc_shifted %>%
  group_by(Species, Phenotype) %>%
  nest() %>%
  mutate(fit = map(data, ~nls_multstart(rate_shifted ~ pawar_2018(temp, r_tref, e, eh, th, tref = 25),
                                        data = .x,
                                        iter = 500,
                                        start_lower = get_start_vals(.x$temp, .x$rate_shifted, model_name = mod_pawar) - 1,
                                        start_upper = get_start_vals(.x$temp, .x$rate_shifted, model_name = mod_pawar) + 1,
                                        supp_errors = 'Y',
                                        convergence_count = 100)))

pawar_params <- fit_pawar %>%
  mutate(params = map(fit, ~calc_params(.x))) %>%
  unnest(params)
```

Bootstrapping TPC metrics
```{r}
# Function to fit model and return a vector of parameters
fit_n_calc <- function(x){
  mod <- 'pawar_2018'
  fit <- nls_multstart(rate_shifted ~ pawar_2018(temp, r_tref, e, eh, th, tref = 25),
                       data = x,
                       iter = 10, 
                       start_lower = get_start_vals(x$temp, x$rate_shifted, model_name = mod) - 1,
                       start_upper = get_start_vals(x$temp, x$rate_shifted, model_name = mod) + 1,
                       supp_errors = 'Y')
  return(calc_params(fit))
}

# Run the bootstrap
n_cores <- detectCores() - 1
boot_results <- calc_shifted %>%
  group_by(Species, Phenotype) %>%
  nest() %>%
  mutate(boot = map(data, ~boot(data = .x, 
                                 statistic = function(d, i) as.numeric(fit_n_calc(d[i,])), 
                                 R = 1000,
                                 parallel = "multicore", 
                                 ncpus = n_cores)))
```

Unnest bootstrap results 
```{r}
boot_stats <- boot_results %>%
  mutate(stats = map(boot, ~tidy(.x, conf.int = TRUE, conf.method = 'perc'))) %>%
  unnest(stats) %>%
  ungroup() %>% 
  mutate(term = rep(names(calc_params(fit_ss$fit[[1]])), 4))
```

Look at specific metrics 
```{r}
# Filter for Topt to check for Phenotype differences
topt_comparison <- boot_stats %>%
  filter(term == "e") %>%
  select(Species, Phenotype, statistic, std.error, conf.low, conf.high)

ggplot(topt_comparison, aes(x = Phenotype, y = statistic, fill = Phenotype)) +
  geom_point() +
  #geom_bar(stat = "identity", width = 0.6, alpha = 0.8) + # Using bars to match your ref
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  facet_wrap(~Species) +
  theme_classic() +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  labs(y = "Acute Thermal Optimum (ºC)", 
       title = "Thermal Optimum Comparison",
       subtitle = "Error bars represent 95% Bootstrap Confidence Intervals")
```


