---
title: "TPCs"
author: "Jill Ashey"
date: "2025-09-15"
output: html_document
---

Interesting: https://www.nature.com/articles/s41467-024-53046-2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(purrr)
library(minpack.lm)  # for nlsLM
library(tidyverse)
library(broom)
library(readr)
library(lubridate)
library(purrr)
library(furrr)
library(fuzzyjoin)
library(future)
library(LoLinR)
library(rTPC)
library(nls.multstart)
```

Testing use of rTPC package from [vignette](https://padpadpadpad.github.io/rTPC/articles/rTPC.html)
```{r}
# load in data
data("chlorella_tpc")

# keep just a single curve
d <- filter(chlorella_tpc, curve_id == 1)

# choose model
mod = 'pawar_2018'

# get start vals
start_vals <- get_start_vals(d$temp, d$rate, model_name = 'pawar_2018')

# get limits
low_lims <- get_lower_lims(d$temp, d$rate, model_name = 'pawar_2018')
upper_lims <- get_upper_lims(d$temp, d$rate, model_name = 'pawar_2018')

start_vals
low_lims
upper_lims

# fit model
fit <- nls_multstart(rate~pawar_2018(temp = temp, r_tref,e,eh,th, tref = 15),
                                                     data = d,
                                                     iter = 500,
                                                     start_lower = start_vals - 10,
                                                     start_upper = start_vals + 10,
                                                     lower = low_lims,
                                                     upper = upper_lims,
                                                     supp_errors = 'Y')

fit

# calculate additional traits
calc_params(fit) %>%
  # round for easy viewing
  mutate_all(round, 2)

# predict new data
new_data <- data.frame(temp = seq(min(d$temp), max(d$temp), 0.5))
preds <- augment(fit, newdata = new_data)

# plot data and model fit
ggplot(d, aes(temp, rate)) +
  geom_point() +
  geom_line(aes(temp, .fitted), preds, col = 'blue') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Respiration across temperatures')
```



# Photosynthesis
## Sept 2023

### Prepare data 

Load run metadata 
```{r}
run_meta <- read_csv("../data/TPC/TPC_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
```

Load sample metadata 
```{r}
sample_meta <- read_csv("../data/TPC/TPC_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```

List all raw files for Sept 2023 
```{r}
files <- list.files(path = "../data/TPC/raw_data/Sept2023/", pattern = "*.csv", full.names = TRUE)
```

Join O2 data with metadata files based on start and stop times 
```{r}
all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "channel\\d+") %>% str_remove("channel") %>% as.numeric()
  
  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )
  
  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )
  
  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )
  
  return(raw_joined)
})
```

Look at joined data and save
```{r}
all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

write_csv(all_data, "../output/TPC/Sept2023/TPC_all_joined_data_Sept2023.csv")
```

### Run localized linear regressions 

The interval fitting code will take about an hour to run and is computationally intense. If needed, run as a background job or on a remote server.

Make list of all Colony_ID / Run combinations (including blanks)
```{r}
colony_runs <- all_data %>%
  filter(!is.na(Colony_ID)) %>%
  distinct(Colony_ID, Run)
```

For this timepoint, only two out of the five runs had blanks. Therefore, I am going to calculate the blank intervals, average them, and then apply them to the regression for each colony. 

Calculate interval slopes for the blanks
```{r}
blank_data <- all_data %>%
  filter(Colony_ID == "blank")

blank_intervals <- blank_data %>%
  group_by(Run, Temperature.y) %>%
  filter(!is.na(Colony_ID)) %>%
  group_split()
```

Fit regressions and extract slopes for each blank interval
```{r}
fit_reg <- function(df) {
  rankLocReg(
    xall = as.numeric(df$`Delta T [min]`),
    yall = df$Oxygen,
    alpha = 0.2,
    method = "pc",
    verbose = FALSE
  )
}

blank_fits <- map(blank_intervals, fit_reg)
blank_slopes <- map_dbl(blank_fits, ~ pluck(., "allRegs", "b1", 1))
blank_summary <- map2_dfr(
  blank_intervals,
  blank_slopes,
  ~ tibble(
    Temperature.y = unique(.x$Temperature.y),
    Blank_Slope = .y
  )
)
```

Read in interval data 
```{r}
all_int_tpc <- read_csv("../output/TPC/Sept2023/all_interval_summaries_Sept2023_TPC.csv")
all_int_tpc$Colony_ID <- as.character(all_int_tpc$Colony_ID)
```

Filter sample metadata for Sept 2023
```{r}
sept_2023 <- sample_meta %>%
  filter(grepl("^2023-09", Date))
sept_2023$Surface_Area_cm2 <- as.numeric(sept_2023$Surface_Area_cm2)
```

Merge sample metadata with interval data 
```{r}
all_int_tpc <- all_int_tpc %>%
  inner_join(sample_meta, by = c("Colony_ID" = "Colony_ID", "Temperature" = "Target_Temperature")) %>%
  filter(!is.na(Phenotype))
length(unique(all_int_tpc$Colony_ID))
```

Normalize interval data by chamber volume and surface area and save
```{r}
# normalize 
all_int_tpc <- all_int_tpc %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Blank_Corrected_Slope * Chamber_Volume_L / Surface_Area_cm2) * 60) # units are umol O2 cm-2 h-1
write.csv(all_int_tpc, "../output/TPC/Sept2023/all_interval_summaries_normalized_Sept2023_TPC.csv")
```

Read data back in so I don't have to run above code again
```{r}
all_int_tpc <- read.csv("../output/TPC/Sept2023/all_interval_summaries_normalized_Sept2023_TPC.csv")
```

Plot curves across temperatures
```{r}
ggplot(all_int_tpc, aes(x = Temperature, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "umol O2 cm-2 h-1", color = "Phenotype")

ggplot(all_int_tpc, aes(x = Temperature, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(aes(fill = Phenotype), method = "loess", se = TRUE, linewidth = 1.2) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "umol O2 cm-2 h-1", color = "Phenotype")
```

Log and plot 
```{r}
all_int_tpc$log.rate <- log(all_int_tpc$Slope_norm + 1)  #logging and adding 0.1 because a log of zero does not exist

ggplot(all_int_tpc, aes(x = Temperature, y = log.rate, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "umol O2 cm-2 h-1", color = "Phenotype")

ggplot(all_int_tpc, aes(x = Temperature, y = log.rate, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(aes(fill = Phenotype), method = "loess", se = TRUE, linewidth = 1.2) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "log(umol O2 cm-2 h-1)", color = "Phenotype")
```




