---
title: "TPCs"
author: "Jill Ashey"
date: "2025-09-15"
output: html_document
---

Interesting: https://www.nature.com/articles/s41467-024-53046-2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(car)
library(tidyverse)
library(broom)
library(readr)
library(lubridate)
library(purrr)
library(furrr)
library(fuzzyjoin)
library(future)
library(rTPC)
library(nls.multstart)
library(emmeans)
library(lme4)
library(nlme)
library(Rmisc)
library(boot)
library(nlme)
library(propagate)
```

# Functions for analysis 

```{r}
sharpe_schoolfield <- function(lnc, E, Eh, Th, temp, Tc) {
  Tc <- 273.15 + Tc
  k <- 8.62e-5
  boltzmann.term <- lnc + log(exp(E/k*(1/Tc - 1/temp))) # units are eV/K, electrovolts/Kelvin
  inactivation.term <- log(1/(1 + exp(Eh/k*(1/Th - 1/temp))))
  return(boltzmann.term + inactivation.term)}

get_topt <- function(E, Th, Eh){
  k <- 8.62e-05
  return((Eh * Th) / (Eh + (k * Th * log((Eh / E) - 1))))}

get_tpc_ci <- function(model, new_data) {
  # predictNLS calculates the intervals based on the Taylor expansion (Delta Method)
  pred <- predictNLS(model, newdata = new_data, interval = "confidence", alpha = 0.05)
  return(as.data.frame(pred$summary))}
```

# Photosynthesis
## Sept 2023

### Prepare data 

Load run metadata 
```{r}
run_meta <- read_csv("../data/TPC/TPC_metadata_run.csv") %>%
  mutate(
    Start = as_datetime(Date) + hours(hour(Start.time)) + minutes(minute(Start.time)),
    Stop  = as_datetime(Date) + hours(hour(Stop.time)) + minutes(minute(Stop.time)))
```

Load sample metadata 
```{r}
sample_meta <- read_csv("../data/TPC/TPC_metadata_sample.csv")

# Make sure Chamber well # is numeric
sample_meta <- sample_meta %>%
  mutate(`Chamber` = as.numeric(`Chamber`))
```

List all raw files for Sept 2023 
```{r}
files <- list.files(path = "../data/TPC/raw_data/Sept2023/", pattern = "*.csv", full.names = TRUE)
```

Join O2 data with metadata files based on start and stop times 
```{r}
all_data <- map_dfr(files, function(f) {
  # Extract channel number from filename, e.g. Ch1 → 1
  channel <- str_extract(f, "channel\\d+") %>% str_remove("channel") %>% as.numeric()
  
  # Read the file
  raw <- read_csv(f) %>%
    mutate(
      # Parse datetime (adjust format if needed)
      DateTime = dmy_hms(Date),
      Channel = channel
    )
  
  # Join to run metadata by fuzzy time match
  raw_joined <- fuzzy_left_join(
    raw,
    run_meta,
    by = c("DateTime" = "Start", "DateTime" = "Stop"),
    match_fun = list(`>=`, `<=`)
  )
  
  # Join to sample metadata by Date, Run #, and Channel
  raw_joined <- raw_joined %>%
    left_join(
      sample_meta,
      by = c(
        "Date.y" = "Date",
        "Run" = "Run",
        "Channel" = "Chamber"
      )
    )
  
  return(raw_joined)
})
```

Look at joined data and save
```{r}
all_data %>% count(is.na(`Colony_ID`))
all_data %>% count(is.na(Run))

write_csv(all_data, "../output/TPC/Sept2023/TPC_all_joined_data_Sept2023.csv")
```

### Run localized linear regressions 

The interval fitting code will take about an hour to run and is computationally intense. If needed, run as a background job or on a remote server.

Make list of all Colony_ID / Run combinations (including blanks)
```{r}
colony_runs <- all_data %>%
  filter(!is.na(Colony_ID)) %>%
  distinct(Colony_ID, Run)
```

For this timepoint, only two out of the five runs had blanks. Therefore, I am going to calculate the blank intervals, average them, and then apply them to the regression for each colony. 

Calculate interval slopes for the blanks
```{r}
blank_data <- all_data %>%
  filter(Colony_ID == "blank")

blank_intervals <- blank_data %>%
  group_by(Run, Temperature.y) %>%
  filter(!is.na(Colony_ID)) %>%
  group_split()
```

Fit regressions and extract slopes for each blank interval
```{r}
fit_reg <- function(df) {
  rankLocReg(
    xall = as.numeric(df$`Delta T [min]`),
    yall = df$Oxygen,
    alpha = 0.2,
    method = "pc",
    verbose = FALSE
  )
}

blank_fits <- map(blank_intervals, fit_reg)
blank_slopes <- map_dbl(blank_fits, ~ pluck(., "allRegs", "b1", 1))
blank_summary <- map2_dfr(
  blank_intervals,
  blank_slopes,
  ~ tibble(
    Temperature.y = unique(.x$Temperature.y),
    Blank_Slope = .y
  )
)
```

Read in interval data 
```{r}
all_int_tpc <- read_csv("../output/TPC/Sept2023/all_interval_summaries_Sept2023_TPC.csv")
all_int_tpc$Colony_ID <- as.character(all_int_tpc$Colony_ID)
```

Filter sample metadata for Sept 2023
```{r}
sept_2023 <- sample_meta %>%
  filter(grepl("^2023-09", Date))
sept_2023$Surface_Area_cm2 <- as.numeric(sept_2023$Surface_Area_cm2)
```

Merge sample metadata with interval data 
```{r}
all_int_tpc <- all_int_tpc %>%
  inner_join(sample_meta, by = c("Colony_ID" = "Colony_ID", "Temperature" = "Target_Temperature")) %>%
  filter(!is.na(Phenotype))
length(unique(all_int_tpc$Colony_ID))
```

Normalize interval data by chamber volume and surface area and save
```{r}
# normalize 
all_int_tpc <- all_int_tpc %>%
  mutate(
    Chamber_Volume_L = Chamber_Volume_mL / 1000,  # mL → L
    Slope_norm = (Blank_Corrected_Slope * Chamber_Volume_L / Surface_Area_cm2) * 60) # units are umol O2 cm-2 h-1
write.csv(all_int_tpc, "../output/TPC/Sept2023/all_interval_summaries_normalized_Sept2023_TPC.csv")
```

Read data back in so I don't have to run above code again
```{r}
all_int_tpc <- read.csv("../output/TPC/Sept2023/all_interval_summaries_normalized_Sept2023_TPC.csv")
```

Plot curves across temperatures
```{r}
ggplot(all_int_tpc, aes(x = Temperature, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "umol O2 cm-2 h-1", color = "Phenotype")

ggplot(all_int_tpc, aes(x = Temperature, y = Slope_norm, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(aes(fill = Phenotype), method = "loess", se = TRUE, linewidth = 1.2) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "umol O2 cm-2 h-1", color = "Phenotype")
```

Log and plot 
```{r}
all_int_tpc$log.rate <- log(all_int_tpc$Slope_norm + 1)  #logging and adding 0.1 because a log of zero does not exist

ggplot(all_int_tpc, aes(x = Temperature, y = log.rate, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point() +
  geom_line(aes(group = Colony_ID)) + 
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "umol O2 cm-2 h-1", color = "Phenotype")

ggplot(all_int_tpc, aes(x = Temperature, y = log.rate, color = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(aes(fill = Phenotype), method = "loess", se = TRUE, linewidth = 1.2) +
  scale_color_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(
    values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  labs(x = "Temperature (C)", y = "log(umol O2 cm-2 h-1)", color = "Phenotype")
```

## Fit TPC curve

Set target temperature to numeric and convert temperature to Kelvin
```{r}
photo_2023 <- all_int_tpc %>%
  mutate(Temperature = as.numeric(Temperature),
         Temperature.K = Temperature + 273.15)
```

Fit curves for all colonies using `sharpe_schoolfield` function defined above
```{r}
fits_2023 <- photo_2023 %>%
  group_by(Phenotype,Species,Colony_ID) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(log.rate ~ sharpe_schoolfield(lnc, E, Eh, Th, temp = Temperature.K, Tc = 27),
                                                data = .x,
                                                iter = 1000,
                                                start_lower = c(lnc = -10, E = 0.1, Eh = 0.2, Th = 285),
                                                start_upper = c(lnc = 10, E = 2, Eh = 5, Th = 330),
                                                supp_errors = 'Y',
                                                na.action = na.omit,
                                                lower = c(lnc = -10, E = 0, Eh = 0, Th = 0))))
```

Get TPC parameters for each colony 
```{r}
params_2023 <- fits_2023 %>%
  unnest_legacy(fit %>% purrr::map(tidy)) %>%
  filter(!Colony_ID == "233")
```

Calculate Topt and Rmax for each colony 
```{r}
Topt_Rmax_2023 <- params_2023 %>%
  dplyr::select(Colony_ID, Phenotype, Species, term, estimate) %>%
  tidyr::pivot_wider(names_from = term, values_from = estimate) %>%
  mutate(
    Topt_K = get_topt(E, Th, Eh),
    Topt_C = Topt_K - 273.15,
    Rmax_log = sharpe_schoolfield(lnc, E, Eh, Th, temp = Topt_K, Tc = 27),
    Rmax_linear = exp(Rmax_log) - 1)
```

Pivot data so that each trait is a row 
```{r}
traits_long_2023 <- Topt_Rmax_2023 %>%
  dplyr::select(Colony_ID, Phenotype, Species, lnc, E, Eh, Topt_C, Rmax_log) %>%
  pivot_longer(cols = c(lnc, E, Eh, Topt_C, Rmax_log), 
               names_to = "trait", 
               values_to = "estimate")
```

## lnc (r_tref) - 2023

Subset by lnc data
```{r}
lnc_long_2023 <- traits_long_2023 %>%
  filter(trait == "lnc")
```

Check for normality
```{r}
hist(lnc_long_2023$estimate)

# Q-Q Plot
qqnorm(lnc_long_2023$estimate)
qqline(lnc_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(lnc_long_2023$estimate)
```

Data is normal! 

Run linear mixed effects model 
```{r}
lnc.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony_ID, data=lnc_long_2023, na.action=na.exclude)
summary(lnc.lme)
anova(lnc.lme)
```

Species and Phenotpye are significant. 

## E - 2023

Subset by E data
```{r}
e_long_2023 <- traits_long_2023 %>%
  filter(trait == "E")
```

Check for normality
```{r}
hist(e_long_2023$estimate)

# Q-Q Plot
qqnorm(e_long_2023$estimate)
qqline(e_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(e_long_2023$estimate)
```

Data is not normal. Deal with this!

Run linear mixed effects model 
```{r}
e.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony_ID, data=e_long_2023, na.action=na.exclude)
summary(e.lme)
anova(e.lme)
```

Phenotype is significant. 

## Eh - 2023

Subset by Eh data
```{r}
eh_long_2023 <- traits_long_2023 %>%
  filter(trait == "Eh")
```

Check for normality
```{r}
hist(eh_long_2023$estimate)

# Q-Q Plot
qqnorm(eh_long_2023$estimate)
qqline(eh_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(eh_long_2023$estimate)
```

Data is normal!

Run linear mixed effects model 
```{r}
eh.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony_ID, data=eh_long_2023, na.action=na.exclude)
summary(eh.lme)
anova(eh.lme)
```

Species is significant. 

## Topt - 2023

Subset by Topt data
```{r}
topt_long_2023 <- traits_long_2023 %>%
  filter(trait == "Topt_C")
```

Check for normality
```{r}
hist(topt_long_2023$estimate)

# Q-Q Plot
qqnorm(topt_long_2023$estimate)
qqline(topt_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(topt_long_2023$estimate)
```

Data is not normal, come back to this. 

Run linear mixed effects model 
```{r}
topt.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony_ID, data=topt_long_2023, na.action=na.exclude)
summary(topt.lme)
anova(topt.lme)
```

No significant differences. 

## Rmax_log - 2023

Subset by Rmax_log data
```{r}
rmax_long_2023 <- traits_long_2023 %>%
  filter(trait == "Rmax_log")
```

Check for normality
```{r}
hist(rmax_long_2023$estimate)

# Q-Q Plot
qqnorm(rmax_long_2023$estimate)
qqline(rmax_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(rmax_long_2023$estimate)
```

Data is normal! 

Run linear mixed effects model 
```{r}
rmax.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony_ID, data=rmax_long_2023, na.action=na.exclude)
summary(rmax.lme)
anova(rmax.lme)
```

Species and Phenotype are significant. 

## Visualization for 2023

Plot
```{r}
# Create the point plot with Mean and SE
ggplot(traits_long_2023, aes(x = Phenotype, y = estimate, color = Phenotype)) +
  # Add raw colony data as small, transparent points to show the spread
  geom_jitter(width = 0.1, alpha = 0.3, size = 1.5) +
  stat_summary(fun = mean, geom = "point", size = 4) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, linewidth = 1) +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "2023 Thermal Performance Traits",
    y = "Estimate (Mean ± SE)",
    x = "Phenotype"
  )

## Box plot 
ggplot(traits_long_2023, aes(x = Phenotype, y = estimate, fill = Phenotype)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.6, color = "black") +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.2, color = "black") +
  facet_grid(trait ~ Species, scales = "free_y") +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 11),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold", size = 13),
    panel.grid.major.x = element_blank() # Clean up vertical lines
  ) +
  labs(
    y = "Estimated Parameter Value",
    x = "Phenotype"
  )
```

Multidimensional visualization via PCA
```{r}
# Select only numeric traits for PCA
pca_prep <- Topt_Rmax_2023 %>%
  ungroup() %>%
  dplyr::select(lnc, E, Eh, Topt_C, Rmax_log)

# Run PCA (scale = TRUE is essential here)
pca_fit <- prcomp(pca_prep, scale. = TRUE)

# Plot
autoplot(pca_fit, data = Topt_Rmax_2023, colour = 'Phenotype', 
         shape = 'Species', frame = TRUE, frame.type = 'norm',
         loadings = TRUE, loadings.label = TRUE, 
         loadings.label.size = 3, loadings.colour = "black", loadings.label.colour = "black") +
  scale_color_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  scale_fill_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  theme_bw() +
  labs(title = "Multivariate Thermal Phenotype PCA",
       caption = "Arrows indicate the direction of trait influence.")
```

## Plot with confidence intervals 

Generate and aggregate predictions
```{r}
# 1. Define the temperature range for the smooth curves
new_data <- data.frame(
  Temperature.K = seq(18 + 273.15 - 1,  # 17°C
                      38 + 273.15 + 1,  # 39°C
                      length.out = 200))

# 2. Generate predictions for every colony model
# This creates a "long" dataframe of all potential curves
preds_all <- fits_2023 %>%
  mutate(preds = map(fit, ~{
    # We use a safe predict call in case some models didn't converge
    p <- try(predict(.x, newdata = new_data), silent = TRUE)
    if(inherits(p, "try-error")) return(NULL)
    data.frame(Temperature.K = new_data$Temperature.K, pred_val = p)
  })) %>%
  filter(!map_lgl(preds, is.null)) %>% # Remove failed fits
  unnest(preds)

# 3. Calculate the MEAN and CI of the curves per group
# At each temperature step, find the mean performance across all colonies
preds_mean_summary <- preds_all %>%
  group_by(Phenotype, Species, Temperature.K) %>%
  dplyr::summarise(
    mean_val = mean(pred_val, na.rm = TRUE),
    sd_val = sd(pred_val, na.rm = TRUE),
    n = n(),
    se_val = sd_val / sqrt(n),
    # 95% Confidence Interval of the Mean
    lower = mean_val - (1.96 * se_val),
    upper = mean_val + (1.96 * se_val),
    .groups = "drop"
  )
```

Plot TPC with fitted curves 
```{r}
ggplot(preds_mean_summary, aes(x = Temperature.K - 273.15, y = mean_val, color = Phenotype, fill = Phenotype)) +
  geom_hline(yintercept = 0, alpha = 0.5, linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5, color = NA) +
  geom_line(linewidth = 1.5) +
  scale_x_continuous(
    breaks = seq(18, 40, by = 2), 
    limits = c(17, 40)) +
  geom_point(data = photo_2023, aes(x = Temperature.K - 273.15, y = Slope_norm), 
             alpha = 0.2, size = 1, inherit.aes = TRUE) +
  facet_wrap(~ Species) +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 12),
    panel.grid = element_blank()
  ) +
  labs(
    x = "Temperature (°C)", 
    y = expression(bold("Net photosynthesis (log(" *mu*"mol " *cm^-2 *hr^-1*"))")),
    title = "Photosynthesis Fitted Thermal Performance Curves (2023)"
  )
```

## Plot with confidence intervals and the environmental data 

```{r}
temps <- read.csv("../output/environmental/NSF_CAREER_KBay_PR13_Temps.csv")
```

Subset for the year prior to Sept 23, 2023
```{r}
temps_sub <- temps %>%
  filter(DateTime <= "2023-09-23 01:00:00")
```

```{r}
temp_freq <- temps_sub %>%
  # Rounding to the nearest whole degree centers the bar on that number
  mutate(temp_bin = round(Temp_C, 0)) %>% 
  group_by(temp_bin) %>%
  dplyr::summarise(count = n(), .groups = "drop") %>%
  # Calculate the percentage of total time spent at that temperature
  mutate(percent = (count / sum(count)) * 100)
```

Plot with temp bars
```{r}
# Axis Parameters
p_min <- -0.5  # Hard floor of the plot
p_max <- 1.5   # Peak of the Y-axis
p_total <- p_max - p_min
s_max <- 30    # Max percent on the right axis
scale_factor <- p_total / s_max

ggplot() +
  geom_rect(data = temp_freq, 
            aes(xmin = temp_bin - 0.45, xmax = temp_bin + 0.45, 
                ymin = p_min, ymax = p_min + (percent * scale_factor)), 
            fill = "#8DBBCC", color = "#727E8B", linewidth = 0.2) +
  geom_point(data = photo_2023, 
             aes(x = Temperature.K - 273.15, y = log.rate, color = Phenotype), 
             alpha = 0.5, size = 1.2, position = position_jitter(width = 0.1)) +
  geom_ribbon(data = preds_mean_summary, 
              aes(x = Temperature.K - 273.15, ymin = lower, ymax = upper, fill = Phenotype), 
              alpha = 0.3, color = NA) +
  geom_hline(yintercept = 0, color = "grey", linewidth = 0.5, linetype = "dashed") +
  geom_line(data = preds_mean_summary, 
            aes(x = Temperature.K - 273.15, y = mean_val, color = Phenotype), 
            linewidth = 1.3) +
  scale_y_continuous(
    name = expression("Net photosynthesis (log(" *mu*"mol " *cm^-2 *hr^-1*"))"),
    limits = c(p_min, p_max),
    expand = c(0, 0),
    sec.axis = sec_axis(~ (. - p_min) / scale_factor, name = "Percent days (%)", breaks = seq(0, 40, 5))
  ) +
  scale_x_continuous(breaks = seq(18, 40, by = 2), limits = c(17.5, 40.5), expand = c(0,0)) +
  facet_wrap(~ Species, nrow = 1) +
  scale_color_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  scale_fill_manual(values = c("Resistant" = "#3b244d", "Susceptible" = "#f2916b")) +
  theme_classic() +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 13),
    axis.line = element_line(color = "black"),
    panel.spacing = unit(2, "lines")
  ) +
  labs(x = "Temperature (°C)")
```






