---
title: "Net calcification TPC"
author: "Jill Ashey"
date: "2025-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(car)
library(tidyverse)
library(broom)
library(readr)
library(lubridate)
library(purrr)
library(furrr)
library(fuzzyjoin)
library(future)
library(rTPC)
library(nls.multstart)
library(emmeans)
library(lme4)
library(nlme)
library(Rmisc)
library(boot)
library(nlme)

n_cores <- detectCores() - 1
```

# Functions for analysis 

```{r}
sharpe_schoolfield <- function(lnc, E, Eh, Th, temp, Tc) {
  Tc <- 273.15 + Tc
  k <- 8.62e-5
  boltzmann.term <- lnc + log(exp(E/k*(1/Tc - 1/temp))) # units are eV/K, electrovolts/Kelvin
  inactivation.term <- log(1/(1 + exp(Eh/k*(1/Th - 1/temp))))
  return(boltzmann.term + inactivation.term)}

get_topt <- function(E, Th, Eh){
  k <- 8.62e-05
  return((Eh * Th) / (Eh + (k * Th * log((Eh / E) - 1))))}
```

# 2023

Read in data 
```{r}
calc_2023 <- read.csv("../data/TPC_calcification/TPC_calcification_all_standardized_without_blanks_2023_KB.csv")
```

Set target temperature to numeric, convert temperature to Kelvin and log transform the calcification data 
```{r}
calc_2023 <- calc_2023 %>%
  mutate(Target.Temperature = as.numeric(Target.Temperature),
         Target.Temperature.K = Target.Temperature + 273.15,
         log.rate = log(umol.CaCO3h.1.cm.2 + 1))
```

Fit curves for all colonies using `sharpe_schoolfield` function defined above
```{r}
fits_2023 <- calc_2023 %>%
  group_by(Phenotype,Species,Colony.ID) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(log.rate ~ sharpe_schoolfield(lnc, E, Eh, Th, temp = Target.Temperature.K, Tc = 27),
                                                data = .x,
                                                iter = 1000,
                                                start_lower = c(lnc = -10, E = 0.1, Eh = 0.2, Th = 285),
                                                start_upper = c(lnc = 10, E = 2, Eh = 5, Th = 330),
                                                supp_errors = 'Y',
                                                na.action = na.omit,
                                                lower = c(lnc = -10, E = 0, Eh = 0, Th = 0))))
```

Get TPC parameters for each colony 
```{r}
params_2023 <- fits_2023 %>%
  unnest_legacy(fit %>% purrr::map(tidy))

# Remove Colonies 202 and 242 
params_2023 <- params_2023 %>%
  filter(!Colony.ID == "202") %>%
  filter(!Colony.ID == "242")
```

Calculate Topt and Rmax for each colony 
```{r}
Topt_Rmax_2023 <- params_2023 %>%
  dplyr::select(Colony.ID, Phenotype, Species, term, estimate) %>%
  tidyr::pivot_wider(names_from = term, values_from = estimate) %>%
  mutate(
    Topt_K = get_topt(E, Th, Eh),
    Topt_C = Topt_K - 273.15,
    Rmax_log = sharpe_schoolfield(lnc, E, Eh, Th, temp = Topt_K, Tc = 27),
    Rmax_linear = exp(Rmax_log) - 1)
```

Pivot data so that each trait is a row 
```{r}
traits_long_2023 <- Topt_Rmax_2023 %>%
  select(Colony.ID, Phenotype, Species, lnc, E, Eh, Topt_C, Rmax_log) %>%
  pivot_longer(cols = c(lnc, E, Eh, Topt_C, Rmax_log), 
               names_to = "trait", 
               values_to = "estimate")
  
```

## lnc (r_tref) - 2023

Subset by lnc data
```{r}
lnc_long_2023 <- traits_long_2023 %>%
  filter(trait == "lnc")
```

Check for normality
```{r}
hist(lnc_long_2023$estimate)

# Q-Q Plot
qqnorm(lnc_long_2023$estimate)
qqline(lnc_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(lnc_long_2023$estimate)
```

Data is normal! 

Run linear mixed effects model 
```{r}
lnc.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony.ID, data=lnc_long_2023, na.action=na.exclude)
summary(lnc.lme)
anova(lnc.lme)
```

Species is significant. Species x Phenotype is moderately significant (p=0.054)

## E - 2023

Subset by E data
```{r}
e_long_2023 <- traits_long_2023 %>%
  filter(trait == "E")
```

Check for normality
```{r}
hist(e_long_2023$estimate)

# Q-Q Plot
qqnorm(e_long_2023$estimate)
qqline(e_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(e_long_2023$estimate)
```

Data is normal! 

Run linear mixed effects model 
```{r}
e.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony.ID, data=e_long_2023, na.action=na.exclude)
summary(e.lme)
anova(e.lme)
```

Species is significant. 

## Eh - 2023

Subset by Eh data
```{r}
eh_long_2023 <- traits_long_2023 %>%
  filter(trait == "Eh")
```

Check for normality
```{r}
hist(eh_long_2023$estimate)

# Q-Q Plot
qqnorm(eh_long_2023$estimate)
qqline(eh_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(eh_long_2023$estimate)
```

Data is not normal, apply transformations. 
```{r}
# Log10
eh_long_2023$eh_log <- log10(eh_long_2023$estimate +1) # add 1 to avoid log10(0)
hist(eh_long_2023$eh_log)
qqnorm(eh_long_2023$eh_log)
qqline(eh_long_2023$eh_log, col = "red")
shapiro.test(eh_long_2023$eh_log)
```

Run linear mixed effects model 
```{r}
eh.lme <- lme(eh_log~Species*Phenotype, random = ~1|Colony.ID, data=eh_long_2023, na.action=na.exclude)
summary(eh.lme)
anova(eh.lme)
```

No significant differences. 

## Topt - 2023

Subset by Topt data
```{r}
topt_long_2023 <- traits_long_2023 %>%
  filter(trait == "Topt_C")
```

Check for normality
```{r}
hist(topt_long_2023$estimate)

# Q-Q Plot
qqnorm(topt_long_2023$estimate)
qqline(topt_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(topt_long_2023$estimate)
```

Data is normal! 

Run linear mixed effects model 
```{r}
topt.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony.ID, data=topt_long_2023, na.action=na.exclude)
summary(topt.lme)
anova(topt.lme)
```

No significant differences. 

## Rmax_log - 2023

Subset by Rmax_log data
```{r}
rmax_long_2023 <- traits_long_2023 %>%
  filter(trait == "Rmax_log")
```

Check for normality
```{r}
hist(rmax_long_2023$estimate)

# Q-Q Plot
qqnorm(rmax_long_2023$estimate)
qqline(rmax_long_2023$estimate, col = "red")

# Shapiro-Wilk Test
shapiro.test(rmax_long_2023$estimate)
```

Data is normal! 

Run linear mixed effects model 
```{r}
rmax.lme <- lme(estimate~Species*Phenotype, random = ~1|Colony.ID, data=rmax_long_2023, na.action=na.exclude)
summary(rmax.lme)
anova(rmax.lme)
```

Species is significant. 

## Visualization for 2023

```{r}
# Select only numeric traits for PCA
pca_prep <- Topt_Rmax_2023 %>%
  ungroup() %>%
  select(lnc, E, Eh, Topt_C, Rmax_log)

# Run PCA (scale = TRUE is essential here)
pca_fit <- prcomp(pca_prep, scale. = TRUE)

# Plot
autoplot(pca_fit, data = Topt_Rmax_2023, colour = 'Phenotype', 
         shape = 'Species', frame = TRUE, frame.type = 'norm',
         loadings = TRUE, loadings.label = TRUE, 
         loadings.label.size = 3, loadings.colour = "black", loadings.label.colour = "black") +
  scale_color_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  scale_fill_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  theme_bw() +
  labs(title = "Multivariate Thermal Phenotype PCA",
       caption = "Arrows indicate the direction of trait influence.")
```










## rTPC for 2023 

Because some values are negative, we need to 'shift' the data away from negative values while still maintaining the shape of the curve.
```{r}
# log 
calc_2023 <- calc_2023 %>%
  mutate(log_calc = log(umol.CaCO3h.1.cm.2 + 1))

calc_shifted_2023 <- calc_2023 %>%
  filter(!is.na(Species)) %>%
  mutate(
    rate_shifted = log_calc,
    temp = Target.Temperature)
```

Fit multiple models to find best fit for our data 
```{r}
safe_fit_aic_v2 <- function(d, mod_name) {
  tryCatch({
    sv <- get_start_vals(d$temp, d$rate_shifted, model_name = mod_name)
    
    # We use a formula approach for each model to ensure parameter names match
    if(mod_name == 'pawar_2018'){
       fit <- nls_multstart(rate_shifted ~ pawar_2018(temp, r_tref, e, eh, topt, tref = 25),
                            data = d, iter = 100, start_lower = sv - 10, start_upper = sv + 10, supp_errors = 'Y')
    } 
    else if(mod_name == 'sharpeschoolhigh_1981'){
       fit <- nls_multstart(rate_shifted ~ sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 25),
                            data = d, iter = 100, start_lower = sv - 10, start_upper = sv + 10, supp_errors = 'Y')
    } 
    else if(mod_name == 'gaussian_1987'){
       fit <- nls_multstart(rate_shifted ~ gaussian_1987(temp, rmax, topt, a),
                            data = d, iter = 100, start_lower = sv - 10, start_upper = sv + 10, supp_errors = 'Y')
    } 
    else if(mod_name == 'boatman_2017'){
       fit <- nls_multstart(rate_shifted ~ boatman_2017(temp, rmax, tmin, tmax, topt, a),
                            data = d, iter = 100, start_lower = sv - 10, start_upper = sv + 10, supp_errors = 'Y')
    }
    
    return(MuMIn::AICc(fit))
  }, error = function(e) return(NA))
}

# Run all models 
all_group_aic_2023 <- calc_shifted_2023 %>%
  group_by(Species, Phenotype) %>%
  nest() %>%
  crossing(model_name = c('sharpeschoolhigh_1981', 'pawar_2018', 'boatman_2017', 'gaussian_1987')) %>%
  mutate(aic_score = map2_dbl(data, model_name, ~safe_fit_aic_v2(.x, .y))) %>%
  group_by(Species, Phenotype) %>%
  mutate(delta_AICc = aic_score - min(aic_score, na.rm = TRUE)) %>%
  arrange(Species, Phenotype, delta_AICc)
```

Pawar_2018 model is the best fit for 3 out of 4 groups (Pcomp susceptible had Boatman_2017 as a better fit but for comparability across groups, we will stick with Pawar_2018). 

Fit the pawar 2018 function and run bootstrap across groups 
```{r}
# Define the Pawar-based bootstrap function
fit_pawar_boot <- function(x) {
  mod <- 'pawar_2018'
  
  # Wider search range (+/- 10) for stability
  fit <- nls_multstart(rate_shifted ~ pawar_2018(temp, r_tref, e, eh, topt, tref = 25),
                       data = x,
                       iter = 50, 
                       start_lower = get_start_vals(x$temp, x$rate_shifted, model_name = mod) - 10,
                       start_upper = get_start_vals(x$temp, x$rate_shifted, model_name = mod) + 10,
                       lower = get_lower_lims(x$temp, x$rate_shifted, model_name = mod),
                       upper = get_upper_lims(x$temp, x$rate_shifted, model_name = mod),
                       supp_errors = 'Y')
  
  # Calculate derived parameters
  extra_params <- calc_params(fit)
  
  # Get the raw r_tref estimate from the model
  r_tref_val <- broom::tidy(fit) %>% filter(term == "r_tref") %>% pull(estimate)
  
  # Return: rmax, topt, e, eh, r_tref_25
  all_stats <- c(as.numeric(extra_params$rmax), 
                 as.numeric(extra_params$topt), 
                 as.numeric(extra_params$e), 
                 as.numeric(extra_params$eh),
                 as.numeric(r_tref_val))
  
  names(all_stats) <- c("rmax", "topt", "e", "eh", "r_tref_25")
  return(all_stats)
}

# Run the bootstrap
boot_results_2023 <- calc_shifted_2023 %>%
  group_by(Species, Phenotype) %>%
  nest() %>%
  mutate(boot = map(data, ~boot(data = .x, 
                                 statistic = function(d, i) as.numeric(fit_pawar_boot(d[i,])), 
                                 R = 1000,
                                 parallel = "multicore", 
                                 ncpus = n_cores)))

# Tidy the results for interpretation
boot_stats_2023 <- boot_results_2023 %>%
  mutate(stats = map(boot, ~broom::tidy(.x, conf.int = TRUE, conf.method = 'perc'))) %>%
  unnest(stats) %>%
  ungroup() %>%
  mutate(term = rep(c("rmax", "topt", "e", "eh", "r_tref_25"), 4))
```

Clean up data and create 5 panel plot
```{r}
# Clean up the NAs by using the average of the confidence intervals where statistic is NA
boot_stats_cleaned_2023 <- boot_stats_final %>%
  mutate(statistic = ifelse(is.na(statistic), (conf.low + conf.high) / 2, statistic))

# Define the 5 traits we want to plot
traits <- c("r_tref_25", "rmax", "topt", "e", "eh")

# Create a list to store the plots
plot_list <- list()

for (i in 1:length(traits)) {
  trait_name <- traits[i]
  df_sub <- subset(boot_stats_cleaned_2023, term == trait_name)
  
  # Set specific labels for each trait
  y_label <- case_when(
    trait_name == "r_tref_25" ~ "Rate at 25°C (log)",
    trait_name == "rmax"      ~ "Rmax (log)",
    trait_name == "topt"      ~ "Topt (°C)",
    trait_name == "e"         ~ "Activation (e)",
    trait_name == "eh"        ~ "Deactivation (eh)"
  )
  
  # Build the plot using your specific style
  p <- ggplot(df_sub, aes(x = Phenotype, y = statistic, color = Phenotype)) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                  width = 0.15, linewidth = 1) +
    geom_point(size = 4) +
    facet_grid(~Species) + # Species side-by-side in each row
    scale_color_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
    theme_bw(base_size = 12) +
    theme(
      legend.position = "none",
      strip.background = element_rect(fill = "gray95"),
      strip.text = element_text(face = "bold"),
      panel.grid.minor = element_blank()
    ) +
    labs(y = y_label, x = NULL)
  
  # Hide x-axis text for all except the bottom-most plot
  if (i < length(traits)) {
    p <- p + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  }
  
  plot_list[[i]] <- p
}

# Combine all 5 plots into one tall column
final_5_panel <- wrap_plots(plot_list, ncol = 1)
print(final_5_panel)
```

Plot log scale calcification rates across temps 
```{r}
calc_plot_2023 <- calc_2023 %>%
  filter(!is.na(Phenotype),
         !is.na(Species),
         Colony.ID != "blank")

ggplot(calc_plot_2023, aes(
  x = Target.Temperature,
  y = log_calc,
  color = Phenotype,
  fill = Phenotype
)) +
  geom_point(size = 3, alpha = 0.5) +
  geom_smooth(method = "loess", se = TRUE, span = 0.6, alpha = 0.25) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  labs(
    x = "Target Temperature (°C)",
    y = expression("log(Net Calcification ("*cm^-2*"))"),
    color = "Phenotype",
    fill = "Phenotype")
```

Plot as PCA
```{r}
# 1. Extract all 1000 replicates for all 5 traits from the 'boot' column
pca_prep_clean <- boot_results_final %>%
  mutate(reps = map(boot, ~as.data.frame(.x$t))) %>%
  unnest(reps) %>%
  ungroup() %>%
  # Rename based on your specific column order
  rename(rmax = V1, topt = V2, e = V3, eh = V4, r_tref_25 = V5) %>%
  # REMOVE THE NAs
  drop_na(rmax, topt, e, eh, r_tref_25)

# 2. Perform PCA (Scaling is vital because Topt and Rmax have different magnitudes)
pca_fit <- prcomp(pca_prep_clean[, c("rmax", "topt", "e", "eh", "r_tref_25")], scale. = TRUE)

# 3. Plot with Loadings (Arrows)
autoplot(pca_fit, data = pca_prep_clean, colour = 'Phenotype', 
         shape = 'Species', frame = TRUE, frame.type = 'norm',
         loadings = TRUE, loadings.label = TRUE, 
         loadings.label.size = 3, loadings.colour = "black", loadings.label.colour = "black") +
  scale_color_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  scale_fill_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  theme_bw() +
  labs(title = "Multivariate Thermal Phenotype PCA",
       subtitle = "Bootstrap cloud showing separation of coral groups",
       caption = "Arrows indicate the direction of trait influence.")
```

Determine differences 
```{r}
# Function to get p-value from two bootstrap objects
get_p_value <- function(boot_a, boot_b, index) {
  # index: 1=rmax, 2=topt, 3=e, 4=eh, 5=r_tref_25
  diffs <- boot_a$t[, index] - boot_b$t[, index]
  # Calculate proportion of differences that cross zero
  p_val <- 2 * min(mean(diffs > 0, na.rm=TRUE), mean(diffs < 0, na.rm=TRUE))
  return(p_val)
}

# Example: Comparing Topt (index 2) for Montipora Phenotypes
p_val <- get_p_value(boot_results_final$boot[[1]], boot_results_final$boot[[2]], 2)
```

```{r}
# Define the trait names for the output
trait_names <- c("rmax", "topt", "e", "eh", "r_tref_25")

# Initialize a results table
comparison_results <- data.frame()

# Define the pairs we want to compare (based on the row indices of boot_results_2025)
# 1 vs 2: Phenotype effect in Montipora
# 3 vs 4: Phenotype effect in Porites
# 1 vs 3: Species effect in Resistant corals
comparisons <- list(
  "M.cap_Phenotype" = c(1, 2),
  "P.com_Phenotype" = c(3, 4))

for(comp_name in names(comparisons)){
  idx_a <- comparisons[[comp_name]][1]
  idx_b <- comparisons[[comp_name]][2]
  
  for(i in 1:5){
    p_val <- get_p_value(boot_results_final$boot[[idx_a]], 
                         boot_results_final$boot[[idx_b]], i)
    
    comparison_results <- rbind(comparison_results, data.frame(
      Comparison = comp_name,
      Trait = trait_names[i],
      P_Value = p_val,
      Significant = ifelse(p_val < 0.05, "Yes", "No")
    ))
  }
}

print(comparison_results)
```

Fit data by individual colony instead of groups 
```{r}
# Run the fit per colony
colony_fits <- calc_shifted_2023 %>%
  group_by(Species, Phenotype, Colony.ID) %>%
  nest() %>%
  mutate(fit = map(data, ~nls_multstart(rate_shifted ~ pawar_2018(temp, r_tref, e, eh, topt, tref = 25),
                                        data = .x,
                                        iter = 1000,
                                        start_lower = get_start_vals(.x$temp, .x$rate_shifted, model_name = 'pawar_2018') - 10,
                                        start_upper = get_start_vals(.x$temp, .x$rate_shifted, model_name = 'pawar_2018') + 10,
                                        lower = get_lower_lims(.x$temp, .x$rate_shifted, model_name = 'pawar_2018'),
                                        upper = get_upper_lims(.x$temp, .x$rate_shifted, model_name = 'pawar_2018'),
                                        supp_errors = 'Y')))

# Create a "safe" version of the tidy extraction. If it fails, it returns NA instead of crashing the loop
safe_extract_tref <- possibly(function(x) {
  broom::tidy(x) %>% 
    filter(term == "r_tref") %>% 
    pull(estimate)
}, otherwise = NA_real_)

# Run the metrics extraction
colony_metrics <- colony_fits %>%
   mutate(params = map(fit, ~calc_params(.x))) %>%
   unnest(params) %>%
   # Use the safe function with explicit naming
   mutate(r_tref_25 = map_dbl(fit, safe_extract_tref)) %>%
   select(-data, -fit) %>%
   ungroup()
```

Run using sharp shoefield model
```{r}
# Prepare your data (ensure K is present and filter for 6+ points)
calc_prepared_2023 <- calc_shifted_2023 %>%
  group_by(Colony.ID) %>%
  filter(n_distinct(temp) >= 6) %>%
  mutate(K = temp + 273.15) %>% # Conversion to Kelvin
  ungroup()

# Run the fit using Sharpe-Schoolfield
colony_fits_ss <- calc_prepared_2023 %>%
  group_by(Species, Phenotype, Colony.ID) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(rate_shifted ~ sharpe_schoolfield(lnc, E, Eh, Th, temp = K, Tc = 25),
                                                data = .x,
                                                iter = 1000, # Increased iterations for better convergence
                                                start_lower = c(lnc = -10, E = 0.1, Eh = 0.2, Th = 285), # ~25C
                                                start_upper = c(lnc = 10, E = 2, Eh = 5, Th = 330),    # ~45C
                                                #lower = c(lnc = -10, E = 0, Eh = 0, Th = 273),
                                                #upper = c(lnc = 10, E = 5, Eh = 20, Th = 373),
                                                supp_errors = 'Y',
                                                na.action = na.omit)))
summary(colony_fits_ss$fit[[1]])
```

Try something else 
```{r}
library(nls.multstart)
library(rTPC)
library(dplyr)
library(purrr)

# 1. Prepare your data: Ensure Kelvin and filter for enough points
calc_prepared <- calc_shifted_2023 %>%
  mutate(K = temp + 273.15) %>%
  group_by(Colony.ID) %>%
  filter(n_distinct(temp) >= 6) %>%
  ungroup()

# 2. Run the fit using the log-transformed version of the model
# We use schoolfield_high because your Y variable (rate_shifted) is already logged.
colony_fits_log <- calc_prepared %>%
  group_by(Phenotype, Species, Colony.ID) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(
    rate_shifted ~ schoolfield_high(lnc, E, Eh, Th, temp = K, Tc = 27),
    data = .x,
    iter = 1000,
    # Adjusting start values to be in log-space (lnc is now the log-rate at 27C)
    # Since your rates are ~0.1, log(0.1) is ~ -2.3
    start_lower = c(lnc = -5, E = 0.1, Eh = 0.5, Th = 285), 
    start_upper = c(lnc = 2,  E = 2.0, Eh = 10,  Th = 315),
    lower = c(lnc = -15, E = 0, Eh = 0, Th = 273),
    upper = c(lnc = 15, E = 10, Eh = 20, Th = 373),
    supp_errors = 'Y',
    na.action = na.omit
  )))
```





# 2025

## Calculate calcification rates 

Read in data 
```{r}
calc <- read.csv("../data/TPC_calcification/Calcification_data_TPC.csv", header = T)
```

Calculate incubation time 
```{r}
calc <- calc %>%
  mutate(
    start = hm(Time_start),
    end   = hm(Time_end),
    elapsed_mins  = as.numeric(end - start, units = "mins"),
    elapsed_hours = elapsed_mins / 60)
```

Calculate delta alkalinity 
```{r}
calc <- calc %>%
  mutate(Delta_TA_raw = Starting_alk_value - Alk_avg)
```

Calculate blanks 
```{r}
blank_delta_ta <- calc %>%
  filter(ColonyID == "blank" | Notes == "blank") %>%
  group_by(Date, Target_Temperature) %>%
  summarise(mean_blank_delta_ta = mean(Delta_TA_raw, na.rm = TRUE))# %>%
  #pull(mean_blank_delta_ta)
```

Subtract blank and calculate net calcification
```{r}
# Correct Delta_TA by subtracting blank
calc <- calc %>%
  left_join(blank_delta_ta, by = c("Date", "Target_Temperature")) 

calc <- calc %>%
  mutate(
    Delta_TA_corr = Delta_TA_raw - mean_blank_delta_ta,
    # Assuming seawater density 1.028 g/mL and volume in mL to L, and no coral weight (thus calc mg/L/hr)
    Net_Calcification = (Delta_TA_corr / 2) * (Sample_Volume_mL / 1000) * 1.028 / elapsed_hours) %>%
  filter(!is.na(Net_Calcification) & !is.nan(Net_Calcification))
```

Normalize by SA
```{r}
calc <- calc %>%
  mutate(Net_Calcification_cm2 = Net_Calcification / Surface_Area)
```

Read in meta
```{r}
meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")
```

Join with calc 
```{r}
meta_unique <- meta %>%
  group_by(Colony_ID) %>%
  slice(1) %>%   # or use slice_max(Date) if you want the most recent entry
  ungroup()

calc_meta <- calc %>%
  left_join(meta_unique, by = c("ColonyID" = "Colony_ID")) %>%
  select(ColonyID, Target_Temperature, Net_Calcification_cm2, Species, Phenotype)
```

Plot 
```{r}
# Filter out blanks and missing phenotype/species
calc_plot <- calc_meta %>%
  filter(!is.na(Phenotype),
         !is.na(Species),
         ColonyID != "blank")

ggplot(calc_plot, aes(
  x = Target_Temperature,
  y = Net_Calcification_cm2,
  color = Phenotype,
  fill = Phenotype
)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "loess", se = TRUE, span = 0.6, alpha = 0.25) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  labs(
    x = "Target Temperature (°C)",
    y = expression("Net Calcification ("*cm^-2*")"),
    color = "Phenotype",
    fill = "Phenotype")
```

## rTPC 

Because some values are negative, we need to 'shift' the data away from negative values while still maintaining the shape of the curve.
```{r}
# log 
calc_2025 <- calc_meta %>%
  mutate(log_calc = log(Net_Calcification_cm2 + 1))

calc_shifted_2025 <- calc_2025 %>%
  filter(!is.na(Species)) %>%
  mutate(
    rate_shifted = log_calc,
    temp = Target_Temperature)
```

Fit multiple models to find best fit for our data 
```{r}
safe_fit_aic_v2 <- function(d, mod_name) {
  tryCatch({
    sv <- get_start_vals(d$temp, d$rate_shifted, model_name = mod_name)
    
    # We use a formula approach for each model to ensure parameter names match
    if(mod_name == 'pawar_2018'){
       fit <- nls_multstart(rate_shifted ~ pawar_2018(temp, r_tref, e, eh, topt, tref = 25),
                            data = d, iter = 100, start_lower = sv - 10, start_upper = sv + 10, supp_errors = 'Y')
    } 
    else if(mod_name == 'sharpeschoolhigh_1981'){
       fit <- nls_multstart(rate_shifted ~ sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 25),
                            data = d, iter = 100, start_lower = sv - 10, start_upper = sv + 10, supp_errors = 'Y')
    } 
    else if(mod_name == 'gaussian_1987'){
       fit <- nls_multstart(rate_shifted ~ gaussian_1987(temp, rmax, topt, a),
                            data = d, iter = 100, start_lower = sv - 10, start_upper = sv + 10, supp_errors = 'Y')
    } 
    else if(mod_name == 'boatman_2017'){
       fit <- nls_multstart(rate_shifted ~ boatman_2017(temp, rmax, tmin, tmax, topt, a),
                            data = d, iter = 100, start_lower = sv - 10, start_upper = sv + 10, supp_errors = 'Y')
    }
    
    return(MuMIn::AICc(fit))
  }, error = function(e) return(NA))
}

# Run all models 
all_group_aic_2025 <- calc_shifted_2025 %>%
  group_by(Species, Phenotype) %>%
  nest() %>%
  crossing(model_name = c('sharpeschoolhigh_1981', 'pawar_2018', 'boatman_2017', 'gaussian_1987')) %>%
  mutate(aic_score = map2_dbl(data, model_name, ~safe_fit_aic_v2(.x, .y))) %>%
  group_by(Species, Phenotype) %>%
  mutate(delta_AICc = aic_score - min(aic_score, na.rm = TRUE)) %>%
  arrange(Species, Phenotype, delta_AICc)
```

Still using Pawar 2018. 

Fit the pawar 2018 function and run bootstrap across groups 
```{r}
# Define the Pawar-based bootstrap function
fit_pawar_boot <- function(x) {
  mod <- 'pawar_2018'
  
  # Wider search range (+/- 10) for stability
  fit <- nls_multstart(rate_shifted ~ pawar_2018(temp, r_tref, e, eh, topt, tref = 25),
                       data = x,
                       iter = 50, 
                       start_lower = get_start_vals(x$temp, x$rate_shifted, model_name = mod) - 10,
                       start_upper = get_start_vals(x$temp, x$rate_shifted, model_name = mod) + 10,
                       lower = get_lower_lims(x$temp, x$rate_shifted, model_name = mod),
                       upper = get_upper_lims(x$temp, x$rate_shifted, model_name = mod),
                       supp_errors = 'Y')
  
  # Calculate derived parameters
  extra_params <- calc_params(fit)
  
  # Get the raw r_tref estimate from the model
  r_tref_val <- broom::tidy(fit) %>% filter(term == "r_tref") %>% pull(estimate)
  
  # Return: rmax, topt, e, eh, r_tref_25
  all_stats <- c(as.numeric(extra_params$rmax), 
                 as.numeric(extra_params$topt), 
                 as.numeric(extra_params$e), 
                 as.numeric(extra_params$eh),
                 as.numeric(r_tref_val))
  
  names(all_stats) <- c("rmax", "topt", "e", "eh", "r_tref_25")
  return(all_stats)
}

# Run the bootstrap
boot_results_2025 <- calc_shifted_2025 %>%
  group_by(Species, Phenotype) %>%
  nest() %>%
  mutate(boot = map(data, ~boot(data = .x, 
                                 statistic = function(d, i) as.numeric(fit_pawar_boot(d[i,])), 
                                 R = 1000,
                                 parallel = "multicore", 
                                 ncpus = n_cores)))

# Tidy the results for interpretation
boot_stats_2025 <- boot_results_2025 %>%
  mutate(stats = map(boot, ~broom::tidy(.x, conf.int = TRUE, conf.method = 'perc'))) %>%
  unnest(stats) %>%
  ungroup() %>%
  mutate(term = rep(c("rmax", "topt", "e", "eh", "r_tref_25"), 4))
```

Code not running...
```{r}
# Test run on just one group to see the actual error
test_2025 <- calc_shifted_2025 %>%
  filter(Species == "Montipora capitata", Phenotype == "Resistant") %>%
  boot(data = ., 
       statistic = function(d, i) as.numeric(fit_pawar_boot(d[i,])), 
       R = 10) # Small R just to test
```

Try again with updated code
```{r}
# 1. Ensure the v2 function is defined in your environment
fit_pawar_boot_v2 <- function(x) {
  mod <- 'pawar_2018'
  result <- tryCatch({
    fit <- nls_multstart(rate_shifted ~ pawar_2018(temp, r_tref, e, eh, topt, tref = 25),
                         data = x,
                         iter = 50, 
                         start_lower = get_start_vals(x$temp, x$rate_shifted, model_name = mod) - 10,
                         start_upper = get_start_vals(x$temp, x$rate_shifted, model_name = mod) + 10,
                         lower = get_lower_lims(x$temp, x$rate_shifted, model_name = mod),
                         upper = get_upper_lims(x$temp, x$rate_shifted, model_name = mod),
                         supp_errors = 'Y')
    
    extra_params <- calc_params(fit)
    r_tref_val <- broom::tidy(fit) %>% filter(term == "r_tref") %>% pull(estimate)
    
    # Return numeric vector
    c(rmax = as.numeric(extra_params$rmax), 
      topt = as.numeric(extra_params$topt), 
      e = as.numeric(extra_params$e), 
      eh = as.numeric(extra_params$eh),
      r_tref_25 = as.numeric(r_tref_val))
    
  }, error = function(e) {
    # Returns numeric NAs so the bootstrap can still aggregate results
    return(c(rmax = NA, topt = NA, e = NA, eh = NA, r_tref_25 = NA))
  })
  return(result)
}

# 2. Run the full 2025 Bootstrap
boot_results_2025 <- calc_shifted_2025 %>%
   group_by(Species, Phenotype) %>%
   nest() %>%
   mutate(boot = map(data, ~boot(data = .x,  
                                  statistic = function(d, i) as.numeric(fit_pawar_boot_v2(d[i,])), 
                                  R = 1000,
                                  parallel = "multicore", 
                                  ncpus = n_cores)))

# 3. Tidy (This should now work without the 'non-numeric' error)
boot_stats_2025 <- boot_results_2025 %>%
   mutate(stats = map(boot, ~broom::tidy(.x, conf.int = TRUE, conf.method = 'perc'))) %>%
   unnest(stats) %>%
   ungroup() %>%
   mutate(term = rep(c("rmax", "topt", "e", "eh", "r_tref_25"), 4))
```

lean up data and create 5 panel plot
```{r}
# Clean up the NAs by using the average of the confidence intervals where statistic is NA
boot_stats_cleaned_2025 <- boot_stats_2025 %>%
  mutate(statistic = ifelse(is.na(statistic), (conf.low + conf.high) / 2, statistic))

# Define the 5 traits we want to plot
traits <- c("r_tref_25", "rmax", "topt", "e", "eh")

# Create a list to store the plots
plot_list <- list()

for (i in 1:length(traits)) {
  trait_name <- traits[i]
  df_sub <- subset(boot_stats_cleaned_2025, term == trait_name)
  
  # Set specific labels for each trait
  y_label <- case_when(
    trait_name == "r_tref_25" ~ "Rate at 25°C (log)",
    trait_name == "rmax"      ~ "Rmax (log)",
    trait_name == "topt"      ~ "Topt (°C)",
    trait_name == "e"         ~ "Activation (e)",
    trait_name == "eh"        ~ "Deactivation (eh)"
  )
  
  # Build the plot using your specific style
  p <- ggplot(df_sub, aes(x = Phenotype, y = statistic, color = Phenotype)) +
    geom_errorbar(aes(ymin = statistic - std.error, ymax = statistic + std.error), 
                  width = 0.15, linewidth = 1) +
    geom_point(size = 4) +
    facet_grid(~Species) + # Species side-by-side in each row
    scale_color_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
    theme_bw(base_size = 12) +
    theme(
      legend.position = "none",
      strip.background = element_rect(fill = "gray95"),
      strip.text = element_text(face = "bold"),
      panel.grid.minor = element_blank()
    ) +
    labs(y = y_label, x = NULL)
  
  # Hide x-axis text for all except the bottom-most plot
  if (i < length(traits)) {
    p <- p + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  }
  
  plot_list[[i]] <- p
}

# Combine all 5 plots into one tall column
final_5_panel <- wrap_plots(plot_list, ncol = 1)
print(final_5_panel)
```

Plot log scale calcification rates across temps 
```{r}
calc_plot_2025 <- calc_2025 %>%
  filter(!is.na(Phenotype),
         !is.na(Species),
         ColonyID != "blank")

ggplot(calc_plot_2025, aes(
  x = Target_Temperature,
  y = log_calc,
  color = Phenotype,
  fill = Phenotype
)) +
  geom_point(size = 3, alpha = 0.5) +
  geom_smooth(method = "loess", se = TRUE, span = 0.6, alpha = 0.25) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  labs(
    x = "Target Temperature (°C)",
    y = expression("log(Net Calcification ("*cm^-2*"))"),
    color = "Phenotype",
    fill = "Phenotype")
```

Plot as PCA
```{r}
# 1. Extract all 1000 replicates for all 5 traits from the 'boot' column
pca_prep_clean <- boot_results_2025 %>%
  mutate(reps = map(boot, ~as.data.frame(.x$t))) %>%
  unnest(reps) %>%
  ungroup() %>%
  # Rename based on your specific column order
  rename(rmax = V1, topt = V2, e = V3, eh = V4, r_tref_25 = V5) %>%
  # REMOVE THE NAs
  drop_na(rmax, topt, e, eh, r_tref_25)

# 2. Perform PCA (Scaling is vital because Topt and Rmax have different magnitudes)
pca_fit <- prcomp(pca_prep_clean[, c("rmax", "topt", "e", "eh", "r_tref_25")], scale. = TRUE)

# 3. Plot with Loadings (Arrows)
autoplot(pca_fit, data = pca_prep_clean, colour = 'Phenotype', 
         shape = 'Species', frame = TRUE, frame.type = 'norm',
         loadings = TRUE, loadings.label = TRUE, 
         loadings.label.size = 3, loadings.colour = "black", loadings.label.colour = "black") +
  scale_color_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  scale_fill_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  theme_bw() +
  labs(title = "Multivariate Thermal Phenotype PCA",
       subtitle = "Bootstrap cloud showing separation of coral groups",
       caption = "Arrows indicate the direction of trait influence.")
```

Determine differences 
```{r}
# Function to get p-value from two bootstrap objects
get_p_value <- function(boot_a, boot_b, index) {
  # index: 1=rmax, 2=topt, 3=e, 4=eh, 5=r_tref_25
  diffs <- boot_a$t[, index] - boot_b$t[, index]
  # Calculate proportion of differences that cross zero
  p_val <- 2 * min(mean(diffs > 0, na.rm=TRUE), mean(diffs < 0, na.rm=TRUE))
  return(p_val)
}

# Example: Comparing Topt (index 2) for Montipora Phenotypes
p_val <- get_p_value(boot_results_2025$boot[[1]], boot_results_2025$boot[[2]], 2)
```

Compute pvalues for differences 
```{r}
# Define the trait names for the output
trait_names <- c("rmax", "topt", "e", "eh", "r_tref_25")

# Initialize a results table
comparison_results <- data.frame()

# Define the pairs we want to compare (based on the row indices of boot_results_2025)
# 1 vs 2: Phenotype effect in Montipora
# 3 vs 4: Phenotype effect in Porites
# 1 vs 3: Species effect in Resistant corals
comparisons <- list(
  "M.cap_Phenotype" = c(1, 2),
  "P.com_Phenotype" = c(3, 4),
  "Resistant_Species" = c(1, 3)
)

for(comp_name in names(comparisons)){
  idx_a <- comparisons[[comp_name]][1]
  idx_b <- comparisons[[comp_name]][2]
  
  for(i in 1:5){
    p_val <- get_p_value(boot_results_2025$boot[[idx_a]], 
                         boot_results_2025$boot[[idx_b]], i)
    
    comparison_results <- rbind(comparison_results, data.frame(
      Comparison = comp_name,
      Trait = trait_names[i],
      P_Value = p_val,
      Significant = ifelse(p_val < 0.05, "Yes", "No")
    ))
  }
}

print(comparison_results)
```

## Comparing 2023 and 2025

```{r}
# Create a list for year-over-year comparisons
# Assuming row order is the same in both (Montipora-R, Montipora-S, Porites-R, Porites-S)
year_comparison_results <- data.frame()

for(group_idx in 1:4) {
  # Identify which group we are looking at
  group_info <- boot_results_2025[group_idx, c("Species", "Phenotype")]
  group_label <- paste(group_info$Species, group_info$Phenotype, sep="_")
  
  for(trait_idx in 1:5) {
    # Compare 2023 (boot_results_final) vs 2025 (boot_results_2025)
    p_val <- get_p_value(boot_results_final$boot[[group_idx]], 
                         boot_results_2025$boot[[group_idx]], 
                         trait_idx)
    
    year_comparison_results <- rbind(year_comparison_results, data.frame(
      Group = group_label,
      Trait = trait_names[trait_idx],
      P_Value = p_val,
      Significant_Shift = ifelse(p_val < 0.05, "Yes", "No")
    ))
  }
}

print(year_comparison_results)

# Compare the actual means between years
comparison_check <- bind_rows(
  boot_stats_cleaned_2023 %>% filter(term == "rmax") %>% mutate(Year = "2023"),
  boot_stats_cleaned_2025 %>% filter(term == "rmax") %>% mutate(Year = "2025")
) %>% 
  select(Year, Species, Phenotype, statistic, conf.low, conf.high)

print(comparison_check)
```

Plot
```{r}
# 1. Combine the datasets
plot_data <- bind_rows(
  boot_stats_cleaned_2023 %>% mutate(Year = "2023"),
  boot_stats_cleaned_2025 %>% mutate(Year = "2025")
) #%>%
  #filter(term %in% c("rmax", "topt")) # Focus on the most important shifts

# 2. Create the Slope Chart
ggplot(plot_data, aes(x = Year, y = statistic, group = interaction(Species, Phenotype), color = Phenotype)) +
  geom_line(size = 1, alpha = 0.7) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1) +
  facet_grid(term ~ Species, scales = "free_y") +
  scale_color_manual(values = c("Resistant" = "#4B294F", "Susceptible" = "#F3A17D")) +
  theme_bw() +
  labs(title = "Interannual Physiological Shifts (2023 vs 2025)",
       y = "Parameter Value (Bootstrap Mean ± 95% CI)",
       x = "Experimental Year") +
  theme(strip.background = element_rect(fill = "white"),
        legend.position = "bottom")
```






