---
title: "Net calcification TPC"
author: "Jill Ashey"
date: "2025-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(emmeans)
library(lme4)
library(lmerTest)
library(lsmeans)
library(rTPC)
```

Read in data 
```{r}
calc <- read.csv("../data/TPC_calcification/Calcification_data_TPC.csv", header = T)
```

Calculate incubation time 
```{r}
calc <- calc %>%
  mutate(
    start = hm(Time_start),
    end   = hm(Time_end),
    elapsed_mins  = as.numeric(end - start, units = "mins"),
    elapsed_hours = elapsed_mins / 60)
```

Calculate delta alkalinity 
```{r}
calc <- calc %>%
  mutate(Delta_TA_raw = Starting_alk_value - Alk_avg)
```

Calculate blanks 
```{r}
blank_delta_ta <- calc %>%
  filter(ColonyID == "blank" | Notes == "blank") %>%
  group_by(Date, Target_Temperature) %>%
  summarise(mean_blank_delta_ta = mean(Delta_TA_raw, na.rm = TRUE))# %>%
  #pull(mean_blank_delta_ta)
```

Subtract blank and calculate net calcification
```{r}
# Correct Delta_TA by subtracting blank
calc <- calc %>%
  left_join(blank_delta_ta, by = c("Date", "Target_Temperature")) 

calc <- calc %>%
  mutate(
    Delta_TA_corr = Delta_TA_raw - mean_blank_delta_ta,
    # Assuming seawater density 1.028 g/mL and volume in mL to L, and no coral weight (thus calc mg/L/hr)
    Net_Calcification = (Delta_TA_corr / 2) * (Sample_Volume_mL / 1000) * 1.028 / elapsed_hours) %>%
  filter(!is.na(Net_Calcification) & !is.nan(Net_Calcification))
```

Normalize by SA
```{r}
calc <- calc %>%
  mutate(Net_Calcification_cm2 = Net_Calcification / Surface_Area)
```

Read in meta
```{r}
meta <- read_csv("../data/pi_curves/pi_curve_metadata_sample.csv")
```

Join with calc 
```{r}
meta_unique <- meta %>%
  group_by(Colony_ID) %>%
  slice(1) %>%   # or use slice_max(Date) if you want the most recent entry
  ungroup()

calc_meta <- calc %>%
  left_join(meta_unique, by = c("ColonyID" = "Colony_ID"))
```

Plot 
```{r}
# Filter out blanks and missing phenotype/species
calc_plot <- calc_meta %>%
  filter(!is.na(Phenotype),
         !is.na(Species),
         ColonyID != "blank")

ggplot(calc_plot, aes(
  x = Target_Temperature,
  y = Net_Calcification_cm2,
  color = Phenotype,
  fill = Phenotype
)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "loess", se = TRUE, span = 0.6, alpha = 0.25) +
  facet_wrap(~Species, scales = "fixed") +
  theme_classic(base_size = 14) +
  scale_color_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  scale_fill_manual(values = c("Resistant" = "#4b294fff", "Susceptible" = "#f3a17dff")) +
  labs(
    x = "Target Temperature (°C)",
    y = expression("Net Calcification ("*cm^-2*")"),
    color = "Phenotype",
    fill = "Phenotype"
  )


```

# rTPC 

## Prep data

```{r}
# Remove blank samples and ensure we have complete data
tpc_data <- calc_meta %>%
  filter(!is.na(Species), ColonyID != "blank") %>%
  # Use the corrected net calcification rate
  select(ColonyID, Species, Phenotype, Target_Temperature, 
         Net_Calcification_cm2, Light, Run.x) %>%
  rename(temp = Target_Temperature,
         rate = Net_Calcification_cm2) %>%
  # Remove any missing values
  filter(!is.na(rate), !is.na(temp))
```

## Fit TPC models

```{r}
# Function to fit TPC for each colony
fit_tpc_colony <- function(data) {
  
  # Get starting values using get_start_vals
  start_vals <- get_start_vals(data$temp, data$rate, model_name = 'sharpeschoolhigh_1981')
  
  # Fit the model using nls_multstart for robust parameter estimation
  fit <- nls_multstart(rate ~ sharpeschoolhigh_1981(temp = temp, r_tref, e, eh, th, tref = 18),
                       data = data,
                       iter = 500,
                       start_lower = start_vals - 1,
                       start_upper = start_vals + 1,
                       lower = get_lower_lims(data$temp, data$rate, model_name = 'sharpeschoolhigh_1981'),
                       upper = get_upper_lims(data$temp, data$rate, model_name = 'sharpeschoolhigh_1981'),
                       supp_errors = 'Y')
  
  return(fit)
}

# Group by colony and fit models
colony_fits <- tpc_data %>%
  group_by(ColonyID, Species, Phenotype) %>%
  nest() %>%
  mutate(
    # Fit model
    fit = map(data, possibly(fit_tpc_colony, otherwise = NULL)),
    # Extract model parameters
    params = map(fit, possibly(tidy, otherwise = NULL))
  ) %>%
  filter(!is.null(fit))

# Extract key thermal performance metrics for each colony
tpc_metrics <- colony_fits %>%
  mutate(
    # Calculate thermal optimum (Topt)
    topt = map_dbl(fit, ~get_topt(.x)),
    # Calculate maximum performance (rmax)
    rmax = map_dbl(fit, ~get_rmax(.x)),
    # Calculate breadth (temperature range at 50% rmax)
    breadth = map_dbl(fit, ~get_breadth(.x)),
    # Calculate critical thermal maximum (CTmax)
    ctmax = map_dbl(fit, ~get_ctmax(.x)),
    # Calculate Q10 (temperature sensitivity)
    q10 = map_dbl(fit, possibly(~get_q10(.x), otherwise = NA))
  )
```










# rTPC attempt 2

Data prep
```{r}
tpc_data <- calc_meta %>%
  filter(!is.na(Species), ColonyID != "blank") %>%
  # Use the corrected net calcification rate
  select(ColonyID, Species, Phenotype, Target_Temperature, 
         Net_Calcification_cm2, Light, Run.x) %>%
  rename(temp = Target_Temperature,
         rate = Net_Calcification_cm2) %>%
  # Remove any missing values
  filter(!is.na(rate), !is.na(temp))
```

 Check colonies 
```{r}
diagnostic_check <- tpc_data %>%
  group_by(ColonyID, Species, Phenotype) %>%
  summarise(
    n_points = n(),
    temp_range = max(temp) - min(temp),
    min_temp = min(temp),
    max_temp = max(temp),
    mean_rate = mean(rate),
    sd_rate = sd(rate),
    any_negative = any(rate < 0),
    all_zero = all(rate == 0),
    .groups = 'drop'
  )

print("Data diagnostic for each colony:")
print(diagnostic_check)

# Check specific problematic colonies
problem_colonies <- c(3, 233, 226, 211, 213, 221)
print("\nDetailed check for problematic colonies:")
print(diagnostic_check %>% filter(ColonyID %in% problem_colonies))

```
 
Fit TPC models
```{r}
# Function to fit TPC for each colony with better error handling
fit_tpc_colony <- function(data, colony_id = NULL) {
  
  # Check for minimum data requirements
  if(nrow(data) < 5) {
    warning(paste("Colony", colony_id, ": Insufficient data points (n =", nrow(data), ")"))
    return(NULL)
  }
  
  # Check for variation in response
  if(sd(data$rate, na.rm = TRUE) == 0) {
    warning(paste("Colony", colony_id, ": No variation in rates"))
    return(NULL)
  }
  
  # Check for all negative or zero rates
  if(all(data$rate <= 0)) {
    warning(paste("Colony", colony_id, ": All rates are zero or negative"))
    return(NULL)
  }
  
  tryCatch({
    # Get starting values using get_start_vals
    start_vals <- get_start_vals(data$temp, data$rate, model_name = 'pawar_2018')
    
    # Fit the model using nls_multstart for robust parameter estimation
    fit <- nls_multstart(rate ~ sharpeschoolhigh_1981(temp = temp, r_tref, e, eh, th, tref = 18),
                         data = data,
                         iter = 1000,  # Increased iterations
                         start_lower = start_vals - 2,  # Wider search range
                         start_upper = start_vals + 2,
                         lower = get_lower_lims(data$temp, data$rate, model_name = 'pawar_2018'),
                         upper = get_upper_lims(data$temp, data$rate, model_name = 'pawar_2018'),
                         supp_errors = 'Y',
                         convergence_count = 100)  # More stringent convergence
    
    return(fit)
  }, error = function(e) {
    warning(paste("Colony", colony_id, ": Model fitting failed -", e$message))
    return(NULL)
  })
}

# Group by colony and fit models
colony_fits <- tpc_data %>%
  group_by(ColonyID, Species, Phenotype) %>%
  nest() %>%
  mutate(
    # Fit model with colony ID for better diagnostics
    fit = map2(data, ColonyID, ~fit_tpc_colony(.x, .y)),
    # Check if fit succeeded
    fit_success = !map_lgl(fit, is.null),
    # Extract model parameters
    params = map(fit, possibly(tidy, otherwise = NULL))
  )

```

In Federica's paper, she used the pawar model. 

Check fits
```{r}
# Report fitting success
fit_summary <- colony_fits %>%
  group_by(fit_success) %>%
  summarise(n_colonies = n(), .groups = 'drop')

print("\nModel Fitting Summary:")
print(fit_summary)

# Identify failed fits
failed_fits <- colony_fits %>%
  filter(!fit_success) %>%
  select(ColonyID, Species, Phenotype)

if(nrow(failed_fits) > 0) {
  print("\nColonies with failed fits:")
  print(failed_fits)
  
  # Plot data for failed colonies to diagnose
  failed_data <- tpc_data %>%
    filter(ColonyID %in% failed_fits$ColonyID)
  
  if(nrow(failed_data) > 0) {
    print("\nPlotting data for failed colonies:")
    p_failed <- ggplot(failed_data, aes(x = temp, y = rate, color = as.factor(ColonyID))) +
      geom_point(size = 3) +
      geom_line() +
      facet_wrap(~ColonyID, scales = "free_y") +
      theme_bw() +
      labs(x = "Temperature (°C)", 
           y = "Net Calcification",
           title = "Data for Colonies with Failed Model Fits") +
      theme(legend.position = "none")
    print(p_failed)
  }
}
```

Come back and check these failed colony values...6 colonies failing total (211, 213, 221, 226, 233, 3) --with failing meaning not being successfully fit for TPC. 

Continue with successful fits
```{r}
# Continue with only successful fits
colony_fits <- colony_fits %>%
  filter(fit_success)
```

Calculate TPC metrics 
```{r}
# Extract key thermal performance metrics for each colony
tpc_metrics <- colony_fits %>%
  mutate(
    # Calculate thermal optimum (Topt) - temperature at maximum rate
    topt = map_dbl(fit, ~get_topt(.x)),
    
    # Calculate maximum performance (rmax) - maximum rate
    rmax = map_dbl(fit, ~get_rmax(.x)),
    
    # Calculate breadth (temperature range at 50% rmax)
    breadth = map_dbl(fit, possibly(~get_breadth(.x), otherwise = NA)),
    
    # Calculate critical thermal maximum (CTmax)
    ctmax = map_dbl(fit, possibly(~get_ctmax(.x), otherwise = NA)),
    
    # Calculate critical thermal minimum (CTmin)
    ctmin = map_dbl(fit, possibly(~get_ctmin(.x), otherwise = NA)),
    
    # Extract model parameters directly
    # rtref - rate at reference temperature (18°C)
    rtref = map_dbl(fit, ~coef(.x)["r_tref"]),
    
    # e - activation energy (controls rate rise with temp)
    e = map_dbl(fit, ~coef(.x)["e"]),
    
    # eh - deactivation energy (controls rate decline at high temp)
    eh = map_dbl(fit, ~coef(.x)["eh"]),
    
    # th - temperature at which enzyme is 50% deactivated
    th = map_dbl(fit, ~coef(.x)["th"]),
    
    # Calculate Q10 (temperature sensitivity) between 18-28°C
    q10 = map_dbl(fit, possibly(~get_q10(.x, temp_1 = 18, temp_2 = 28), otherwise = NA)),
    
    # Model fit quality metrics
    #r_squared = map_dbl(fit, ~summary(.x)$r.squared),
    aic = map_dbl(fit, ~AIC(.x)),
    residual_se = map_dbl(fit, ~summary(.x)$sigma)
  )

# Create comprehensive metrics table
metrics_table <- tpc_metrics %>%
  select(ColonyID, Species, Phenotype, 
         topt, rmax, ctmax, ctmin, breadth,
         rtref, e, eh, th, q10,
         #r_squared, 
         aic, residual_se) %>%
  arrange(Species, Phenotype, ColonyID)

print("\nComprehensive TPC Metrics:")
print(metrics_table)
```

Summarize 
```{r}
tpc_summary <- tpc_metrics %>%
  group_by(Species, Phenotype) %>%
  summarise(
    n = n(),
    # Thermal optimum
    mean_topt = mean(topt, na.rm = TRUE),
    sd_topt = sd(topt, na.rm = TRUE),
    se_topt = sd_topt / sqrt(n),
    # Maximum rate
    mean_rmax = mean(rmax, na.rm = TRUE),
    sd_rmax = sd(rmax, na.rm = TRUE),
    se_rmax = sd_rmax / sqrt(n),
    # Thermal breadth
    mean_breadth = mean(breadth, na.rm = TRUE),
    sd_breadth = sd(breadth, na.rm = TRUE),
    se_breadth = sd_breadth / sqrt(n),
    # CTmax
    mean_ctmax = mean(ctmax, na.rm = TRUE),
    sd_ctmax = sd(ctmax, na.rm = TRUE),
    se_ctmax = sd_ctmax / sqrt(n),
    # CTmin
    mean_ctmin = mean(ctmin, na.rm = TRUE),
    sd_ctmin = sd(ctmin, na.rm = TRUE),
    se_ctmin = sd_ctmin / sqrt(n),
    # Rate at reference temp
    mean_rtref = mean(rtref, na.rm = TRUE),
    sd_rtref = sd(rtref, na.rm = TRUE),
    # Activation energy
    mean_e = mean(e, na.rm = TRUE),
    sd_e = sd(e, na.rm = TRUE),
    # Deactivation energy
    mean_eh = mean(eh, na.rm = TRUE),
    sd_eh = sd(eh, na.rm = TRUE),
    # Q10
    mean_q10 = mean(q10, na.rm = TRUE),
    sd_q10 = sd(q10, na.rm = TRUE),
    .groups = 'drop'
  )

```

Plot
```{r}
# Create predictions for smooth curves
new_data <- tpc_data %>%
  group_by(ColonyID) %>%
  summarise(temp = seq(min(temp), max(temp), length.out = 100))

# Add predictions from models
predictions <- colony_fits %>%
  select(ColonyID, Species, Phenotype, fit) %>%
  left_join(new_data, by = "ColonyID") %>%
  mutate(pred_rate = map2_dbl(fit, temp, ~predict(.x, newdata = data.frame(temp = .y))))

# Plot fitted curves with data points
ggplot() +
  geom_point(data = tpc_data, aes(x = temp, y = rate, color = Species), 
             alpha = 0.5, size = 2) +
  geom_line(data = predictions, aes(x = temp, y = pred_rate, 
                                     color = Species, group = ColonyID),
            linewidth = 1) +
  facet_wrap(~Phenotype) +
  theme_bw() +
  labs(x = "Temperature (°C)",
       y = "Net Calcification (μmol CaCO₃ cm⁻² h⁻¹)",
       title = "Fitted Thermal Performance Curves by Colony") +
  theme(legend.position = "bottom")
```

Trying another way
```{r}
# TPC Model Diagnostics and Alternative Fitting Approaches
# Diagnose issues with thermal performance curves

library(rTPC)
library(nls.multstart)
library(broom)
library(tidyverse)

# 1. VISUALIZE INDIVIDUAL CURVES TO CHECK MODEL APPROPRIATENESS
# Plot each colony's data to see the shape
plot_individual_tpcs <- function(data, n_cols = 4) {
  ggplot(data, aes(x = temp, y = rate)) +
    geom_point(size = 2, alpha = 0.6) +
    geom_line(alpha = 0.3) +
    facet_wrap(~paste(ColonyID, Species, Phenotype, sep = "\n"), 
               scales = "free_y", ncol = n_cols) +
    theme_bw() +
    labs(x = "Temperature (°C)", 
         y = "Net Calcification Rate",
         title = "Individual Colony TPCs - Check for Thermal Optima") +
    theme(strip.text = element_text(size = 7))
}

# Plot all colonies
print(plot_individual_tpcs(tpc_data, n_cols = 5))

# 2. CHECK TEMPERATURE RANGE FOR EACH COLONY
temp_range_check <- tpc_data %>%
  group_by(ColonyID, Species, Phenotype) %>%
  summarise(
    n_temps = n_distinct(temp),
    min_temp = min(temp),
    max_temp = max(temp),
    temp_span = max_temp - min_temp,
    max_rate = max(rate),
    temp_at_max = temp[which.max(rate)],
    max_at_edge = temp_at_max %in% c(min_temp, max_temp),
    .groups = 'drop'
  )

print("\nTemperature Range Summary:")
print(temp_range_check)

# Flag colonies where max is at edge (suggests incomplete curve)
edge_max <- temp_range_check %>%
  filter(max_at_edge)

if(nrow(edge_max) > 0) {
  cat("\nWARNING: These colonies have peak performance at edge of temp range:\n")
  print(edge_max %>% select(ColonyID, Species, min_temp, max_temp, temp_at_max))
  cat("These may not have a true thermal optimum in the data range.\n")
}

# 3. ASSESS CURVE SHAPE - Does it match Sharpe-Schoolfield assumptions?
assess_curve_shape <- function(data) {
  data %>%
    group_by(ColonyID) %>%
    summarise(
      # Check if there's a peak (unimodal)
      has_peak = any(rate > rate[temp == min(temp)] & 
                     rate > rate[temp == max(temp)]),
      # Check for monotonic increase
      monotonic_increase = all(diff(rate[order(temp)]) >= 0),
      # Check for monotonic decrease  
      monotonic_decrease = all(diff(rate[order(temp)]) <= 0),
      # Coefficient of variation
      cv = sd(rate) / mean(rate),
      .groups = 'drop'
    )
}

curve_shapes <- assess_curve_shape(tpc_data)
print("\nCurve Shape Assessment:")
print(curve_shapes)

cat("\nCurves without clear peak:", sum(!curve_shapes$has_peak), "\n")
cat("Monotonically increasing:", sum(curve_shapes$monotonic_increase), "\n")
cat("Monotonically decreasing:", sum(curve_shapes$monotonic_decrease), "\n")

# 4. ALTERNATIVE MODEL: Try Gaussian model (simpler, more robust)
# Better for curves without strong high-temp decline

fit_gaussian_tpc <- function(data, colony_id = NULL) {
  if(nrow(data) < 4) return(NULL)
  if(sd(data$rate, na.rm = TRUE) == 0) return(NULL)
  if(all(data$rate <= 0)) return(NULL)
  
  tryCatch({
    start_vals <- get_start_vals(data$temp, data$rate, model_name = 'gaussian_1987')
    
    fit <- nls_multstart(rate ~ gaussian_1987(temp = temp, rmax, topt, a),
                         data = data,
                         iter = 1000,
                         start_lower = start_vals - 2,
                         start_upper = start_vals + 2,
                         lower = get_lower_lims(data$temp, data$rate, 
                                                model_name = 'gaussian_1987'),
                         upper = get_upper_lims(data$temp, data$rate, 
                                                model_name = 'gaussian_1987'),
                         supp_errors = 'Y')
    return(fit)
  }, error = function(e) {
    warning(paste("Colony", colony_id, "- Gaussian fit failed"))
    return(NULL)
  })
}

# 5. FIT BOTH MODELS AND COMPARE
colony_fits_comparison <- tpc_data %>%
  group_by(ColonyID, Species, Phenotype) %>%
  nest() %>%
  mutate(
    # Sharpe-Schoolfield fit
    fit_ss = map2(data, ColonyID, ~fit_tpc_colony(.x, .y)),
    ss_success = !map_lgl(fit_ss, is.null),
    ss_aic = map_dbl(fit_ss, possibly(~AIC(.x), otherwise = NA)),
    
    # Gaussian fit
    fit_gauss = map2(data, ColonyID, ~fit_gaussian_tpc(.x, .y)),
    gauss_success = !map_lgl(fit_gauss, is.null),
    gauss_aic = map_dbl(fit_gauss, possibly(~AIC(.x), otherwise = NA)),
    
    # Choose best model by AIC (lower is better)
    best_model = case_when(
      !ss_success & !gauss_success ~ "none",
      !ss_success ~ "gaussian",
      !gauss_success ~ "sharpeschool",
      gauss_aic < ss_aic - 2 ~ "gaussian",  # AIC difference > 2
      TRUE ~ "sharpeschool"
    ),
    
    # Use best fitting model
    fit_best = case_when(
      best_model == "gaussian" ~ fit_gauss,
      best_model == "sharpeschool" ~ fit_ss,
      TRUE ~ list(NULL)
    )
  )

# Report model comparison
model_comparison_summary <- colony_fits_comparison %>%
  group_by(best_model) %>%
  summarise(n = n(), .groups = 'drop')

print("\nModel Selection Summary:")
print(model_comparison_summary)

# 6. CALCULATE METRICS BASED ON BEST MODEL
# For Sharpe-Schoolfield models
calc_ss_metrics <- function(fit) {
  if(is.null(fit)) return(NULL)
  
  tibble(
    topt = get_topt(fit),
    rmax = get_rmax(fit),
    ctmax = tryCatch(get_ctmax(fit), error = function(e) NA),
    ctmin = tryCatch(get_ctmin(fit), error = function(e) NA),
    breadth = tryCatch(get_breadth(fit), error = function(e) NA),
    rtref = coef(fit)["r_tref"],
    e = coef(fit)["e"],
    eh = coef(fit)["eh"],
    th = coef(fit)["th"],
    aic = AIC(fit),
    model = "sharpeschool"
  )
}

# For Gaussian models (fewer parameters)
calc_gauss_metrics <- function(fit) {
  if(is.null(fit)) return(NULL)
  
  tibble(
    topt = coef(fit)["topt"],
    rmax = coef(fit)["rmax"],
    ctmax = NA,  # Gaussian doesn't estimate this
    ctmin = NA,  # Gaussian doesn't estimate this
    breadth = tryCatch(get_breadth(fit), error = function(e) NA),
    rtref = NA,
    e = NA,
    eh = NA,
    th = NA,
    aic = AIC(fit),
    model = "gaussian"
  )
}

# Calculate metrics for all colonies
metrics_improved <- colony_fits_comparison %>%
  mutate(
    metrics = case_when(
      best_model == "sharpeschool" ~ map(fit_ss, calc_ss_metrics),
      best_model == "gaussian" ~ map(fit_gauss, calc_gauss_metrics),
      TRUE ~ list(NULL)
    )
  ) %>%
  filter(!is.null(metrics[[1]])) %>%
  select(ColonyID, Species, Phenotype, best_model, metrics) %>%
  unnest(metrics)

print("\nImproved Metrics (first 20):")
print(head(metrics_improved, 20))

# 7. COMPARE METRICS TO IDENTIFY OUTLIERS
outlier_check <- metrics_improved %>%
  group_by(Species, Phenotype) %>%
  mutate(
    topt_z = abs(topt - mean(topt, na.rm = TRUE)) / sd(topt, na.rm = TRUE),
    ctmax_z = abs(ctmax - mean(ctmax, na.rm = TRUE)) / sd(ctmax, na.rm = TRUE),
    is_outlier = topt_z > 3 | ctmax_z > 3
  ) %>%
  ungroup()

outliers <- outlier_check %>%
  filter(is_outlier) %>%
  select(ColonyID, Species, Phenotype, topt, ctmax, topt_z, ctmax_z)

if(nrow(outliers) > 0) {
  cat("\nStatistical outliers detected:\n")
  print(outliers)
}

# 8. EXPORT IMPROVED METRICS
write.csv(metrics_improved, "tpc_metrics_improved.csv", row.names = FALSE)
write.csv(model_comparison_summary, "model_selection_summary.csv", row.names = FALSE)

# 9. RECOMMENDATIONS
cat("\n=== RECOMMENDATIONS ===\n")
cat("1. Inspect the individual TPC plots above\n")
cat("2. Colonies with max at edge may need more extreme temperatures\n")
cat("3. Consider using Gaussian model for colonies without clear thermal decline\n")
cat("4. Suspect CTmin < 10°C and CTmax > 40°C for tropical corals\n")
cat("5. Check 'tpc_metrics_improved.csv' for model-appropriate metrics\n")
```






